### 2025-05-09 hotOS/sample-sigconf-authordraft.tex
hotOS paper

%%
%% This is file `sample-sigconf-authordraft.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `all,proceedings,bibtex,authordraft')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf-authordraft.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigconf,10pt]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmlicensed}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}
%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation email}{June 03--05,
  2018}{Woodstock, NY}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
\acmISBN{978-1-4503-XXXX-X/2018/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%\usepackage{todonotes}
\usepackage{booktabs}
\usepackage{multirow}
 \usepackage{proof}
\usepackage{array}
\newcolumntype{O}[1]{>{\raggedright\arraybackslash}m{#1}}
\newcolumntype{P}[1]{>{\centering\arraybackslash}m{#1}}
\newcolumntype{Q}[1]{>{\raggedleft\arraybackslash}m{#1}}
\newcolumntype{S}{>{\centering\arraybackslash}m{2cm}}
% Useful macros from Isaac Oscar Gariano: a very simple library for author-specific change notations:
%  After calling \newAuthor{<name>}{<colour>}
% You can use \<name>{<text>} to add a comment
% Or \<name>A{<text>} to  write stuff in your colour (to denote additional text)
% Or \<name>B{<text>} is similar, but changes the background (to highlight stuff?)
% I may add more commands later
% (there's a 'changes' package with similar functionality, but it has a habit of breaking when you use it with  the Define environment)
\newcommand\newAuthor[2]{%
        \ExpandArgs{c}\newcommand{#1}[1]{\todo[color=#2!10, bordercolor=#2, inline]{\textbf{#1:}~##1}}
	\ExpandArgs{c}\newcommand{#1A}[1]{\textcolor{#2}{##1}}
        \ExpandArgs{c}\newcommand{#1B}[1]{\colorbox{#2!30}{\textcolor{black}{##1}}}}
\newtheorem*{remark}{Remark}
\newcommand{\blackdiamond}{\rotatebox[origin=c]{45}{$\vcenter{\hbox{$\scriptscriptstyle\blacksquare$}}$}}

\newAuthor{Ismail}{blue}
\newAuthor{Colin}{cyan}

%% The "title" command has an optional parameter,
\usepackage{amsthm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{On Verification Patterns for Low-Level Systems via Modal Abstractions }

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
% Commentout
%\author{Ben Trovato}
%\authornote{Both authors contributed equally to this research.}
%\email{trovato@corporation.com}
%\orcid{1234-5678-9012}
%\author{G.K.M. Tobin}
%\authornotemark[1]
%\email{webmaster@marysville-ohio.com}
%\affiliation{%
 % \institution{Institute for Clarity in Documentation}
 % \city{Dublin}
 % \state{Ohio}
 % \country{USA}
%}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Kuru and Gordon}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
 Although they differ in the functionality they offer, low-level systems exhibit certain patterns of design and utilization of computing resources.

  In this paper, we argue the position that \emph{modalities}, in the sense of modal logic, 
  should be a go-to approach when specifying and verifying low-level systems code.
  We explain how the concept of a \emph{resource context} helps guide the design of new modalities
  for verification of systems code, and
  we justify our perspective by discussing prior systems that have used modalities for systems verification
  successfully, arguing that they fit into the verification design pattern we articulate,
  and explaining how this approach might apply to other systems verification challenges.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
%\begin{CCSXML}
%<ccs2012>
% <concept>
%  <concept_id>00000000.0000000.0000000</concept_id>
%  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
%  <concept_significance>500</concept_significance>
% </concept>
% <concept>
%  <concept_id>00000000.00000000.00000000</concept_id>
%  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
%  <concept_significance>300</concept_significance>
% </concept>
% <concept>
%  <concept_id>00000000.00000000.00000000</concept_id>
%  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
%  <concept_significance>100</concept_significance>
% </concept>
% <concept>
%  <concept_id>00000000.00000000.00000000</concept_id>
%  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
%  <concept_significance>100</concept_significance>
% </concept>
%</ccs2012>
%\end{CCSXML}
%
%\ccsdesc[500]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
%\ccsdesc[300]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
%\ccsdesc{Do Not Use This Code~Generate the Correct Terms for Your Paper}
%\ccsdesc[100]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
%
%%%
%%% Keywords. The author(s) should pick words that accurately describe
%%% the work being presented. Separate the keywords with commas.
%\keywords{Do, Not, Us, This, Code, Put, the, Correct, Terms, for, Your, Paper}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
%\begin{teaserfigure}
%  \includegraphics[width=\textwidth]{sampleteaser}
%  \caption{Seattle Mariners at Spring Training, 2010.}
%  \Description{Enjoying the baseball game from the third-base
%  seats. Ichiro Suzuki preparing to bat.}
%  \label{fig:teaser}
%\end{teaserfigure}

%\received{20 February 2007}
%\received[revised]{12 March 2009}
%\received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle
\section{Introduction}
\label{sec:intro}
Systems verification has attracted much attention from both the verification community and the systems community over the past couple of decades. %There are two main streams of verification efforts on systems. In one class of verification, \emph{automated} verification tools are used. They allow people to specify systems via specific annotation languages they offer, and incorporate translation of these specifications into SMT-like formulas for which proofs are searched by the SMT solvers at the backend ~\cite{}. 
Most of this effort has been build on top of \emph{program logics}~\cite{reynolds2002separation,hoare1969axiomatic,smans2012implicit,pratt1976semantical},
which are formal logics of reasoning about program behaviour --- given any starting state
satisfying a particular precondition, running a certain code segment of interest should not fail and should produce a final state satisfying a desired postcondition.
This requires an interacting family of pieces, including a specification language
to describe those desired pre- and postconditions, as well as rules of the logic to deal with
how program operations like variable or register assignments and memory accesses interact with pre- and postconditions.
Since most low-level systems are stateful, \emph{separation logic} has been the focus 
of most recent work.
%has been either the foundational or direct choice of logic in these verification efforts. 
%\Colin{Talking about separation logic makes lots of sense. Do we care about Iris/Coq/Lean/etc. vs. automation? Using the modality idea
%for something like Dafny would certainly look different from the Iris work, but can't it work?
%e.g. \url{https://csgordon.github.io/publications/agere19/}
%I don't want to push to plug my own work if you don't think it makes sense or are making a different point. But I think it's worth
%considering how broadly the big ideas can be applied.
%}

%Advancements in separation logic have been drastic ~\cite{}. These advancements, such as the capability of introducing \emph{ghost} (logical) bookkeeping and associating this bookkeeping with the concept of \emph{invariant} in Iris separation logic, increased the expressivity of any logic that is built on top. Expectedly, many logics enjoying these strong \emph{separation logic} capabilities have been built to verify low-level systems~\cite{}.  %All these program logics built on top of Iris customize and utilize two main aspects of a separation logic like Iris.
%%\paragraph{Custom-Tailored Weakest-Precondition (WP)}
%%\todo[inline]{speak about crash modality etc, encoding how the state should get changed ... GC as well}
%%\paragraph{An Essential Step in Operational Semantics}
%%\todo[inline]{speak about crash step in operational semantics. GC as well}
%%Regarding generalization of customized weakest-preconditions, there has been foundational generalizations defined over incorporating the knowledge of how the resources change, i.e.,transformation of resources, ~\cite{nextgenmodality}, into the separation logic resource assertions. This allows encoding the many customized weakest-preconditions as an instance of a generalized modality definition and lifting of the composition of resource transformation knowledge, which itself is defined as a resource algebra, to the ordinary separation logic resource assertions so that all the underlying resource algebra capabilities can be used for the assertions decorated with the resource-transformation modality. However, even with this kind of generalization effort, it still customizes both the weakest-precondition and resource assertions with the version value (\emph{transformation-generation} value) which most probably lacks being generalizable regarding the real systems to be specified with it as it is not always the case that systems take \emph{versioning} as its key aspect. In general, modalities used in program verification ~\cite{}, are either still tightly coupled with the separation logic they are hosted in, i.e., modalities as definitions for hiding which state a certain truth on resources holds, or as part of the underlying logic's machinery, e.g., later modality as logical machinery encoding step-indexing mechanism.
Modern separation logics are often implemented as frameworks (notably \textsc{Iris}~\cite{jung2018iris}), which offers a collection of tools for building custom separation logic variants for resources of interest, modularly. This makes it possible to
reuse significant amounts of work across verifications using memory~\cite{jung2018iris},
disks~\cite{chajed2019verifying}, and other diverse resources.
%Highly customizable logic ~\cite{iris,..} being the basis for building custom-tailored logics for specific verification tasks ~\cite{some sl refs} enabled building clean and intuitive specifications and proof rules. 
When not using such a framework and instead using an existing tool~\cite{leino2010dafny,vazou2014refinement} as a basis, one can often customize the assertion language by defining libraries of specification definitions and lemmas to effectively have a custom logic.

With these options comes the question: if you can customize your verification tool and how it expresses program states, \emph{how should} you customize it?
We argue that a go-to methodology should be to define \emph{modalities},
which it turns out are in active use in a variety of systems papers because that approach seems to naturally fit many problems that arise in verifying systems code
~\cite{tejperennial19,tejthesis, derekrustbelt20, jung2018iris, kuru2024modalabstractionsvirtualizingmemory, garbagecollect, amalreal2024,fsl, fsl++}.

This paper  
 \begin{itemize}
 \item builds an intuition from a systems-verification perspective about the modal concepts that have been used in recent years with modern program logic -- Section \ref{sec:motivation}
\item describes the common structure underlying the modalities used in systems verification work, effectively a high-level design pattern and process for designing a customized specification language, which has not previously been described --- Section \ref{sec:definitions} % to understand the structure of modal abstractions given for different kinds of systems in Table \ref{table:decomposition}
%\item identifies the common patterns in different modal abstractions for different systems
\item gives example specifications to critical steps of different systems, and discuss how modal abstractions look in these specifications
\end{itemize}

\section{Motivation, Background, and Related Work}
\label{sec:motivation}
To explain our ideas clearly, we briefly recap some of the background and themes our ideas build on,
casting them in a certain way to bring out the relevance of our philosophy.
%\begin{table}[h]
%\centering
%\caption{Decomposing Systems}
%\label{table:decompositionsys}
%\begin{tabular}{@{}lcScP{2cm}@{}}
%\toprule
%   & Column & Column                    & Column                   & Column \\ \midrule
%Row & 1      & 2                         & 3                        & $k=a+b+c+d+e+f+g$      \\
%Row & 5      & \multicolumn{2}{c}{\multirow{2}{*}{Because why not}} & 8      \\
%Row & 9      & \multicolumn{2}{c}{}                                 & 12     \\
%Row & 13     & $k=a+b+c+d+e$                        & 15                       & 16     \\ \bottomrule
%\end{tabular}
%\end{table}

\subsection{Resources in Systems Software}
\label{sec:systemsoft}
Systems software, in general, interfaces with an underlying computing architecture such that any software system at any higher level in the software stack can (at least indirectly) utilize the resources of the machine.
The last layer of software before the hardware is naturally critical to the correctness of an overall system,
as essentially all software built on top of it assumes its correctness.
%Assuming verified hardware, this makes the correctness of the system software critical to the rest of the software stack. 
%Because they interface with the hardware, abstractions and their implementations are intricate and error-prone. 
And because hardware is complex and highly diverse, the implementation of those lowest layers
of the software stack is typically intricate and naturally error-prone, despite how critical its correctness is.
Typically systems software has, as a primary focus, the task of \emph{abstracting} from hardware details to simplify 
the construction of higher layers of the stack.

\paragraph{Virtualization}
One of the most common forms of abstraction provided by systems software is \emph{virtualization},
which abstracts the relationship between conceptual and physical computing resources. 
Operating System (OS) kernels virtualize memory locations and quantity (via virtual memory and paging~\cite{denning1970virtual}).
Distributed language runtimes may virtualize addresses, even when processes may migrate across machines~\cite{jul1988fine}.
Filesystems virtualize locations on disk~\cite{rodeh2013btrfs,hitz1994file,rosenblum1992design,bonwick2003zettabyte}.
Programs built on top of the corresponding systems software layer work logically at the level of these virtualized resources,
and it makes sense to specify the systems software directly in terms of those abstractions.

\paragraph{Sharing} is one of the fundamental reasons for kernels to {virtualize} resources,
for creating the illusion of exclusive access to virtualized versions of shared resources, including CPU time (via preemption).
An implementation of sharing relies on two concepts: \emph{indirection} and \emph{coordination}. Indirection is a method that allows the mapping of resources of one type into another. By doing so, it allows treating the limited resources (e.g. hardware resources such as physical memory) as if there are more than they actually are (e.g. via virtual memory addresses and paging). Consequently, virtualization for the sake of ensuring availability requires \emph{sharing} of the resources that exist in reality.

\paragraph{Access by Translation}
Accessing virtualized resources via translation is a common way to virtualize notions of location (e.g.,
virtual memory addresses, inodes or object IDs instead of disk addresses~\cite{bonwick2003zettabyte,hitz1994file,rosenblum1992design,rodeh2013btrfs}).
%Accessing the hardware resources in the presence of sharing means having more resource references (references to the virtualized resources) to access the \emph{mapped} resources that are physical and potentially fewer. 
B-trees~\cite{}, page tables~\cite{}, and related structures both behave like maps, when the corresponding physical resources exist as such
just in a different location.
Control over the lookup process (e.g, handling the case of a missing translation entry) allows for additional flexibility,
such as filling holes in sparse files, or demand paging (both from disk or lazily populating anonymous initially-zeroed mappings).
Although the realization of these maps may differ from a system to system based on the context (and sometimes hardware details),
they are semantically --- logically --- partial maps, worth treating as such in verification.

%\paragraph{Isolation Hand-in-Hand with Sharing}
%\Ismail{put a few words on abstractions such as address space, stack region, snapshots ,etc.}
%\Colin{Not sure what you want here?}

\subsection{Logic of Resources \& Systems Verification }
\label{sec:seplogic}
Separation logic~\cite{reynolds2002separation} is a specialization of Hoare Logic~\cite{hoare1969axiomatic}
to better deal with reasoning about disjoint resources.
Classic Hoare Logic, as used in prominent systems verifications like \textsc{seL4}~\cite{klein2014comprehensive},
deals with pre- and post-conditions. A so-called ``triple''
\[
\{P\}\;C\;\{Q\}
\]
represents an assertion that if the current program state is accurately described by assertion $P$,
then executing the ``command'' (program) $C$ will not crash, and if it terminates then $Q$ will accurately
describe the ending program state.

\paragraph{Ownership} Separation Logic extends Hoare Logic 
with an additional logical connective $\ast$ (called \emph{separating conjunction}) such that $P \ast Q$ asserts that $P$ and $Q$ not only both hold, but apply to
\emph{disjoint} portions of the program state.\footnote{Though much like systems design, modern separation logics~\cite{jung2018iris}
allow this separation to apply to a kind of virtualization of physical resources~\cite{jensen2012fictional},
or even resources that only exist in the logic~\cite{jung2016higher}.
}
This disjointness assumption leads to discussing separation logic assertions as asserting \emph{ownership}
of the (physical or virtual) resources they hold on.
The classic \emph{points-to} assertion $\ell\mapsto v$ asserts ownership of a portion of memory containing memory address
$\ell$, and that the memory at that location contains value $v$. The disjointness from separating conjunction
is useful: if $\ell\mapsto v \ast \ell'\mapsto v'$ is true, it means $\ell$ and $\ell'$ are distinct memory addresses,
because they must hold on disjoint parts of memory.

Modern separation logics~\cite{jung2018iris} allow definition of more nuanced versions of this idea.
A common form of this, which we revisit in Section \ref{sec:definitions}, is the parameterization of this and
other assertions by a kind of context --- a \emph{resource context} --- resulting in assertions like  $\ell \mapsto_{\textsf{Ctx}} v$, conveying the ownership of the resource defined by the reference $\ell$ and the value ($v$) it's referring to, but where the \emph{resource context}
of the assertion may have the name ($\ell$) and value ($v$) exist not in a partial map of physical memory locations to values,
but of register names to register values~\cite{chlipala2011mostly,ni2007using},
filesystem block indices to block contents~\cite{chajed2019verifying},
or other kinds of associations.
Thus much of the machinery of modern separation logics comes from shifting the notion of which resource context 
assertions are relative to.

%\Colin{Regarding next paragraph: separation logics are not necessarily foundational (early work was all pen-and-paper,
%automated work with SMT solvers clearly isn't foundational). Unless you mean ``foundational'' in a non-technical sense (in which case, a different word should be used, to avoid confusion).}
%\paragraph{Specification} Separation logics~\cite{} rely on foundational Hoare-Logic when introducing their abstractions: make specific assertions and abstractions appear in the form of pre- and post-conditions, $\{P\} e \{Q\}$, called Hoare Triple. It is also important to note that program logics also enjoy encoding Hoare Triples in terms of weakest-precondition~\cite{dijkstra1975guarded} program logics~\cite{pratt1976semantical}, which is roughly
%\[\textsf{P} \; -\ast \;\textsf{wp}\; e \; \{Q\}\]
%where $-\ast$ can  be intuitively thought of as \textit{"when P is added into the resource context we can have the resource represented with Q"}
%\Colin{\emph{Is} it important to note this? I'm genuinely not sure it is.}

\subsection{Modal Reasoning}
\label{sec:modalbackground}
Modal logics are logics of contingent truth, enabling reasoning about cases where something is or may
be true in alternate circumstances. These alternate circumstances are expressed via modal operators,
which decorate a proposition with those alternate circumstances of its truth. This can be express
conditions related to time~\cite{pnueli1977temporal} (e.g., $\diamond P$ for $P$ being true at some point in the future in \textsc{LTL}), necessity and/or possibility~\cite{harel1979first,pratt1976semantical} (e.g., $[C]P$ indicating that $P$ is true after execution
of the program command $C$ terminates), or knowledge and belief~\cite{halpern1990knowledge,hintikka1962knowledge}
(e.g., $\mathcal{K}_a P$ indicating that node $a$ in a distributed system possesses knowledge of $P$),
or the location \emph{where} something is true~\cite{murphy2007type,gordon2019modal}.

\paragraph{Nominals} 
Some classes of assertions benefit specifically from \emph{naming} the explicit conditions where they are true (as opposed to simply requiring them to be true \emph{somewhere} or \emph{everywhere} as in the most classic modal operators).
This naming generally resembles
the \emph{satisfaction} operator of 
\emph{hybrid logic}~\cite{blackburn1995hybrid,areces2001hybrid}: $@_\iota P$ which evaluates the truth of $P$ at the named (Kripke model) state $\iota$. For this reason we refer to the general idea of naming circumstances explicitly as \emph{nominalization},
even though the examples we discuss are not necessarily actually hybrid logics.

\subsection{Modal Operators in Program Specifications}
The value of modal operators in program specifications is that they permit describing something that is true
within some context(s), without needing to fully describe the context itself. Yet unlike shorthands,
abbreviations, or general definitions, modal operators interact with the logic in systematic --- rather than ad hoc --- ways.

For example, most modal operators $M$ have the property that if $P \Rightarrow Q$, then $M(P) \Rightarrow M(Q)$.
Many modal operators distribute over conjunction, so $M(P\wedge Q)\Rightarrow (M(P) \wedge M(Q))$.
So when a program state or behaviour can be captured by a modality, this interacts nicely with the rest of the logic
in ways that simplify both specifications and proofs. Here we give a few examples of how this is true in
classic work; the next section argues that this is \emph{particularly} true for systems software, because
the concepts and reasoning that arise in low-level software are naturally captured by modalities.

Although we focus on understanding the modal patterns in the program logics themselves for verifying the client programs (Section \ref{sec:definitions}), not in the logical machinery they implement, we think that it sheds light on how the modal operator utilized in the machinery of the program logics shows resemblance and can be lifted to the task of client verification.

\paragraph{Temporal Operators}
The best known class of modal logics in the systems community is undoubtedly temporal logics (whether \textsc{LTL}~\cite{pnueli1977temporal}, \textsc{CTL}~\cite{emerson1982using}, \textsc{TLA}~\cite{lamport1994temporal}, or others),
due in part to Lamport's influence~\cite{lamport2002specifying} leading to its not-infrequent use in specifying
distributed algorithms (e.g., \cite{ongaro2014search,ailijiang2019wpaxos}).

\paragraph{Nominalization}
An example utilization of state naming explicitly on the assertion appears in program logics such as Iris~\cite{jung2018iris}, which enables encoding of usage protocols (e.g. state transition system) resembling typestate~\cite{strom1986typestate,garcia2014foundations} as specifications. Protocol assertions are \emph{annotated with the name of the last (abstract) state at which the protocol is ensured}.
An example more familiar to the systems community is Halpern et al.'s history of adapting modal logics of
knowledge to deal with distributed systems~\cite{halpern1990knowledge,halpern1989modelling,halpern2017reasoning,halpern1990knowledge}.
In most of that line of work, $\mathcal{K}_a(P)$ indicates that the node $a$ in the system \emph{knows} or \emph{possesses knowledge of} $P$
(for example, a Raft node may ``know'' a lower bound on the commit index).
Alternatively, a modality $@_i(P)$ may represent that $P$ is true of/at the specific node $i$~\cite{gordon2019modal} (e.g., that node $i$ has stored a certain
piece of data to reliable storage).
These permit capturing specific concepts relevant to the correctness (and reasoning about correctness) of a certain class of systems,
involving facts about specific named entities in the system.
\\

In each of these cases, the fact that these facts are described using modalities means, for example, that if a
process $p$ knows that the commit index is greater than 5 (e.g., $\mathcal{K}_p(\mathsf{commitIndex} > 5)$)
no extra work is required to conclude that the node knows it is greater than 3 (i.e., that $\mathcal{K}_p(\mathsf{commitIndex} > 3)$),
because this follows from standard properties of modal operators as described above.
In contrast, if verification instead used a custom assertion $\mathsf{minCommitIndex}(5)$ to represent the former knowledge,
one would need to separately provide custom reasoning to conclude $\mathsf{minCommitIndex}(3)$.

%\paragraph{Implementing the Logic of Impredicativity} Modal operators such as \emph{later} modality speculating over the fact that modal assertion holds at some state later in the execution are used in the logical machinery of program logics to handle the impredicative terms in the logic. Interestingly, similar reasoning principles using \emph{later} modality are also used in the verification of programs against store-buffers ~\cite{} which also requires speculation over the state in which stored writes appear in the memory.
%
%\paragraph{Persistent Weakest-Precondition} We already gave a rough encoding of Hoare Triple in terms of weakest-precondition, but a thorough version of our definition for the program logics of resources would be 
%\[
%\Box (\textsf{P} \; -\ast \;\textsf{wp}\; e \; \{Q\})
%\]
%decorated with a necessity modality ensures that the weakest-precondition assertion 
%$
%    \textsf{P} \; -\ast \;\textsf{wp}\; e \; \{Q\}
%$
%inside the necessity modality is always valid by the exclusive ownership of the weakest-precondition; as a consequence, the whole specification 
%$
%\Box (\textsf{P} \; -\ast \;\textsf{wp}\; e \; \{Q\})
%$ can be used multiple times -- i.e., a \emph{duplicable resource}.

\section{Logical Decomposition of a System into its Constituents \emph{Contingently}}
\label{sec:definitions}
\begin{table*}[h]
\centering
\caption{Modal Decomposition of Program-Logics.}
\label{table:decomposition}
 %$\{\Diamond (P) \}$ write  $\{ P\}_{n}$
 %$\{\overset{ICut^{n}}{\hookrightarrow} (P) \}$ return $\{ P\}_{n} $
\begin{tabular}{@{}cSSSp{5cm}@{}}
\toprule
& Resource Context & Resource Elements  &  Nominalization & Resource Context Steps \\ \midrule
Post-Crash Modality ~\cite{tejthesis, perennialgit,tejperennial19} & $\Diamond \; P $  & $  \ell \mapsto_{n}^{\overline{\gamma}} v $ & Strong &  Crash Recovery   \\ 
%GC-Modality &      & \multicolumn{2}{c}{\multirow{2}{*}{Because why not}} & 8                                   & & ~\cite{}\\
NextGen Modality~\cite{larsnextgen25} &  $\overset{t}{\hookrightarrow} \; P$  & Own (t(a))  & Strong & Determined Based on the Model$^{*}$   \\
StackRegion  Modality$^{*}$  ~\cite{larsnextgen25} & $\overset{ICut^{n}}{\hookrightarrow} \; P$      & $\fbox{n} \; \ell \mapsto v$                                 & Strong & Alloc and Return to/from stack\\
Memory-Fence  Modality ~\cite{fsl,fsl++,derekrustbelt20} & $\triangle_{\pi}$ and $\triangledown_{\pi}$      & $\ell \mapsto v$                                 & Weak & Fence Acquire and Release  \\
Address-Space Modality ~\cite{kuru2024modalabstractionsvirtualizingmemory} & [r]P     &  $\ell \mapsto v$ & Weak                         & Address-Space Switch  \\ 
Ref-Count Modality ~\cite{amalreal2024}& @$_{\ell}$ P    &  $\ell_1 \mapsto v$ & Weak  & Allocating, Dropping and Sharing a Reference \\ \bottomrule
\end{tabular}

*The StackRegion Modality is an instance of NextGen (called the Independence Modality in \cite{larsnextgen25}).
%and the $\Diamond$ Figure for Post-Crash Modality uses notationis taken from NextGen Modality paper ~\cite{larsnextgen25}.
\end{table*}

In this section, we explain how we can systematically propose new modal abstractions for different types of systems.
We summarize our discussion of these logics in Table \ref{table:decomposition}.
We discuss, based on examples, common aspects of how we intuitively think about correctness of systems code in many contexts,
articulate those pieces, and call out the commonalities across a range of systems.

Consider first the address-space abstraction in an OS kernel. An address-space of a process is a \textsf{container} of \textsf{virtual} addresses referencing data in memory. One would expect to have \textit{points-to} assertion from separation logic to specify \textit{ownership} of a memory reference pointing to some data. But that ownership is relative to a specific
address space --- a specific container. We tend to think directly about what is true \emph{in an address space},
with the simplest piece being an association between a virtual address and the data it points to.
We call the simplest piece, in this and other examples, the \emph{resource element}:

\begin{definition}[Resource Elements]
%Inherited from separation logic, we identify the resource element as the indivisible logical element on which contingency defined by the modality holds. 
The simplest atomic facts we want to work with in a particular setting, specific to that setting.
\end{definition}
By definition, the resource elements are specific to some limited domain or setting.
For example, knowing that a certain address points to a 32-bit signed integer representing 3 is knowledge restricted to a certain address space.
In general,  we call these domains that any resource element is tied to \emph{resource contexts}:
\begin{definition}[Resource Context]
A resource context is an abstraction, context, or container of resource elements of the same type, e.g., an address space of a process.
\end{definition}
We discuss a range of examples for each of these in turn.

\subsection{Resource Elements}
Table \ref{table:decomposition} gives additional examples of systems and their corresponding resource elements,
though none of the work in that table analyzes itself according to the structure we are giving.
When reasoning about stack frame contents, the resource element would be a stack-memory points-to assertion ( $\fbox{n} \; \ell \mapsto v$)
indicating that a certain offset into stack region $n$ holds value $v$.
%which is an instance of ownership assertion ($\textsf{Own(t(a))}$) for which a well-behaved transformation ($\textsf{t}$), e.g. $\textsf{ICut}^{n}$ algebra quantifying over valid stack region ids. 
For virtual memory management, a \textit{virtual-points-to} ownership assertion pairing a virtual address ($\ell$) with data ($v$) in an address space is natural.
When considering weak memory models, we also want points-to information (address-value mappings).
When dealing with reference-counting APIs, we may care to specify reachability of memory nodes ($\ell \mapsto v$) in a certain context defined by a shared root address .
A generalization of separation logic to deal with arbitrary transformations (the \emph{NextGen} modality) 
%\Colin{Ismail, not sure how to describe this. We don't want to go into the weeds on the NextGen}
%that can be jumped from a shared memory referenced whose reference count is at least 1 are also example resource elements from memory management task at different abstraction levels. 
The resource element of Post-Crash Modality is not obvious in Table \ref{table:decomposition}, and needs a bit of explanation. Perennial, based on the Iris logic, has both disk-points-to assertions $d[\textsf{p}] \mapsto_{n} v$ (for a specific disk $d$) and in-memory points-to assertions $\ell \mapsto_{n}^{\overline{\gamma}}$. Perennial crash-recovery logic book-keeps resource names (can be thought of as logical variables) $\overline{\gamma}$ to identify 
which assertions (resource elements) remain valid after a crash --- these assertions
are only usable while the names in $\overline{\gamma}$ are valid, and a crash resets
them, discarding assumptions about volatile state.
%whose capability is considered to be relevant to be kept in the crash invariant $\fbox{C}_n$, and to be used when a crash occurs.\Colin{Ismail, not sure about last line above. Can you better explain gamma?}
%regions of state (including auxiliary/ghost state) which persist after a crash (anything not identified

%Perennial keeps the consistent state of persistent storage in a crash invariant at : multiple disk-points-to assertions $disk[\textsf{p}] \mapsto_{n} v$ in the \emph{crash} invariant $\fbox{C}_n$ became consistent at the n$^{\textsf{th}}$ successful completion of writes, i.e., generation number, from in-memory to disk. The in-memory updates ($p \mapsto_{n} v$) are also indexed with the generation number $n$. Since Perennial is built on top of Iris Separation Logic which implicitly decorates ownership assertions -- both disk and in-memory points-to assertions -- with ghost names ($\gamma$), Perennial utilizes this and bookkeeps points-to assertions ($\ell \mapsto_{n}^{\overline{\gamma}}$) of a certain version number $n$ as \emph{a resource element lifted by resource names} needed in case of a crash to distinguish the resources to be invalidated, i.e., discard the piece of the physical state that resides in the memory and preserve the resources that reside in the persistent disk.

%\Colin{I don't understand this next bit, we need to discuss:}
%\begin{remark}[Resource Element Names]
%Regarding the names, like Iris resource names, all the resource elements in the second column have names; they are not subject to change --- \emph{static} --- when they are not parameterized by names and mentioned explicitly in the element when they do not matter for the contingent truth.
%\end{remark}
%\Colin{This in particular would benefit from an earlier intuition discussion}

\subsection{Resource Contexts}
Except for Post-Crash-Modality, one can think of the resource contexts in the first column in Table \ref{table:decomposition} as containers for the corresponding resource elements in the second column. For our example, a process's address space with the root address \textsf{r} is an abstraction that is treated as a container for virtual address mappings, $\ell \mapsto v$; likewise, a shared memory address $\ell$ can be the root of the graph that can be a container of memory nodes ($\ell_1 \mapsto v$) that are reachable from the root $\ell$. 
\looseness=-1

A subtlety of the notion of a resource context is that, unlike the earlier examples, the context does not need to be a literal data structure. It can instead be (various forms of) a set of executions, as in the Post-Crash and NextGen modalities.
The Post-Crash modality $\diamond$ expresses that the assertion $P$ will be true after a crash discards
all unstable storage (i.e., RAM).
This was the inspiration for the NextGen modality, which is in fact a framework for defining ``after-$t$'' modalities, where $t$ is an transformation of the global state.\footnote{The transformations are subject to some
technical constraints that are unrelated to our point here.}

%\begin{definition}[Transforming Resource Elements]
%    Any resource element that is parameterized w.r.t to a parameter such as resource \emph{names} $\overline{\gamma}$ in Post-Crash Modality is subject to the change, and this change has to be acknowledged in the resource context.
%\end{definition}
% Since we have the resource elements of Post-Crash modality parameterized by names as shown in Table \ref{table:decomposition}, the modality has to acknowledge how the resource names plugged into the resource elements relate to the names that already exist globally --- Perennial utilizes this for invalidating the points-to in-memory assertions whose names are not part of the plugged-in resource names. To do so, Post-Crash Modality resource context $\Diamond$ has a relation (ignoring the universally quantified machine states Perennial needs as it utilizes $\Diamond$ for the crash-aware Hoare triple not client specification/proof) with universally quantified new names, $\overline{\gamma'}$, to be related to the old global names and plugged into the parameterized resource elements to obtain post-crash values of the elements $\lambda \; \overline{\gamma} \ldotp P $ via disjunctive choice on the new and old values presented by the resource-element transform relation ($R$)
%\[\Diamond \; P \triangleq  \forall_{\overline{\gamma'}} \ldotp R(\_,\overline{\gamma},\overline{\gamma'}) -\ast P (\overline{\gamma'}) \ast R(\_,\overline{\gamma},\overline{\gamma'})\]
%This, in fact, inspires NextGen Modality to encode resource-element transform relation as a resource algebra, and make the resource element generalized w.r.t. a \emph{well-behaved} transformations defined by the rules of resource algebra. 
%\paragraph{Dynamically Naming Resources} Names are global and each resource element has a name. Unlike others in Table \ref{table:decomposition} whose resources are fixed (inspired by Iris, each points-to relation has a hidden name  $\gamma$), Post-Condition Modality resource element is parameterized by the underlying resources' names. Unlike fixed %naming, dynamic naming in which resource elements are parameterized by the resource names, the knowledge of which resource name is transforming because an
%any other resource elements named with the ones th

\subsection{Resource Ambiance}
While the central aspects of using a modality to capture contextual reasoning are the elements and contexts, there is \emph{sometimes} more to it.
Sometimes there are multiple modes of execution (e.g., crashing, running in a particular address space),
so not only state descriptions but also rules of the program logic itself must be decorated with
a \emph{resource ambiance}:
\begin{definition}[Resource Ambiance]
Resource ambiance is a dependency of the logic's description of program behaviour on
a particular mode of execution tied to the modality.
%\emph{general form of contingent} status of resource elements of the same type against their resource context. The crucial aspect is to understand how \emph{identifiable} both the context and its elements should be: \emph{nominalization} of the context and its elements in the ambient logic, e.g., how should stack-points-to relations together with the StackRegion Modality should be \emph{referred} in a Hoare triple for Return step in the ambient separation logic.
\end{definition}
In formal presentations, this typically shows up with modified Hoare triples $\{P\}\;C\;\{Q\}_A$,
where $P$, $C$, and $Q$ are the precondition, code, and postcondition that code satisfies assuming that precondition (as before) and $A$ is essentially a marker of how the reasoning about the program interacts with the modality --- such as a post-crash assertion, or an indication of what address space the verification is assuming.
This means the proof rules for deciding how program operations affect knowledge of program state
can depend on, restrict, or manipulate the context assumptions of reasoning about program behaviour, not just
descriptions of program state at some moment in time.
Not all use cases for modal assertions require an ambiance, only those where the modality corresponds to modes of execution.

\subsection{Nominalization}
Finally, another design point is the question of whether or not resource elements must explicitly track
their corresponding context, or if they implicitly pick up their context from where they are used.

\emph{Strong nominalization} is the case where resource elements must explicitly include
the identity of their intended context, while \emph{weak nominalization} occurs when the resource
elements implicitly pick up the relevant context from how they are used.
The first three modalities in Table \ref{table:decomposition} are strongly nominalized,
with the resource elements generally carrying identifiers of a specific modality usage.\footnote{The Post-Crash modality does not look like this in the presentation here; technically the definition of $\diamond$ quantifies over names $\overline{\gamma}$ internally, dealing with sets of possible contexts.}
The last three are weakly nominalized.

This choice trades off complexity against flexibility and scoping constraints.
Strongly nominalized elements track additional specifier/prover-visible book-keeping data.
But in exchange for carrying those identifiers of their context with them, 
strongly nominalized elements can be used together under any modality. For example, one can
use the StackRegion modality to talk about two different stack frames simultaneously for code
which accesses multiple stack frames: $\fbox{n} \; \ell \mapsto v \ast \fbox{n+1} \; \ell' \mapsto v'$.
Using a given strongly nominalized assertion element under different modalities for different
frames does not change its meaning.

By contrast, weakly nominalized elements are more concise, but then make talking about multiple 
contexts together marginally more complex: changing which modality an assertion is used with drastically changes its meaning. In the case of the Ref-Count modality, $@_\ell(\ell_1\mapsto v)$
says that $\ell$ points to a reference count wrapping $\ell_1$, while placing the $\ell_q\mapsto v$ under a jump modality for a different location entails talking about a different region of memory.

In general, use cases where code frequently manipulates small parts of multiple contexts together
should prefer strong nominalization, while use cases where usually larger portions of a single context
are at issue should prefer weak nominalization.

%$\paragraph{Strong Nominalization} Some resource elements require strong \emph{association} with their contexts as \emph{up-to} proper identification of resource context. That means, not only does the resource context have to be identified with a nominal, but also resource elements need to be associated with some nominals that are related to the context nominal, because certain actions in operational semantics require \emph{identification} of the resource context and elements together. For a stack with $m$ regions, returning from a i$^{th}$ index roughly (ignoring the evaluation context issues related to the underlying logic):
%\[
%\infer{\textsf{wp} \; return \; cont^{i}(K') \{Q\}_n}{%
%\begin{array}{l}
%    \textsf{ExistsRegion} \; m  \quad\quad   n <= m - i  \quad\quad i <= m  \\
%\textsf{ExistsRegion} \; (m-i) -\ast \; \overset{.}{\hookrightarrow}^{m-i} (\textsf{wp} \; K'[shift \; i] \;\{Q\}_n)  \end{array}
%}
%\]
%asserting that with proper stack region lower bound (n), and the index falling to the proper range, knowing that the next required precondition ($(\textsf{wp} \; K'[shift \; i] \;\{Q\}_n)$) is independent of any stack points-to ($\fbox{m-i} \;\ell \mapsto v$) above the region $m-i$ by the StackRegion Modality (as an instance of NextGen Modality)  $\overset{.}{\hookrightarrow}^{m-i}$ (Independence Modality ~\cite{larsnextgen25}). Here, we see that both the resource context (Region Modality) and the resource elements (stack points-to assertions) nominalized, and they are aware of each other. Regarding the Post-Crash Modality, since Perennial utilizes it for extending Hoare triple encoding of Perennial with the crash recovery ( \textsf{wpc} ~\cite{tejperennial19, tejthesis}) such that in case a crash occurs on a disk-access, Hoare triple decorated with the crash recovery condition utilizes the capabilities in the crash invariant for the n$^{th}$ successful disk state (what NextGen ~\cite{larsnextgen25} calls \emph{generation} number). Since Post-Crash Modality is parameterized with the resource names ($\overline{\gamma}$), based on these resource names, the in-memory updates ($p \mapsto^{\gamma}_n v$) are ignored and only relevant successful disk updates ($d[p] \mapsto^{\gamma'}_n v$) are considered to create the consistent crash invariant n+1, $\fbox{C}_{n+1}$. Since NextGen Modality ~\cite{larsnextgen25} generalizes Perennial, lifting Post-Crash Modality into the client specification would require \emph{strong nominalization} as the proof would require the guarantees similar to the StackRegion Modality (Independence Modality ~\cite{larsnextgen25}) asserting the updates on disk are \emph{independent} of any in-memory updates.
%
%%When we look at the third column, we see the modal program logics that require strong normalization. Thinking of in-memory updates to be persistent on the disk, Perennial ~\cite{tejperennial19} decorates the in-memory ownership capabilities, which is resource elements of the  (e.g. $p \mapsto_{n} v_0$) with generation numberbrs 
%%For example, the resource elements in the separation logic utilizing \textsf{NextGen} modality and the modality itself are subject to strong normalization: both weakest-precondition and the ownership assertions are nominalized with region index such that the ownerships are valid as long as the constraint on the region index is satisfied -- e.g., all the ownership assertions with region index greater than the one on the weakest-precondition are valid against a crash-step in a filesystem or garbage collecting step in another system updating the consistent disk or reclaimed memory layout. \todo[inline]{maybe paraphrase more on referring them in the ambient logic}.
%
%\paragraph{Weak Context Nominalization} Unlike the resources we discussed so far, some resource elements are \emph{agnostic} to the resource context they belong. That means the resource context needs a nominalization as it needs to be distinguished from other resource contexts when certain actions in the operational semantics are taken (e.g., address-space switch in Table \ref{table:decomposition}), but the resource elements of the context are unaware of this context nominalization, and they may even not be labeled with any other nominal as well. For example, virtual memory references in an address space do not know which address space they belong, but the address spaces are distinguished w.r.t. the unique memory address of the root page-table tree therefore, a virtual-address-points-to ownership assertion does not need to be aware (or nominalized with) of the nominal of the address-space which is the unique memory address of the address table. Logically, specifying an address-space switch ~\cite{kuru2024modalabstractionsvirtualizingmemory} from the address space with the root address $r2$ to the one with the root address $r1$
%\[
%\{[r1] P \ast Q \} cr3 := r1 \{P \ast [r2]Q\}_{\textsf{cr3}}
%\]
%requires having the resource elements to be loaded into the memory 
%\[P \triangleq \ell_1 \mapsto v ...\]
%to be in the resource context of the address-space modality with the nominal $r1$, and, once the new address space is loaded --- i.e., cr3 is loaded with the other address-space root address ---, the resource elements ($Q$) that were loaded in the previous view of the memory are now introduced to the resource context with the nominal $r2$. Unlike what we discuss in the strong normalization (e.g. StackModality in Return specification) in which nominalized resource elements (stack points-to assertions) are guarded by Independence Modality ~\cite{larsnextgen25} (StackModality row in Table \ref{table:decomposition}) to preserve the soundness of the proof rule. However, with weak-nominalization, resource elements don't have nominals (Address-Space Modality Resource Element Column in Table \ref{table:decomposition}), and most importantly, address-space resource context switch, unlike shifted stack region in Return Specification, does not require any explicit independence check because all the resource elements at any time are either in one of two resource contexts or hosted in the ambient logic.

%\begin{definition}[Transforming Resource Context in Resource Ambience]
%    For the examples in this paper, a Resource ambiance is always a Hoare Triple indexed with the resource context nominals, which hosts the resource contexts. Unlike the transformation of resource elements hosted by a resource context (\textbf{Definition 3.3.}) handled by a relation in the modality definition, e.g. the resource element transformation relation in Post-Crash modality, updating the resource contexts themselves depends on the operational semantics action ( the last column in Table \ref{table:decomposition}) a resource ambiance is operating on.
%\end{definition}
%The address-space switch updating the cr3 register that holds the nominal value for address-space modality's resource contexts and the resource ambiance's (Hoare Triple indexed with resource context nominal) is an example of updating a resource context in a resource ambiance.
%\paragraph{Composition of Contingent Truth} Another example of weak-nominalization is Ref-Counter Modality (jump modality ($@_{l} \; P$ ~\cite{amalreal2024}) shown in Table \ref{table:decomposition}. The resource context $@_{l} \; P$ is a container of memory nodes (i.e. the resource elements ($\ell \mapsto v$ in Table \ref{table:decomposition}) that are \emph{reachable} by a jump over the shared/duplicable root address $\ell$. In addition to being weakly nominalized due to its agnostic resource elements, it exhibits a good example of how one contingent truth \emph{entails} another. Wagner et.al., also define \emph{reachability} modality --- a monad style modality for specification of non-destructive actions ($\blackdiamond$) ~\cite{amalreal2024} --- to assert the reachability to a certain memory node. Knowing $P$ satisfied under \emph{jump} ~\cite{amalreal2024} modality (\textsf{Ref-Counter Modality} in \ref{table:decomposition}) entails the satisfaction of $P$ under non-destructive reachability modality
%\[ 
%@_{\ell} \;P \vdash  \blackdiamond \; P 
%\]
%\begin{definition}[Resource Interaction]
%A resource interaction rule defines a particular way to interact with a resource context to access resource elements. A resource interaction is defined by a proof rule with the relevant resource context in the specification. 
%\end{definition}
%Although it is almost always the case that when a modality is defined, classical proof rules such as introduction and elimination are also defined and proven. However, an interesting perspective emerges when we realize that, in fact, the proof rule for switching address spaces resembles an introduction (also elimination) rule for adding resource elements into an address-space modality's resource context.
%\Colin{So far I like how this is shaping up as a kind of ``design recipe'' for approaching this style of verification!}

\subsection{Making It Work}
While the choice of resource elements, contexts, ambiance, and nominalization to capture
the setting and contextually-relevant pieces of information help lay out the 
shape of specifications based on how we would like to think about them, there are a few more steps
beyond designing the specifications.

It is necessary to work out how proof rules interact with the modalities --- which program
operations or hardware primitives update machine state in a way that interacts with the modality?
This could be all or some, depending on the use case.

It is also necessary to implement the approach. This design has been most thoroughly explored
with logics embedded in proof assistants (everything in Table \ref{table:decomposition} uses \textsf{Coq}), in which case a framework like \textsc{Iris}~\cite{jung2018iris} (used for all of our examples) will require proofs that each rule interacting with the modality is sound, but in exchange
for that work and a few supplementary definitions, modal assertions can work naturally with existing tools for the logic~\cite{krebbers2018mosel}.
However, it is also suitable for use with automated techniques, for example by writing
new axioms corresponding to proof rules~\cite{gordon2019modal}, though better automation may require more work.

%\section{Related Work}
%\label{sec:relatedwork}
%We discussed

\section{Conclusion}
\label{sec:conclusion}

We have argued that modalities should be a go-to design when working out how to specify different kinds of systems code based on recent successes, and articulated an informal design process that seems to 
map out the degrees of variation present among those designs.
It also captures (a slightly more organized version of) our own thought processes while
coming up with designs for specifying distributed systems~\cite{gordon2019modal}, virtual memory managers~\cite{kuru2024modalabstractionsvirtualizingmemory}, and in ongoing work using different modalities for different parts of a copy-on-write filesystem (e.g., for assertions true in different snapshots). So not only are the modalities useful
for conducting proofs, the pieces we have identified seem to be an effective way of working out
a specific modality for a specific use-case.

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
%\begin{acks}
%To Robert, for the bagels and explaining CMYK and color spaces.
%\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{software}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\end{document}
\endinput
%%
%% End of file `sample-sigconf-authordraft.tex'.

### 2025-05-09 hotOS/sample-sigconf.tex
hotOS paper

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `all,proceedings,bibtex,sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigconf]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{\let\Bbbk\relax
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmlicensed}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}
%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation email}{June 03--05,
  2018}{Woodstock, NY}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
\acmISBN{978-1-4503-XXXX-X/2018/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{The Name of the Title Is Hope}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Ben Trovato}
\authornote{Both authors contributed equally to this research.}
\email{trovato@corporation.com}
\orcid{1234-5678-9012}
\author{G.K.M. Tobin}
\authornotemark[1]
\email{webmaster@marysville-ohio.com}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \city{Dublin}
  \state{Ohio}
  \country{USA}
}

\author{Lars Th{\o}rv{\"a}ld}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \city{Hekla}
  \country{Iceland}}
\email{larst@affiliation.org}

\author{Valerie B\'eranger}
\affiliation{%
  \institution{Inria Paris-Rocquencourt}
  \city{Rocquencourt}
  \country{France}
}

\author{Aparna Patel}
\affiliation{%
 \institution{Rajiv Gandhi University}
 \city{Doimukh}
 \state{Arunachal Pradesh}
 \country{India}}

\author{Huifen Chan}
\affiliation{%
  \institution{Tsinghua University}
  \city{Haidian Qu}
  \state{Beijing Shi}
  \country{China}}

\author{Charles Palmer}
\affiliation{%
  \institution{Palmer Research Laboratories}
  \city{San Antonio}
  \state{Texas}
  \country{USA}}
\email{cpalmer@prl.com}

\author{John Smith}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \city{Hekla}
  \country{Iceland}}
\email{jsmith@affiliation.org}

\author{Julius P. Kumquat}
\affiliation{%
  \institution{The Kumquat Consortium}
  \city{New York}
  \country{USA}}
\email{jpkumquat@consortium.net}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  A clear and well-documented \LaTeX\ document is presented as an
  article formatted for publication by ACM in a conference proceedings
  or journal publication. Based on the ``acmart'' document class, this
  article presents and explains many of the common variations, as well
  as many of the formatting elements an author may use in the
  preparation of the documentation of their work.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>00000000.0000000.0000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc[300]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc[100]{Do Not Use This Code~Generate the Correct Terms for Your Paper}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Do, Not, Us, This, Code, Put, the, Correct, Terms, for,
  Your, Paper}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \includegraphics[width=\textwidth]{sampleteaser}
  \caption{Seattle Mariners at Spring Training, 2010.}
  \Description{Enjoying the baseball game from the third-base
  seats. Ichiro Suzuki preparing to bat.}
  \label{fig:teaser}
\end{teaserfigure}

\received{20 February 2007}
\received[revised]{12 March 2009}
\received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
ACM's consolidated article template, introduced in 2017, provides a
consistent \LaTeX\ style for use across ACM publications, and
incorporates accessibility and metadata-extraction functionality
necessary for future Digital Library endeavors. Numerous ACM and
SIG-specific \LaTeX\ templates have been examined, and their unique
features incorporated into this single new template.

If you are new to publishing with ACM, this document is a valuable
guide to the process of preparing your work for publication. If you
have published with ACM before, this document provides insight and
instruction into more recent changes to the article template.

The ``\verb|acmart|'' document class can be used to prepare articles
for any ACM publication --- conference or journal, and for any stage
of publication, from review to final ``camera-ready'' copy, to the
author's own version, with {\itshape very} few changes to the source.

\section{Template Overview}
As noted in the introduction, the ``\verb|acmart|'' document class can
be used to prepare many different kinds of documentation --- a
double-anonymous initial submission of a full-length technical paper, a
two-page SIGGRAPH Emerging Technologies abstract, a ``camera-ready''
journal article, a SIGCHI Extended Abstract, and more --- all by
selecting the appropriate {\itshape template style} and {\itshape
  template parameters}.

This document will explain the major features of the document
class. For further information, the {\itshape \LaTeX\ User's Guide} is
available from
\url{https://www.acm.org/publications/proceedings-template}.

\subsection{Template Styles}

The primary parameter given to the ``\verb|acmart|'' document class is
the {\itshape template style} which corresponds to the kind of publication
or SIG publishing the work. This parameter is enclosed in square
brackets and is a part of the {\verb|documentclass|} command:
\begin{verbatim}
  \documentclass[STYLE]{acmart}
\end{verbatim}

Journals use one of three template styles. All but three ACM journals
use the {\verb|acmsmall|} template style:
\begin{itemize}
\item {\texttt{acmsmall}}: The default journal template style.
\item {\texttt{acmlarge}}: Used by JOCCH and TAP.
\item {\texttt{acmtog}}: Used by TOG.
\end{itemize}

The majority of conference proceedings documentation will use the {\verb|acmconf|} template style.
\begin{itemize}
\item {\texttt{sigconf}}: The default proceedings template style.
\item{\texttt{sigchi}}: Used for SIGCHI conference articles.
\item{\texttt{sigplan}}: Used for SIGPLAN conference articles.
\end{itemize}

\subsection{Template Parameters}

In addition to specifying the {\itshape template style} to be used in
formatting your work, there are a number of {\itshape template parameters}
which modify some part of the applied template style. A complete list
of these parameters can be found in the {\itshape \LaTeX\ User's Guide.}

Frequently-used parameters, or combinations of parameters, include:
\begin{itemize}
\item {\texttt{anonymous,review}}: Suitable for a ``double-anonymous''
  conference submission. Anonymizes the work and includes line
  numbers. Use with the \texttt{\string\acmSubmissionID} command to print the
  submission's unique ID on each page of the work.
\item{\texttt{authorversion}}: Produces a version of the work suitable
  for posting by the author.
\item{\texttt{screen}}: Produces colored hyperlinks.
\end{itemize}

This document uses the following string as the first command in the
source file:
\begin{verbatim}
\documentclass[sigconf]{acmart}
\end{verbatim}

\section{Modifications}

Modifying the template --- including but not limited to: adjusting
margins, typeface sizes, line spacing, paragraph and list definitions,
and the use of the \verb|\vspace| command to manually adjust the
vertical spacing between elements of your work --- is not allowed.

{\bfseries Your document will be returned to you for revision if
  modifications are discovered.}

\section{Typefaces}

The ``\verb|acmart|'' document class requires the use of the
``Libertine'' typeface family. Your \TeX\ installation should include
this set of packages. Please do not substitute other typefaces. The
``\verb|lmodern|'' and ``\verb|ltimes|'' packages should not be used,
as they will override the built-in typeface families.

\section{Title Information}

The title of your work should use capital letters appropriately -
\url{https://capitalizemytitle.com/} has useful rules for
capitalization. Use the {\verb|title|} command to define the title of
your work. If your work has a subtitle, define it with the
{\verb|subtitle|} command.  Do not insert line breaks in your title.

If your title is lengthy, you must define a short version to be used
in the page headers, to prevent overlapping text. The \verb|title|
command has a ``short title'' parameter:
\begin{verbatim}
  \title[short title]{full title}
\end{verbatim}

\section{Authors and Affiliations}

Each author must be defined separately for accurate metadata
identification.  As an exception, multiple authors may share one
affiliation. Authors' names should not be abbreviated; use full first
names wherever possible. Include authors' e-mail addresses whenever
possible.

Grouping authors' names or e-mail addresses, or providing an ``e-mail
alias,'' as shown below, is not acceptable:
\begin{verbatim}
  \author{Brooke Aster, David Mehldau}
  \email{dave,judy,steve@university.edu}
  \email{firstname.lastname@phillips.org}
\end{verbatim}

The \verb|authornote| and \verb|authornotemark| commands allow a note
to apply to multiple authors --- for example, if the first two authors
of an article contributed equally to the work.

If your author list is lengthy, you must define a shortened version of
the list of authors to be used in the page headers, to prevent
overlapping text. The following command should be placed just after
the last \verb|\author{}| definition:
\begin{verbatim}
  \renewcommand{\shortauthors}{McCartney, et al.}
\end{verbatim}
Omitting this command will force the use of a concatenated list of all
of the authors' names, which may result in overlapping text in the
page headers.

The article template's documentation, available at
\url{https://www.acm.org/publications/proceedings-template}, has a
complete explanation of these commands and tips for their effective
use.

Note that authors' addresses are mandatory for journal articles.

\section{Rights Information}

Authors of any work published by ACM will need to complete a rights
form. Depending on the kind of work, and the rights management choice
made by the author, this may be copyright transfer, permission,
license, or an OA (open access) agreement.

Regardless of the rights management choice, the author will receive a
copy of the completed rights form once it has been submitted. This
form contains \LaTeX\ commands that must be copied into the source
document. When the document source is compiled, these commands and
their parameters add formatted text to several areas of the final
document:
\begin{itemize}
\item the ``ACM Reference Format'' text on the first page.
\item the ``rights management'' text on the first page.
\item the conference information in the page header(s).
\end{itemize}

Rights information is unique to the work; if you are preparing several
works for an event, make sure to use the correct set of commands with
each of the works.

The ACM Reference Format text is required for all articles over one
page in length, and is optional for one-page articles (abstracts).

\section{CCS Concepts and User-Defined Keywords}

Two elements of the ``acmart'' document class provide powerful
taxonomic tools for you to help readers find your work in an online
search.

The ACM Computing Classification System ---
\url{https://www.acm.org/publications/class-2012} --- is a set of
classifiers and concepts that describe the computing
discipline. Authors can select entries from this classification
system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
commands to be included in the \LaTeX\ source.

User-defined keywords are a comma-separated list of words and phrases
of the authors' choosing, providing a more flexible way of describing
the research being presented.

CCS concepts and user-defined keywords are required for for all
articles over two pages in length, and are optional for one- and
two-page articles (or abstracts).

\section{Sectioning Commands}

Your work should use standard \LaTeX\ sectioning commands:
\verb|\section|, \verb|\subsection|, \verb|\subsubsection|,
\verb|\paragraph|, and \verb|\subparagraph|. The sectioning levels up to
\verb|\subsusection| should be numbered; do not remove the numbering
from the commands.

Simulating a sectioning command by setting the first word or words of
a paragraph in boldface or italicized text is {\bfseries not allowed.}

Below are examples of sectioning commands.

\subsection{Subsection}
\label{sec:subsection}

This is a subsection.

\subsubsection{Subsubsection}
\label{sec:subsubsection}

This is a subsubsection.

\paragraph{Paragraph}

This is a paragraph.

\subparagraph{Subparagraph}

This is a subparagraph.

\section{Tables}

The ``\verb|acmart|'' document class includes the ``\verb|booktabs|''
package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
high-quality tables.

Table captions are placed {\itshape above} the table.

Because tables cannot be split across pages, the best placement for
them is typically the top of the page nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and the
table caption.  The contents of the table itself must go in the
\textbf{tabular} environment, to be aligned properly in rows and
columns, with the desired horizontal and vertical rules.  Again,
detailed instructions on \textbf{tabular} material are found in the
\textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table~\ref{tab:freq} is included in the input file; compare the
placement of the table here with the table in the printed output of
this document.

\begin{table}
  \caption{Frequency of Special Characters}
  \label{tab:freq}
  \begin{tabular}{ccl}
    \toprule
    Non-English or Math&Frequency&Comments\\
    \midrule
    \O & 1 in 1,000& For Swedish names\\
    $\pi$ & 1 in 5& Common in math\\
    \$ & 4 in 5 & Used in business\\
    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  \bottomrule
\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of the page's
live area, use the environment \textbf{table*} to enclose the table's
contents and the table caption.  As with a single-column table, this
wide table will ``float'' to a location deemed more
desirable. Immediately following this sentence is the point at which
Table~\ref{tab:commands} is included in the input file; again, it is
instructive to compare the placement of the table here with the table
in the printed output of this document.

\begin{table*}
  \caption{Some Typical Commands}
  \label{tab:commands}
  \begin{tabular}{ccl}
    \toprule
    Command &A Number & Comments\\
    \midrule
    \texttt{{\char'134}author} & 100& Author \\
    \texttt{{\char'134}table}& 300 & For tables\\
    \texttt{{\char'134}table*}& 400& For wider tables\\
    \bottomrule
  \end{tabular}
\end{table*}

Always use midrule to separate table header rows from data rows, and
use it only for this purpose. This enables assistive technologies to
recognise table headers and support their users in navigating tables
more easily.

\section{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of the three are
discussed in the next sections.

\subsection{Inline (In-text) Equations}
A formula that appears in the running text is called an inline or
in-text formula.  It is produced by the \textbf{math} environment,
which can be invoked with the usual
\texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
and structures, from $\alpha$ to $\omega$, available in
\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
examples of in-text equations in context. Notice how this equation:
\begin{math}
  \lim_{n\rightarrow \infty}x=0
\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsection{Display Equations}
A numbered display equation---one set off by vertical space from the
text and centered horizontally---is produced by the \textbf{equation}
environment. An unnumbered display equation is produced by the
\textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols and
structures available in \LaTeX\@; this section will just give a couple
of examples of display equations in context.  First, consider the
equation, shown as an inline equation above:
\begin{equation}
  \lim_{n\rightarrow \infty}x=0
\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}
  \sum_{i=0}^{\infty} x + 1
\end{displaymath}
and follow it with another numbered equation:
\begin{equation}
  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\section{Figures}

The ``\verb|figure|'' environment should be used for figures. One or
more images can be placed within a figure. If your figure contains
third-party material, you must clearly identify it as such, as shown
in the example below.
\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{sample-franklin}
  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
    Ewing, Inc. [Public domain], via Wikimedia
    Commons. (\url{https://goo.gl/VLCRBB}).}
  \Description{A woman and a girl in white dresses sit in an open car.}
\end{figure}

Your figures should contain a caption which describes the figure to
the reader.

Figure captions are placed {\itshape below} the figure.

Every figure should also have a figure description unless it is purely
decorative. These descriptions convey what’s in the image to someone
who cannot see it. They are also used by search engine crawlers for
indexing images, and when images cannot be loaded.

A figure description must be unformatted plain text less than 2000
characters long (including spaces).  {\bfseries Figure descriptions
  should not repeat the figure caption – their purpose is to capture
  important information that is not already provided in the caption or
  the main text of the paper.} For figures that convey important and
complex new information, a short text description may not be
adequate. More complex alternative descriptions can be placed in an
appendix and referenced in a short figure description. For example,
provide a data table capturing the information in a bar chart, or a
structured list representing a graph.  For additional information
regarding how best to write figure descriptions and why doing this is
so important, please see
\url{https://www.acm.org/publications/taps/describing-figures/}.

\subsection{The ``Teaser Figure''}

A ``teaser figure'' is an image, or set of images in one figure, that
are placed after all author and affiliation information, and before
the body of the article, spanning the page. If you wish to have such a
figure in your article, place the command immediately before the
\verb|\maketitle| command:
\begin{verbatim}
  \begin{teaserfigure}
    \includegraphics[width=\textwidth]{sampleteaser}
    \caption{figure caption}
    \Description{figure description}
  \end{teaserfigure}
\end{verbatim}

\section{Citations and Bibliographies}

The use of \BibTeX\ for the preparation and formatting of one's
references is strongly recommended. Authors' names should be complete
--- use full first names (``Donald E. Knuth'') not initials
(``D. E. Knuth'') --- and the salient identifying features of a
reference should be included: title, year, volume, number, pages,
article DOI, etc.

The bibliography is included in your source document with these two
commands, placed just before the \verb|\end{document}| command:
\begin{verbatim}
  \bibliographystyle{ACM-Reference-Format}
  \bibliography{bibfile}
\end{verbatim}
where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
suffix, of the \BibTeX\ file.

Citations and references are numbered by default. A small number of
ACM publications have citations and references formatted in the
``author year'' style; for these exceptions, please include this
command in the {\bfseries preamble} (before the command
``\verb|\begin{document}|'') of your \LaTeX\ source:
\begin{verbatim}
  \citestyle{acmauthoryear}
\end{verbatim}

  Some examples.  A paginated journal article \cite{Abril07}, an
  enumerated journal article \cite{Cohen07}, a reference to an entire
  issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
  monograph/whole book in a series (see 2a in spec. document)
  \cite{Harel79}, a divisible-book such as an anthology or compilation
  \cite{Editor00} followed by the same example, however we only output
  the series if the volume number is given \cite{Editor00a} (so
  Editor00a's series should NOT be present since it has no vol. no.),
  a chapter in a divisible book \cite{Spector90}, a chapter in a
  divisible book in a series \cite{Douglass98}, a multi-volume work as
  book \cite{Knuth97}, a couple of articles in a proceedings (of a
  conference, symposium, workshop for example) (paginated proceedings
  article) \cite{Andler79, Hagerup1993}, a proceedings article with
  all possible elements \cite{Smith10}, an example of an enumerated
  proceedings article \cite{VanGundy07}, an informally published work
  \cite{Harel78}, a couple of preprints \cite{Bornmann2019,
    AnzarootPBM14}, a doctoral dissertation \cite{Clarkson85}, a
  master's thesis: \cite{anisi03}, an online document / world wide web
  resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game
  (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05}
  and (Case 3) a patent \cite{JoeScientist001}, work accepted for
  publication \cite{rous08}, 'YYYYb'-test for prolific author
  \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
  contain 'duplicate' DOI and URLs (some SIAM articles)
  \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
  multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
  couple of citations with DOIs:
  \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
  citations: \cite{TUGInstmem, Thornburg01, CTANacmart}.
  Artifacts: \cite{R} and \cite{UMassCitations}.

\section{Acknowledgments}

Identification of funding sources and other support, and thanks to
individuals and groups that assisted in the research and the
preparation of the work should be included in an acknowledgment
section, which is placed just before the reference section in your
document.

This section has a special environment:
\begin{verbatim}
  \begin{acks}
  ...
  \end{acks}
\end{verbatim}
so that the information contained therein can be more easily collected
during the article metadata extraction phase, and to ensure
consistency in the spelling of the section heading.

Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.

\section{Appendices}

If your work needs an appendix, add it before the
``\verb|\end{document}|'' command at the conclusion of your source
document.

Start the appendix with the ``\verb|appendix|'' command:
\begin{verbatim}
  \appendix
\end{verbatim}
and note that in the appendix, sections are lettered, not
numbered. This document has two appendices, demonstrating the section
and subsection identification method.

\section{Multi-language papers}

Papers may be written in languages other than English or include
titles, subtitles, keywords and abstracts in different languages (as a
rule, a paper in a language other than English should include an
English title and an English abstract).  Use \verb|language=...| for
every language used in the paper.  The last language indicated is the
main language of the paper.  For example, a French paper with
additional titles and abstracts in English and German may start with
the following command
\begin{verbatim}
\documentclass[sigconf, language=english, language=german,
               language=french]{acmart}
\end{verbatim}

The title, subtitle, keywords and abstract will be typeset in the main
language of the paper.  The commands \verb|\translatedXXX|, \verb|XXX|
begin title, subtitle and keywords, can be used to set these elements
in the other languages.  The environment \verb|translatedabstract| is
used to set the translation of the abstract.  These commands and
environment have a mandatory first argument: the language of the
second argument.  See \verb|sample-sigconf-i13n.tex| file for examples
of their usage.

\section{SIGCHI Extended Abstracts}

The ``\verb|sigchi-a|'' template style (available only in \LaTeX\ and
not in Word) produces a landscape-orientation formatted article, with
a wide left margin. Three environments are available for use with the
``\verb|sigchi-a|'' template style, and produce formatted output in
the margin:
\begin{description}
\item[\texttt{sidebar}:]  Place formatted text in the margin.
\item[\texttt{marginfigure}:] Place a figure in the margin.
\item[\texttt{margintable}:] Place a table in the margin.
\end{description}

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
To Robert, for the bagels and explaining CMYK and color spaces.
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\section{Research Methods}

\subsection{Part One}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
lacinia dolor. Integer ultricies commodo sem nec semper.

\subsection{Part Two}

Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
eros. Vivamus non purus placerat, scelerisque diam eu, cursus
ante. Etiam aliquam tortor auctor efficitur mattis.

\section{Online Resources}

Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
enim maximus. Vestibulum gravida massa ut felis suscipit
congue. Quisque mattis elit a risus ultrices commodo venenatis eget
dui. Etiam sagittis eleifend elementum.

Nam interdum magna at lectus dignissim, ac dignissim lorem
rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
massa et mattis lacinia.

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.

### 2025-06-02 paper.tex
springer

--- 2025-05-09 hotOS/sample-sigconf-authordraft.tex
@@
-%%
-%% This is file `sample-sigconf-authordraft.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `all,proceedings,bibtex,authordraft')
-%% 
-%% IMPORTANT NOTICE:
-%% 
-%% For the copyright see the source file.
-%% 
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigconf-authordraft.tex.
-%% 
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%% 
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%%
-%% Commands for TeXCount
-%TC:macro \cite [option:text,text]
-%TC:macro \citep [option:text,text]
-%TC:macro \citet [option:text,text]
-%TC:envir table 0 1
-%TC:envir table* 0 1
-%TC:envir tabular [ignore] word
-%TC:envir displaymath 0 word
-%TC:envir math 0 word
-%TC:envir comment 0 0
-%%
-%% The first command in your LaTeX source must be the \documentclass
-%% command.
-%%
-%% For submission and review of your manuscript please change the
-%% command to \documentclass[manuscript, screen, review]{acmart}.
-%%
-%% When submitting camera ready or to TAPS, please change the command
-%% to \documentclass[sigconf]{acmart} or whichever template is required
-%% for your publication.
-%%
-%%
-\documentclass[sigconf,10pt]{acmart}
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    Bib\TeX}}}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmlicensed}
-\copyrightyear{2018}
-\acmYear{2018}
-\acmDOI{XXXXXXX.XXXXXXX}
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[Conference acronym 'XX]{Make sure to enter the correct
-  conference title from your rights confirmation email}{June 03--05,
-  2018}{Woodstock, NY}
-%%
-%%  Uncomment \acmBooktitle if the title of the proceedings is different
-%%  from ``Proceedings of ...''!
-%%
-%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%%  June 03--05, 2018, Woodstock, NY}
-\acmISBN{978-1-4503-XXXX-X/2018/06}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% For managing citations, it is recommended to use bibliography
-%% files in BibTeX format.
-%%
-%% You can then either use BibTeX with the ACM-Reference-Format style,
-%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
-%% support for advanced citation of software artefact from the
-%% biblatex-software package, also separately available on CTAN.
-%%
-%% Look at the sample-*-biblatex.tex files for templates showcasing
-%% the biblatex styles.
-%%
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-%\usepackage{todonotes}
+%\documentclass[a4paper,UKenglish]{lipics-v2016}
+% For two-column layout for PPoPP, change 'acmsmall' to 'sigplan'.  Probably also turn on the
+% 'review' option, and the 'anonymous' option if it's DBR
+\documentclass[runningheads]{llncs}
+%This is a template for producing LIPIcs articles.
+%See lipics-manual.pdf for further information.
+%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
+%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
+% for section-numbered lemmas etc., use "numberwithinsect"
+\usepackage[export]{adjustbox}
+
 \usepackage{booktabs}
@@
 \newcolumntype{S}{>{\centering\arraybackslash}m{2cm}}
-% Useful macros from Isaac Oscar Gariano: a very simple library for author-specific change notations:
-%  After calling \newAuthor{<name>}{<colour>}
-% You can use \<name>{<text>} to add a comment
-% Or \<name>A{<text>} to  write stuff in your colour (to denote additional text)
-% Or \<name>B{<text>} is similar, but changes the background (to highlight stuff?)
-% I may add more commands later
-% (there's a 'changes' package with similar functionality, but it has a habit of breaking when you use it with  the Define environment)
-\newcommand\newAuthor[2]{%
-        \ExpandArgs{c}\newcommand{#1}[1]{\todo[color=#2!10, bordercolor=#2, inline]{\textbf{#1:}~##1}}
-	\ExpandArgs{c}\newcommand{#1A}[1]{\textcolor{#2}{##1}}
-        \ExpandArgs{c}\newcommand{#1B}[1]{\colorbox{#2!30}{\textcolor{black}{##1}}}}
-\newtheorem*{remark}{Remark}
-\newcommand{\blackdiamond}{\rotatebox[origin=c]{45}{$\vcenter{\hbox{$\scriptscriptstyle\blacksquare$}}$}}
-
-\newAuthor{Ismail}{blue}
-\newAuthor{Colin}{cyan}
-
-%% The "title" command has an optional parameter,
-\usepackage{amsthm}
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
+
+\usepackage{microtype}%if unwanted, comment out or use option "draft"
+%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
+%\usepackage{t1enc}
+%\usepackage[strings]{underscore}
+%%%%%\usepackage{geometry}
+
+\usepackage[rawfloats=true]{floatrow} 
+\restylefloat{figure}    
+\DeclareMathSizes{4}{4}{4}{4}
+\usepackage{varwidth}
+\usepackage{lmodern}
+\usepackage{amsfonts}
+\usepackage[fleqn]{amsmath}
+\usepackage{mathtools}
+\usepackage{amssymb}
+\usepackage{hyperref}
+\usepackage{proof}
+\usepackage{stmaryrd}
+\usepackage{xspace}
+\usepackage{listings}
+\usepackage{comment}
+\usepackage{longtable}
+\usepackage{float}
+\usepackage[section]{placeins}
+\usepackage[toc,page]{appendix}
+\usepackage[export]{adjustbox}
+\usepackage{amssymb}
+\usepackage{graphics,graphicx}
+\usepackage{mathpartir}
+\usepackage{array}
+\usepackage{tikz}
+\usetikzlibrary{shapes.geometric}
+\usetikzlibrary{arrows.meta,arrows}
+\usetikzlibrary{positioning}
+\usetikzlibrary{calc, shapes, backgrounds}
+\usepackage{verbatim}
+\usepackage{tikz-qtree}
+\usepackage{pgfplots}
+\usepackage{caption,capt-of}
+\usepackage{color}
+\usepackage{soul}
+\setul{1ex}{0.8ex}
+\definecolor{orange}{rgb}{1,0.5,0}
+\setulcolor{orange}
+\usepackage{subcaption}
+\usepackage{lipsum}
+\usepackage{afterpage}
+%%sequeezeeee
+\addtolength{\textfloatsep}{-10pt}
+\addtolength{\dbltextfloatsep}{-10pt}
+
+\newcommand{\defeq}{\mathbin{\stackrel{\text{def}}{=}}}
+% some stuff for prettier proof outlines
+\newcommand{\havoc}{\textsf{havoc}}
+\newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
+\newcommand{\assertion}[1]{\textsf{assert} {\color{red}\left\{#1\right\}}}
+\newcommand{\proofstepnr}[5][1pt/3pt]{\small \left.
+\begin{array}{l}
+\hdashline[#1]
+\ensuremath{\specline{#2}}\\
+\begin{array}{l}
+#3
+\end{array}\\
+
+\ensuremath{\specline{#5}}\\
+\hdashline[1pt/3pt]
+\end{array}
+\color{green}\right)\hspace{-6pt}{\color{green}-}
+\begin{array}{@{}l@{}}
+#4
+\end{array}
+}
+\newcommand{\proofoutstepnr}[5][1pt/3pt]{\small \left.
+\begin{array}{@{}l@{}}
+\hdashline[#1]
+\ensuremath{\specline{#2}}\\
+\begin{array}{@{\;\;\;}l}
+#3
+\end{array}\\
+\ensuremath{\specline{#5}}\\
+\hdashline[1pt/3pt]
+\end{array}
+\color{green}\right)\hspace{-6pt}{\color{green}-}
+\begin{array}{@{}l@{}}
+#4
+\end{array}
+}
+% \proofoutstep[dash stuff]{pre}{C}{rule}{post}
+\newcommand{\proofoutstep}[5][1pt/3pt]{
+\proofoutstepnr[#1]{#2}
+{#3}
+{\rotatebox[origin=c]{90}{$\begin{array}{@{}c@{}}{}#4\end{array}$}}
+{#5}
+}
+
+\newcommand{\NULL}{\texttt{NULL}}
+\newcommand{\NEW}{\texttt{new}}
+\newcommand{\SKIP}{\texttt{skip}}
+\newcommand{\RESULT}{\texttt{result}}
+\newcommand{\class}{\texttt{class}}
+\newcommand{\where}{\texttt{where}}
+\newcommand{\readable}{\texttt{readable}}
+\newcommand{\writable}{\texttt{writable}}
+\newcommand{\isolated}{\texttt{isolated}}
+\newcommand{\immutable}{\texttt{immutable}}
+\newcommand{\consume}[1]{\texttt{consume}(#1)}
+\newcommand{\consumeiso}[1]{\mathsf{RemIso}(#1)}
+\newcommand{\INT}{\texttt{int}}
+\newcommand{\bool}{\texttt{bool}}
+\newcommand{\this}{\texttt{this}}
+\newcommand{\immutdecl}[1]{\mathsf{ImmutableDeclared}(#1)}
+\newcommand{\IsoOrImm}{\textsf{IsoOrImm}}
+\newcommand{\NonWrit}{\textsf{NoWrit}}
+\newcommand{\db}[1]{\llbracket#1\rrbracket}
+\newcommand{\nil}{\textsf{null}}
+\newcommand{\atom}{\textsf{atom} \mapsto \mstatename \mapsto \mstatename}
+\newcommand{\loc}{\textsf{Loc}}
+\newcommand{\val}{\textsf{Val}}
+\newcommand{\mstatename}{\textsf{MState}}
+\newcommand{\superset}{\mathcal{P}}
+\newcommand{\mstatedecl}{\sigma \in \mstatename \overset{def}{=} s \times h \times l \times rt \times R \times B  \times F }
+\newcommand{\stack}{s \in \var \times \tid \rightharpoonup \loc}
+\newcommand{\heap}{h \in \loc \times \fname \rightharpoonup \val}
+\newcommand{\wlock}{l \in \tid \uplus \textsf{unlocked}}
+\newcommand{\rlock}{R \in \superset(\tid)}
+\newcommand{\fset}{F : \loc \mapsto \superset(\tid) }
+\newcommand{\atomeval}{\db{-}:\atom}
+\newcommand{\rcuwrite}{\textsf{RCUWrite} \text{ x.f as y in }\, \{ C \}}
+\newcommand{\rcuread}{\textsf{RCURead} \text{ x.f as y in }\, \{ C \}}
+  \newcommand{\desugarread}{ \rcuread \overset{\Delta}{=} \texttt{\textsf{ReadBegin}}; y = x.f;\{C'\}  ;\textsf{ReadEnd} }
+  \newcommand{\desugarwrite}{\rcuwrite \overset{\Delta}{=} \texttt{\textsf{WriteBegin}} ; y = x.f ; \{C'\}  ;\textsf{WriteEnd}}
+    \newcommand{\desugarsync}{\textsf{Sync} \overset{\Delta}{=} \texttt{\textsf{SyncStart};\textsf{SyncStop}}}
+  %\newcommand{\boundpath}{\bar f_{i}^{k} = \forall_{k \in \mathbb{N},i \in \fnameenum} \ldotp  \underbrace{f_{i}^{0} \ldots f_{i}^{k}}_\text{$\mid$ $f_{i}^{0}$ \ldots $f_{i}^{k}$ $\mid$ = k} }
+\newcommand{\subt}{\prec:}
+%typing rules
+ \newcommand{\fname}{\textsf{FName}}
+\newcommand{\fmname}{\textsf{FMap}}
+\newcommand{\uiglobal}{\textsf{UIGlobal}}
+\newcommand{\udef}{\textsf{undef}}
+\newcommand{\unlinked}{\textsf{unlinked}}
+\newcommand{\fresh}{\textsf{fresh}}
+\newcommand{\var}{\textsf{Var}}
+\newcommand{\tid}{\textsf{TID}}
+
+\newcommand{\fmapfieldemp}{\N_{f,\emptyset} = \N \setminus \{f \rightharpoonup \_ \}}
+\newcommand{\fmapname}{\N = \{ f_0|\ldots|f_n\rightharpoonup  \{y\} \mid f_{i} \in \fname \land \; 0\leq i\leq n  \land \; (y \in \var \lor y \in \{null\}) \}}
+\newcommand{\fmapextend}{  \N(f\rightharpoonup x \setminus y) = \N \setminus \{f \rightharpoonup x\} \cup \{f \rightharpoonup y\} }
+\newcommand{\fmapextendn}{  \N(\cup_{f\rightharpoonup y}) = \N \cup \{f \rightharpoonup y\} }
+\newcommand{\fmapdiff}{\N(\setminus_{f\rightharpoonup y}) = \N - \{f\rightharpoonup y\}}
+\newcommand{\fmapwithf}{\N([f\rightharpoonup  y]) = \N \text{   where  } f \rightharpoonup y \in \N }
+\newcommand{\fmapempty}{\N_{\emptyset} =  \{ \}}
+\newcommand{\codom}{\cod(\N)=\dom(\Gamma)}
+\newcommand{\cod}{\texttt{cod}}
+\newcommand{\dom}{\texttt{dom}}
+%\newcommand{\types}{\tau ::= \textsf{rcu} | \textsf{normal}}
+
+\newcommand{\rpth}{f \in \textsf{FName} 
+     \left\{\begin{array}{cl}
+     \epsilon &    \\
+     f  &  f \in \fname\\
+   f_1 | \ldots | f_n &\forall_{1 \leq i \leq n} \ldotp f_i \in \fname 
+     \end{array}\right\}
+     \quad \text{  where } \textsf{FType(f) = \textsf{RCU}}
+}
+\newcommand{\rextendswith}{\rho^{k}[i] \mapsto f_i \text{  where  } \forall_{0\leq i \leq k} \ldotp f_i \in \fname }
+\newcommand{\extendswith}{\rho[i] \mapsto \rho_{i}^{k} \text{  where  } \forall_{0\leq i \leq n} \ldotp \rho_i^{k} \in \textsf{RelativePath}}
+
+\newcommand{\pth}{\rho \in \textsf{Path}  \qquad \rho ::= \rho_1 \ldots\rho_n  \text{   where  } \forall_{0\leq i \leq n}\ldotp \rho_i \in \textsf{RelativePath} \text{ or } \rho_i \in \textsf{ConcretePath}}
+\newcommand{\cpath}{\forall_{0 \leq i \leq n} \ldotp f_i \in \mathsf{FName} \land l_i \in \textsf{Label} \land \exists_{k\geq0} \ldotp f_i^{l_i} = \overbrace{f_i\ldots f_i}^{k} \land (\forall_{0\leq j \leq n} \ldotp j\neq i \implies l_i \neq l_j) }
+
+\newcommand{\rpath}{\rho \in \textsf{RelativePath} \quad \rho = (f)^{*} \text{   where  } f \in \textsf{FName} \text{  and  } \exists_{k>=0} \ldotp (f)^{*} = \overbrace{f\ldots f}^\text{k} }
+\newcommand{\falt}{f \in \textsf{AlternatingFName} = f_1 | f_2 | \ldots | f_{n-1} | f_n \text{  where  } \forall_{0\leq i \leq n}\ldotp f_i \in \textsf{FName} }
+
+\newcommand{\lenmap}{\mathcal{L} \in \textsf{LPath} : \rho \mapsto \mathbb{N}}
+\newcommand{\etadecl}{\eta = \{f \mapsto y \mid f \in \fname \land y\in \var \}  }
+\newcommand{\rcunfd}{\textsf{rcuFresh} \,  \N}
+\newcommand{\rcunfdecl}{x : \textsf{rcuFresh} \;\; \N }
+\newcommand{\rcunf}[3]{#1 : \textsf{rcuFresh} \;\; \N(\cup_{ #2 \rightharpoonup #3}) } %x, f , y
+\newcommand{\rcuf}[1]{#1 : \textsf{rcuFresh} \;\;\N } %x, f , y
+\newcommand{\rcunextfresh}[1]{\textsf{rcuFresh} \;\; \N #1 }
+\newcommand{\ftypedecl}{\textsf{FType} : \fname \mapsto \tau}
+\newcommand{\ftype}[2]{\textsf{FType}(f_{#1}) \mapsto \textsf{#2}}
+%%%
+\newcommand{\typedepthdecl}{\textsf{TypeDepth}(x:\text{T}) =  \mid\rho\mid \;\; \texttt{where T} = \textsf{rcuItr} \underset{y}{\mid} \rho \,\, \N }
+%%%
+\newcommand{\rcuonly}{\textsf{RCUOnly}(x) \overset{\texttt{def}}{=} x:T \text{ where } T = \textsf{rcuItr}\_}
+\newcommand{\rcumodonlydecl}{\textsf{RCUModifiedOnly}(C) \overset{\texttt{def}}{=} \forall_{x \in \textsf{Modifies}(C)}\ldotp \textsf{RcuOnly}(x) }
+\newcommand{\rcufilterb}{\textsf{RemRcu} : \Gamma \mapsto \Gamma}
+\newcommand{\rcufilterr}[1]{\textsf{RemRcu} ( #1 ) = \texttt{filter} \, ( \lambda x \ldotp x \neq \textsf{rcuItr \_ } ) \,  #1 }
+\newcommand{\rcufiltera}[1]{\textsf{RemRcu} ( #1 ) }
+%%%Alias
+\newcommand{\aliaswith}[2]{#1 \in \textsf{AliasWith}(#2)}
+\newcommand{\aliaswithdecl}{\textsf{AliasWith}(x) = \{y \mid y:\textsf{rcuItr}\underset{r',k'}{\mid} \rho' \;\; \N' \land \rho = \rho' \land \N = \N' \}  \texttt{  where  }  \rcuitrT{x}{r}{k}{kk}{\_} }
+%%%Framing Type Contexts
+    \newcommand{\wellframed}{
+      \mathsf{WellFramed}(\Gamma,\Delta)\overset{def}{=}
+      \left\{\begin{array}{l}
+
+        \forall_{z\in dom(\Gamma),f,\rho,y }\ldotp                           z\notin dom(\Delta) \land            z:\textsf{rcuItr}\,\rho\,\N([f\rightharpoonup y]) \implies \\
+                      \left\{ 
+	              \begin{array}{l}
+                        y\notin dom(\Delta) \land y\in dom(\Gamma) \land \forall_{n\in dom(\Delta),\rho',f',m}\ldotp n\notin dom(\Gamma) \land 	\\
+            \left\{
+            \begin{array}{l}
+              n:\textsf{rcuItr}\,\rho'\,\N([f'\rightharpoonup m]) \implies \\
+              \left\{\begin{array}{l}
+              m \notin dom(\Gamma) \land \\
+              m \in dom(\Delta) \land
+              \rho \neq \rho'                    
+              \end{array} \right\}
+            \end{array}\right\}
+     \end{array}\right\}
+         \end{array}\right.
+
+}
+
+%%%Framing Type Contexts ends
+\newcommand{\deltaepsilon}{\Delta >= 0 \qquad 0\leq \epsilon \leq \Delta }
+\newcommand{\reduce}[2]{\textsf{InReduction}(#1,#2) = \{ \Gamma' \mid  \Gamma' \preceq_{\varepsilon} \Gamma \} }
+\newcommand{\rename}[2]{\textsf{ContextRename}(#1,#2) = #2 \mapsto #1 }
+%%Sep RCU
+\newcommand{\seprcu}[4]{\textsf{SepLoop}(#1,#2,#3,#4) \overset{\texttt{def}}{=} \text{dom}(#1)\cap \text{dom}(#2) = \emptyset \, \land \, #1\uplus#2= #4 \, \land \, \textsf{RcuModifiedOnly}(#2,#3)}
+\newcommand{\seprcud}[4]{\textsf{SepLoop}(#1,#2,#3,#4)}
 %%
-%% end of the preamble, start of the body of the document source.
+\newcommand{\seprcub}[5]{\textsf{SepBranch}(#1,#2,#3,#4,#5) \overset{\texttt{def}}{=}
+   \left\{\begin{array}{cl}
+  \text{dom}(#2)\cap \text{dom}(#3) = \emptyset \, \land \, #2\uplus#3=#1 \land \\
+  \textsf{RcuModifiedOnly}(#2,#4) \,\land\, \textsf{RcuModifiedOnly}(#2,#5)\,\land \\
+
+ \end{array}\right\}
+}
+\newcommand{\seprcubd}[5]{\textsf{SepBranch}(#1,#2,#3,#4,#5)}
+%%
+\newcommand{\precgamma}[2]{#1  \preceq #2 \overset{\texttt{def}}{=} \forall_{(x)\in #1}\ldotp x \in #2 \implies  #1 \preceq_{x} #2}
+%%%if branches
+\newcommand{\iffree}{\textsf{if}(\text{e})\{\text{C}\}}
+\newcommand{\ifelse}{\textsf{if}(\text{e})\{\text{C}_1\}\textsf{else}\{\text{C}_2\}}
+%%var  root,k  , k+1 mappedvar
+\newcommand{\rcuitertype}[4]{\textsf{rcuItr}(#1,#2)\;#3\;#4}
+\newcommand{\rcuit}[4]{#1:\textsf{rcuItr}\;\; #3 \;\; #4 }
+\newcommand{\rcuitrT}[5]{#1:\textsf{rcuItr}\;\; \rho \;\; \N }
+\newcommand{\rcuitrn}[5]{#1:\textsf{rcuItr}\;\; \rho \, f \;\; \N }
+\newcommand{\rcuitr}[5]{#1:\textsf{rcuItr}\;\; \rho \;\; \N[f  \rightharpoonup #5] }
+\newcommand{\rcustack}[2]{#1 : \textsf{rcuItr}\underset{#2}{\mid}}
+\newcommand{\afree}[1]{\textsf{asyncDelayedFree}(#1)}
+\newcommand{\allocate}{\textsf{asyncDelayedFree}}
+\newcommand{\typelocal}{\tau_{local}}
+\newcommand{\typeglobal}{\tau_{global}::=\textsf{rcuRoot}  }
+\newcommand{\rcutypes}{ \typelocal ::=   \textsf{undef} \mid \textsf{rcuItr} \,\, \rho \,\, \N \mid \rcunfd \mid \textsf{unlinked}  \mid \textsf{freeable}}
+\newcommand{\typelocalmap}{\textsf{TypeLocal} : \var \mapsto \typelocal}
+\newcommand{\command}[1]{\db{#1} \mstate \commandeq}
+\newcommand{\commandm}[2]{\db{#1} #2 \commandeq}
+
+\newcommand{\mstate}{(s,h,l,R,F)}
+\newcommand{\commandeq}{\;\;\Downarrow_{\mathrm{tid}}\;\;\overset{\Delta}{=}}
+\newcommand{\op}[2]{\begin{array}{rll}\command{#1} & #2\end{array}}
+\newcommand{\opm}[3]{\begin{array}{rll}\commandm{#1}{#2} & #3\end{array}}
+%%%
+\newcommand{\return}[1]{\mathsf{return~\mathit{#1}}}
+\newcommand{\bind}[2]{\mathsf{Bind}(#1;#2)}
+\newcommand{\result}[1]{\mathsf{y = \mathsf{result}}}
+\newcommand{\mbody}[2]{\mathsf{mbody}(#1,#2)}
+\newcommand{\pre}{\textsf{pre}}
+\newcommand{\assume}[1]{\textsf{assume}(#1)}
+\newcommand{\lift}[1]{\lceil #1\rceil}
+\newcommand{\erase}[1]{\lfloor #1\rfloor}
+%%%
+\newcommand{\M}{\ensuremath{\mathcal{M}}}
+\newcommand{\s}{\ensuremath{\mathcal{S}}}
+\newcommand{\R}{\ensuremath{\mathcal{R}}}
+\newcommand{\N}{\ensuremath{\mathcal{N}}}
+\newcommand{\condexpr}{e \in \textsf{CondExpr}}
+%function definitions for paths
+
+\newcommand{\lengthofpath}{$\mid \rho\mid \;=\;
+  \begin{cases}
+      0+ \mid \rho'\mid &  \rho = \rho_{i}^{k}\rho' \text{   where  } 0\leq i  \text{   and   } \rho_{i}^{k} = \epsilon \\
+         k + \mid \rho'\mid & \rho = \rho_{i}^{k}\rho' \text{   where  } 0\leq i
+  \end{cases}$
+
+}
+%%\textsf{Stabilize}(\Gamma,e) = \forall x \in \N[\fname] \ldotp \exists_{f_1,f_2 \in \fname}\ldotp \N[f_1] =  \N[f_2] \implies \N[f_1,f^{*}] \land \N[f_2,f^{*}]
+
+\newcommand{\stabilizenext}{\textsf{StabilizeNext}(a,b,f)}
+\newcommand{\stabilizepath}{\textsf{StabilzePath}(b,f,k)}
+\newcommand{\stabilizecontext}{\textsf{StabilizeContext}(\Gamma,e) =
+    \left\{\begin{array}{cl}
+    e \in \textsf{PointsToExpr} \land
+    \exists_{a,b,f}\ldotp  (a,b,f) = \textsf{PointsToExprMap}(e) \land \\
+  \pi_{\N}(a) = \textsf{StabilizeNextMap}(a,b,f) \land \\
+  \pi_{\rho}(b) = \textsf{StabilizePath}(b,f, \mid \pi_{\rho}(b)\mid ) \land\\
+  \forall_{c \in \Gamma \land \textsf{TypeDepth}(c) >= \textsf{TypeDepth}(b)} \ldotp \pi_{\rho}(c) = \textsf{StabilizePath}(c,f,\mid\pi_{\rho}(c)\mid)
+  \end{array}\right\}
+}
+%%%frame branch
+\newcommand{\alternatingvars}[1]{\textsf{AbstractVars}(#1)}
+\newcommand{\alternatingvarsdecl}{\textsf{AbstractVars}(\Gamma) \;=\; \{ x,y \mid  (x,y)\in \textsf{AbstractPairs}(\Gamma)\}}
+\newcommand{\alternatetype}{ \text{x:}\textsf{rcuItr}\;\; \rho \;\; \N[f^{*}\mapsto y] \qquad   \text{y:}\textsf{rcuItr}\;\; \rho.\rho^{k} \;\; \N \text{    where    } \rho^{k}[k]=f^{*}}
+\newcommand{\alternatingpath}[1]{f^{*} \; = \; f_1 \mid f_2  \; \text{where} \;  (f=f_1) \;\oplus \; ( f=f_2 )  }
+\newcommand{\alternatingpathsdecl}[1]{\textsf{AbstractPairs}(#1) = \; \{ (\text{x},\text{y}) \mid \text{x} \in #1 \land \text{y} \in #1\}}
+\newcommand{\alternatingpaths}[1]{\textsf{AbstractPairs}(#1)}
+\newcommand{\pointstotuple}{\textsf{PointsToTuple}}
+\newcommand{\pointstoexpr}{\textsf{PointsToExpr} = \{ x.f==y  \in \textsf{Expr} \mid \text{x} \in \var \; \land \; \text{y} \in \var \; \land \; f \in \fname \}}
+\newcommand{\pointstopair}[1]{\textsf{PointsTo}(\text{#1})}
+\newcommand{\projectfieldecl}{\pi_{f} : (\text{x},\text{y},f) \mapsto f}
+\newcommand{\projectfield}[1]{\pi_{f}(#1)}
+\newcommand{\projectpairdecl}{\pi_{\var} :(\text{x},\text{y},f) \mapsto (\text{x},\text{y})}
+\newcommand{\projectpair}[1]{\pi_{\var}(#1)} % tuple,f.
+\newcommand{\pointstoexprdecl}{\textsf{PointsToExpr} = \{e \mid  e = (a.f == b) \land \condexpr \land a,b \in \var \land f \in \fname  \} }
+\newcommand{\pointstostmtdecl}{\textsf{PointsToExprMap}: \textsf{PointsToExpr}\mapsto \var \times \var \times \fname}
+\newcommand{\pointstostmt}[1]{\textsf{PointsToExprMap}(#1) }
+\newcommand{\freductionindexdecl}{\textsf{FieldReductionIndex}(\text{C},\text{x},\text{y},\text{irdc})\; = \;
+  \left\{\begin{array}{cl}
+    \forall_{0 \leq  i \leq n} \ldotp &  (x',y',f')_{i}  \in \pointstostmt{C} \land \\
+    &(\text{x},\text{y})= \projectpair{(\text{x}',\text{y}',f')_{i}} \implies\\
+    & \left\{\begin{array}{cl}
+    \exists_{ firdc} \ldotp  & firdc \leq i \land  (firdc = 0 \implies firdc=irdc) \land \\
+     & \left\{\begin{array}{cl}
+    \forall_{0 \leq none \le firdc} \ldotp & (\text{x},\text{y}) \neq \projectpair{(\text{x}',\text{y}',f')_{none})}  \land \\
+    &  firdc = \text{irdc}
+      \end{array}\right\}
+    \end{array}\right\}
+  \end{array}\right\}
+}
+
+\newcommand{\projectsharpdecl}{\Pi_{\varphi_{e}}(f^{*}) \mapsto f_1 \qquad \Pi_{\lnot\varphi_{e}}(f^{*}) \mapsto f_2 }
+\newcommand{\projectsharpdeclback}{\Pi(\varphi_{e})\mapsto e}
+
+\newcommand{\frenamefielddecl}{\textsf{FieldRename}(\text{x},f,f')  \; = \; \text{x:T}[f/f'] }
+
+\newcommand{\freductiondecl}{\textsf{FieldReduction}(\Gamma,\varphi_{e}) \; = \;
+\left\{\begin{array}{cl}
+  \textsf{FieldRename}(\text{x},f^{*},\Pi_{\varphi_{e}}(f^{*})),
+  \textsf{FieldRename}(\text{y},f^{*},\Pi_{\varphi_{e}}(f^{*})) \mid (\text{x},\text{y}) =\projectpair{\pointstostmt{\Pi(\varphi_{e})}}
+\end{array}\right\}
+}
+
+\newcommand{\freductiondeclneq}{\textsf{FieldReduction}(\Gamma,\lnot\varphi_{e}) \; = \;
+\left\{\begin{array}{cl}
+  \textsf{FieldRename}(\text{x},f^{*},\Pi_{\lnot\varphi_{e}}(f^{*})),
+  \textsf{FieldRename}(\text{y},f^{*},\Pi_{\lnot\varphi_{e}}(f^{*})) \mid (\text{x},\text{y}) = \projectpair{\pointstostmt{\Pi(\varphi_{e})}}
+\end{array}\right\}
+  }
+
+\newcommand{\gammafieldreduction}[4]{\textsf{ReduceAbstractPaths}(#3,#4,#2,#1) = \forall_{x,y \in \textsf{AbstractVars}(#4)} \ldotp \textsf{FieldReduction}(\text{x},\text{y},\varphi_{#3}) = #2 \land \textsf{FieldReduction}(\text{x},\text{y},\varphi_{#3}) = #1  } % x,y,gamma,c
+\newcommand{\alterinxys}{\text{a:}\textsf{rcuItr}\;\; \rho \;\; \N[f_1\mapsto b] \qquad
+  \text{b:}\textsf{rcuItr}\;\; \rho.f_1 \;\; \N  \text{  or }\text{b:}\textsf{rcuItr}\underset{r}{\mid} \rho.\{...f_1\} \;\; \N \\
+  \text{a:}\textsf{rcuItr}\;\; \rho \;\; \N[f_2\mapsto b] \qquad
+  \text{b:}\textsf{rcuItr}\;\; \rho.f_2 \;\; \N \text{  or  } \text{b:}\textsf{rcuItr}\;\; \rho.\{...f_2\} \;\; \N }
+\newcommand{\reducedpathsdecl}[2]{\textsf{ReducedPaths}(#1,#2)\; = \; \left\{\begin{array}{cll}
+  (a,b) \mid  &\text{a:}\textsf{rcuItr}\;\; \rho \;\; \N[f_1\mapsto b] \in #1 & \\
+  &  \land  \text{b:}\textsf{rcuItr}\;\; \rho.\rho^{k} \;\; \N \in #1  & \text{     where    } \rho^{k}[k]=f_1 \\
+  &  \land  \text{a:}\textsf{rcuItr}\;\; \rho \;\; \N[f_2\mapsto b] \in #2 & \\
+  &  \land \text{b:}\textsf{rcuItr}\;\; \rho.\rho^{k} \;\; \N  \in #2 & \text{     where    } \rho^{k}[k]=f_2
+  \end{array}\right\} }
+\newcommand{\exprTopredDecl}{\textsf{ExprToPredicate}(e) \mapsto \varphi_{e}}
+\newcommand{\reducedvars}{\textsf{ReducedVars}(\Gamma_1,\Gamma_2)}
+\newcommand{\reducedvarsdecl}{\textsf{ReducedVars}(\Gamma_1,\Gamma_2)\;=\; \{\forall_{a,b \in \Gamma_1 \cup \Gamma_2} \ldotp (a,b) \in \textsf{ReducedPaths}(\Gamma_1,\Gamma_2) \}}
+\newcommand{\reducedpaths}[2]{\textsf{ReducedPaths}(#1,#2)}
+\newcommand{\abstractpaths}[4]{\textsf{AbstractPaths}(#1,#2,#3,#4) \; = \; \left\{\begin{array}{cl} a ,b \mid
+
+  \begin{array}{cl}
+ \forall_{a,b \in #1}\ldotp & \textsf{FieldRename}(a,#2,\rewritealternate{#2}{#3}{#4}) \land \\
+  &\textsf{FieldRename}(a,#3,\rewritealternate{#2}{#3}{#4}) \land\\
+  &\textsf{FieldRename}(b,#2,\rewritealternate{#2}{#3}{#4}) \land \\
+  &\textsf{FieldRename}(b,#3,\rewritealternate{#2}{#3}{#4})
+  \end{array}
+   \end{array}\right\}
+ }
+\newcommand{\rewritealternatedecl}[3]{\textsf{AbstractField}(#1,#2,#3) = f^{*} \; \texttt{    where   } f^{*} \;=\;  \frac{\varphi_{#3}}{#1 \mid #2} }
+  \newcommand{\rewritealternate}[3]{\textsf{AbstractField}(#1,#2,#3)}
+\newcommand{\composerename}{\circ \text{  provides composing renaming of types ... removes duplicates? [TODO : May change this, keep this simple, now it is nothing, same as ;]. }}
+
+%%%%%%%
+\newcommand{\p}{\rho}
+
+\newcommand{\ptriple}[3]{\{#1\}#2\{#3\}}
+\newcommand{\atriple}[3]{\begin{array}{l}\{#1\}\\#2\\\{#3\}\end{array}}
+\newcommand{\assert}[1]{\{#1\}}
+\newcommand{\modifies}[1]{\textsf{Modifies}(#1)}
+
+%\usepackage[usenames,dvipsnames]{color}
+\definecolor{framed}{gray}{0.50}
+\newcommand{\framed}[1]{\textcolor{framed}{\assert{#1}}}
+
+% Make sure these are all done
+\usepackage{todonotes}
+
+\newcommand{\iso}[1]{\textrm{\color{magenta}\ensuremath{\llbracket}\textrm{Iso says:} #1\ensuremath{\rrbracket}}}
+\newcommand{\colin}[1]{\textrm{\color{cyan}\ensuremath{\llbracket}\textrm{Colin says:} #1\ensuremath{\rrbracket}}}
+%\newcommand{\tocite}[1]{\textrm{\color{green}$\llbracket$CITE: #1$\rrbracket$}}
+
+\newcommand{\mypar}[1]{\textbf{#1}~\xspace}
+
+\newcommand{\hallocprovided}{
+  \texttt{ provided } s[(y,tid)\mapsto o], \;  h[o \mapsto \texttt{new}] \text{ and } \texttt{ default }(\textsf{FType}(f))\\
+  \lambda o', f  \ldotp  \text{ if }  o = o' \;  \text{ skip } \text{ else }  h(o',f)
+}
+\newcommand{\gcprovided}{
+  \texttt{ provided }  s[(x,tid)\mapsto o], \; (\forall f, o' \ldotp o\neq o' \Rightarrow h(o',f) = h'(o',f))) \\
+  \text{ and } \forall f \ldotp h'(o,f)\; \textsf{ undefined }
+}
+\newcommand{\rcuwbeginprovided}{
+ \textrm{provided }tid \notin R
+}
+\newcommand{\rcurbeginprovided}{
+ \textrm{provided }tid \neq l
+}
+\newcommand{\rcurendprovided}{
+ \textrm{provided } \; F \setminus tid \; \textrm{ as } \; \{ (T \setminus tid \},o) \mid (T,o) \in F \}
+}
+\newcommand{\gc}{
+  \inferrule*[left={\scriptsize(Garbage-Collect)}]{}
+           {
+             \op{\texttt{gcStep}}{(s,h',l,R,F)}
+           }
+}
+%%%
+\newcommand{\rcuwbegin}{
+  \inferrule*[left={\scriptsize(RCU-WriteBegin)}]{}
+           {
+            \opm{\texttt{WriteBegin}}{(s,h,\mathrm{unlocked},R,F)}{(s,h,\mathrm{tid},R,F)}
+           }
+}
+\newcommand{\rcuwend}{
+  \inferrule*[left={\scriptsize(RCU-WriteEnd)}]{}
+           {
+            \opm{\texttt{WriteEnd}}{(s,h,\mathrm{tid},R,F)}{(s,h,\textsf{unlocked},R,F)}
+           }
+}
+\newcommand{\rcurbegin}{
+  \inferrule*[left={\scriptsize(RCU-ReadBegin)}]{}{
+    \opm{\texttt{ReadBegin}}{(s,h,l,R,F)}{(s,h,l,R \uplus \{ \mathrm{tid}\},F)}
+  }
+}
+\newcommand{\rcurend}{
+  \inferrule*[left={\scriptsize(RCU-ReadEnd)}]{}{
+    \opm{\texttt{ReadEnd}}{( s,h,l,R \uplus \{ \mathrm{tid} \},F)}{(s,h,l,R,F \setminus \mathrm{tid}) }
+  }
+}
+
+%%%
+\newcommand{\asyncfree}{
+\inferrule*[left={\scriptsize(Async-Free)}]{}
+           {
+            \op{\texttt{asyncDelayedFree}(x)}{(s,h,l,R,F \uplus \{ (R,S(x))\} )}
+           }
+
+}
+\newcommand{\hallocate}{
+  \inferrule*[left={\scriptsize(Heap-Allocate)}]{}
+             {
+                \op{\texttt{y=new}}{(s,h,l,R,F)}
+             }
+}
+\newcommand{\supdate}{
+  \inferrule*[left={\scriptsize(Stack-Update)}]{}
+           {
+             \op{\texttt{y=x}}{(s[(y,tid) \mapsto (x,tid)],h,l,R,F)}
+           }
+}
+\newcommand{\hread}{
+  \inferrule*[left={\scriptsize(Heap-Read)}]{}
+             {
+               \op{\texttt{y=x.f}}{((s[(y,tid) \mapsto h(s(x,tid),f)],h,l,R,F)}
+             }
+}
+\newcommand{\hupdate}{
+  \inferrule*[left={\scriptsize(Heap-Update)}]{}
+             {
+               \op{\texttt{x.f=y}}{(s,h[s(x,tid),f \mapsto s(y,tid)],l,R,F)}
+             }
+}
+\newcommand{\tuigread}[1][T-UniqueGlobalRead]{
+\infer[\textsc{#1}]{
+  \Gamma \, , y:\_ \vdash_{M,R} y = G \dashv y: \textsf{rcuItr}\;\; \rho \;\; \N_{\emptyset} , \Gamma
+}{
+\begin{array}{c}
+   G : \uiglobal \\
+\end{array}
+}
+}
+
+\newcommand{\tuisread}[1][T-UniqueStackRead]{
+\infer[\textsc{#1}]{
+  \begin{array}{c}
+  \Gamma,\rcuitrn{y}{G}{0}{k+1}{\_} , x:\textsf{\_}  \vdash_{M,R} x = y \dashv  \rcuitrn{x}{y}{1}{2}{\_} , \rcuitrn{y}{G}{0}{1}{\_}, \Gamma \\
+  \end{array}
+}{
+\begin{array}{c}
+   G : \uiglobal \\
+\end{array}
+}
+}
+
+\newcommand{\toverwc}[1][T-OverWrite]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma, x:\textsf{rcuItr}\;\; \rho \;\; \N([f \rightharpoonup \_])  \vdash_{M,R} x = x.f \dashv   x:\textsf{rcuItr}\;\; \rho' \;\; \N(\setminus_{f\rightharpoonup \_})  ,\,  \Gamma \\
+\end{array}
+}{
+\begin{array}{c}
+  \ftype{ }{\textsf{rcu}}  \qquad \mid \rho \mid +1 = \mid \rho' \mid  \\
+\end{array}
+}
+}
+%\qquad \aliaswith{z}{x} \rcuitr{z}{y}{k}{k+1}{x}\,
+
+\newcommand{\treadstack}[1][T-ReadStack]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma, z:{\_}, \rcuitrT{x}{y}{k}{k+1}{\_}   \vdash_{M,R} z=x \dashv  \rcuitrT{x}{y}{k}{k+1}{\_}, \rcuitrT{z}{y}{k}{k+1}{\_} , \Gamma \\
+\end{array}
+}{
+\begin{array}{c}
+ \\
+\end{array}
+}
+}
+
+\newcommand{\treadnext}[1][T-ReadNext]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma, z:{\_},  x:\textsf{rcuItr}\;\; \rho \;\; \N([f\rightharpoonup \_])   \vdash_{M,R} z=x.f \dashv  x:\textsf{rcuItr}\;\; \rho \;\; \N(f\rightharpoonup \_ \setminus z),\;\;z:\textsf{rcuItr}\;\; \rho' \;\; \N \;, \Gamma \\
+\end{array}
+}{
+\begin{array}{c}
+  \ftype{ }{\textsf{rcu}} \qquad \mid \rho \mid + 1 = \mid \rho' \mid \\
+\end{array}
+}
+}
+%%%This is read and assignment for loop
+
+\newcommand{\tloverwc}[1][T-OverLoopWrite]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma,  x:\textsf{rcuItr}\;\; \rho\{...f_{loop}\}_{\Delta} \,\, \N   \vdash_{M,R} x = x.f \dashv   \textsf{rcuItr}\;\; \rho\{...f\}_{\Delta} \,\, \N ,\,  \Gamma \\
+\end{array}
+}{
+\begin{array}{c}
+  \ftype{ }{\textsf{rcu}}  \\
+\end{array}
+}
+}
+\newcommand{\tlreadstackext}[1][T-ReadLoopStack1]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma, z:{\_}, x:\textsf{rcuItr}\;\; \rho\{...f_{loop}\}_{\Delta} \,\, \N \\  \vdash_{M,R} z=x \dashv \\ x:\textsf{rcuItr}\;\; \rho\{...f_{loop}\}_{\Delta} \,\, \N, z:\textsf{rcuItr}\;\; \rho\{...f_{loop}\}_{\Delta} \,\, \N \, , \Gamma \\
+\end{array}
+}{
+\begin{array}{c}
+ \\
+\end{array}
+}
+}
+\newcommand{\tlreadstackinit}[1][T-ReadLoopStackInit]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma, z:\textsf{rcuItr}\;\; \rho\{...f_{loop}\}_{\Delta} \,\, \N[f \mapsto x], x:\textsf{rcuItr}\;\; \rho f \{...f_{loop}\}_{\Delta} \,\, \N \\  \vdash_{M,R} z=x \dashv \\ x:\textsf{rcuItr}\;\; \rho f \{...f_{loop}\}_{\Delta} \,\, \N, z:\textsf{rcuItr}\;\; \rho\{...f\}_{\Delta} \,\, \N \, , \Gamma \\
+\end{array}
+}{
+\begin{array}{c}
+  \ftype{ }{\textsf{rcu}} \qquad f_{loop} = \epsilon \\
+\end{array}
+}
+}
+\newcommand{\tlreadnextone}[1][T-ReadLoopNextAbstract]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma, z:\textsf{rcuItr}\;\; \rho f_1\{...f_{loop}\}_{\Delta} \,\, \N, x:\textsf{rcuItr}\;\; \rho\{...f_1\}_{\Delta} \,\, \N[f_1\mapsto z] \\  \vdash_{M,R} z=x.f_2 \dashv \\  x:\textsf{rcuItr}\;\; \rho\{...f_1\}_{\Delta} \,\, \N[\frac{}{f_1 \mid f_2}\mapsto z],\, z:\textsf{rcuItr}\;\; \rho\{...f_2\}_{\Delta} \,\, \N \,, \Gamma \\
+\end{array}
+}{
+\begin{array}{c}
+  \ftype{ }{\textsf{rcu}} \qquad f_{loop}=\epsilon  \\
+\end{array}
+}
+}
+\newcommand{\tlreadnexttwo}[1][T-ReadLoopNextExtend]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma, z:\textsf{rcuItr}\;\; \rho f_1\{...f_{loop}\}_{\Delta} \,\, \N, x:\textsf{rcuItr}\;\; \rho\{...f_1\}_{\Delta} \,\, \N[f_1\mapsto z] \\  \vdash_{M,R} z=x.f_1 \dashv \\  x:\textsf{rcuItr}\;\; \rho\{...f_1\}_{\Delta} \,\, \N[f_1\mapsto z],\, z:\textsf{rcuItr}\;\; \rho\{...f_1\}_{\Delta} \,\, \N \,, \Gamma \\
+\end{array}
+}{
+\begin{array}{c}
+  \ftype{ }{\textsf{rcu}} \qquad f_{loop}=\epsilon  \\
+\end{array}
+}
+}
+\newcommand{\tlreadnextthree}[1][T-ReadLoopNext3]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma, z:{\_}, x:\textsf{rcuItr}\;\; \rho\{...f_{loop}\}_{\Delta} \,\, \N \\  \vdash_{M,R} z=x.f \dashv \\  x:\textsf{rcuItr}\;\; \rho\{...f_{loop}\}_{\Delta} \,\, \N[f\mapsto z],\, z:\textsf{rcuItr}\;\; \rho\{...f\}_{\Delta} \,\, \N \,, \Gamma \\
+\end{array}
+}{
+\begin{array}{c}
+  \ftype{ }{\textsf{rcu}} \\
+\end{array}
+}
+}
+\newcommand{\approximatedecl}{\textsf{Approximate}(x:\textsf{rcuItr}\;\; \rho \;\; \N, \Delta)\mapsto x:\textsf{rcuItr} \;\;\rho\{...\} \;\;\N}
+\newcommand{\approxtype}[2]{\textsf{Approximate}(#1,#2)}
+\newcommand{\gammaapproxdecl}{\Gamma_1 \approx_{\Delta} \Gamma_2 \text{   where  } \Gamma_2 = \{\approxtype{x:\text{T}}{\Delta} \mid \forall_{x:\text{T}\in\Gamma_1}  \}}
+\newcommand{\gammapprox}[3]{#1 \approx_{#3} #2}
+% \qquad \aliaswith{t}{x} \rcuitr{t}{y}{k}{k+1}{z}
+\newcommand{\tunlinknode}[1][T-UnlinkANode]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+\Gamma, \,
+x:\textsf{rcuItr}\;\; \rho \;\; \N([f_1\rightharpoonup z]) , \;\;\;
+z:\textsf{rcuItr} \;\; \rho' \;\; \N'([f_2\rightharpoonup r]) , \;\;\; r:\textsf{rcuItr}\;\; \rho'' \;\; \N'' \\
+\vdash_{M} x.f=r \dashv \\
+z:\unlinked, \,
+x:\textsf{rcuItr}\;\; \rho \;\; \N(f_1\rightharpoonup z \setminus r), \;\;\;
+r:\textsf{rcuItr}\;\; \rho' \;\; \N'' \;\;
+\Gamma \\
+\end{array}
+}{
+\begin{array}{c}
+  \ftype{ }{\textsf{rcu}} \qquad \mid\rho\mid + 1= \mid \rho' \mid  \qquad \mid \rho' \mid + 1 = \mid \rho'' \mid  \\ \forall_{f\in dom(\N')} \ldotp f\neq f_2 \implies \N'([f\rightharpoonup \textsf{Null}])  \\
+\end{array}
+}
+}
+%\qquad  \aliaswith{t}{z} t: \unlinked , \,
+\newcommand{\tafree}[1][T-AsyncFree]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma, x:\unlinked \vdash_{M} \afree{x} \dashv x:\udef , \Gamma \\
+\end{array}
+}{
+\begin{array}{c}
+   \\
+\end{array}
+}
+}
+
+\newcommand{\tallocate}[1][T-Allocate]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma, x:\_ \vdash_{M} x = \NEW \dashv x:\textsf{rcuFresh} \;\; \N_{\emptyset} , \, \Gamma \\
+\end{array}
+}{
+\begin{array}{c}
+   \\
+\end{array}
+}
+}
+
+\newcommand{\tsetfieldfresh}[1][T-SetFieldFresh]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma, \rcuf{p} \;\;, x:\textsf{rcuItr}\;\; \rho \;\; \N([f\rightharpoonup z]) \;\; \\
+  \vdash_{M} p.f = x.f \dashv \\
+  \rcunf{p}{f}{z} ,  \;\;, x:\textsf{rcuItr}\;\; \rho \;\; \N([f\rightharpoonup z])\;\;, \Gamma \\
+\end{array}
+}{
+\begin{array}{c}
+   \\ f \notin \N
+\end{array}
+}
+}
+
+\newcommand{\tlinkfreshf}[1][T-LinkFreshNext]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma, \:\;
+z:\textsf{rcuItr}\;\; \rho \;\; \N([f\rightharpoonup x]) \;\;,
+x:\textsf{rcuItr}\;\; \rho' \;\; \N \;\;,
+r:\textsf{rcuFresh} \;\; \N([f\rightharpoonup x]) \\
+  \vdash_{M} z.f = r \dashv \\
+z:\textsf{rcuItr}\;\; \rho \;\; \N(f\rightharpoonup x\setminus r) \;\;,
+r:\textsf{rcuItr}\;\; \rho' \;\; \N([f\rightharpoonup x]) \;\; ,
+x:\textsf{rcuItr}\;\; \rho'' \;\; \N
+ \;\; , \Gamma \\
+\end{array}
+}{
+\begin{array}{c}
+   \ftype{ }{\textsf{rcu}} \qquad \mid\rho\mid + 1= \mid \rho' \mid  \qquad \mid \rho' \mid + 1 = \mid \rho'' \mid \\
+\end{array}
+}
+}
+
+\newcommand{\tlinkfresh}[1][T-LinkFresh]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma, \;\;
+  p:\textsf{rcuItr}\;\; \rho \;\; \N([f\rightharpoonup o]) \;\;,
+  o:\textsf{rcuItr}\;\; \rho' \;\; \N' \;\;, n:\textsf{rcuFresh} \;\; \N'' \\
+  \vdash_{M} p.f = n \dashv \\
+  p:\textsf{rcuItr}\;\; \rho \;\; \N(f \rightharpoonup o \setminus n) \;\;,
+  n:\textsf{rcuItr}\;\; \rho' \;\; \N'' \;\;,
+  o:\unlinked, \;\;  \Gamma
+\end{array}
+}{
+\begin{array}{c}
+   \ftype{ }{\textsf{rcu}}  \qquad \mid \rho \mid + 1= \mid \rho' \mid \qquad \N' = \N'' \\
+\end{array}
+}
+}
+\newcommand{\tloop}[1][T-Loop]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+   \Gamma \vdash \text{C}^{*} \dashv  \Gamma \\
+\end{array}
+}{
+  \begin{array}{c}
+     \Gamma \vdash C \dashv \Gamma
+\end{array}
+}
+}
+\newcommand{\tloopold}[1][T-LoopOld]{
+\infer[\textsc{#1}]{
+  \begin{array}{c}
+ \Gamma_1',\Gamma_1'' \vdash \text{C}^{\Delta} \dashv   \Gamma_1'''',\Gamma_1''    \\
+%  \tloopweaken\\
+\end{array}
+}{
+  \begin{array}{c}
+    \tframeloop \\
+       \Delta = 0 \implies \Gamma_1' =\Gamma_1''''  \land \\
+       \Delta >0 \implies
+       \left\{\begin{array}{cl}
+       \Gamma_1'\approx_{\Delta}\Gamma_1''' \land \left\{\begin{array}{cl}
+       \exists_{0<\varepsilon \leq\Delta} \ldotp 0<\varepsilon<\Delta  \implies  \Gamma_1'''\vdash C \dashv \textsf{ContexRename}(\Delta,\Gamma_1''',\Gamma_1'''') \land \\
+       \varepsilon = \Delta  \implies \Gamma_1'''\vdash C \dashv \Gamma_1''''
+       \end{array}\right\}
+       \end{array}\right\} \\
+\end{array}
+}
+}
+
+% \qquad \rcuitrT{y}{G}{0}{k+1}{\_} \\  \rcuonly{\rcufiltera{\Gamma_1}}{k}{y} \\ \rcuonly{\rcufiltera{\Gamma_2}}{k'}{y} \\   \Gamma_1 \vdash C \dashv   \Gamma_2
+
+\newcommand{\tpar}[1][T-Par]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma_1 \; \Gamma_2  \vdash_{M,R} \bar{s_1} || \bar{s_2} \dashv  \Gamma'_1 \; , \Gamma'_2 \\
+\end{array}
+}{
+\begin{array}{c}
+   \Gamma_1 \vdash_{R} \bar{s_1} \dashv   \Gamma'_1  \qquad   \Gamma_2 \vdash_{M,R} \bar{s_2} \dashv   \Gamma'_2\\
+\end{array}
+}
+}
+\newcommand{\tllinkfresh}[1][T-LoopExtLinkFresh]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \{\text{TODO}\} \\
+\end{array}
+}{
+\begin{array}{c}
+ \{\text{TODO}\}\\
+\end{array}
+}
+}
+\newcommand{\tllinkfreshnext}[1][T-LoopExtLinkFreshNext]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+\{\text{TODO}\}\\
+\end{array}
+}{
+\begin{array}{c}
+\{\text{TODO}\} \\
+\end{array}
+}
+}
+\newcommand{\tlunlink}[1][T-LoopExtUnlink]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+   \{\text{TODO}\} \\
+\end{array}
+}{
+\begin{array}{c}
+ \{\text{TODO}\}\\
+\end{array}
+}
+}
+\newcommand{\tskip}[1][T-Skip]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+   \Gamma \vdash \textsf{skip} \dashv   \Gamma \\
+\end{array}
+}{
+\begin{array}{c}
+ \\
+\end{array}
+}
+}
+\begin{comment}
+\newcommand{\tframebranch}[1][T-FrameRCUBranch]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+ \Gamma_{1\text{C}_1}, \, \Gamma'_{1\text{C}_1}  \vdash C_1 \dashv \Gamma'_{23R},\, \Gamma_{abs} \qquad  \Gamma_{1\text{C}_1}, \, \Gamma'_{1\text{C}_2}  \vdash C_2 \dashv  \Gamma'_{23R},\, \Gamma_{abs}
+\end{array}
+}{
+  \begin{array}{c}
+    \Gamma_1  \vdash \text{if e C_1 else C_2}  \dashv  \Gamma_4 \\
+    \exists_{\Gamma'_1} \ldotp \Gamma'_1= \textsf{AbstractVars}(\Gamma_1) \qquad \exists_{\Gamma''_1}\ldotp \Gamma_1 = \Gamma'_1 \uplus \Gamma''_1\\
+    \Gamma''_1,\Gamma'_1\vdash C_1 \dashv   \Gamma_2 \qquad \Gamma''_1,\Gamma'_1 \vdash C_2 \dashv \Gamma_3
+
+\end{array}
+}
+}
+\end{comment}
+
+\newcommand{\tbranch}[1][T-Branch]{
+\infer[\textsc{#1}]{
+  \begin{array}{c}
+ \Gamma,x:\textsf{rcuItr}\;\; \rho \;\; \N([f_1\mid f_2 \rightharpoonup z]) \vdash \text{if e C$_1$ else C$_2$} \dashv \Gamma_4
+\end{array}
+}{
+  \begin{array}{c}
+    e \in \textsf{PointsToExpr} \qquad e \text{ is } x.f_1==z \\
+ \Gamma,x:\textsf{rcuItr}\;\; \rho \;\; \N([f_1 \rightharpoonup  z]) \vdash C_1 \dashv\Gamma_4 \qquad \Gamma,x:\textsf{rcuItr}\;\; \rho \;\; \N([f_2 \rightharpoonup z]) \vdash C_4 \dashv \Gamma_4
+\end{array}
+}
+}
+
+\newcommand{\tframe}[1][T-Frame]{
+\infer[\textsc{#1}]{
+  \begin{array}{c}
+ \Gamma, \; \Gamma_1 \vdash C \dashv \Gamma, \; \Gamma_2
+\end{array}
+}{
+  \begin{array}{c}
+ \Gamma_1 \vdash C\dashv \Gamma_2  \qquad \textsf{RcuModifiedOnly}(C)
+\end{array}
+}
+}
+
+%%\seprcubd{\Gamma_1}{\Gamma_2}{\Gamma_3}{\Gamma_4}{C_1}{C_2}{\Gamma}
+
+%\Gamma''_1,\Gamma'_3\vdash \text{C$_1$} \dashv   \Gamma_2 \qquad \Gamma''_1,\Gamma'_1 \vdash \text{C$_2$} \dashv \Gamma_3 \qquad \Gamma_{23} = \Gamma_2 \cup \Gamma_3  \\
+   %  \Gamma_{23R} = \textsf{ReducedVars}(\Gamma_2,\Gamma_3) \qquad \Gamma'_{23R}=\Gamma_{23} - \Gamma_{23R} \qquad \Gamma_{abs} = \textsf{AbstractPaths}(\Gamma_{23R},\text{f$_1$},\text{f$_2$},e) \qquad \Gamma_4 =  \Gamma'_{23R},\, \Gamma_{abs}
+
+\newcommand{\tchoosef}[1][T-Choose-First]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma_1  \vdash C_1 \dashv  \Gamma_2 \\
+\end{array}
+}{
+\begin{array}{c}
+   \Gamma_1 \vdash C_1 \dashv   \Gamma_2  \qquad   \Gamma_1 \vdash C_2 \dashv   \Gamma_3\\
+\end{array}
+}
+}
+\newcommand{\tchooses}[1][T-Choose-Second]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma_1  \vdash C_2 \dashv  \Gamma_3 \\
+\end{array}
+}{
+\begin{array}{c}
+   \Gamma_1 \vdash C_1 \dashv   \Gamma_2  \qquad   \Gamma_1 \vdash C_2 \dashv   \Gamma_3\\
+\end{array}
+}
+}
+\newcommand{\tseq}[1][T-Seq]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \Gamma_1  \vdash \bar{s_1} \; ; \; \bar{s_2} \dashv  \Gamma_3 \\
+\end{array}
+}{
+\begin{array}{c}
+   \Gamma_1 \vdash \bar{s_1} \dashv   \Gamma_2  \qquad   \Gamma_2 \vdash \bar{s_2} \dashv   \Gamma_3\\
+\end{array}
+}
+}
+
+\newcommand{\tshuffle}[1][T-Exchange]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+   \Gamma, \, x:T , \, y:T' , \, \Gamma' \vdash \bar{s} \dashv   \Gamma'' \\
+\end{array}
+}{
+\begin{array}{c}
+   \Gamma, \, y:T' , \, x:T, \, \Gamma' \vdash \bar{s} \dashv   \Gamma'' \\
+\end{array}
+}
+}
+\newcommand{\tconsequence}[1][T-Conseq]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+   \Gamma \vdash C \dashv \Gamma''' \\
+\end{array}
+}{
+\begin{array}{c}
+   \Gamma \subt \Gamma'  \qquad \Gamma' \vdash C \dashv \Gamma'' \qquad \Gamma'' \subt \Gamma''' \\
+\end{array}
+}
+}
+
+%%Sub typing
+\newcommand{\nsubone}[1][N-Sub-F1]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+   \vdash \, \N \subt \N'  \\
+\end{array}
+}{
+\begin{array}{c}
+   \text{T } = \textsf{rcuItr}\;\; \rho \;\; \N([f_1 \rightharpoonup y]) \\
+   \text{T}'=\textsf{rcuItr}\;\; \rho \;\; \N'([f^{*} \rightharpoonup y])
+\end{array}
+}
+}
+
+\newcommand{\nsubtwo}[1][N-Sub-F2]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+   \vdash \, \N \subt \N'  \\
+\end{array}
+}{
+\begin{array}{c}
+   \text{T} = \textsf{rcuItr}\;\; \rho \;\; \N([f_2 \rightharpoonup y]) \\
+   \text{T}'=\textsf{rcuItr}\;\; \rho \;\; \N'([f^{*} \rightharpoonup y])
+\end{array}
+}
+}
+\newcommand{\nsubthree}[1][N-Sub]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+   \vdash \, \N_1 \subt \N \, \land \, \N_2 \subt \N  \\
+\end{array}
+}{
+\begin{array}{c}
+  \N_1[f_2\mapsto x] \qquad  \N_2[f_1\mapsto x] \qquad  \N[f^{*} \mapsto x] \qquad f^{*} = \frac{\varphi_{e}}{f_1 \mid f_2}\\
+\end{array}
+}
+}
+
+\newcommand{\psubone}[1][P-Sub]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+   \vdash  \rho' \subt \rho'' \\
+\end{array}
+}{
+  \begin{array}{c}
+  \rho'' = \rho'\rho^{k}
+\end{array}
+}
+}
+
+\newcommand{\psubtwo}[1][P-Sub-Ref]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+   \vdash  \rho \subt \rho  \\
+\end{array}
+}{
+  \begin{array}{c}
+  \\
+\end{array}
+}
+}
+
+\newcommand{\psubthree}[1][P-Sub-Trans]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+   \vdash  \rho' \subt \rho'''  \\
+\end{array}
+}{
+  \begin{array}{c}
+    \vdash  \rho' \subt \rho'' \qquad
+    \vdash  \rho'' \subt \rho'''
+\end{array}
+}
+}
+
+\newcommand{\csubfour}[1][C-Sub-Emp]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+   \vdash  \epsilon \subt \epsilon  \\
+\end{array}
+}{
+  \begin{array}{c}
+\end{array}
+}
+}
+
+\newcommand{\tsubase}[1][S-Base]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+ \vdash \, \text{T} \prec_{\varepsilon = 0} \text{T'} \qquad \text{T} =_{\varepsilon = 0} \text{T'} \\
+\end{array}
+}{
+\begin{array}{c}
+ \varepsilon = 0 \\
+\end{array}
+}
+}
+\newcommand{\tsubreflexive}[1][S-Reflexive]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+ \vdash \, \text{T} \preceq_{\varepsilon} \text{T} \\
+\end{array}
+}{
+\begin{array}{c}
+  \\
+\end{array}
+}
+}
+\newcommand{\tsubtransitive}[1][S-Transitive]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+  \vdash \, \text{T}_1 \preceq_{2\varepsilon} \text{T}_3 \\
+\end{array}
+}{
+\begin{array}{c}
+    \vdash \, \text{T}_1 \preceq_{\varepsilon} \text{T}_2 \qquad \vdash \, \text{T}_2 \preceq_{\varepsilon} \text{T}_3 \\
+\end{array}
+}
+}
+\newcommand{\tsubempty}[1][S-Empty]{
+\infer[\textsc{#1}]{
+\begin{array}{c}
+   \vdash \, \epsilon \preceq_{\varepsilon} \epsilon \\
+\end{array}
+}{
+\begin{array}{c}
+    \\
+\end{array}
+}
+}
+\newcommand{\subcontextone}[1][C-Sub-Red-notsure]{
+\infer[\textsc{#1}]{
+  \begin{array}{c}
+        \vdash \Gamma_1  \subt \Gamma \, \land \, \Gamma_2 \subt \Gamma \\
+\end{array}
+}{
+  \begin{array}{c}
+    \textsf{ReduceAbstractPaths}(e,\Gamma,\Gamma_1,\Gamma_2) \\
+\end{array}
+}
+}
+\newcommand{\subeqcontexttwo}[1][C-Sub-Abs-notsure]{
+\infer[\textsc{#1}]{
+  \begin{array}{c}
+        \vdash \Gamma'  \subt \Gamma \\
+\end{array}
+}{
+  \begin{array}{c}
+    \textsf{ReducedVars}(\Gamma_1,\Gamma_2) = \Gamma' \qquad \textsf{AbstractPaths}(\Gamma',f_1,f_2,e) =\Gamma
+\end{array}
+}
+}
+\newcommand{\subeqcontextthree}[1][C-Sub]{
+\infer[\textsc{#1}]{
+  \begin{array}{c}
+        \vdash \Gamma,\, x:\text{T}  \subt \Gamma',\, x:\text{T'} \\
+\end{array}
+}{
+  \begin{array}{c}
+    \vdash\Gamma \subt \Gamma' \qquad
+    \vdash \text{T}  \subt \text{T'}
+\end{array}
+}
+}
+
+\begin{comment}
+\newcommand{\tsubcons}[1][S-Cons]{
+\infer[\textsc{#1}]{
+  \begin{array}{c}
+        \vdash \Gamma, \, \text{x:T}  \prec \Gamma',\, \text{x':T'}  \\
+\end{array}
+}{
+\begin{array}{c}
+      \vdash \Gamma \prec \Gamma'  \qquad \text{T} \prec \text{T'}  \\
+\end{array}
+}
+}
+
+\newcommand{\tsubdrop}[1][S-Drop]{
+\infer[\textsc{#1}]{
+  \begin{array}{c}
+      \vdash \Gamma \prec \Gamma' \\
+\end{array}
+}{
+\begin{array}{c}
+    \vdash \Gamma, \, \text{x:T}  \prec \Gamma'  \\
+\end{array}
+}
+}
+\end{comment}
+%%%%
+\newcommand{\tsmovewrite}{
+\infer[\textsc{Move-To-RCUWrite}]{
+\begin{array}{c}
+  \Gamma \vdash \textsf{RCUWrite}\, x.f \textsf{ as }y\textsf{ in }\{C\} \\
+\end{array}
+}{
+ \begin{array}{c}
+ \Gamma, y : \textsf{rcuItr} \vdash_M C \dashv \Gamma' \qquad
+ \ftype{}{\textsf{RCU}} \\
+ \textsf{NoFresh}(\Gamma') \qquad
+\textsf{NoUnlinked}(\Gamma')
+\end{array}
+}
+}
+\newcommand{\tsmoveread}{
+  \infer[\textsc{Move-To-RCURead}]{
+\begin{array}{c}
+\Gamma \vdash \textsf{RCURead}\, x.f \textsf{ as }y\textsf{ in }\{C\} \\
+\end{array}
+}{
+\begin{array}{c}
+\ftype{}{\textsf{rcu}} \\
+\Gamma, y : \textsf{rcuItr} \vdash_R C \dashv \Gamma' \\
+\end{array}
+}
+}
+
+\newcommand{\meta}{
+\[
+\begin{array}{rclcl}
+a & \textsf{atoms} \\
+s & \textsf{statement}  & & s::= & a \mid skip \mid s;s \mid s+s \mid s \parallel s \mid s^{*} \\
+n,x,y,z,o,p & \textsf{variables}  & &  \Gamma::= & . \mid \Gamma,x:t \\
+t & \textsf{types}  & & t::= & \mathsf{int} \mid \mathsf{bool} \mid \mathsf{struct} \; sn \{ \overline{fld}\} \mid T\\
+T & \textsf{Qualified types} & & T::= & t<rt> \\
+f & \textsf{field name}  & & fld::= & t \; f \mid  t<rt> \; f\\
+rb & \textsf{rcu blocks}  & & rb::= & \mathsf{ReadBegin} \, ; \bar{s}  \, ; \mathsf{ReadEnd} \mid \mathsf{WriteBegin} \, ; \bar{s}  \, ;  \mathsf{WriteEnd} \\
+fld & \textsf{field decl}  & & rt::= & \mathsf{rcu} \mid \mathsf{rcuItr} \mid \mathsf{rcuRoot} \\
+sn & \textsf{struct name}   & &  i,j & \mathsf{nat (indices)}  \\
+
+\end{array}
+\]\\
+\[
+\begin{array}{rclcl}
+\alpha &::= & \mathsf{skip} \mid \mathsf{x.f = y} \mid \mathsf{y=x} \mid \mathsf{y =x.f}  \mid \mathsf{y =new}\mathsf{Free}\mathsf{(\textsf{x})} \mid \mathsf{Sync}
+ \end{array}
+\]
+}
+
+\newcommand{\grammar}{
+\[
+\begin{array}{rclcl}
+\alpha &::= & \mathsf{skip} \mid \mathsf{x.f = y} \mid \mathsf{y=x} \mid \mathsf{y =x.f}  \mid \mathsf{y =new} \mid  \mathsf{Free}\mathsf{(\textsf{x})} \mid \mathsf{Sync}                   
+\end{array}\quad
+\textsf{Sync} \overset{\Delta}{=} \texttt{\textsf{SyncStart};\textsf{SyncStop}}
+\]
+}
+
+\lstset{numbersep=1.7pt,xrightmargin=-0.7cm,xleftmargin=0.7em,numbers=left,columns=fullflexible,escapechar=`,mathescape,language=Java,basicstyle=\small\ttfamily,keywordstyle=\color{blue},morekeywords={ReadBegin,ReadEnd,WriteBegin,WriteEnd,SyncStart,SyncStop,Free,foreach,in,var,consume,readable,r,isolated,writable,immutable}}
+
+%\lstset{
+%  xleftmargin=1em,      % <-- change this to a suitable length
+%  xrightmargin=-0.5cm,  % <-- change this to a suitable negative length
+%  numbersep=2pt,        % default 10pt
+%  numbers=left,
+%  numberstyle=\footnotesize,
+%  basicstyle=\ttfamily,
+%  breaklines=true}
+\usepackage{lipsum} % for dummy text
+
+\renewcommand{\and}{\\}
+
+\usepackage{microtype}%if unwanted, comment out or use option "draft"
+%\bibliographystyle{plainurl}% the recommended bibstyle
+
+%\bibliography{references/paper}
 \begin{document}
-
-%%
-%% The "title" command has an optional parameter,
-%% allowing the author to define a "short title" to be used in page headers.
-\title{On Verification Patterns for Low-Level Systems via Modal Abstractions }
-
-%%
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
-% Commentout
-%\author{Ben Trovato}
-%\authornote{Both authors contributed equally to this research.}
-%\email{trovato@corporation.com}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
-%\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
-%\affiliation{%
- % \institution{Institute for Clarity in Documentation}
- % \city{Dublin}
- % \state{Ohio}
- % \country{USA}
-%}
-
-%%
-%% By default, the full list of authors will be used in the page
-%% headers. Often, this list is too long, and will overlap
-%% other information printed in the page headers. This command allows
-%% the author to define a more concise list
-%% of authors' names for this purpose.
-\renewcommand{\shortauthors}{Kuru and Gordon}
-
-%%
-%% The abstract is a short summary of the work to be presented in the
-%% article.
+%
+\title{Safe Deferred Memory Reclamation with Types}
+%
+%\titlerunning{RCU Types}
+% If the paper title is too long for the running head, you can set
+% an abbreviated paper title here
+%
+\author{Ismail Kuru \orcidID{0000-0002-5796-2150} \and
+Colin S. Gordon \orcidID{0000-0002-9012-4490}}
+%%%Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
+%
+\authorrunning{I. Kuru et al.}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
+\institute{Drexel University \\ %\and
+%%%Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
+\email{\{ik335,csgordon\}@drexel.edu}}%\\
+%%%\url{http://www.springer.com/gp/computer-science/lncs} \and
+%%%ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
+%%%\email{\{abc,lncs\}@uni-heidelberg.de}}
+%
+\maketitle              % typeset the header of the contribution
+
+%\input{chapters/squeezing.tex}
 \begin{abstract}
@@
 
-%%
-%% The next two lines define the bibliography style to be used, and
-%% the bibliography file.
-\bibliographystyle{ACM-Reference-Format}
+\bibliographystyle{splncs04}
 \bibliography{software}
-
-%%
-%% If your work has an appendix, this is the place to put it.
+\newpage
 \appendix
@@
 \end{document}
-\endinput
-%%
-%% End of file `sample-sigconf-authordraft.tex'.
+\grid

### 2025-06-02 paper.tex
ready for dali

@@
 %
-\title{Safe Deferred Memory Reclamation with Types}
+\title{Modal Verification Patterns for Systems}
 %
@@
 %% information and builds the first part of the formatted document.
-\maketitle
-\section{Introduction}
-\label{sec:intro}
-Systems verification has attracted much attention from both the verification community and the systems community over the past couple of decades.
-%There are two main streams of verification efforts on systems.
-In one class of verification, \emph{automated} verification tools are used.
-They allow people to specify systems via specific annotation languages they offer, and incorporate translation of these specifications into SMT-like formulas for which proofs are searched by the SMT solvers at the backend ~\cite{}.
-Most of this effort has been build on top of \emph{program logics}~\cite{reynolds2002separation,hoare1969axiomatic,smans2012implicit,pratt1976semantical},
-which are formal logics of reasoning about program behaviour --- given any starting state
-satisfying a particular precondition, running a certain code segment of interest should not fail and should produce a final state satisfying a desired postcondition.
-This requires an interacting family of pieces, including a specification language
-to describe those desired pre- and postconditions, as well as rules of the logic to deal with
-how program operations like variable or register assignments and memory accesses interact with pre- and postconditions.
-Since most low-level systems are stateful, \emph{separation logic} has been the focus 
-of most recent work.
-%has been either the foundational or direct choice of logic in these verification efforts. 
-%\Colin{Talking about separation logic makes lots of sense. Do we care about Iris/Coq/Lean/etc. vs. automation? Using the modality idea
-%for something like Dafny would certainly look different from the Iris work, but can't it work?
-%e.g. \url{https://csgordon.github.io/publications/agere19/}
-%I don't want to push to plug my own work if you don't think it makes sense or are making a different point. But I think it's worth
-%considering how broadly the big ideas can be applied.
-%}
-
-%Advancements in separation logic have been drastic ~\cite{}.
-These advancements, such as the capability of introducing \emph{ghost} (logical) bookkeeping and associating this bookkeeping with the concept of \emph{invariant} in Iris separation logic, increased the expressivity of any logic that is built on top.
-Expectedly, many logics enjoying these strong \emph{separation logic} capabilities have been built to verify low-level systems~\cite{}.
-%All these program logics built on top of Iris customize and utilize two main aspects of a separation logic like Iris.
-%%\paragraph{Custom-Tailored Weakest-Precondition (WP)}
-%%\todo[inline]{speak about crash modality etc, encoding how the state should get changed ... GC as well}
-%%\paragraph{An Essential Step in Operational Semantics}
-%%\todo[inline]{speak about crash step in operational semantics. GC as well}
-%%Regarding generalization of customized weakest-preconditions, there has been foundational generalizations defined over incorporating the knowledge of how the resources change, i.e.,transformation of resources, ~\cite{nextgenmodality}, into the separation logic resource assertions.
-This allows encoding the many customized weakest-preconditions as an instance of a generalized modality definition and lifting of the composition of resource transformation knowledge, which itself is defined as a resource algebra, to the ordinary separation logic resource assertions so that all the underlying resource algebra capabilities can be used for the assertions decorated with the resource-transformation modality.
-However, even with this kind of generalization effort, it still customizes both the weakest-precondition and resource assertions with the version value (\emph{transformation-generation} value) which most probably lacks being generalizable regarding the real systems to be specified with it as it is not always the case that systems take \emph{versioning} as its key aspect.
-In general, modalities used in program verification ~\cite{}, are either still tightly coupled with the separation logic they are hosted in, i.e., modalities as definitions for hiding which state a certain truth on resources holds, or as part of the underlying logic's machinery, e.g., later modality as logical machinery encoding step-indexing mechanism.
-Modern separation logics are often implemented as frameworks (notably \textsc{Iris}~\cite{jung2018iris}), which offers a collection of tools for building custom separation logic variants for resources of interest, modularly.
-This makes it possible to
-reuse significant amounts of work across verifications using memory~\cite{jung2018iris},
-disks~\cite{chajed2019verifying}, and other diverse resources.
-%Highly customizable logic ~\cite{iris,..} being the basis for building custom-tailored logics for specific verification tasks ~\cite{some sl refs} enabled building clean and intuitive specifications and proof rules.
-When not using such a framework and instead using an existing tool~\cite{leino2010dafny,vazou2014refinement} as a basis, one can often customize the assertion language by defining libraries of specification definitions and lemmas to effectively have a custom logic.
-
-With these options comes the question: if you can customize your verification tool and how it expresses program states, \emph{how should} you customize it?
-We argue that a go-to methodology should be to define \emph{modalities},
-which it turns out are in active use in a variety of systems papers because that approach seems to naturally fit many problems that arise in verifying systems code
-~\cite{tejperennial19,tejthesis, derekrustbelt20, jung2018iris, kuru2024modalabstractionsvirtualizingmemory, garbagecollect, amalreal2024,fsl, fsl++}.
-
-This paper  
- \begin{itemize}
- \item builds an intuition from a systems-verification perspective about the modal concepts that have been used in recent years with modern program logic -- Section \ref{sec:motivation}
-\item describes the common structure underlying the modalities used in systems verification work, effectively a high-level design pattern and process for designing a customized specification language, which has not previously been described --- Section \ref{sec:definitions} % to understand the structure of modal abstractions given for different kinds of systems in Table \ref{table:decomposition}
-%\item identifies the common patterns in different modal abstractions for different systems
-\item gives example specifications to critical steps of different systems, and discuss how modal abstractions look in these specifications
-\end{itemize}
-
-\section{Motivation, Background, and Related Work}
-\label{sec:motivation}
-To explain our ideas clearly, we briefly recap some of the background and themes our ideas build on,
-casting them in a certain way to bring out the relevance of our philosophy.
-%\begin{table}[h]
-%\centering
-%\caption{Decomposing Systems}
-%\label{table:decompositionsys}
-%\begin{tabular}{@{}lcScP{2cm}@{}}
-%\toprule
-%   & Column & Column                    & Column                   & Column \\ \midrule
-%Row & 1      & 2                         & 3                        & $k=a+b+c+d+e+f+g$      \\
-%Row & 5      & \multicolumn{2}{c}{\multirow{2}{*}{Because why not}} & 8      \\
-%Row & 9      & \multicolumn{2}{c}{}                                 & 12     \\
-%Row & 13     & $k=a+b+c+d+e$                        & 15                       & 16     \\ \bottomrule
-%\end{tabular}
-%\end{table}
-
+%\maketitle
+\section{Introduction}\label{sec:intro}
+Low-level systems exhibit certain patterns in their designs, especially when interacting with computing resources.
+Exploiting certain patterns while designing software has been an important field of study.
+In this regard, we think that certain properties of modalities enable us to understand and do the verification challenges that show certain patterns.
+\subsection{Contributions to Identifying the Verification Patterns via Modal Abstractions}
+We argue how modal abstractions can be used to identify and abstract system verification challenges.
+We justify our perspective by discussing prior systems that have successfully used modalities for system verification, arguing that they fit into the verification design pattern we articulate, and explaining how this approach might apply to other systems' verification challenges.
+\paragraph{Identifying System Verification Challenges} We start with identifying common patterns in system verification: \emph{virtualization, sharing, and translation}.
+\paragraph{Introducing the Concept of \emph{Resource}} Then we discuss the concept of \emph{resource} which has already been an essential concept in the design of systems.
+Inspired by the concept of resource in the systems, we define what a resource and its context are in the modal abstractions.
+\paragraph{Introducing the Concept of \emph{Nominals}} Nominalization enables identifying a resource in a context.
+For example, a transaction is a context of resources of in-memory updated disk blocks.
+The transaction identifier is used to associate a transaction with a disk-block to be persisted so that, in case of a crash while persisting updated disk-blocks, the filesystem can rollback the \emph{already persisted} disk-blocks of the transaction, and reach to the previous consistent disk state.
+To be able to do so, both the updated in-memory disk blocks and the transaction must refer to the transaction identifier -- \emph{strong nominalization}.
+In another example, virtual memory references (resources) in an address space (resource context), which can be uniquely identified with a root address (the nominal) of its page-table tree, are \emph{agnostic} to the address space that they are in.
+However, they can only be accessed (be valid) in the address space to which they are agnostic.
+However, then an address space switch happens, the virtual memory references of the previous address space must be made inaccessible.
+To be able to do so, although virtual memory references do not hold any piece of information related to their address space, we still have to associate them.
+We call this kind of unilateral nominalization -- \emph{weak nominalization}.
+\paragraph{Taxonomy of the Current Modal Approaches in System Verification} Based on these concepts defined, we summarize contemporary verification efforts using modal abstractions.
+We choose them from different domains, for example, weak memory verification, storage persistence ~\cite{tejthesis, perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2024modalabstractionsvirtualizingmemory,amalreal2024} because we would like to justify that our \emph{definitions} are not domain dependent.
+
+\section{Definitions}
+Although they differ in the functionality they offer, low-level systems exhibit certain patterns of design and utilization of computing resources.
+  In this section, we argue the position that \emph{modalities}  
+  should be a go-to approach when specifying and verifying low-level systems code.
+  We explain how the concept of a \emph{resource context} helps guide the design of new modalities
+  for verification of systems code, and
+  we justify our perspective by discussing prior systems that have used modalities for systems verification
+  successfully, arguing that they fit into the verification design pattern we articulate,
+  and explaining how this approach might apply to other systems verification challenges.
+
+  To explain our ideas in the general systems understanding, we briefly recap some of the background and themes our ideas build on again,casting them in a certain way to bring out the relevance of our philosophy.
   \subsection{Resources in Systems Software}
@@
 
-\paragraph{Virtualization}
+\subsubsection{Virtualization}
 One of the most common forms of abstraction provided by systems software is \emph{virtualization},
 which abstracts the relationship between conceptual and physical computing resources. 
-Operating System (OS) kernels virtualize memory locations and quantity (via virtual memory and paging~\cite{denning1970virtual}).
+Operating System (OS) kernels virtualize memory locations and quantity (via virtual memory and paging~\cite{Denning1970VM}).
 Distributed language runtimes may virtualize addresses, even when processes may migrate across machines~\cite{jul1988fine}.
-Filesystems virtualize locations on disk~\cite{rodeh2013btrfs,hitz1994file,rosenblum1992design,bonwick2003zettabyte}.
+Filesystems virtualize locations on disk~\cite{rodeh2013btrfs,hitz1994file,Rosenblum1992LFS,bonwick2003zettabyte}.
 Programs built on top of the corresponding systems software layer work logically at the level of these virtualized resources,
 and it makes sense to specify the systems software directly in terms of those abstractions.
-
-\paragraph{Sharing} is one of the fundamental reasons for kernels to {virtualize} resources,
-for creating the illusion of exclusive access to virtualized versions of shared resources, including CPU time (via preemption).
-An implementation of sharing relies on two concepts: \emph{indirection} and \emph{coordination}.
-Indirection is a method that allows the mapping of resources of one type into another.
-By doing so, it allows treating the limited resources (e.g.
-hardware resources such as physical memory) as if there are more than they actually are (e.g.
-via virtual memory addresses and paging).
-Consequently, virtualization for the sake of ensuring availability requires \emph{sharing} of the resources that exist in reality.
-
 \paragraph{Access by Translation}
 Accessing virtualized resources via translation is a common way to virtualize notions of location (e.g.,
-virtual memory addresses, inodes or object IDs instead of disk addresses~\cite{bonwick2003zettabyte,hitz1994file,rosenblum1992design,rodeh2013btrfs}).
+virtual memory addresses, inodes or object IDs instead of disk addresses~\cite{bonwick2003zettabyte,hitz1994file,Rosenblum1992LFS,rodeh2013btrfs}).
 %Accessing the hardware resources in the presence of sharing means having more resource references (references to the virtualized resources) to access the \emph{mapped} resources that are physical and potentially fewer.
@@
 
-%\paragraph{Isolation Hand-in-Hand with Sharing}
-%\Ismail{put a few words on abstractions such as address space, stack region, snapshots ,etc.}
-%\Colin{Not sure what you want here?}
-
-\subsection{Logic of Resources \& Systems Verification }
-\label{sec:seplogic}
-Separation logic~\cite{reynolds2002separation} is a specialization of Hoare Logic~\cite{hoare1969axiomatic}
-to better deal with reasoning about disjoint resources.
-Classic Hoare Logic, as used in prominent systems verifications like \textsc{seL4}~\cite{klein2014comprehensive},
-deals with pre- and post-conditions. A so-called ``triple''
-\[
-\{P\}\;C\;\{Q\}
-\]
-represents an assertion that if the current program state is accurately described by assertion $P$,
-then executing the ``command'' (program) $C$ will not crash, and if it terminates then $Q$ will accurately
-describe the ending program state.
-
-\paragraph{Ownership} Separation Logic extends Hoare Logic 
-with an additional logical connective $\ast$ (called \emph{separating conjunction}) such that $P \ast Q$ asserts that $P$ and $Q$ not only both hold, but apply to
-\emph{disjoint} portions of the program state.\footnote{Though much like systems design, modern separation logics~\cite{jung2018iris}
-allow this separation to apply to a kind of virtualization of physical resources~\cite{jensen2012fictional},
-or even resources that only exist in the logic~\cite{jung2016higher}.
-}
-This disjointness assumption leads to discussing separation logic assertions as asserting \emph{ownership}
-of the (physical or virtual) resources they hold on.
-The classic \emph{points-to} assertion $\ell\mapsto v$ asserts ownership of a portion of memory containing memory address
-$\ell$, and that the memory at that location contains value $v$. The disjointness from separating conjunction
-is useful: if $\ell\mapsto v \ast \ell'\mapsto v'$ is true, it means $\ell$ and $\ell'$ are distinct memory addresses,
-because they must hold on disjoint parts of memory.
-
-Modern separation logics~\cite{jung2018iris} allow definition of more nuanced versions of this idea.
-A common form of this, which we revisit in Section \ref{sec:definitions}, is the parameterization of this and
-other assertions by a kind of context --- a \emph{resource context} --- resulting in assertions like  $\ell \mapsto_{\textsf{Ctx}} v$, conveying the ownership of the resource defined by the reference $\ell$ and the value ($v$) it's referring to, but where the \emph{resource context}
-of the assertion may have the name ($\ell$) and value ($v$) exist not in a partial map of physical memory locations to values,
-but of register names to register values~\cite{chlipala2011mostly,ni2007using},
-filesystem block indices to block contents~\cite{chajed2019verifying},
-or other kinds of associations.
-Thus much of the machinery of modern separation logics comes from shifting the notion of which resource context 
-assertions are relative to.
-
-%\Colin{Regarding next paragraph: separation logics are not necessarily foundational (early work was all pen-and-paper,
-%automated work with SMT solvers clearly isn't foundational). Unless you mean ``foundational'' in a non-technical sense (in which case, a different word should be used, to avoid confusion).}
-%\paragraph{Specification} Separation logics~\cite{} rely on foundational Hoare-Logic when introducing their abstractions: make specific assertions and abstractions appear in the form of pre- and post-conditions, $\{P\} e \{Q\}$, called Hoare Triple.
-It is also important to note that program logics also enjoy encoding Hoare Triples in terms of weakest-precondition~\cite{dijkstra1975guarded} program logics~\cite{pratt1976semantical}, which is roughly
-%\[\textsf{P} \; -\ast \;\textsf{wp}\; e \; \{Q\}\]
-%where $-\ast$ can  be intuitively thought of as \textit{"when P is added into the resource context we can have the resource represented with Q"}
-%\Colin{\emph{Is} it important to note this? I'm genuinely not sure it is.}
-
-\subsection{Modal Reasoning}
+%\subsection{Sharing} Sharing is one of the fundamental reasons for kernels to {virtualize} resources,
+%for creating the illusion of exclusive access to virtualized versions of shared resources, including CPU time (via preemption).
+%An implementation of sharing relies on two concepts: \emph{indirection} and \emph{coordination}.
+Indirection is a method that allows the mapping of resources of one type into another.
+By doing so, it allows treating the limited resources (e.g.
+hardware resources such as physical memory) as if there are more than they actually are (e.g.
+via virtual memory addresses and paging).
+Consequently, virtualization for the sake of ensuring availability requires \emph{sharing} of the resources that exist in reality.
+%%\paragraph{Isolation Hand-in-Hand with Sharing}
+%%\Ismail{put a few words on abstractions such as address space, stack region, snapshots ,etc.}
+%%\Colin{Not sure what you want here?}
+
+\subsection{Nominals}
 \label{sec:modalbackground}
-Modal logics are logics of contingent truth, enabling reasoning about cases where something is or may
-be true in alternate circumstances. These alternate circumstances are expressed via modal operators,
-which decorate a proposition with those alternate circumstances of its truth. This can be express
-conditions related to time~\cite{pnueli1977temporal} (e.g., $\diamond P$ for $P$ being true at some point in the future in \textsc{LTL}), necessity and/or possibility~\cite{harel1979first,pratt1976semantical} (e.g., $[C]P$ indicating that $P$ is true after execution
-of the program command $C$ terminates), or knowledge and belief~\cite{halpern1990knowledge,hintikka1962knowledge}
-(e.g., $\mathcal{K}_a P$ indicating that node $a$ in a distributed system possesses knowledge of $P$),
-or the location \emph{where} something is true~\cite{murphy2007type,gordon2019modal}.
-
-\paragraph{Nominals} 
+
+\subsubsection{Recapping Modal Operators in Program Specifications: Systems Perspective}
+The value of modal operators in program specifications is that they permit describing something that is true
+within some context(s), without needing to fully describe the context itself. Yet unlike shorthands,
+abbreviations, or general definitions, modal operators interact with the logic in systematic --- rather than ad hoc --- ways.
+
+For example, most modal operators $M$ have the property that if $P \Rightarrow Q$, then $M(P) \Rightarrow M(Q)$.
+Many modal operators distribute over conjunction, so $M(P\wedge Q)\Rightarrow (M(P) \wedge M(Q))$.
+So when a program state or behaviour can be captured by a modality, this interacts nicely with the rest of the logic
+in ways that simplify both specifications and proofs. Here we give a few examples of how this is true in
+classic work; the next section argues that this is \emph{particularly} true for systems software, because
+the concepts and reasoning that arise in low-level software are naturally captured by modalities.
+
+Although we focus on understanding the modal patterns in the program logics themselves for verifying the client programs (Section \ref{sec:definitions}), not in the logical machinery they implement, we think that it sheds light on how the modal operator utilized in the machinery of the program logics shows resemblance and can be lifted to the task of client verification.
+\paragraph{Temporal Operators}
+The best known class of modal logics in the systems community is undoubtedly temporal logics (whether \textsc{LTL}~\cite{pnueli1977temporal}, \textsc{CTL}~\cite{emerson1982using}, \textsc{TLA}~\cite{lamport1994temporal}, or others),
+due in part to Lamport's influence~\cite{lamport2002specifying} leading to its not-infrequent use in specifying
+distributed algorithms \cite{ongaro2014search,ailijiang2019wpaxos}.
+\subsubsection{Nominalization}
 Some classes of assertions benefit specifically from \emph{naming} the explicit conditions where they are true (as opposed to simply requiring them to be true \emph{somewhere} or \emph{everywhere} as in the most classic modal operators).
@@
 
-\subsection{Modal Operators in Program Specifications}
-The value of modal operators in program specifications is that they permit describing something that is true
-within some context(s), without needing to fully describe the context itself. Yet unlike shorthands,
-abbreviations, or general definitions, modal operators interact with the logic in systematic --- rather than ad hoc --- ways.
-
-For example, most modal operators $M$ have the property that if $P \Rightarrow Q$, then $M(P) \Rightarrow M(Q)$.
-Many modal operators distribute over conjunction, so $M(P\wedge Q)\Rightarrow (M(P) \wedge M(Q))$.
-So when a program state or behaviour can be captured by a modality, this interacts nicely with the rest of the logic
-in ways that simplify both specifications and proofs. Here we give a few examples of how this is true in
-classic work; the next section argues that this is \emph{particularly} true for systems software, because
-the concepts and reasoning that arise in low-level software are naturally captured by modalities.
-
-Although we focus on understanding the modal patterns in the program logics themselves for verifying the client programs (Section \ref{sec:definitions}), not in the logical machinery they implement, we think that it sheds light on how the modal operator utilized in the machinery of the program logics shows resemblance and can be lifted to the task of client verification.
-
-\paragraph{Temporal Operators}
-The best known class of modal logics in the systems community is undoubtedly temporal logics (whether \textsc{LTL}~\cite{pnueli1977temporal}, \textsc{CTL}~\cite{emerson1982using}, \textsc{TLA}~\cite{lamport1994temporal}, or others),
-due in part to Lamport's influence~\cite{lamport2002specifying} leading to its not-infrequent use in specifying
-distributed algorithms (e.g., \cite{ongaro2014search,ailijiang2019wpaxos}).
-
-\paragraph{Nominalization}
 An example utilization of state naming explicitly on the assertion appears in program logics such as Iris~\cite{jung2018iris}, which enables encoding of usage protocols (e.g.
@@
 An example more familiar to the systems community is Halpern et al.'s history of adapting modal logics of
-knowledge to deal with distributed systems~\cite{halpern1990knowledge,halpern1989modelling,halpern2017reasoning,halpern1990knowledge}.
+knowledge to deal with distributed systems~\cite{halpern1990knowledge,halpern1989modelling,halpern2017reasoning}.
 In most of that line of work, $\mathcal{K}_a(P)$ indicates that the node $a$ in the system \emph{knows} or \emph{possesses knowledge of} $P$
@@
 
-In each of these cases, the fact that these facts are described using modalities means, for example, that if a
+In each of these cases, the fact that these facts are described using modalities with the core modal property $P \Rightarrow Q$, then $M(P) \Rightarrow M(Q)$ means, for example, that if a
 process $p$ knows that the commit index is greater than 5 (e.g., $\mathcal{K}_p(\mathsf{commitIndex} > 5)$)
@@
 one would need to separately provide custom reasoning to conclude $\mathsf{minCommitIndex}(3)$.
-
-%\paragraph{Implementing the Logic of Impredicativity} Modal operators such as \emph{later} modality speculating over the fact that modal assertion holds at some state later in the execution are used in the logical machinery of program logics to handle the impredicative terms in the logic.
-Interestingly, similar reasoning principles using \emph{later} modality are also used in the verification of programs against store-buffers ~\cite{} which also requires speculation over the state in which stored writes appear in the memory.
-%
-%\paragraph{Persistent Weakest-Precondition} We already gave a rough encoding of Hoare Triple in terms of weakest-precondition, but a thorough version of our definition for the program logics of resources would be
-%\[
-%\Box (\textsf{P} \; -\ast \;\textsf{wp}\; e \; \{Q\})
-%\]
-%decorated with a necessity modality ensures that the weakest-precondition assertion 
-%$
-%    \textsf{P} \; -\ast \;\textsf{wp}\; e \; \{Q\}
-%$
-%inside the necessity modality is always valid by the exclusive ownership of the weakest-precondition; as a consequence, the whole specification 
-%$
-%\Box (\textsf{P} \; -\ast \;\textsf{wp}\; e \; \{Q\})
-%$ can be used multiple times -- i.e., a \emph{duplicable resource}.
-
-\section{Logical Decomposition of a System into its Constituents \emph{Contingently}}
+\section{Contingency Decomposition of a System}
+
+\subsection{Decomposing a System into its Constituents \emph{Contingently}}
 \label{sec:definitions}
@@
 %$\{\overset{ICut^{n}}{\hookrightarrow} (P) \}$ return $\{ P\}_{n} $
-\begin{tabular}{@{}cSSSp{5cm}@{}}
+\begin{tabular}{@{}clllp{3.25cm}@{}}
   \toprule
-& Resource Context & Resource Elements  &  Nominalization & Resource Context Steps \\ \midrule
-Post-Crash Modality ~\cite{tejthesis, perennialgit,tejperennial19} & $\Diamond \; P $  & $  \ell \mapsto_{n}^{\overline{\gamma}} v $ & Strong &  Crash Recovery   \\ 
+& {\scriptsize Resource Context} & {\scriptsize Resource Elements}  &  {\scriptsize Nominalization} & {\scriptsize Context Steps} \\ \midrule
+\scriptsize Post-Crash Modality ~\cite{tejthesis, perennialgit,tejperennial19} & \scriptsize $\Diamond \; P $  & \scriptsize $  \ell \mapsto_{n}^{\overline{\gamma}} v $ & \scriptsize Strong &  \scriptsize Crash Recovery    \\
 %GC-Modality &      & \multicolumn{2}{c}{\multirow{2}{*}{Because why not}} & 8                                   & & ~\cite{}\\
-NextGen Modality~\cite{larsnextgen25} &  $\overset{t}{\hookrightarrow} \; P$  & Own (t(a))  & Strong & Determined Based on the Model$^{*}$   \\
-StackRegion  Modality$^{*}$  ~\cite{larsnextgen25} & $\overset{ICut^{n}}{\hookrightarrow} \; P$      & $\fbox{n} \; \ell \mapsto v$                                 & Strong & Alloc and Return to/from stack\\
-Memory-Fence  Modality ~\cite{fsl,fsl++,derekrustbelt20} & $\triangle_{\pi}$ and $\triangledown_{\pi}$      & $\ell \mapsto v$                                 & Weak & Fence Acquire and Release  \\
-Address-Space Modality ~\cite{kuru2024modalabstractionsvirtualizingmemory} & [r]P     &  $\ell \mapsto v$ & Weak                         & Address-Space Switch  \\ 
-Ref-Count Modality ~\cite{amalreal2024}& @$_{\ell}$ P    &  $\ell_1 \mapsto v$ & Weak  & Allocating, Dropping and Sharing a Reference \\ \bottomrule
+\scriptsize NextGen Modality~\cite{larsnextgen25} &  \scriptsize $\overset{t}{\hookrightarrow} \; P$  &\scriptsize  Own (t(a))  & \scriptsize Strong & \scriptsize Determined Based on the Model$^{*}$   \\
+\scriptsize StackRegion  Modality$^{*}$  ~\cite{larsnextgen25} & \scriptsize $\overset{ICut^{n}}{\hookrightarrow} \; P$      & \scriptsize $\fbox{n} \; \ell \mapsto v$                                 & \scriptsize Strong & \scriptsize Alloc and Return to/from stack\\
+\scriptsize Memory-Fence  Modality ~\cite{fsl,fsl++,derekrustbelt20} & \scriptsize $\triangle_{\pi}$ and $\triangledown_{\pi}$      & \scriptsize $\ell \mapsto v$                                 & \scriptsize Weak & \scriptsize Fence Acquire and Release  \\
+\scriptsize Address-Space Modality ~\cite{kuru2024modalabstractionsvirtualizingmemory} & \scriptsize [r]P     &  \scriptsize $\ell \mapsto v$ & \scriptsize Weak                         & \scriptsize Address-Space Switch  \\
+\scriptsize Ref-Count Modality ~\cite{amalreal2024}&\scriptsize @$_{\ell}$ P    & \scriptsize $\ell_1 \mapsto v$ & \scriptsize Weak  & \scriptsize Allocating, Dropping and Sharing a Reference \\ \bottomrule
 \end{tabular}
 
-*The StackRegion Modality is an instance of NextGen (called the Independence Modality in \cite{larsnextgen25}).
+{\scriptsize *The StackRegion Modality is an instance of NextGen (called the Independence Modality in \cite{larsnextgen25})}.
 %and the $\Diamond$ Figure for Post-Crash Modality uses notationis taken from NextGen Modality paper ~\cite{larsnextgen25}.
@@
 
-In this section, we explain how we can systematically propose new modal abstractions for different types of systems.
+Lots of existing program logics for system verifications have a common structure, which maps to modalities with a couple extra dimensions of design.
 We summarize our discussion of these logics in Table \ref{table:decomposition}.
@@
 articulate those pieces, and call out the commonalities across a range of systems.
-
+\subsection{Resource}
 Consider first the address-space abstraction in an OS kernel.
@@
 
-\subsection{Resource Elements}
-Table \ref{table:decomposition} gives additional examples of systems and their corresponding resource elements,
-though none of the work in that table analyzes itself according to the structure we are giving.
-When reasoning about stack frame contents, the resource element would be a stack-memory points-to assertion ( $\fbox{n} \; \ell \mapsto v$)
+Table \ref{table:decomposition} gives additional examples of systems and their corresponding resource elements and contexts where these elements reside, though none of the work in that table analyzes itself according to the structure we are giving.
+
+Except for Post-Crash-Modality, one can think of the resource contexts in the first column in Table \ref{table:decomposition} as containers for the corresponding resource elements in the second column.
+
+\paragraph{Stack Regions} When reasoning about stack frame contents, the resource element would be a stack-memory points-to assertion ( $\fbox{n} \; \ell \mapsto v$)
 indicating that a certain offset into stack region $n$ holds value $v$. 
 %which is an instance of ownership assertion ($\textsf{Own(t(a))}$) for which a well-behaved transformation ($\textsf{t}$), e.g. $\textsf{ICut}^{n}$ algebra quantifying over valid stack region ids. 
-For virtual memory management, a \textit{virtual-points-to} ownership assertion pairing a virtual address ($\ell$) with data ($v$) in an address space is natural.
-When considering weak memory models, we also want points-to information (address-value mappings).
-When dealing with reference-counting APIs, we may care to specify reachability of memory nodes ($\ell \mapsto v$) in a certain context defined by a shared root address .
-A generalization of separation logic to deal with arbitrary transformations (the \emph{NextGen} modality) 
+\paragraph{Virtual Memory} For virtual memory management, a \textit{virtual-points-to} ownership assertion pairing a virtual address ($\ell$) with data ($v$) in an address space is natural.
+A process's address space with the root address \textsf{r} is an abstraction that is treated as a container for virtual address mappings, $\ell \mapsto v$;
+
+\paragraph{Weak-Memory} When considering weak memory models, we also want points-to information (address-value mappings).
+
+\paragraph{Reference Counting} When dealing with reference-counting APIs, we may care to specify reachability of memory nodes ($\ell \mapsto v$) in a certain context defined by a shared root address.
+A shared memory address $\ell$ can be the root of the graph that can be a container of memory nodes ($\ell_1 \mapsto v$) that are reachable from the root $\ell$.
+
+\paragraph{Post-Crash}
 %\Colin{Ismail, not sure how to describe this. We don't want to go into the weeds on the NextGen}
@@
 them, discarding assumptions about volatile state.
-%whose capability is considered to be relevant to be kept in the crash invariant $\fbox{C}_n$, and to be used when a crash occurs.\Colin{Ismail, not sure about last line above.
-Can you better explain gamma?}
-%regions of state (including auxiliary/ghost state) which persist after a crash (anything not identified
-
-%Perennial keeps the consistent state of persistent storage in a crash invariant at : multiple disk-points-to assertions $disk[\textsf{p}] \mapsto_{n} v$ in the \emph{crash} invariant $\fbox{C}_n$ became consistent at the n$^{\textsf{th}}$ successful completion of writes, i.e., generation number, from in-memory to disk.
-The in-memory updates ($p \mapsto_{n} v$) are also indexed with the generation number $n$.
-Since Perennial is built on top of Iris Separation Logic which implicitly decorates ownership assertions -- both disk and in-memory points-to assertions -- with ghost names ($\gamma$), Perennial utilizes this and bookkeeps points-to assertions ($\ell \mapsto_{n}^{\overline{\gamma}}$) of a certain version number $n$ as \emph{a resource element lifted by resource names} needed in case of a crash to distinguish the resources to be invalidated, i.e., discard the piece of the physical state that resides in the memory and preserve the resources that reside in the persistent disk.
-
-%\Colin{I don't understand this next bit, we need to discuss:}
-%\begin{remark}[Resource Element Names]
-%Regarding the names, like Iris resource names, all the resource elements in the second column have names; they are not subject to change --- \emph{static} --- when they are not parameterized by names and mentioned explicitly in the element when they do not matter for the contingent truth.
-%\end{remark}
-%\Colin{This in particular would benefit from an earlier intuition discussion}
-
-\subsection{Resource Contexts}
-Except for Post-Crash-Modality, one can think of the resource contexts in the first column in Table \ref{table:decomposition} as containers for the corresponding resource elements in the second column.
-For our example, a process's address space with the root address \textsf{r} is an abstraction that is treated as a container for virtual address mappings, $\ell \mapsto v$; likewise, a shared memory address $\ell$ can be the root of the graph that can be a container of memory nodes ($\ell_1 \mapsto v$) that are reachable from the root $\ell$.
-\looseness=-1
-
 A subtlety of the notion of a resource context is that, unlike the earlier examples, the context does not need to be a literal data structure.
@@
 technical constraints that are unrelated to our point here.}
-
+%whose capability is considered to be relevant to be kept in the crash invariant $\fbox{C}_n$, and to be used when a crash occurs.\Colin{Ismail, not sure about last line above.
+Can you better explain gamma?}
+%regions of state (including auxiliary/ghost state) which persist after a crash (anything not identified
+
+%Perennial keeps the consistent state of persistent storage in a crash invariant at : multiple disk-points-to assertions $disk[\textsf{p}] \mapsto_{n} v$ in the \emph{crash} invariant $\fbox{C}_n$ became consistent at the n$^{\textsf{th}}$ successful completion of writes, i.e., generation number, from in-memory to disk.
+The in-memory updates ($p \mapsto_{n} v$) are also indexed with the generation number $n$.
+Since Perennial is built on top of Iris Separation Logic which implicitly decorates ownership assertions -- both disk and in-memory points-to assertions -- with ghost names ($\gamma$), Perennial utilizes this and bookkeeps points-to assertions ($\ell \mapsto_{n}^{\overline{\gamma}}$) of a certain version number $n$ as \emph{a resource element lifted by resource names} needed in case of a crash to distinguish the resources to be invalidated, i.e., discard the piece of the physical state that resides in the memory and preserve the resources that reside in the persistent disk.
+
+%\Colin{I don't understand this next bit, we need to discuss:}
+%\begin{remark}[Resource Element Names]
+%Regarding the names, like Iris resource names, all the resource elements in the second column have names; they are not subject to change --- \emph{static} --- when they are not parameterized by names and mentioned explicitly in the element when they do not matter for the contingent truth.
+%\end{remark}
+%\Colin{This in particular would benefit from an earlier intuition discussion}
+
+%\section{Resource Contexts}
+%
+%\looseness=-1
+%
+%
+%
 %\begin{definition}[Transforming Resource Elements]
@@
 %any other resource elements named with the ones th
-
-\subsection{Resource Ambiance}
-While the central aspects of using a modality to capture contextual reasoning are the elements and contexts, there is \emph{sometimes} more to it.
-Sometimes there are multiple modes of execution (e.g., crashing, running in a particular address space),
-so not only state descriptions but also rules of the program logic itself must be decorated with
-a \emph{resource ambiance}:
-\begin{definition}[Resource Ambiance]
-Resource ambiance is a dependency of the logic's description of program behaviour on
-a particular mode of execution tied to the modality.
+%\section{Resource Ambiance}
+%While the central aspects of using a modality to capture contextual reasoning are the elements and contexts, there is \emph{sometimes} more to it.
+%Sometimes there are multiple modes of execution (e.g., crashing, running in a particular address space),
+%so not only state descriptions but also rules of the program logic itself must be decorated with
+%a \emph{resource ambiance}:
+%\begin{definition}[Resource Ambiance]
+%Resource ambiance is a dependency of the logic's description of program behaviour on
+%a particular mode of execution tied to the modality.
 %\emph{general form of contingent} status of resource elements of the same type against their resource context.
-The crucial aspect is to understand how \emph{identifiable} both the context and its elements should be: \emph{nominalization} of the context and its elements in the ambient logic, e.g., how should stack-points-to relations together with the StackRegion Modality should be \emph{referred} in a Hoare triple for Return step in the ambient separation logic.
-\end{definition}
-In formal presentations, this typically shows up with modified Hoare triples $\{P\}\;C\;\{Q\}_A$,
-where $P$, $C$, and $Q$ are the precondition, code, and postcondition that code satisfies assuming that precondition (as before) and $A$ is essentially a marker of how the reasoning about the program interacts with the modality --- such as a post-crash assertion, or an indication of what address space the verification is assuming.
-This means the proof rules for deciding how program operations affect knowledge of program state
-can depend on, restrict, or manipulate the context assumptions of reasoning about program behaviour, not just
-descriptions of program state at some moment in time.
-Not all use cases for modal assertions require an ambiance, only those where the modality corresponds to modes of execution.
-
+The crucial aspect is to understand how \emph{identifiable} both the context and its elements should be: \emph{nominalization} of the contex%t and its elements in the ambient logic, e.g., how should stack-points-to relations together with the StackRegion Modality should be \emph{referred} in a Hoare triple for Return step in the ambient separation logic.
+%\end{definition}
+%In formal presentations, this typically shows up with modified Hoare triples $\{P\}\;C\;\{Q\}_A$,
+%where $P$, $C$, and $Q$ are the precondition, code, and postcondition that code satisfies assuming that precondition (as before) and $A$ is essentially a marker of how the reasoning about the program interacts with the modality --- such as a post-crash assertion, or an indication of what address space the verification is assuming.
+%This means the proof rules for deciding how program operations affect knowledge of program state
+%can depend on, restrict, or manipulate the context assumptions of reasoning about program behaviour, not just
+%descriptions of program state at some moment in time.
+%Not all use cases for modal assertions require an ambiance, only those where the modality corresponds to modes of execution.
+%
 \subsection{Nominalization}
-Finally, another design point is the question of whether or not resource elements must explicitly track
+Finally, another design point is the question of whether or not resource element assertions must explicitly track
 their corresponding context, or if they implicitly pick up their context from where they are used.
@@
 contexts together marginally more complex: changing which modality an assertion is used with drastically changes its meaning. In the case of the Ref-Count modality, $@_\ell(\ell_1\mapsto v)$
-says that $\ell$ points to a reference count wrapping $\ell_1$, while placing the $\ell_q\mapsto v$ under a jump modality for a different location entails talking about a different region of memory.
+says that $\ell$ points to a reference count wrapping $\ell_1$, while placing the $\ell_1\mapsto v$ under a jump modality for a different location entails talking about a different region of memory.
 
@@
 
-\subsection{Making It Work}
-While the choice of resource elements, contexts, ambiance, and nominalization to capture
-the setting and contextually-relevant pieces of information help lay out the 
-shape of specifications based on how we would like to think about them, there are a few more steps
-beyond designing the specifications.
-
-It is necessary to work out how proof rules interact with the modalities --- which program
-operations or hardware primitives update machine state in a way that interacts with the modality?
-This could be all or some, depending on the use case.
-
-It is also necessary to implement the approach. This design has been most thoroughly explored
-with logics embedded in proof assistants (everything in Table \ref{table:decomposition} uses \textsf{Coq}), in which case a framework like \textsc{Iris}~\cite{jung2018iris} (used for all of our examples) will require proofs that each rule interacting with the modality is sound, but in exchange
-for that work and a few supplementary definitions, modal assertions can work naturally with existing tools for the logic~\cite{krebbers2018mosel}.
-However, it is also suitable for use with automated techniques, for example by writing
-new axioms corresponding to proof rules~\cite{gordon2019modal}, though better automation may require more work.
-
-%\section{Related Work}
-%\label{sec:relatedwork}
-%We discussed
-
 \section{Conclusion}
@@
 
-We have argued that modalities should be a go-to design when working out how to specify different kinds of systems code based on recent successes, and articulated an informal design process that seems to
-map out the degrees of variation present among those designs.
-It also captures (a slightly more organized version of) our own thought processes while
-coming up with designs for specifying distributed systems~\cite{gordon2019modal}, virtual memory managers~\cite{kuru2024modalabstractionsvirtualizingmemory}, and in ongoing work using different modalities for different parts of a copy-on-write filesystem (e.g., for assertions true in different snapshots).
+We have argued the essential patterns that help to choose the modalities when working out how to specify different kinds of systems code based on recent successes.
+It also captures (a slightly more organized version of) our own thinking in coming up with designs for specifying distributed systems~\cite{gordon2019modal}, virtual memory managers~\cite{kuru2024modalabstractionsvirtualizingmemory}, and in ongoing work using different modalities for different parts of a copy-on-write filesystem (e.g., for assertions true in different snapshots).
 So not only are the modalities useful
@@
 
-%%
-%% The acknowledgments section is defined using the "acks" environment
-%% (and NOT an unnumbered section). This ensures the proper
-%% identification of the section in the article metadata, and the
-%% consistent spelling of the heading.
-%\begin{acks}
-%To Robert, for the bagels and explaining CMYK and color spaces.
-%\end{acks}
-
+As our main conclusion and future work, we think that these essential patterns in the designs of modal abstractions discussed constitute the fundamentals of \emph{verification patterns} when working out how to specify different kinds of systems code.
 \bibliographystyle{splncs04}

### 2025-06-02 paper.tex
arxiv

@@
 %Accessing the hardware resources in the presence of sharing means having more resource references (references to the virtualized resources) to access the \emph{mapped} resources that are physical and potentially fewer.
-B-trees~\cite{}, page tables~\cite{}, and related structures both behave like maps, when the corresponding physical resources exist as such
+B-trees, page tables, and related structures both behave like maps, when the corresponding physical resources exist as such
 just in a different location.

### 2025-06-09 paper.tex
notes for Dali 2025 resubmission

@@
 %
-\authorrunning{I. Kuru et al.}
+\authorrunning{Kuru and Gordon}
 % First names are abbreviated in the running head.
@@
   and explaining how this approach might apply to other systems verification challenges.
+
+  \todo[inline]{Colin says: how about this framing instead of the second paragraph above:}
+  In this paper we examine how modalities have emerged as a common structure in formal verification
+  of low-level software, and explain how many recent examples naturally share 
+  common structure in the relationship between the modalities and software features they are used
+  to reason about.
+  We explain how the concept of a \emph{resource context} (a class of system resources
+  to reason about) naturally corresponds to families of modal operators indexed by
+  system data, and how this naturally leads to using modal assertions to describe
+  \emph{resource elements} (data in the relevant context).
+  We also describe extensions of this idea to ongoing work in formal verification
+  of filesystems code.
 \end{abstract}
@@
 \section{Introduction}\label{sec:intro}
+\todo[inline]{Colin says: These comments aren't the only places things need to change, but they're
+specific places I have suggestions.
+One thing to keep in mind is that the CfP lists two kinds of papers:
+short papers are for "original research or case studies, maybe in the incubation phase,"
+while long papers are for "original research, case studies, or surveys."
+This is clearly in the survey category, so we (a) should take advantage of the extra space to
+explain things clearly (15 pages, plus 2 pages for bibliography), and (b) lean into the survey nature explicitly.
+}
 Low-level systems exhibit certain patterns in their designs, especially when interacting with computing resources.
@@
 In this regard, we think that certain properties of modalities enable us to understand and do the verification challenges that show certain patterns.
+\todo[inline]{I think one of the biggest things we can do to reframe this is to explain the broader vision
+here before jumping into contributions.
+This is an audience that knows dynamic logic and modal logic more broadly, but perhaps not everyone is familiar
+with hybrid logic, and many will not be familiar with systems concepts.
+This is the place to provide high-level intuition about the domain we're surveying, tailored for this audience.
+I'd talk about dynamic logic in general, but then highlight the need for reasoning specific to certain
+compute resources to produce manageable specifications and proofs.
+Think big picture: why is this space interesting to logicians?
+}
+
 \subsection{Contributions to Identifying the Verification Patterns via Modal Abstractions}
@@
 We justify our perspective by discussing prior systems that have successfully used modalities for system verification, arguing that they fit into the verification design pattern we articulate, and explaining how this approach might apply to other systems' verification challenges.
+
 \paragraph{Identifying System Verification Challenges} We start with identifying common patterns in system verification: \emph{virtualization, sharing, and translation}.
@@
 Inspired by the concept of resource in the systems, we define what a resource and its context are in the modal abstractions.
-\paragraph{Introducing the Concept of \emph{Nominals}} Nominalization enables identifying a resource in a context.
+\paragraph{Introducing the Concept of \emph{Nominals}}
+\todo[inline]{Colin says: I'd reframe the way nominals are introduced. This discussion is good,
+but is a "systems-first" explanation. At least some of this audience will already know
+what classic nominals are, so need to be shown naming specific resources as an application
+of what they already know. (And some readers will need a very short intro/refresher.)
+}
+Nominalization enables identifying a resource in a context.
 For example, a transaction is a context of resources of in-memory updated disk blocks.
@@
 \paragraph{Taxonomy of the Current Modal Approaches in System Verification} Based on these concepts defined, we summarize contemporary verification efforts using modal abstractions.
-We choose them from different domains, for example, weak memory verification, storage persistence ~\cite{tejthesis, perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2024modalabstractionsvirtualizingmemory,amalreal2024} because we would like to justify that our \emph{definitions} are not domain dependent.
+We choose them from different domains, for example, weak memory verification, storage persistence ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2024modalabstractionsvirtualizingmemory,amalreal2024} because we would like to justify that our \emph{definitions} are not domain dependent.
 
@@
   \label{sec:systemsoft}
+  \todo[inline]{this section is already nice for introducing the ideas and what you mean by "systems software."
+  This section mostly needs a little more background on systems concepts like virtual memory, or what it means
+  to virtualize locations of blocks on disk}
 Systems software, in general, interfaces with an underlying computing architecture such that any software system at any higher level in the software stack can (at least indirectly) utilize the resources of the machine.
@@
 
+\todo[inline]{Colin says: the nominals section needs another shift for the new audience:
+You don't need to define modalities here, but do need to give initially the classic notion of nominal
+(name for a Kripke world) before bridging the gap to using the concept of naming
+for something a little different.
+}
 \subsection{Nominals}
@@
 \section{Contingency Decomposition of a System}
+\todo[inline]{Colin says: I don't think this section needs to change much,
+but it should be read carefully with the new audience in mind,
+possibly giving more details on what systems concepts each system is dealing with, for a non-systems audience.}
 
@@
 \label{sec:definitions}
+\todo[inline]{Colin says: I dropped the strong/weak nominalization column for space and readability;
+it can be explicit in the text, or maybe we can mark the system names with daggers or similar to indicate.}
 \begin{table*}[h]
@@
 %$\{\overset{ICut^{n}}{\hookrightarrow} (P) \}$ return $\{ P\}_{n} $
-\begin{tabular}{@{}clllp{3.25cm}@{}}
+\begin{tabular}{@{}lcccp{4cm}@{}}
   \toprule
-& {\scriptsize Resource Context} & {\scriptsize Resource Elements}  &  {\scriptsize Nominalization} & {\scriptsize Context Steps} \\ \midrule
-\scriptsize Post-Crash Modality ~\cite{tejthesis, perennialgit,tejperennial19} & \scriptsize $\Diamond \; P $  & \scriptsize $  \ell \mapsto_{n}^{\overline{\gamma}} v $ & \scriptsize Strong &  \scriptsize Crash Recovery    \\
+{\scriptsize Modality}& {\scriptsize Context} & {\scriptsize Elements}  &  {\scriptsize Nominalization} & {\scriptsize Context Steps} \\ \midrule
+\scriptsize Post-Crash~\cite{tejthesis,perennialgit,tejperennial19} & \scriptsize $\Diamond \; P $  & \scriptsize $  \ell \mapsto_{n}^{\overline{\gamma}} v $ & \scriptsize Strong &  \scriptsize Crash Recovery    \\
 %GC-Modality &      & \multicolumn{2}{c}{\multirow{2}{*}{Because why not}} & 8                                   & & ~\cite{}\\
-\scriptsize NextGen Modality~\cite{larsnextgen25} &  \scriptsize $\overset{t}{\hookrightarrow} \; P$  &\scriptsize  Own (t(a))  & \scriptsize Strong & \scriptsize Determined Based on the Model$^{*}$   \\
-\scriptsize StackRegion  Modality$^{*}$  ~\cite{larsnextgen25} & \scriptsize $\overset{ICut^{n}}{\hookrightarrow} \; P$      & \scriptsize $\fbox{n} \; \ell \mapsto v$                                 & \scriptsize Strong & \scriptsize Alloc and Return to/from stack\\
-\scriptsize Memory-Fence  Modality ~\cite{fsl,fsl++,derekrustbelt20} & \scriptsize $\triangle_{\pi}$ and $\triangledown_{\pi}$      & \scriptsize $\ell \mapsto v$                                 & \scriptsize Weak & \scriptsize Fence Acquire and Release  \\
-\scriptsize Address-Space Modality ~\cite{kuru2024modalabstractionsvirtualizingmemory} & \scriptsize [r]P     &  \scriptsize $\ell \mapsto v$ & \scriptsize Weak                         & \scriptsize Address-Space Switch  \\
-\scriptsize Ref-Count Modality ~\cite{amalreal2024}&\scriptsize @$_{\ell}$ P    & \scriptsize $\ell_1 \mapsto v$ & \scriptsize Weak  & \scriptsize Allocating, Dropping and Sharing a Reference \\ \bottomrule
+\scriptsize NextGen~\cite{larsnextgen25} &  \scriptsize $\overset{t}{\hookrightarrow} \; P$  &\scriptsize  Own (t(a))  & \scriptsize Strong & \scriptsize Determined Based on the Model$^{*}$   \\
+\scriptsize StackRegion$^{*}$  ~\cite{larsnextgen25} & \scriptsize $\overset{ICut^{n}}{\hookrightarrow} \; P$      & \scriptsize $\fbox{n} \; \ell \mapsto v$                                 & \scriptsize Strong & \scriptsize Alloc and Return to/from stack\\
+\scriptsize Memory-Fence~\cite{fsl,fsl++,derekrustbelt20} & \scriptsize $\triangle_{\pi}$ and $\triangledown_{\pi}$      & \scriptsize $\ell \mapsto v$                                 & \scriptsize Weak & \scriptsize Fence Acquire and Release  \\
+\scriptsize Address-Space~\cite{kuru2024modalabstractionsvirtualizingmemory} & \scriptsize [r]P     &  \scriptsize $\ell \mapsto v$ & \scriptsize Weak                         & \scriptsize Address-Space Switch  \\
+\scriptsize Ref-Count~\cite{amalreal2024}&\scriptsize @$_{\ell}$ P    & \scriptsize $\ell_1 \mapsto v$ & \scriptsize Weak  & \scriptsize Allocating, Dropping and Sharing a Reference \\ \bottomrule
 \end{tabular}
@@
 
-Lots of existing program logics for system verifications have a common structure, which maps to modalities with a couple extra dimensions of design.
+Many existing program logics for system verifications have a common structure, which maps to modalities with a couple extra dimensions of design.
 We summarize our discussion of these logics in Table \ref{table:decomposition}.
@@
 \label{sec:conclusion}
-
+\todo[inline]{this definitely needs adjusting for both audience and emphasis on this being a survey}
 We have argued the essential patterns that help to choose the modalities when working out how to specify different kinds of systems code based on recent successes.

### 2025-06-16 paper.tex
some work on colin's comments

@@
 \usepackage{lipsum} % for dummy text
+\usepackage{comment}
 
@@
   \todo[inline]{Colin says: how about this framing instead of the second paragraph above:}
+  \begin{comment}
   In this paper we examine how modalities have emerged as a common structure in formal verification
@@
   of filesystems code.
+  \end{comment}
 \end{abstract}
@@
 
-\section{Definitions}
+\section{Background on Low-Level Systems}
 Although they differ in the functionality they offer, low-level systems exhibit certain patterns of design and utilization of computing resources.
-  In this section, we argue the position that \emph{modalities}  
-  should be a go-to approach when specifying and verifying low-level systems code.
-  We explain how the concept of a \emph{resource context} helps guide the design of new modalities
-  for verification of systems code, and
-  we justify our perspective by discussing prior systems that have used modalities for systems verification
-  successfully, arguing that they fit into the verification design pattern we articulate,
-  and explaining how this approach might apply to other systems verification challenges.
-
-  To explain our ideas in the general systems understanding, we briefly recap some of the background and themes our ideas build on again,casting them in a certain way to bring out the relevance of our philosophy.
-  \subsection{Resources in Systems Software}
-  \label{sec:systemsoft}
   \todo[inline]{this section is already nice for introducing the ideas and what you mean by "systems software."
@@
 the construction of higher layers of the stack.
-
-\subsubsection{Virtualization}
+\subsection{Virtualization in Systems Software}
+\label{sec:background}
+\paragraph{Virtualization of Memory Locations}
+\label{sec:backgroundonmachinemodel}
+\todo{change the text}
+Memory addresses seen by programs running on modern computers are \emph{virtualized}: the address observed by a running program generally will not correspond directly to the physical location in memory, and may not even correspond to a physical location that \emph{exists} in the machine.
+Instead, these \emph{virtual} addresses are translated to \emph{physical} addresses that correspond directly to locations in RAM.
+On most modern architectures, this translation is performed through cooperation of the hardware and OS kernel: while executing an instruction that dereferences a (virtual) address, the CPU's \emph{memory management unit} (\textsc{MMU}) implements an architecture-specific process of \emph{address translation}, resulting in a physical address used to access the cache\footnote{Technically, for performance reasons most caches are indexed with parts of the virtual address, but tagged with the physical data addresses, so cache lookups and address translations can proceed in parallel.} and/or memory-bus.
+
+\begin{figure}
+    \includegraphics[width=\columnwidth]{pagetables2.png}
+    \caption{x86-64 page table lookups.}
+    \label{fig:pagetables}
+\end{figure}
+
+On the x86-64 architecture, the input to the \textsc{MMU}'s address translation is a sparse hierarchical set of tables calle d \emph{page tables} (referring to pages of memory).
+As Figure \ref{fig:pagetables} shows, address translation proceeds by repeatedly taking designated slices of the virtual address and indexing into a table.
+On x86-64, standard configurations use 4 levels of page tables, labelled levels 4 through 1, with lookups in the level 1 page table resulting in the actual page of physical memory holding the requested data, and the low-order 12 bits being used to index into this page.\footnote{Technically levels 1--3 have designated names originating when there were fewer levels.
+We omit those names to avoid confusion, and for greater consistency with the recent addition of a 5th level on the most recent processors, where levels 4 and 5 are simply known by their numbers.}  Because address translation traverses a series of tables, the translation process or algorithm is sometimes referred to as a \emph{page-table walk}.
+While Figure \ref{fig:pagetables} and most of our details are specific to the x86-64 architecture, ARMv8 (a.k.a.\ \texttt{aarch64}), RISC-V, and PowerPC use similar hierarchical page tables for address translation.
+
+The entries of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set), and a range of additional bits which can be used to control caching, write-through, and more.
+This paper will only consider the present bit (0), the read-write bit (1), and the accessed bit (5), which is set to 1 by the hardware every time an entry is used, allowing the OS to determine which pieces of memory are used or unused.
+
+The page tables mapping virtual addresses to physical addresses are managed by the OS.
+Typically each process has its on virtul-to-physical mappings represented by its own page table, which the OS registers with the CPU by storing it in a specific register (\texttt{cr3}) as part of switching to a new process.
+Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section) is how the OS ensures memory isolation between processes.
+
+If an instruction is executed that accesses a virtual address that either has no mapping, or does not have a mapping permitting the kind of access that was performed (e.g., the instruction was a memory write, but the relevant address range was marked read-only in the relevant page table entry), the hardware triggers a \emph{page fault}, transferring control to a \emph{page fault handler} registered with the hardware by the OS, allowing it to take corrective action (described in the next section).
+
+\paragraph{Virtual Memory Managers}
+\label{sec:backgroundonvmm}
+\todo{change the text}
+Most OS kernels have a component called a \emph{Virtual Memory Manager} (VMM)\footnote{Not to be confused with Virtual Machine Monitor.
+In this paper we focus on non-virtualized scenarios, but hardware virtualization extensions for both x86-64 and ARM make use of an additional set of page tables translating what a \emph{guest} considers to be its (virutalized) physical memory to actual physical memory.
+Our contributions should offer value in this scenario as well.}
+which is responsible for setting up page table mappings and for taking action when a page fault occurs.
+Often, a page fault indicates a bug in a program --- for example, a null pointer dereference crashes a program not because the hardware designates it an error, but because most OSes refuse to map the first 4K worth of virtual addresses, so dereferencing \texttt{NULL} results in a page fault.
+In such cases, the OS will terminate the program.
+
+In other cases the OS uses page faults to implement specialized functionality and optimizations.
+A key example of this is (confusingly) called \emph{paging}: saving room in physical RAM by waiting until memory addresses for code are accessed before copying them to RAM (saving process startup time); or copying memory that has not recently been accessed to disk (i.e., in a \emph{swap} file or partition) and marking those page table entries invalid so a future page fault will give the OS the chance to copy the relevant data back into memory when the program tries to access it again.
+
+The key pieces of functionality a VMM must implement itself, or have available in another OS component (depending on design), are
+adding a new page mapping (whether the mapped page contains zeros, file data, or swap data), and removing an existing page mapping.
+While this initially sounds like relatively modest functionality whose implementation may be complicated by hardware subtleties, a number of more serious complexities must be handled.
+
+For one, virtual addresses may \emph{alias} when more than one is mapped to the same physical memory.
+This can occur when a process requests it (e.g., via \texttt{mmap} or \texttt{shm\_open} on UNIX-style kernels).
+It also occurs automatically in most general-purpose kernels: while for normal programs the OS uses the MMU to isolate and abstract physical memory, the kernel's own functionality is often easier to implement if the kernel can directly access memory given its physical address, for example when interacting with hardware devices.
+For this reason, most kernels contain an \emph{identity map} of all physical memory in a machine.
+In some cases this is a literal identity map, where for every virtual address $p$, interpreting $p$ as a virtual address instead will map back to $p$ as a physical address.
+In other cases there is an offset involved, where to access physical address $p$ the kernel can access virtual address $p+\mathit{offset}$.
+This identity map saves the kernel the performance cost of having to continually add and remove transient mappings simply to briefly access a specific physical memory region.
+But this coexists with additional virtual addresses (those used by most kernel code and data structures), in a form of virtual address aliasing.
+Note that virtual address aliasing is both prevalent in general-purpose kernels, and violates the core assumption of most memory models assumed by verified compilers like CompCert, which assume no virtual address aliasing.
+
+Another complicating factor is that portions of page tables can be shared, i.e., trees of page tables can overlap.
+One common way this used to happen was for one subtree of page tables to map the kernel if linked in appropriately to a higher-level page table, and then to reuse that kernel mapping portion in every process's page tables.
+This was mostly phased out in the wake of Spectre
+%\mytodo{Colin finish}
+
+These dove-tail with functionality in the OS scheduler, which since it is responsible for dealing with multiple address spaces, must keep track of which virtual addresses are valid (and in what way) in which address spaces, where some virtual addresses are valid in only a single address space (e.g., a code address for a particular usermode process), while others are valid in all address spaces (e.g., kernel data structure pointers).
+The VMM must maintain some of these assumptions on behalf of the rest of the kernel, for example by guaranteeing that a certain range of virtual addresses (corresponding to the kernel's code and data) are valid in every address space.
+
+\paragraph{Virtualization of Disk-Blocks in Filesystems} Likewise, 
+\paragraph{Consistency of Disk Layout} ...
+\paragraph{Memory Reclamation in Client Code}...
+\paragraph{Consistency of Reclaimed Memory}...
+
+  \section{Resources in Systems Software}
+  \label{sec:systemsoft}
+   In this section, we argue the position that \emph{modalities}  
+  should be a go-to approach when specifying and verifying low-level systems code.
+  We explain how the concept of a \emph{resource context} helps guide the design of new modalities
+  for verification of systems code, and
+  we justify our perspective by discussing prior systems that have used modalities for systems verification
+  successfully, arguing that they fit into the verification design pattern we articulate,
+  and explaining how this approach might apply to other systems verification challenges.
+
+  To explain our ideas in the general systems understanding, we briefly recap some of the background and themes our ideas build on again,casting them in a certain way to bring out the relevance of our philosophy.
+
+\subsection{Virtualization}
 One of the most common forms of abstraction provided by systems software is \emph{virtualization},
@@
 
-\subsubsection{Recapping Modal Operators in Program Specifications: Systems Perspective}
-The value of modal operators in program specifications is that they permit describing something that is true
-within some context(s), without needing to fully describe the context itself. Yet unlike shorthands,
-abbreviations, or general definitions, modal operators interact with the logic in systematic --- rather than ad hoc --- ways.
-
-For example, most modal operators $M$ have the property that if $P \Rightarrow Q$, then $M(P) \Rightarrow M(Q)$.
-Many modal operators distribute over conjunction, so $M(P\wedge Q)\Rightarrow (M(P) \wedge M(Q))$.
-So when a program state or behaviour can be captured by a modality, this interacts nicely with the rest of the logic
-in ways that simplify both specifications and proofs. Here we give a few examples of how this is true in
-classic work; the next section argues that this is \emph{particularly} true for systems software, because
-the concepts and reasoning that arise in low-level software are naturally captured by modalities.
-
-Although we focus on understanding the modal patterns in the program logics themselves for verifying the client programs (Section \ref{sec:definitions}), not in the logical machinery they implement, we think that it sheds light on how the modal operator utilized in the machinery of the program logics shows resemblance and can be lifted to the task of client verification.
+\paragraph{Recapping Modal Operators in Program Specifications: Systems Perspective}
+We believe that for the kinds of informal reasoning and the sorts of data structures discussed in the previous section, ideas from modal logic are a promising approach to formalization and verification.
+Broadly speaking modal logics incorporate \emph{modal operators}, which take as arguments a proposition expected to be true in another time~\cite{pnueli1977temporal}, place~\cite{gordon2019modal,goranko1996hierarchies,areces2001hybrid,gargov1993modal,murphy2004symmetric}, circumstance or point-of-view~\cite{hintikka1962knowledge,halpern1985guide}, and result in a proposition true in the \emph{current} time, place, or circumstance at which the truth of the use of the modal operator is being evaluated.
+Classic examples include modal necessity $\square P$ describing that $P$ is \emph{necessarily} true, $\mathsf{G}~P$ meaning $P$ is true \emph{globally} (i.e., forever from this time onwards in temporal logics), or $K_i(P)$ describing that a particular participant $i$ \emph{knows} that $P$ is true.
+The latter is an example of \emph{multimodal} logic, where there is an family of modalities (modal operators) parameterized by some dimension of interest (there, participants).
+A closely related variant of multimodal logic is \emph{dynamic logic}~\cite{pratt1976semantical}, a logic of weakest preconditions~\cite{dijkstra-75} which works with modalities of the form $[p](P)$, which states that \emph{in the current program state}, \emph{if} program $p$ is run then afterwards $P$ will hold (modulo non-termination).
+This same idea is used to encode Hoare triples in the Iris program logic~\cite{krebbers2017essence}, using the same encoding as in Pratt's original presentation, where a Hoare triple $\{P\}C\{Q\}$ is encoded as $P\rightarrow[C](Q)$.
+A hallmark of a unary logical operator $M$ being a modality is if $M$ satisfies a property akin to $(\varphi\rightarrow\psi)\rightarrow M(\varphi)\rightarrow M(\psi)$, which roughly states that modus ponens holds under the modality.\footnote{Afficionados of modal logic will note that this property is not quite Axiom K (which requires the initial implication to also hold under $M$), but follows from K and a necessitation rule $\varphi\rightarrow M(\varphi)$.
+Non-necessitive modalities typically satisfy the weaker property we call out above.} This and related laws allow transporting general reasoning into the specific contexts or hypotheticals modeled by the modalities.
+
+We build on ideas from \textit{hybrid logic}~\cite{blackburn1995hybrid,goranko1996hierarchies,gargov1993modal,areces2001hybrid,brauner2010hybrid}.
+Hybrid logics are a class of modal logics that add two primary new concepts to a logic.
+Nominals $\loc \in \textsf{Loc}$ uniquely identify points in a model (e.g., a particular state), so there is exactly one point in the model where a nominal $\loc$ is true.
+Satisfaction operators are modal operators indexed by nominals, which enable claims about the truth of another proposition at some arbitrary point in the model identified by a nominal.
+Traditionally $@_l\varphi$ asserts that $\varphi$ is true in the state identified by $l$.
+Conceptually, given our context, we might like $l$ to indicate truth in a particular data structure $l$.
+However, most prior work on hybrid logics works in classical logics rather than a substructural setting like separation logic.
+We adapt these ideas to assertions of the form $\varphi @[l]\sigma$\colinsays{I still think this syntax is weird; why phi on the left?
+-- Iso: no answer :) right:) I will change it.} asserts that $\varphi$ is true in a data structure with \emph{logical} state $\sigma$ \emph{at location} $l$.
+In practice we must design variants of this for each kind of structure we wish to apply this to, e.g. tree nodes, free list, or snapshots.
+This section provides details of and models for each modal abstraction, focusing on how each data structure can act as a model for a data-structure-specific logic.
+We continue to motivate the development with aspects of the write amplification process described in the previous section; Section \ref{sec:evaluation} provides further examples using the modal abstractions in proofs.
+
+\paragraph{Shallow Modalities in Iris}
+\label{sec:modal_in_iris}
+Classically modal operators are given direct semantics in terms of an underlying model, which for \iris would be a collection of \textsc{CMRA}s; this is the case for \iris's one built-in modality for step-indexing~\cite{krebbers2017essence}.
+However in general, many modalities can be given proof-theoretic semantics directly within a logic.
+This is the case for \iris's weakest precondition modality (which is defined within the \iris logic itself), as well as for a wide range of other modalities in the literature~\cite{restall2002introduction}.
+These modalities are typically multi-modal, with a form of roughly $M~\overline{ctxt}~\phi\triangleq\ldots\rightarrow \phi~\overline{ctxt}$.
+In a substructural setting~\cite{restall1993modalities,krebbers2017essence} the implication $\rightarrow$ is typically replaced with substructural implication $\sepimp$.
+
+This definition also reveals a subtlety about the denotation of assertions $\phi$ --- they are effectively \emph{predicates} of some kind, functions from some possible world into some logical algebra such as a BI-algebra~\cite{ohearn1999bunched}.
+Taking this interpretation literally gives rise to a convenient way to embed new modalities in \iris in a way well-suited to our use cases.
+If $B$ is a type with a BI-algebra structure, then any space of predicate functions (e.g., $\mathsf{val}\rightarrow B$ as a simplification) also carries a BI-algebra structure, lifted from $B$, often called the \emph{pointwise lifting} of the algebra.
+We will exploit this repeatedly, lifting the BI-algebra structure from \iris itself to \iris predicates used as assertions within our own modalities.
+Connecting this to \iris's typeclasses for interactive proofs~\cite{krebbers-17} then permits working within the embedded logic as in \iris.
+
+In addition, the modal definitions themselves may represent ownership of resources in a substructural setting~\cite{restall1993modalities,dovsen1992modal,kamide2002kripke}.
+This will allow our modalities to simultaneously represent ownership of certain resources used for interpreting the modality itself, as well as whatever resource ownership is implied by the modal argument.
 \paragraph{Temporal Operators}

### 2025-06-16 paper.tex
more restructuring

@@
 }
-
+\newcommand{\sepimp}{\mathrel{-\mkern-6mu*}}
 \newcommand{\NULL}{\texttt{NULL}}
@@
 \paragraph{Virtualization of Disk-Blocks in Filesystems} Likewise, 
-\paragraph{Consistency of Disk Layout} ...
-\paragraph{Memory Reclamation in Client Code}...
-\paragraph{Consistency of Reclaimed Memory}...
+\paragraph{Consistency of Disk Layout} \todo{}
+\paragraph{Memory Reclamation in Client Code}\todo{iso:speak about Amal's stuff}
+\paragraph{Consistency of Reclaimed Memory}\todo{iso: speak about the operator for which the hybrid logic stuff is usefull }
 
@@
 }
-\subsection{Nominals}
+\subsection{Hybrid Logic Dynamic Logic and Nominals in Program Logics}
 \label{sec:modalbackground}
 
-\paragraph{Recapping Modal Operators in Program Specifications: Systems Perspective}
+\paragraph{Dynamic Logic's Satisfaction Operator for Program Logics}
 We believe that for the kinds of informal reasoning and the sorts of data structures discussed in the previous section, ideas from modal logic are a promising approach to formalization and verification.
@@
 
-We build on ideas from \textit{hybrid logic}~\cite{blackburn1995hybrid,goranko1996hierarchies,gargov1993modal,areces2001hybrid,brauner2010hybrid}.
-Hybrid logics are a class of modal logics that add two primary new concepts to a logic.
+\paragraph{Hybrid Logic} Hybrid logics~\cite{blackburn1995hybrid,goranko1996hierarchies,gargov1993modal,areces2001hybrid,brauner2010hybrid} are a class of modal logics that add two primary new concepts to a logic.
 Nominals $\loc \in \textsf{Loc}$ uniquely identify points in a model (e.g., a particular state), so there is exactly one point in the model where a nominal $\loc$ is true.
@@
 However, most prior work on hybrid logics works in classical logics rather than a substructural setting like separation logic.
-We adapt these ideas to assertions of the form $\varphi @[l]\sigma$\colinsays{I still think this syntax is weird; why phi on the left?
--- Iso: no answer :) right:) I will change it.} asserts that $\varphi$ is true in a data structure with \emph{logical} state $\sigma$ \emph{at location} $l$.
+We adapt these ideas to assertions of the form $@[l]\varphi$asserts that $\varphi$ is true in a data structure with \emph{logical} state $\sigma$ \emph{at location} $l$.
 In practice we must design variants of this for each kind of structure we wish to apply this to, e.g. tree nodes, free list, or snapshots.
-This section provides details of and models for each modal abstraction, focusing on how each data structure can act as a model for a data-structure-specific logic.
-We continue to motivate the development with aspects of the write amplification process described in the previous section; Section \ref{sec:evaluation} provides further examples using the modal abstractions in proofs.
-
-\paragraph{Shallow Modalities in Iris}
+%This section provides details of and models for each modal abstraction, focusing on how each data structure can act as a model for a data-structure-specific logic.
+%We continue to motivate the development with aspects of the write amplification process described in the previous section; Section \ref{sec:evaluation} provides further examples using the modal abstractions in proofs.
+\paragraph{Shallow Modalities in Iris Separation Logic}
 \label{sec:modal_in_iris}
-Classically modal operators are given direct semantics in terms of an underlying model, which for \iris would be a collection of \textsc{CMRA}s; this is the case for \iris's one built-in modality for step-indexing~\cite{krebbers2017essence}.
+Classically modal operators are given direct semantics in terms of an underlying model, which for \textsf{Iris} would be a collection of \textsc{CMRA}s; this is the case for \textsf{Iris}'s one built-in modality for step-indexing~\cite{krebbers2017essence}.
 However in general, many modalities can be given proof-theoretic semantics directly within a logic.
-This is the case for \iris's weakest precondition modality (which is defined within the \iris logic itself), as well as for a wide range of other modalities in the literature~\cite{restall2002introduction}.
+This is the case for \textsf{Iris}'s weakest precondition modality (which is defined within the \textsf{Iris} logic itself), as well as for a wide range of other modalities in the literature~\cite{restall2002introduction}.
 These modalities are typically multi-modal, with a form of roughly $M~\overline{ctxt}~\phi\triangleq\ldots\rightarrow \phi~\overline{ctxt}$.
@@
 This definition also reveals a subtlety about the denotation of assertions $\phi$ --- they are effectively \emph{predicates} of some kind, functions from some possible world into some logical algebra such as a BI-algebra~\cite{ohearn1999bunched}.
-Taking this interpretation literally gives rise to a convenient way to embed new modalities in \iris in a way well-suited to our use cases.
+Taking this interpretation literally gives rise to a convenient way to embed new modalities in \textsf{Iris} in a way well-suited to our use cases.
 If $B$ is a type with a BI-algebra structure, then any space of predicate functions (e.g., $\mathsf{val}\rightarrow B$ as a simplification) also carries a BI-algebra structure, lifted from $B$, often called the \emph{pointwise lifting} of the algebra.
-We will exploit this repeatedly, lifting the BI-algebra structure from \iris itself to \iris predicates used as assertions within our own modalities.
-Connecting this to \iris's typeclasses for interactive proofs~\cite{krebbers-17} then permits working within the embedded logic as in \iris.
+We will exploit this repeatedly, lifting the BI-algebra structure from \textsf{Iris} itself to \textsf{Iris} predicates used as assertions within our own modalities.
+Connecting this to \textsf{Iris}'s typeclasses for interactive proofs~\cite{krebbers-17} then permits working within the embedded logic as in \textsf{Iris}.
 

### 2025-06-16 paper.tex
+

@@
 
-\paragraph{Virtual Memory Managers}
+\paragraph{Virtual Memory Managers with Multiple Address-Spaces}
 \label{sec:backgroundonvmm}
@@
 While this initially sounds like relatively modest functionality whose implementation may be complicated by hardware subtleties, a number of more serious complexities must be handled.
-
-For one, virtual addresses may \emph{alias} when more than one is mapped to the same physical memory.
+%
+%For one, virtual addresses may \emph{alias} when more than one is mapped to the same physical memory.
 This can occur when a process requests it (e.g., via \texttt{mmap} or \texttt{shm\_open} on UNIX-style kernels).
@@
 Note that virtual address aliasing is both prevalent in general-purpose kernels, and violates the core assumption of most memory models assumed by verified compilers like CompCert, which assume no virtual address aliasing.
-
-Another complicating factor is that portions of page tables can be shared, i.e., trees of page tables can overlap.
+%
+%Another complicating factor is that portions of page tables can be shared, i.e., trees of page tables can overlap.
 One common way this used to happen was for one subtree of page tables to map the kernel if linked in appropriately to a higher-level page table, and then to reuse that kernel mapping portion in every process's page tables.
@@
 
-These dove-tail with functionality in the OS scheduler, which since it is responsible for dealing with multiple address spaces, must keep track of which virtual addresses are valid (and in what way) in which address spaces, where some virtual addresses are valid in only a single address space (e.g., a code address for a particular usermode process), while others are valid in all address spaces (e.g., kernel data structure pointers).
+The OS scheduler, since it is responsible for dealing with multiple address spaces, must keep track of which virtual addresses are valid (and in what way) in which address spaces, where some virtual addresses are valid in only a single address space (e.g., a code address for a particular usermode process), while others are valid in all address spaces (e.g., kernel data structure pointers).
 The VMM must maintain some of these assumptions on behalf of the rest of the kernel, for example by guaranteeing that a certain range of virtual addresses (corresponding to the kernel's code and data) are valid in every address space.
@@
 \paragraph{Consistency of Reclaimed Memory}\todo{iso: speak about the operator for which the hybrid logic stuff is usefull }
+
+\subsection{Speculative Observation in Systems Software}
+\label{sec:speculativeobservation}
 

### 2025-06-16 paper.tex
+

@@
 %%%%%\usepackage{geometry}
-
+\usepackage{wrapfig}
 \usepackage[rawfloats=true]{floatrow} 
@@
 \usepackage{afterpage}
+\usepackage{fontawesome}
+\usepackage[utf8]{inputenc}
+
 %%sequeezeeee
@@
 \addtolength{\dbltextfloatsep}{-10pt}
+
+\newcommand{\ptree}{\text{\faSitemap}@[\ell]}
+\newcommand{\mtree}{\varupphi\text{\faSitemap}@[\ell]^{\Pi}_{\pi}(\sigma)}
+\newcommand{\oprtree}[4]{#3\;\text{\faSitemap}@[\textsf{#1_{#2}}]^{\Pi}_{\pi}(#4)}
+
+\newcommand{\osubtree}[4]{#2\;\text{\faSitemap}@[\textsf{node}_{#1}]^{\Pi}_{\pi}(#3_#4)}
+\newcommand{\otree}[3]{#2\;\text{\faSitemap}@[\textsf{node}_{#1}]^{\Pi}_{\pi}(#3)}
+\newcommand{\oltree}[3]{#2\;\text{\faSitemap}@[#1]^{\Pi}_{\pi}(#3)}
 
@@
 
-\paragraph{Virtualization of Disk-Blocks in Filesystems} Likewise, 
+\paragraph{Virtualization of Disk-Blocks in Filesystems}
+
 \paragraph{Consistency of Disk Layout} \todo{}
@@
 they are semantically --- logically --- partial maps, worth treating as such in verification. 
-
+\paragraph{Modality for Recursive Modal Contexts on Trees}
+\label{sec:overviewnode}
+
+To reason about states of trees and subtrees in particular, we work with a modality for stating an assertion that is true \emph{of all nodes within a tree}, identifying the tree by a root note.
+We refer to the resources that provide the context of these modal assertions as the \emph{modal context}; for a recursively defined tree like those in \textsf{Modfs}, these contexts are recursively defined in terms of the tree structure.
+
+%We define the modal context as the resources (e.g.
+a btree node tree, free page id list) which are accessed under some relation defined within the modal definition (e.g.
+being children of a cursor node).
+The choice of modal context can be structurally dependent on the semantics, in our case \textsc{CoW} semantics.
+
+Write amplification traverses the overall tree depth-first, transforming the tree from the bottom up, one full subtree at a time.
+Each time a node is split, it results in disconnecting some nodes from the next snapshot, sharing other nodes, and linking some newly-allocated nodes into the new snapshot, as in Figure \ref{fig:cow_index_tree_update}.
+Nodes are split (or deemed not to require splitting) one at a time, and considered only after all child nodes have been amplified.
+Concepually this means the amplification code's proof must track the fact that at any point in time, there are regions of the tree where each node has already been amplified (e.g., $\phi_{\textsf{amplified}}$ definitely holds for that node), and regions where nodes may or may not require amplification, and this will not be known until those nodes are visited (i.e., some basic validity $\phi_{\textsf{valid}}$ is known to hold, but not the stronger $\phi_{\textsf{amplified}}$).
+
+%\setlength\intextsep{0pt}
+\begin{wrapfigure}{r}{0.25\textwidth}
+  \centering  \includegraphics[width=0.25\textwidth]{entail.png}
+  \caption{Understanding the level of $\varphi$ satisfaction in the \textsc{CoW} tree of modal nodes}
+   \label{fig:nodeentail} 
+\end{wrapfigure}
+We use a modality similar to $\text{\faSitemap}@[\ell]\phi$ to indicate that $\phi$ is true of the tree node rooted at $\ell$, as well as at \emph{all reachable children} of that node.
+This means that establishing a particular $\phi$ throughout a tree must be done in a bottom-up manner (see Figure \ref{fig:nodeentail}).
+Capturing this modally, rather than with more typical ad-hoc hand-crafted invariants, offers advantages in a number of proof steps, in addition to concisely capturing intuitive reasoning about subtrees with all nodes satisfying a particular property.
+One basic example is that while the actual disk sync code requires the full filesystem tree to be in an amplified state (hence write amplification), most of the filesystem works with more general trees.
+Since $\phi_{\textsf{amplified}}\vdash\phi_{\textsf{valid}}$, using our \faSitemap modality we can conclude $\text{\faSitemap}@[\ell]\phi_{\textsf{amplified}}\vdash\text{\faSitemap}@[\ell]\phi_{\textsf{valid}}$ by an analogue of Axiom K.
+There are also other places in the code where a subtree satisfying a stronger assertion at all nodes must be linked into a larger tree satisfying only a weaker assertion overall: predicate representing the valid tree as the tree of nodes which are backed by resources allocated pages, i.e.
+$\phi_{\textsf{valid}}$, is weaker than the predicate representing the tree of nodes being currently amplified in the current snapshot.
+Again, the same reasoning principle makes this straightforward.
+As alluded to earlier, our $\phi_{\textsf{amplified}}$ and company are in fact \emph{predicates} parameterized by a choice of node, though by lifing \textsf{Iris}'s BI structure they can use all of the standard separation logic connectives.
+The rest of this subsection develops this modality in full detail.
+%\colinsaysinl{I cut some text here about the three-way split (commented out) that I wasn't sure how to use, but maybe I deleted something important?}
+%To put it concretely, as shown in Figure \ref{fig:cow_index_tree.JPG}, \textsc{CoW} semantics determine what we can claim on the structure of the filesystem indexing tree:
+%\begin{itemize}
+%\item The bottom-up direction splits the tree into the two half at the cursor node: 1.
+the bottom part accessible from the cursor node and per-node-amplification is reflected in case needed 2.
+the upper part on which amplifcation has not reflected yet.
+%\item Horizontally, amplifcation path splits the bottom part of the tree into three different contexts: untouched tree nodes, nodes on the amplification path, and the ones generated as part of node splitting.
+%\end{itemize}
+%Based on these two observations, in fact, bottom-up amplification determines the choice of modal context by giving the boundries of \textit{any claimable knowledge} (e.g.
+$\varphi_{\textsf{amplified}}$, $\varphi_{\textsf{untouched}}$): a subtree bounded by the cursor node.
+\subsection{Speculation in Consistency}
 %\subsection{Sharing} Sharing is one of the fundamental reasons for kernels to {virtualize} resources,

### 2025-06-16 paper.tex
+

@@
 they are semantically --- logically --- partial maps, worth treating as such in verification.
-\paragraph{Modality for Recursive Modal Contexts on Trees}
+
+\subsection{Speculation in Consistency}
+%\subsection{Sharing} Sharing is one of the fundamental reasons for kernels to {virtualize} resources,
+%for creating the illusion of exclusive access to virtualized versions of shared resources, including CPU time (via preemption).
+%An implementation of sharing relies on two concepts: \emph{indirection} and \emph{coordination}.
+Indirection is a method that allows the mapping of resources of one type into another.
+By doing so, it allows treating the limited resources (e.g.
+hardware resources such as physical memory) as if there are more than they actually are (e.g.
+via virtual memory addresses and paging).
+Consequently, virtualization for the sake of ensuring availability requires \emph{sharing} of the resources that exist in reality.
+%%\paragraph{Isolation Hand-in-Hand with Sharing}
+%%\Ismail{put a few words on abstractions such as address space, stack region, snapshots ,etc.}
+%%\Colin{Not sure what you want here?}
+
+\todo[inline]{Colin says: the nominals section needs another shift for the new audience:
+You don't need to define modalities here, but do need to give initially the classic notion of nominal
+(name for a Kripke world) before bridging the gap to using the concept of naming
+for something a little different.
+}
+\subsection{Hybrid Logic Dynamic Logic and Nominals in Program Logics}
+\label{sec:modalbackground}
+
+\paragraph{Dynamic Logic's Satisfaction Operator for Program Logics}
+We believe that for the kinds of informal reasoning and the sorts of data structures discussed in the previous section, ideas from modal logic are a promising approach to formalization and verification.
+Broadly speaking modal logics incorporate \emph{modal operators}, which take as arguments a proposition expected to be true in another time~\cite{pnueli1977temporal}, place~\cite{gordon2019modal,goranko1996hierarchies,areces2001hybrid,gargov1993modal,murphy2004symmetric}, circumstance or point-of-view~\cite{hintikka1962knowledge,halpern1985guide}, and result in a proposition true in the \emph{current} time, place, or circumstance at which the truth of the use of the modal operator is being evaluated.
+Classic examples include modal necessity $\square P$ describing that $P$ is \emph{necessarily} true, $\mathsf{G}~P$ meaning $P$ is true \emph{globally} (i.e., forever from this time onwards in temporal logics), or $K_i(P)$ describing that a particular participant $i$ \emph{knows} that $P$ is true.
+The latter is an example of \emph{multimodal} logic, where there is an family of modalities (modal operators) parameterized by some dimension of interest (there, participants).
+A closely related variant of multimodal logic is \emph{dynamic logic}~\cite{pratt1976semantical}, a logic of weakest preconditions~\cite{dijkstra-75} which works with modalities of the form $[p](P)$, which states that \emph{in the current program state}, \emph{if} program $p$ is run then afterwards $P$ will hold (modulo non-termination).
+This same idea is used to encode Hoare triples in the Iris program logic~\cite{krebbers2017essence}, using the same encoding as in Pratt's original presentation, where a Hoare triple $\{P\}C\{Q\}$ is encoded as $P\rightarrow[C](Q)$.
+A hallmark of a unary logical operator $M$ being a modality is if $M$ satisfies a property akin to $(\varphi\rightarrow\psi)\rightarrow M(\varphi)\rightarrow M(\psi)$, which roughly states that modus ponens holds under the modality.\footnote{Afficionados of modal logic will note that this property is not quite Axiom K (which requires the initial implication to also hold under $M$), but follows from K and a necessitation rule $\varphi\rightarrow M(\varphi)$.
+Non-necessitive modalities typically satisfy the weaker property we call out above.} This and related laws allow transporting general reasoning into the specific contexts or hypotheticals modeled by the modalities.
+
+\paragraph{Hybrid Logic} Hybrid logics~\cite{blackburn1995hybrid,goranko1996hierarchies,gargov1993modal,areces2001hybrid,brauner2010hybrid} are a class of modal logics that add two primary new concepts to a logic.
+Nominals $\loc \in \textsf{Loc}$ uniquely identify points in a model (e.g., a particular state), so there is exactly one point in the model where a nominal $\loc$ is true.
+Satisfaction operators are modal operators indexed by nominals, which enable claims about the truth of another proposition at some arbitrary point in the model identified by a nominal.
+Traditionally $@_l\varphi$ asserts that $\varphi$ is true in the state identified by $l$.
+Conceptually, given our context, we might like $l$ to indicate truth in a particular data structure $l$.
+However, most prior work on hybrid logics works in classical logics rather than a substructural setting like separation logic.
+We adapt these ideas to assertions of the form $@[l]\varphi$asserts that $\varphi$ is true in a data structure with \emph{logical} state $\sigma$ \emph{at location} $l$.
+In practice we must design variants of this for each kind of structure we wish to apply this to, e.g. tree nodes, free list, or snapshots.
+%This section provides details of and models for each modal abstraction, focusing on how each data structure can act as a model for a data-structure-specific logic.
+%We continue to motivate the development with aspects of the write amplification process described in the previous section; Section \ref{sec:evaluation} provides further examples using the modal abstractions in proofs.
+\paragraph{Temporal Operators}
+The best known class of modal logics in the systems community is undoubtedly temporal logics (whether \textsc{LTL}~\cite{pnueli1977temporal}, \textsc{CTL}~\cite{emerson1982using}, \textsc{TLA}~\cite{lamport1994temporal}, or others),
+due in part to Lamport's influence~\cite{lamport2002specifying} leading to its not-infrequent use in specifying
+distributed algorithms \cite{ongaro2014search,ailijiang2019wpaxos}.
+\paragraph{Nominalization}
+Some classes of assertions benefit specifically from \emph{naming} the explicit conditions where they are true (as opposed to simply requiring them to be true \emph{somewhere} or \emph{everywhere} as in the most classic modal operators).
+This naming generally resembles
+the \emph{satisfaction} operator of 
+\emph{hybrid logic}~\cite{blackburn1995hybrid,areces2001hybrid}: $@_\iota P$ which evaluates the truth of $P$ at the named (Kripke model) state $\iota$.
+For this reason we refer to the general idea of naming circumstances explicitly as \emph{nominalization},
+even though the examples we discuss are not necessarily actually hybrid logics.
+
+An example utilization of state naming explicitly on the assertion appears in program logics such as Iris~\cite{jung2018iris}, which enables encoding of usage protocols (e.g.
+state transition system) resembling typestate~\cite{strom1986typestate,garcia2014foundations} as specifications.
+Protocol assertions are \emph{annotated with the name of the last (abstract) state at which the protocol is ensured}.
+An example more familiar to the systems community is Halpern et al.'s history of adapting modal logics of
+knowledge to deal with distributed systems~\cite{halpern1990knowledge,halpern1989modelling,halpern2017reasoning}.
+In most of that line of work, $\mathcal{K}_a(P)$ indicates that the node $a$ in the system \emph{knows} or \emph{possesses knowledge of} $P$
+(for example, a Raft node may ``know'' a lower bound on the commit index).
+Alternatively, a modality $@_i(P)$ may represent that $P$ is true of/at the specific node $i$~\cite{gordon2019modal} (e.g., that node $i$ has stored a certain
+piece of data to reliable storage).
+These permit capturing specific concepts relevant to the correctness (and reasoning about correctness) of a certain class of systems,
+involving facts about specific named entities in the system.
+\\
+
+In each of these cases, the fact that these facts are described using modalities with the core modal property $P \Rightarrow Q$, then $M(P) \Rightarrow M(Q)$ means, for example, that if a
+process $p$ knows that the commit index is greater than 5 (e.g., $\mathcal{K}_p(\mathsf{commitIndex} > 5)$)
+no extra work is required to conclude that the node knows it is greater than 3 (i.e., that $\mathcal{K}_p(\mathsf{commitIndex} > 3)$),
+because this follows from standard properties of modal operators as described above.
+In contrast, if verification instead used a custom assertion $\mathsf{minCommitIndex}(5)$ to represent the former knowledge,
+one would need to separately provide custom reasoning to conclude $\mathsf{minCommitIndex}(3)$.
+\paragraph{Shallow Modalities in Iris Separation Logic}
+\label{sec:modal_in_iris}
+Classically modal operators are given direct semantics in terms of an underlying model, which for \textsf{Iris} would be a collection of \textsc{CMRA}s; this is the case for \textsf{Iris}'s one built-in modality for step-indexing~\cite{krebbers2017essence}.
+However in general, many modalities can be given proof-theoretic semantics directly within a logic.
+This is the case for \textsf{Iris}'s weakest precondition modality (which is defined within the \textsf{Iris} logic itself), as well as for a wide range of other modalities in the literature~\cite{restall2002introduction}.
+These modalities are typically multi-modal, with a form of roughly $M~\overline{ctxt}~\phi\triangleq\ldots\rightarrow \phi~\overline{ctxt}$.
+In a substructural setting~\cite{restall1993modalities,krebbers2017essence} the implication $\rightarrow$ is typically replaced with substructural implication $\sepimp$.
+
+This definition also reveals a subtlety about the denotation of assertions $\phi$ --- they are effectively \emph{predicates} of some kind, functions from some possible world into some logical algebra such as a BI-algebra~\cite{ohearn1999bunched}.
+Taking this interpretation literally gives rise to a convenient way to embed new modalities in \textsf{Iris} in a way well-suited to our use cases.
+If $B$ is a type with a BI-algebra structure, then any space of predicate functions (e.g., $\mathsf{val}\rightarrow B$ as a simplification) also carries a BI-algebra structure, lifted from $B$, often called the \emph{pointwise lifting} of the algebra.
+We will exploit this repeatedly, lifting the BI-algebra structure from \textsf{Iris} itself to \textsf{Iris} predicates used as assertions within our own modalities.
+Connecting this to \textsf{Iris}'s typeclasses for interactive proofs~\cite{krebbers-17} then permits working within the embedded logic as in \textsf{Iris}.
+
+In addition, the modal definitions themselves may represent ownership of resources in a substructural setting~\cite{restall1993modalities,dovsen1992modal,kamide2002kripke}.
+This will allow our modalities to simultaneously represent ownership of certain resources used for interpreting the modality itself, as well as whatever resource ownership is implied by the modal argument.
+\paragraph{Modal Abstractions for Recursive Trees Structure with Nominals}
 \label{sec:overviewnode}
@@
 $\varphi_{\textsf{amplified}}$, $\varphi_{\textsf{untouched}}$): a subtree bounded by the cursor node.
-\subsection{Speculation in Consistency}
-%\subsection{Sharing} Sharing is one of the fundamental reasons for kernels to {virtualize} resources,
-%for creating the illusion of exclusive access to virtualized versions of shared resources, including CPU time (via preemption).
-%An implementation of sharing relies on two concepts: \emph{indirection} and \emph{coordination}.
-Indirection is a method that allows the mapping of resources of one type into another.
-By doing so, it allows treating the limited resources (e.g.
-hardware resources such as physical memory) as if there are more than they actually are (e.g.
-via virtual memory addresses and paging).
-Consequently, virtualization for the sake of ensuring availability requires \emph{sharing} of the resources that exist in reality.
-%%\paragraph{Isolation Hand-in-Hand with Sharing}
-%%\Ismail{put a few words on abstractions such as address space, stack region, snapshots ,etc.}
-%%\Colin{Not sure what you want here?}
-
-\todo[inline]{Colin says: the nominals section needs another shift for the new audience:
-You don't need to define modalities here, but do need to give initially the classic notion of nominal
-(name for a Kripke world) before bridging the gap to using the concept of naming
-for something a little different.
-}
-\subsection{Hybrid Logic Dynamic Logic and Nominals in Program Logics}
-\label{sec:modalbackground}
-
-\paragraph{Dynamic Logic's Satisfaction Operator for Program Logics}
-We believe that for the kinds of informal reasoning and the sorts of data structures discussed in the previous section, ideas from modal logic are a promising approach to formalization and verification.
-Broadly speaking modal logics incorporate \emph{modal operators}, which take as arguments a proposition expected to be true in another time~\cite{pnueli1977temporal}, place~\cite{gordon2019modal,goranko1996hierarchies,areces2001hybrid,gargov1993modal,murphy2004symmetric}, circumstance or point-of-view~\cite{hintikka1962knowledge,halpern1985guide}, and result in a proposition true in the \emph{current} time, place, or circumstance at which the truth of the use of the modal operator is being evaluated.
-Classic examples include modal necessity $\square P$ describing that $P$ is \emph{necessarily} true, $\mathsf{G}~P$ meaning $P$ is true \emph{globally} (i.e., forever from this time onwards in temporal logics), or $K_i(P)$ describing that a particular participant $i$ \emph{knows} that $P$ is true.
-The latter is an example of \emph{multimodal} logic, where there is an family of modalities (modal operators) parameterized by some dimension of interest (there, participants).
-A closely related variant of multimodal logic is \emph{dynamic logic}~\cite{pratt1976semantical}, a logic of weakest preconditions~\cite{dijkstra-75} which works with modalities of the form $[p](P)$, which states that \emph{in the current program state}, \emph{if} program $p$ is run then afterwards $P$ will hold (modulo non-termination).
-This same idea is used to encode Hoare triples in the Iris program logic~\cite{krebbers2017essence}, using the same encoding as in Pratt's original presentation, where a Hoare triple $\{P\}C\{Q\}$ is encoded as $P\rightarrow[C](Q)$.
-A hallmark of a unary logical operator $M$ being a modality is if $M$ satisfies a property akin to $(\varphi\rightarrow\psi)\rightarrow M(\varphi)\rightarrow M(\psi)$, which roughly states that modus ponens holds under the modality.\footnote{Afficionados of modal logic will note that this property is not quite Axiom K (which requires the initial implication to also hold under $M$), but follows from K and a necessitation rule $\varphi\rightarrow M(\varphi)$.
-Non-necessitive modalities typically satisfy the weaker property we call out above.} This and related laws allow transporting general reasoning into the specific contexts or hypotheticals modeled by the modalities.
-
-\paragraph{Hybrid Logic} Hybrid logics~\cite{blackburn1995hybrid,goranko1996hierarchies,gargov1993modal,areces2001hybrid,brauner2010hybrid} are a class of modal logics that add two primary new concepts to a logic.
-Nominals $\loc \in \textsf{Loc}$ uniquely identify points in a model (e.g., a particular state), so there is exactly one point in the model where a nominal $\loc$ is true.
-Satisfaction operators are modal operators indexed by nominals, which enable claims about the truth of another proposition at some arbitrary point in the model identified by a nominal.
-Traditionally $@_l\varphi$ asserts that $\varphi$ is true in the state identified by $l$.
-Conceptually, given our context, we might like $l$ to indicate truth in a particular data structure $l$.
-However, most prior work on hybrid logics works in classical logics rather than a substructural setting like separation logic.
-We adapt these ideas to assertions of the form $@[l]\varphi$asserts that $\varphi$ is true in a data structure with \emph{logical} state $\sigma$ \emph{at location} $l$.
-In practice we must design variants of this for each kind of structure we wish to apply this to, e.g. tree nodes, free list, or snapshots.
-%This section provides details of and models for each modal abstraction, focusing on how each data structure can act as a model for a data-structure-specific logic.
-%We continue to motivate the development with aspects of the write amplification process described in the previous section; Section \ref{sec:evaluation} provides further examples using the modal abstractions in proofs.
-\paragraph{Shallow Modalities in Iris Separation Logic}
-\label{sec:modal_in_iris}
-Classically modal operators are given direct semantics in terms of an underlying model, which for \textsf{Iris} would be a collection of \textsc{CMRA}s; this is the case for \textsf{Iris}'s one built-in modality for step-indexing~\cite{krebbers2017essence}.
-However in general, many modalities can be given proof-theoretic semantics directly within a logic.
-This is the case for \textsf{Iris}'s weakest precondition modality (which is defined within the \textsf{Iris} logic itself), as well as for a wide range of other modalities in the literature~\cite{restall2002introduction}.
-These modalities are typically multi-modal, with a form of roughly $M~\overline{ctxt}~\phi\triangleq\ldots\rightarrow \phi~\overline{ctxt}$.
-In a substructural setting~\cite{restall1993modalities,krebbers2017essence} the implication $\rightarrow$ is typically replaced with substructural implication $\sepimp$.
-
-This definition also reveals a subtlety about the denotation of assertions $\phi$ --- they are effectively \emph{predicates} of some kind, functions from some possible world into some logical algebra such as a BI-algebra~\cite{ohearn1999bunched}.
-Taking this interpretation literally gives rise to a convenient way to embed new modalities in \textsf{Iris} in a way well-suited to our use cases.
-If $B$ is a type with a BI-algebra structure, then any space of predicate functions (e.g., $\mathsf{val}\rightarrow B$ as a simplification) also carries a BI-algebra structure, lifted from $B$, often called the \emph{pointwise lifting} of the algebra.
-We will exploit this repeatedly, lifting the BI-algebra structure from \textsf{Iris} itself to \textsf{Iris} predicates used as assertions within our own modalities.
-Connecting this to \textsf{Iris}'s typeclasses for interactive proofs~\cite{krebbers-17} then permits working within the embedded logic as in \textsf{Iris}.
-
-In addition, the modal definitions themselves may represent ownership of resources in a substructural setting~\cite{restall1993modalities,dovsen1992modal,kamide2002kripke}.
-This will allow our modalities to simultaneously represent ownership of certain resources used for interpreting the modality itself, as well as whatever resource ownership is implied by the modal argument.
-\paragraph{Temporal Operators}
-The best known class of modal logics in the systems community is undoubtedly temporal logics (whether \textsc{LTL}~\cite{pnueli1977temporal}, \textsc{CTL}~\cite{emerson1982using}, \textsc{TLA}~\cite{lamport1994temporal}, or others),
-due in part to Lamport's influence~\cite{lamport2002specifying} leading to its not-infrequent use in specifying
-distributed algorithms \cite{ongaro2014search,ailijiang2019wpaxos}.
-\subsubsection{Nominalization}
-Some classes of assertions benefit specifically from \emph{naming} the explicit conditions where they are true (as opposed to simply requiring them to be true \emph{somewhere} or \emph{everywhere} as in the most classic modal operators).
-This naming generally resembles
-the \emph{satisfaction} operator of 
-\emph{hybrid logic}~\cite{blackburn1995hybrid,areces2001hybrid}: $@_\iota P$ which evaluates the truth of $P$ at the named (Kripke model) state $\iota$.
-For this reason we refer to the general idea of naming circumstances explicitly as \emph{nominalization},
-even though the examples we discuss are not necessarily actually hybrid logics.
-
-An example utilization of state naming explicitly on the assertion appears in program logics such as Iris~\cite{jung2018iris}, which enables encoding of usage protocols (e.g.
-state transition system) resembling typestate~\cite{strom1986typestate,garcia2014foundations} as specifications.
-Protocol assertions are \emph{annotated with the name of the last (abstract) state at which the protocol is ensured}.
-An example more familiar to the systems community is Halpern et al.'s history of adapting modal logics of
-knowledge to deal with distributed systems~\cite{halpern1990knowledge,halpern1989modelling,halpern2017reasoning}.
-In most of that line of work, $\mathcal{K}_a(P)$ indicates that the node $a$ in the system \emph{knows} or \emph{possesses knowledge of} $P$
-(for example, a Raft node may ``know'' a lower bound on the commit index).
-Alternatively, a modality $@_i(P)$ may represent that $P$ is true of/at the specific node $i$~\cite{gordon2019modal} (e.g., that node $i$ has stored a certain
-piece of data to reliable storage).
-These permit capturing specific concepts relevant to the correctness (and reasoning about correctness) of a certain class of systems,
-involving facts about specific named entities in the system.
-\\
-
-In each of these cases, the fact that these facts are described using modalities with the core modal property $P \Rightarrow Q$, then $M(P) \Rightarrow M(Q)$ means, for example, that if a
-process $p$ knows that the commit index is greater than 5 (e.g., $\mathcal{K}_p(\mathsf{commitIndex} > 5)$)
-no extra work is required to conclude that the node knows it is greater than 3 (i.e., that $\mathcal{K}_p(\mathsf{commitIndex} > 3)$),
-because this follows from standard properties of modal operators as described above.
-In contrast, if verification instead used a custom assertion $\mathsf{minCommitIndex}(5)$ to represent the former knowledge,
-one would need to separately provide custom reasoning to conclude $\mathsf{minCommitIndex}(3)$.
+
 \section{Contingency Decomposition of a System}

### 2025-06-16 paper.tex
+

@@
 %\maketitle
-\section{Introduction}\label{sec:intro}
-\todo[inline]{Colin says: These comments aren't the only places things need to change, but they're
-specific places I have suggestions.
-One thing to keep in mind is that the CfP lists two kinds of papers:
-short papers are for "original research or case studies, maybe in the incubation phase,"
-while long papers are for "original research, case studies, or surveys."
-This is clearly in the survey category, so we (a) should take advantage of the extra space to
-explain things clearly (15 pages, plus 2 pages for bibliography), and (b) lean into the survey nature explicitly.
-}
-Low-level systems exhibit certain patterns in their designs, especially when interacting with computing resources.
-Exploiting certain patterns while designing software has been an important field of study.
-In this regard, we think that certain properties of modalities enable us to understand and do the verification challenges that show certain patterns.
-\todo[inline]{I think one of the biggest things we can do to reframe this is to explain the broader vision
-here before jumping into contributions.
-This is an audience that knows dynamic logic and modal logic more broadly, but perhaps not everyone is familiar
-with hybrid logic, and many will not be familiar with systems concepts.
-This is the place to provide high-level intuition about the domain we're surveying, tailored for this audience.
-I'd talk about dynamic logic in general, but then highlight the need for reasoning specific to certain
-compute resources to produce manageable specifications and proofs.
-Think big picture: why is this space interesting to logicians?
-}
-
-\subsection{Contributions to Identifying the Verification Patterns via Modal Abstractions}
-We argue how modal abstractions can be used to identify and abstract system verification challenges.
-We justify our perspective by discussing prior systems that have successfully used modalities for system verification, arguing that they fit into the verification design pattern we articulate, and explaining how this approach might apply to other systems' verification challenges.
-
-\paragraph{Identifying System Verification Challenges} We start with identifying common patterns in system verification: \emph{virtualization, sharing, and translation}.
-\paragraph{Introducing the Concept of \emph{Resource}} Then we discuss the concept of \emph{resource} which has already been an essential concept in the design of systems.
-Inspired by the concept of resource in the systems, we define what a resource and its context are in the modal abstractions.
-\paragraph{Introducing the Concept of \emph{Nominals}}
-\todo[inline]{Colin says: I'd reframe the way nominals are introduced. This discussion is good,
-but is a "systems-first" explanation. At least some of this audience will already know
-what classic nominals are, so need to be shown naming specific resources as an application
-of what they already know. (And some readers will need a very short intro/refresher.)
-}
-Nominalization enables identifying a resource in a context.
-For example, a transaction is a context of resources of in-memory updated disk blocks.
-The transaction identifier is used to associate a transaction with a disk-block to be persisted so that, in case of a crash while persisting updated disk-blocks, the filesystem can rollback the \emph{already persisted} disk-blocks of the transaction, and reach to the previous consistent disk state.
-To be able to do so, both the updated in-memory disk blocks and the transaction must refer to the transaction identifier -- \emph{strong nominalization}.
-In another example, virtual memory references (resources) in an address space (resource context), which can be uniquely identified with a root address (the nominal) of its page-table tree, are \emph{agnostic} to the address space that they are in.
-However, they can only be accessed (be valid) in the address space to which they are agnostic.
-However, then an address space switch happens, the virtual memory references of the previous address space must be made inaccessible.
-To be able to do so, although virtual memory references do not hold any piece of information related to their address space, we still have to associate them.
-We call this kind of unilateral nominalization -- \emph{weak nominalization}.
-\paragraph{Taxonomy of the Current Modal Approaches in System Verification} Based on these concepts defined, we summarize contemporary verification efforts using modal abstractions.
-We choose them from different domains, for example, weak memory verification, storage persistence ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2024modalabstractionsvirtualizingmemory,amalreal2024} because we would like to justify that our \emph{definitions} are not domain dependent.
-
-\section{Background on Low-Level Systems}
+
+\section{Introduction}
 Although they differ in the functionality they offer, low-level systems exhibit certain patterns of design and utilization of computing resources.
@@
 the construction of higher layers of the stack.
-\subsection{Virtualization in Systems Software}
-\label{sec:background}
-\paragraph{Virtualization of Memory Locations}
-\label{sec:backgroundonmachinemodel}
-\todo{change the text}
-Memory addresses seen by programs running on modern computers are \emph{virtualized}: the address observed by a running program generally will not correspond directly to the physical location in memory, and may not even correspond to a physical location that \emph{exists} in the machine.
-Instead, these \emph{virtual} addresses are translated to \emph{physical} addresses that correspond directly to locations in RAM.
-On most modern architectures, this translation is performed through cooperation of the hardware and OS kernel: while executing an instruction that dereferences a (virtual) address, the CPU's \emph{memory management unit} (\textsc{MMU}) implements an architecture-specific process of \emph{address translation}, resulting in a physical address used to access the cache\footnote{Technically, for performance reasons most caches are indexed with parts of the virtual address, but tagged with the physical data addresses, so cache lookups and address translations can proceed in parallel.} and/or memory-bus.
-
-\begin{figure}
-    \includegraphics[width=\columnwidth]{pagetables2.png}
-    \caption{x86-64 page table lookups.}
-    \label{fig:pagetables}
-\end{figure}
-
-On the x86-64 architecture, the input to the \textsc{MMU}'s address translation is a sparse hierarchical set of tables calle d \emph{page tables} (referring to pages of memory).
-As Figure \ref{fig:pagetables} shows, address translation proceeds by repeatedly taking designated slices of the virtual address and indexing into a table.
-On x86-64, standard configurations use 4 levels of page tables, labelled levels 4 through 1, with lookups in the level 1 page table resulting in the actual page of physical memory holding the requested data, and the low-order 12 bits being used to index into this page.\footnote{Technically levels 1--3 have designated names originating when there were fewer levels.
-We omit those names to avoid confusion, and for greater consistency with the recent addition of a 5th level on the most recent processors, where levels 4 and 5 are simply known by their numbers.}  Because address translation traverses a series of tables, the translation process or algorithm is sometimes referred to as a \emph{page-table walk}.
-While Figure \ref{fig:pagetables} and most of our details are specific to the x86-64 architecture, ARMv8 (a.k.a.\ \texttt{aarch64}), RISC-V, and PowerPC use similar hierarchical page tables for address translation.
-
-The entries of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set), and a range of additional bits which can be used to control caching, write-through, and more.
-This paper will only consider the present bit (0), the read-write bit (1), and the accessed bit (5), which is set to 1 by the hardware every time an entry is used, allowing the OS to determine which pieces of memory are used or unused.
-
-The page tables mapping virtual addresses to physical addresses are managed by the OS.
-Typically each process has its on virtul-to-physical mappings represented by its own page table, which the OS registers with the CPU by storing it in a specific register (\texttt{cr3}) as part of switching to a new process.
-Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section) is how the OS ensures memory isolation between processes.
-
-If an instruction is executed that accesses a virtual address that either has no mapping, or does not have a mapping permitting the kind of access that was performed (e.g., the instruction was a memory write, but the relevant address range was marked read-only in the relevant page table entry), the hardware triggers a \emph{page fault}, transferring control to a \emph{page fault handler} registered with the hardware by the OS, allowing it to take corrective action (described in the next section).
-
-\paragraph{Virtual Memory Managers with Multiple Address-Spaces}
-\label{sec:backgroundonvmm}
-\todo{change the text}
-Most OS kernels have a component called a \emph{Virtual Memory Manager} (VMM)\footnote{Not to be confused with Virtual Machine Monitor.
-In this paper we focus on non-virtualized scenarios, but hardware virtualization extensions for both x86-64 and ARM make use of an additional set of page tables translating what a \emph{guest} considers to be its (virutalized) physical memory to actual physical memory.
-Our contributions should offer value in this scenario as well.}
-which is responsible for setting up page table mappings and for taking action when a page fault occurs.
-Often, a page fault indicates a bug in a program --- for example, a null pointer dereference crashes a program not because the hardware designates it an error, but because most OSes refuse to map the first 4K worth of virtual addresses, so dereferencing \texttt{NULL} results in a page fault.
-In such cases, the OS will terminate the program.
-
-In other cases the OS uses page faults to implement specialized functionality and optimizations.
-A key example of this is (confusingly) called \emph{paging}: saving room in physical RAM by waiting until memory addresses for code are accessed before copying them to RAM (saving process startup time); or copying memory that has not recently been accessed to disk (i.e., in a \emph{swap} file or partition) and marking those page table entries invalid so a future page fault will give the OS the chance to copy the relevant data back into memory when the program tries to access it again.
-
-The key pieces of functionality a VMM must implement itself, or have available in another OS component (depending on design), are
-adding a new page mapping (whether the mapped page contains zeros, file data, or swap data), and removing an existing page mapping.
-While this initially sounds like relatively modest functionality whose implementation may be complicated by hardware subtleties, a number of more serious complexities must be handled.
-%
-%For one, virtual addresses may \emph{alias} when more than one is mapped to the same physical memory.
-This can occur when a process requests it (e.g., via \texttt{mmap} or \texttt{shm\_open} on UNIX-style kernels).
-It also occurs automatically in most general-purpose kernels: while for normal programs the OS uses the MMU to isolate and abstract physical memory, the kernel's own functionality is often easier to implement if the kernel can directly access memory given its physical address, for example when interacting with hardware devices.
-For this reason, most kernels contain an \emph{identity map} of all physical memory in a machine.
-In some cases this is a literal identity map, where for every virtual address $p$, interpreting $p$ as a virtual address instead will map back to $p$ as a physical address.
-In other cases there is an offset involved, where to access physical address $p$ the kernel can access virtual address $p+\mathit{offset}$.
-This identity map saves the kernel the performance cost of having to continually add and remove transient mappings simply to briefly access a specific physical memory region.
-But this coexists with additional virtual addresses (those used by most kernel code and data structures), in a form of virtual address aliasing.
-Note that virtual address aliasing is both prevalent in general-purpose kernels, and violates the core assumption of most memory models assumed by verified compilers like CompCert, which assume no virtual address aliasing.
-%
-%Another complicating factor is that portions of page tables can be shared, i.e., trees of page tables can overlap.
-One common way this used to happen was for one subtree of page tables to map the kernel if linked in appropriately to a higher-level page table, and then to reuse that kernel mapping portion in every process's page tables.
-This was mostly phased out in the wake of Spectre
-%\mytodo{Colin finish}
-
-The OS scheduler, since it is responsible for dealing with multiple address spaces, must keep track of which virtual addresses are valid (and in what way) in which address spaces, where some virtual addresses are valid in only a single address space (e.g., a code address for a particular usermode process), while others are valid in all address spaces (e.g., kernel data structure pointers).
-The VMM must maintain some of these assumptions on behalf of the rest of the kernel, for example by guaranteeing that a certain range of virtual addresses (corresponding to the kernel's code and data) are valid in every address space.
-
-\paragraph{Virtualization of Disk-Blocks in Filesystems}
-
-\paragraph{Consistency of Disk Layout} \todo{}
-\paragraph{Memory Reclamation in Client Code}\todo{iso:speak about Amal's stuff}
-\paragraph{Consistency of Reclaimed Memory}\todo{iso: speak about the operator for which the hybrid logic stuff is usefull }
-
-\subsection{Speculative Observation in Systems Software}
-\label{sec:speculativeobservation}
-
-\section{Resources in Systems Software}
-\label{sec:systemsoft}
-In this section, we argue the position that \emph{modalities}  
-should be a go-to approach when specifying and verifying low-level systems code.
-We explain how the concept of a \emph{resource context} helps guide the design of new modalities
-for verification of systems code, and
-we justify our perspective by discussing prior systems that have used modalities for systems verification
-successfully, arguing that they fit into the verification design pattern we articulate,
-and explaining how this approach might apply to other systems verification challenges.
-
-To explain our ideas in the general systems understanding, we briefly recap some of the background and themes our ideas build on again,casting them in a certain way to bring out the relevance of our philosophy.
-
-\subsection{Virtualization}
-One of the most common forms of abstraction provided by systems software is \emph{virtualization},
-which abstracts the relationship between conceptual and physical computing resources. 
-Operating System (OS) kernels virtualize memory locations and quantity (via virtual memory and paging~\cite{Denning1970VM}).
-Distributed language runtimes may virtualize addresses, even when processes may migrate across machines~\cite{jul1988fine}.
-Filesystems virtualize locations on disk~\cite{rodeh2013btrfs,hitz1994file,Rosenblum1992LFS,bonwick2003zettabyte}.
-Programs built on top of the corresponding systems software layer work logically at the level of these virtualized resources,
-and it makes sense to specify the systems software directly in terms of those abstractions.
-\paragraph{Access by Translation}
-Accessing virtualized resources via translation is a common way to virtualize notions of location (e.g.,
-virtual memory addresses, inodes or object IDs instead of disk addresses~\cite{bonwick2003zettabyte,hitz1994file,Rosenblum1992LFS,rodeh2013btrfs}).
-%Accessing the hardware resources in the presence of sharing means having more resource references (references to the virtualized resources) to access the \emph{mapped} resources that are physical and potentially fewer.
-B-trees, page tables, and related structures both behave like maps, when the corresponding physical resources exist as such
-just in a different location.
-Control over the lookup process (e.g, handling the case of a missing translation entry) allows for additional flexibility,
-such as filling holes in sparse files, or demand paging (both from disk or lazily populating anonymous initially-zeroed mappings).
-Although the realization of these maps may differ from a system to system based on the context (and sometimes hardware details),
-they are semantically --- logically --- partial maps, worth treating as such in verification.
-
-\subsection{Speculation in Consistency}
-%\subsection{Sharing} Sharing is one of the fundamental reasons for kernels to {virtualize} resources,
-%for creating the illusion of exclusive access to virtualized versions of shared resources, including CPU time (via preemption).
-%An implementation of sharing relies on two concepts: \emph{indirection} and \emph{coordination}.
-Indirection is a method that allows the mapping of resources of one type into another.
-By doing so, it allows treating the limited resources (e.g.
-hardware resources such as physical memory) as if there are more than they actually are (e.g.
-via virtual memory addresses and paging).
-Consequently, virtualization for the sake of ensuring availability requires \emph{sharing} of the resources that exist in reality.
-%%\paragraph{Isolation Hand-in-Hand with Sharing}
-%%\Ismail{put a few words on abstractions such as address space, stack region, snapshots ,etc.}
-%%\Colin{Not sure what you want here?}
-
-\todo[inline]{Colin says: the nominals section needs another shift for the new audience:
-You don't need to define modalities here, but do need to give initially the classic notion of nominal
-(name for a Kripke world) before bridging the gap to using the concept of naming
-for something a little different.
-}
 \subsection{Hybrid Logic Dynamic Logic and Nominals in Program Logics}
@@
 one would need to separately provide custom reasoning to conclude $\mathsf{minCommitIndex}(3)$.
+
+\subsection{Virtualization in Systems Software}
+\label{sec:background}
+\paragraph{Virtualization of Memory Locations}
+\label{sec:backgroundonmachinemodel}
+\todo{change the text}
+Memory addresses seen by programs running on modern computers are \emph{virtualized}: the address observed by a running program generally will not correspond directly to the physical location in memory, and may not even correspond to a physical location that \emph{exists} in the machine.
+Instead, these \emph{virtual} addresses are translated to \emph{physical} addresses that correspond directly to locations in RAM.
+On most modern architectures, this translation is performed through cooperation of the hardware and OS kernel: while executing an instruction that dereferences a (virtual) address, the CPU's \emph{memory management unit} (\textsc{MMU}) implements an architecture-specific process of \emph{address translation}, resulting in a physical address used to access the cache\footnote{Technically, for performance reasons most caches are indexed with parts of the virtual address, but tagged with the physical data addresses, so cache lookups and address translations can proceed in parallel.} and/or memory-bus.
+
+\begin{figure}
+    \includegraphics[width=\columnwidth]{pagetables2.png}
+    \caption{x86-64 page table lookups.}
+    \label{fig:pagetables}
+\end{figure}
+
+On the x86-64 architecture, the input to the \textsc{MMU}'s address translation is a sparse hierarchical set of tables calle d \emph{page tables} (referring to pages of memory).
+As Figure \ref{fig:pagetables} shows, address translation proceeds by repeatedly taking designated slices of the virtual address and indexing into a table.
+On x86-64, standard configurations use 4 levels of page tables, labelled levels 4 through 1, with lookups in the level 1 page table resulting in the actual page of physical memory holding the requested data, and the low-order 12 bits being used to index into this page.\footnote{Technically levels 1--3 have designated names originating when there were fewer levels.
+We omit those names to avoid confusion, and for greater consistency with the recent addition of a 5th level on the most recent processors, where levels 4 and 5 are simply known by their numbers.}  Because address translation traverses a series of tables, the translation process or algorithm is sometimes referred to as a \emph{page-table walk}.
+While Figure \ref{fig:pagetables} and most of our details are specific to the x86-64 architecture, ARMv8 (a.k.a.\ \texttt{aarch64}), RISC-V, and PowerPC use similar hierarchical page tables for address translation.
+
+The entries of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set), and a range of additional bits which can be used to control caching, write-through, and more.
+This paper will only consider the present bit (0), the read-write bit (1), and the accessed bit (5), which is set to 1 by the hardware every time an entry is used, allowing the OS to determine which pieces of memory are used or unused.
+
+The page tables mapping virtual addresses to physical addresses are managed by the OS.
+Typically each process has its on virtul-to-physical mappings represented by its own page table, which the OS registers with the CPU by storing it in a specific register (\texttt{cr3}) as part of switching to a new process.
+Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section) is how the OS ensures memory isolation between processes.
+
+If an instruction is executed that accesses a virtual address that either has no mapping, or does not have a mapping permitting the kind of access that was performed (e.g., the instruction was a memory write, but the relevant address range was marked read-only in the relevant page table entry), the hardware triggers a \emph{page fault}, transferring control to a \emph{page fault handler} registered with the hardware by the OS, allowing it to take corrective action (described in the next section).
+
+\paragraph{Virtual Memory Managers with Multiple Address-Spaces}
+\label{sec:backgroundonvmm}
+\todo{change the text}
+Most OS kernels have a component called a \emph{Virtual Memory Manager} (VMM)\footnote{Not to be confused with Virtual Machine Monitor.
+In this paper we focus on non-virtualized scenarios, but hardware virtualization extensions for both x86-64 and ARM make use of an additional set of page tables translating what a \emph{guest} considers to be its (virutalized) physical memory to actual physical memory.
+Our contributions should offer value in this scenario as well.}
+which is responsible for setting up page table mappings and for taking action when a page fault occurs.
+Often, a page fault indicates a bug in a program --- for example, a null pointer dereference crashes a program not because the hardware designates it an error, but because most OSes refuse to map the first 4K worth of virtual addresses, so dereferencing \texttt{NULL} results in a page fault.
+In such cases, the OS will terminate the program.
+
+In other cases the OS uses page faults to implement specialized functionality and optimizations.
+A key example of this is (confusingly) called \emph{paging}: saving room in physical RAM by waiting until memory addresses for code are accessed before copying them to RAM (saving process startup time); or copying memory that has not recently been accessed to disk (i.e., in a \emph{swap} file or partition) and marking those page table entries invalid so a future page fault will give the OS the chance to copy the relevant data back into memory when the program tries to access it again.
+
+The key pieces of functionality a VMM must implement itself, or have available in another OS component (depending on design), are
+adding a new page mapping (whether the mapped page contains zeros, file data, or swap data), and removing an existing page mapping.
+While this initially sounds like relatively modest functionality whose implementation may be complicated by hardware subtleties, a number of more serious complexities must be handled.
+%
+%For one, virtual addresses may \emph{alias} when more than one is mapped to the same physical memory.
+This can occur when a process requests it (e.g., via \texttt{mmap} or \texttt{shm\_open} on UNIX-style kernels).
+It also occurs automatically in most general-purpose kernels: while for normal programs the OS uses the MMU to isolate and abstract physical memory, the kernel's own functionality is often easier to implement if the kernel can directly access memory given its physical address, for example when interacting with hardware devices.
+For this reason, most kernels contain an \emph{identity map} of all physical memory in a machine.
+In some cases this is a literal identity map, where for every virtual address $p$, interpreting $p$ as a virtual address instead will map back to $p$ as a physical address.
+In other cases there is an offset involved, where to access physical address $p$ the kernel can access virtual address $p+\mathit{offset}$.
+This identity map saves the kernel the performance cost of having to continually add and remove transient mappings simply to briefly access a specific physical memory region.
+But this coexists with additional virtual addresses (those used by most kernel code and data structures), in a form of virtual address aliasing.
+Note that virtual address aliasing is both prevalent in general-purpose kernels, and violates the core assumption of most memory models assumed by verified compilers like CompCert, which assume no virtual address aliasing.
+%
+%Another complicating factor is that portions of page tables can be shared, i.e., trees of page tables can overlap.
+One common way this used to happen was for one subtree of page tables to map the kernel if linked in appropriately to a higher-level page table, and then to reuse that kernel mapping portion in every process's page tables.
+This was mostly phased out in the wake of Spectre
+%\mytodo{Colin finish}
+
+The OS scheduler, since it is responsible for dealing with multiple address spaces, must keep track of which virtual addresses are valid (and in what way) in which address spaces, where some virtual addresses are valid in only a single address space (e.g., a code address for a particular usermode process), while others are valid in all address spaces (e.g., kernel data structure pointers).
+The VMM must maintain some of these assumptions on behalf of the rest of the kernel, for example by guaranteeing that a certain range of virtual addresses (corresponding to the kernel's code and data) are valid in every address space.
+
+\paragraph{Virtualization of Disk-Blocks in Filesystems}
+
+\paragraph{Consistency of Disk Layout} \todo{}
+\paragraph{Memory Reclamation in Client Code}\todo{iso:speak about Amal's stuff}
+\paragraph{Consistency of Reclaimed Memory}\todo{iso: speak about the operator for which the hybrid logic stuff is usefull }
+
+\subsection{Speculative Observation in Systems Software}
+\label{sec:speculativeobservation}
+\section{Introduction}\label{sec:intro}
+\todo[inline]{Colin says: These comments aren't the only places things need to change, but they're
+specific places I have suggestions.
+One thing to keep in mind is that the CfP lists two kinds of papers:
+short papers are for "original research or case studies, maybe in the incubation phase,"
+while long papers are for "original research, case studies, or surveys."
+This is clearly in the survey category, so we (a) should take advantage of the extra space to
+explain things clearly (15 pages, plus 2 pages for bibliography), and (b) lean into the survey nature explicitly.
+}
+
+\todo[inline]{I think one of the biggest things we can do to reframe this is to explain the broader vision
+here before jumping into contributions.
+This is an audience that knows dynamic logic and modal logic more broadly, but perhaps not everyone is familiar
+with hybrid logic, and many will not be familiar with systems concepts.
+This is the place to provide high-level intuition about the domain we're surveying, tailored for this audience.
+I'd talk about dynamic logic in general, but then highlight the need for reasoning specific to certain
+compute resources to produce manageable specifications and proofs.
+Think big picture: why is this space interesting to logicians?
+}
+
+\subsection{Conclusion \& Contributions}
+
+\paragraph{Conclusion} Low-level systems exhibit certain patterns in their designs, especially when interacting with computing resources.
+Exploiting certain patterns while designing software has been an important field of study.
+In this regard, we think that certain properties of modalities enable us to understand and do the verification challenges that show certain patterns.
+
+\paragraph{Contributions} We argue how modal abstractions can be used to identify and abstract system verification challenges.
+We justify our perspective by discussing prior systems that have successfully used modalities for system verification, arguing that they fit into the verification design pattern we articulate, and explaining how this approach might apply to other systems' verification challenges.
+\begin{itemize}
+\item \textbf{Identifying System Verification Challenges}: We start with identifying common patterns in system verification: \emph{virtualization, sharing, and translation}.
+\item \textbf{Introducing the Concept of \emph{Resource}}: Then we discuss the concept of \emph{resource} which has already been an essential concept in the design of systems.
+Inspired by the concept of resource in the systems, we define what a resource and its context are in the modal abstractions.
+\item \textbf{Introducing the Concept of \emph{Nominals}}:
+\todo[inline]{Colin says: I'd reframe the way nominals are introduced. This discussion is good,
+but is a "systems-first" explanation. At least some of this audience will already know
+what classic nominals are, so need to be shown naming specific resources as an application
+of what they already know. (And some readers will need a very short intro/refresher.)
+}
+Nominalization enables identifying a resource in a context.
+For example, a transaction is a context of resources of in-memory updated disk blocks.
+The transaction identifier is used to associate a transaction with a disk-block to be persisted so that, in case of a crash while persisting updated disk-blocks, the filesystem can rollback the \emph{already persisted} disk-blocks of the transaction, and reach to the previous consistent disk state.
+To be able to do so, both the updated in-memory disk blocks and the transaction must refer to the transaction identifier -- \emph{strong nominalization}.
+In another example, virtual memory references (resources) in an address space (resource context), which can be uniquely identified with a root address (the nominal) of its page-table tree, are \emph{agnostic} to the address space that they are in.
+However, they can only be accessed (be valid) in the address space to which they are agnostic.
+However, then an address space switch happens, the virtual memory references of the previous address space must be made inaccessible.
+To be able to do so, although virtual memory references do not hold any piece of information related to their address space, we still have to associate them.
+We call this kind of unilateral nominalization -- \emph{weak nominalization}.
+\item \textbf{Taxonomy of the Current Modal Approaches in System Verification}: Based on these concepts defined, we summarize contemporary verification efforts using modal abstractions.
+We choose them from different domains, for example, weak memory verification, storage persistence ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2024modalabstractionsvirtualizingmemory,amalreal2024} because we would like to justify that our \emph{definitions} are not domain dependent.
+\end{itemize}
+\section{Resources in Systems Software}
+\label{sec:systemsoft}
+In this section, we argue the position that \emph{modalities}  
+should be a go-to approach when specifying and verifying low-level systems code.
+We explain how the concept of a \emph{resource context} helps guide the design of new modalities
+for verification of systems code, and
+we justify our perspective by discussing prior systems that have used modalities for systems verification
+successfully, arguing that they fit into the verification design pattern we articulate,
+and explaining how this approach might apply to other systems verification challenges.
+
+To explain our ideas in the general systems understanding, we briefly recap some of the background and themes our ideas build on again,casting them in a certain way to bring out the relevance of our philosophy.
+
+\subsection{Virtualization}
+One of the most common forms of abstraction provided by systems software is \emph{virtualization},
+which abstracts the relationship between conceptual and physical computing resources. 
+Operating System (OS) kernels virtualize memory locations and quantity (via virtual memory and paging~\cite{Denning1970VM}).
+Distributed language runtimes may virtualize addresses, even when processes may migrate across machines~\cite{jul1988fine}.
+Filesystems virtualize locations on disk~\cite{rodeh2013btrfs,hitz1994file,Rosenblum1992LFS,bonwick2003zettabyte}.
+Programs built on top of the corresponding systems software layer work logically at the level of these virtualized resources,
+and it makes sense to specify the systems software directly in terms of those abstractions.
+\paragraph{Access by Translation}
+Accessing virtualized resources via translation is a common way to virtualize notions of location (e.g.,
+virtual memory addresses, inodes or object IDs instead of disk addresses~\cite{bonwick2003zettabyte,hitz1994file,Rosenblum1992LFS,rodeh2013btrfs}).
+%Accessing the hardware resources in the presence of sharing means having more resource references (references to the virtualized resources) to access the \emph{mapped} resources that are physical and potentially fewer.
+B-trees, page tables, and related structures both behave like maps, when the corresponding physical resources exist as such
+just in a different location.
+Control over the lookup process (e.g, handling the case of a missing translation entry) allows for additional flexibility,
+such as filling holes in sparse files, or demand paging (both from disk or lazily populating anonymous initially-zeroed mappings).
+Although the realization of these maps may differ from a system to system based on the context (and sometimes hardware details),
+they are semantically --- logically --- partial maps, worth treating as such in verification.
+
+\subsection{Speculation in Consistency}
+%\subsection{Sharing} Sharing is one of the fundamental reasons for kernels to {virtualize} resources,
+%for creating the illusion of exclusive access to virtualized versions of shared resources, including CPU time (via preemption).
+%An implementation of sharing relies on two concepts: \emph{indirection} and \emph{coordination}.
+Indirection is a method that allows the mapping of resources of one type into another.
+By doing so, it allows treating the limited resources (e.g.
+hardware resources such as physical memory) as if there are more than they actually are (e.g.
+via virtual memory addresses and paging).
+Consequently, virtualization for the sake of ensuring availability requires \emph{sharing} of the resources that exist in reality.
+%%\paragraph{Isolation Hand-in-Hand with Sharing}
+%%\Ismail{put a few words on abstractions such as address space, stack region, snapshots ,etc.}
+%%\Colin{Not sure what you want here?}
+
+\todo[inline]{Colin says: the nominals section needs another shift for the new audience:
+You don't need to define modalities here, but do need to give initially the classic notion of nominal
+(name for a Kripke world) before bridging the gap to using the concept of naming
+for something a little different.
+}
+
+\section{Contingency Decomposition of a System}
+\todo[inline]{Colin says: I don't think this section needs to change much,
+but it should be read carefully with the new audience in mind,
+possibly giving more details on what systems concepts each system is dealing with, for a non-systems audience.}
 \paragraph{Shallow Modalities in Iris Separation Logic}
@@
 This will allow our modalities to simultaneously represent ownership of certain resources used for interpreting the modality itself, as well as whatever resource ownership is implied by the modal argument.
-\paragraph{Modal Abstractions for Recursive Trees Structure with Nominals}
-\label{sec:overviewnode}
-
-To reason about states of trees and subtrees in particular, we work with a modality for stating an assertion that is true \emph{of all nodes within a tree}, identifying the tree by a root note.
-We refer to the resources that provide the context of these modal assertions as the \emph{modal context}; for a recursively defined tree like those in \textsf{Modfs}, these contexts are recursively defined in terms of the tree structure.
-
-%We define the modal context as the resources (e.g.
-a btree node tree, free page id list) which are accessed under some relation defined within the modal definition (e.g.
-being children of a cursor node).
-The choice of modal context can be structurally dependent on the semantics, in our case \textsc{CoW} semantics.
-
-Write amplification traverses the overall tree depth-first, transforming the tree from the bottom up, one full subtree at a time.
-Each time a node is split, it results in disconnecting some nodes from the next snapshot, sharing other nodes, and linking some newly-allocated nodes into the new snapshot, as in Figure \ref{fig:cow_index_tree_update}.
-Nodes are split (or deemed not to require splitting) one at a time, and considered only after all child nodes have been amplified.
-Concepually this means the amplification code's proof must track the fact that at any point in time, there are regions of the tree where each node has already been amplified (e.g., $\phi_{\textsf{amplified}}$ definitely holds for that node), and regions where nodes may or may not require amplification, and this will not be known until those nodes are visited (i.e., some basic validity $\phi_{\textsf{valid}}$ is known to hold, but not the stronger $\phi_{\textsf{amplified}}$).
-
-%\setlength\intextsep{0pt}
-\begin{wrapfigure}{r}{0.25\textwidth}
-  \centering  \includegraphics[width=0.25\textwidth]{entail.png}
-  \caption{Understanding the level of $\varphi$ satisfaction in the \textsc{CoW} tree of modal nodes}
-   \label{fig:nodeentail} 
-\end{wrapfigure}
-We use a modality similar to $\text{\faSitemap}@[\ell]\phi$ to indicate that $\phi$ is true of the tree node rooted at $\ell$, as well as at \emph{all reachable children} of that node.
-This means that establishing a particular $\phi$ throughout a tree must be done in a bottom-up manner (see Figure \ref{fig:nodeentail}).
-Capturing this modally, rather than with more typical ad-hoc hand-crafted invariants, offers advantages in a number of proof steps, in addition to concisely capturing intuitive reasoning about subtrees with all nodes satisfying a particular property.
-One basic example is that while the actual disk sync code requires the full filesystem tree to be in an amplified state (hence write amplification), most of the filesystem works with more general trees.
-Since $\phi_{\textsf{amplified}}\vdash\phi_{\textsf{valid}}$, using our \faSitemap modality we can conclude $\text{\faSitemap}@[\ell]\phi_{\textsf{amplified}}\vdash\text{\faSitemap}@[\ell]\phi_{\textsf{valid}}$ by an analogue of Axiom K.
-There are also other places in the code where a subtree satisfying a stronger assertion at all nodes must be linked into a larger tree satisfying only a weaker assertion overall: predicate representing the valid tree as the tree of nodes which are backed by resources allocated pages, i.e.
-$\phi_{\textsf{valid}}$, is weaker than the predicate representing the tree of nodes being currently amplified in the current snapshot.
-Again, the same reasoning principle makes this straightforward.
-As alluded to earlier, our $\phi_{\textsf{amplified}}$ and company are in fact \emph{predicates} parameterized by a choice of node, though by lifing \textsf{Iris}'s BI structure they can use all of the standard separation logic connectives.
-The rest of this subsection develops this modality in full detail.
-%\colinsaysinl{I cut some text here about the three-way split (commented out) that I wasn't sure how to use, but maybe I deleted something important?}
-%To put it concretely, as shown in Figure \ref{fig:cow_index_tree.JPG}, \textsc{CoW} semantics determine what we can claim on the structure of the filesystem indexing tree:
-%\begin{itemize}
-%\item The bottom-up direction splits the tree into the two half at the cursor node: 1.
-the bottom part accessible from the cursor node and per-node-amplification is reflected in case needed 2.
-the upper part on which amplifcation has not reflected yet.
-%\item Horizontally, amplifcation path splits the bottom part of the tree into three different contexts: untouched tree nodes, nodes on the amplification path, and the ones generated as part of node splitting.
-%\end{itemize}
-%Based on these two observations, in fact, bottom-up amplification determines the choice of modal context by giving the boundries of \textit{any claimable knowledge} (e.g.
-$\varphi_{\textsf{amplified}}$, $\varphi_{\textsf{untouched}}$): a subtree bounded by the cursor node.
-
-\section{Contingency Decomposition of a System}
-\todo[inline]{Colin says: I don't think this section needs to change much,
-but it should be read carefully with the new audience in mind,
-possibly giving more details on what systems concepts each system is dealing with, for a non-systems audience.}
 
@@
 %
+\paragraph{Modal Abstractions for Recursive Trees Structure with Nominals}
+\label{sec:overviewnode}
+
+To reason about states of trees and subtrees in particular, we work with a modality for stating an assertion that is true \emph{of all nodes within a tree}, identifying the tree by a root note.
+We refer to the resources that provide the context of these modal assertions as the \emph{modal context}; for a recursively defined tree like those in \textsf{Modfs}, these contexts are recursively defined in terms of the tree structure.
+
+%We define the modal context as the resources (e.g.
+a btree node tree, free page id list) which are accessed under some relation defined within the modal definition (e.g.
+being children of a cursor node).
+The choice of modal context can be structurally dependent on the semantics, in our case \textsc{CoW} semantics.
+
+Write amplification traverses the overall tree depth-first, transforming the tree from the bottom up, one full subtree at a time.
+Each time a node is split, it results in disconnecting some nodes from the next snapshot, sharing other nodes, and linking some newly-allocated nodes into the new snapshot, as in Figure \ref{fig:cow_index_tree_update}.
+Nodes are split (or deemed not to require splitting) one at a time, and considered only after all child nodes have been amplified.
+Concepually this means the amplification code's proof must track the fact that at any point in time, there are regions of the tree where each node has already been amplified (e.g., $\phi_{\textsf{amplified}}$ definitely holds for that node), and regions where nodes may or may not require amplification, and this will not be known until those nodes are visited (i.e., some basic validity $\phi_{\textsf{valid}}$ is known to hold, but not the stronger $\phi_{\textsf{amplified}}$).
+
+%\setlength\intextsep{0pt}
+\begin{wrapfigure}{r}{0.25\textwidth}
+  \centering  \includegraphics[width=0.25\textwidth]{entail.png}
+  \caption{Understanding the level of $\varphi$ satisfaction in the \textsc{CoW} tree of modal nodes}
+   \label{fig:nodeentail} 
+\end{wrapfigure}
+We use a modality similar to $\text{\faSitemap}@[\ell]\phi$ to indicate that $\phi$ is true of the tree node rooted at $\ell$, as well as at \emph{all reachable children} of that node.
+This means that establishing a particular $\phi$ throughout a tree must be done in a bottom-up manner (see Figure \ref{fig:nodeentail}).
+Capturing this modally, rather than with more typical ad-hoc hand-crafted invariants, offers advantages in a number of proof steps, in addition to concisely capturing intuitive reasoning about subtrees with all nodes satisfying a particular property.
+One basic example is that while the actual disk sync code requires the full filesystem tree to be in an amplified state (hence write amplification), most of the filesystem works with more general trees.
+Since $\phi_{\textsf{amplified}}\vdash\phi_{\textsf{valid}}$, using our \faSitemap modality we can conclude $\text{\faSitemap}@[\ell]\phi_{\textsf{amplified}}\vdash\text{\faSitemap}@[\ell]\phi_{\textsf{valid}}$ by an analogue of Axiom K.
+There are also other places in the code where a subtree satisfying a stronger assertion at all nodes must be linked into a larger tree satisfying only a weaker assertion overall: predicate representing the valid tree as the tree of nodes which are backed by resources allocated pages, i.e.
+$\phi_{\textsf{valid}}$, is weaker than the predicate representing the tree of nodes being currently amplified in the current snapshot.
+Again, the same reasoning principle makes this straightforward.
+As alluded to earlier, our $\phi_{\textsf{amplified}}$ and company are in fact \emph{predicates} parameterized by a choice of node, though by lifing \textsf{Iris}'s BI structure they can use all of the standard separation logic connectives.
+The rest of this subsection develops this modality in full detail.
+%\colinsaysinl{I cut some text here about the three-way split (commented out) that I wasn't sure how to use, but maybe I deleted something important?}
+%To put it concretely, as shown in Figure \ref{fig:cow_index_tree.JPG}, \textsc{CoW} semantics determine what we can claim on the structure of the filesystem indexing tree:
+%\begin{itemize}
+%\item The bottom-up direction splits the tree into the two half at the cursor node: 1.
+the bottom part accessible from the cursor node and per-node-amplification is reflected in case needed 2.
+the upper part on which amplifcation has not reflected yet.
+%\item Horizontally, amplifcation path splits the bottom part of the tree into three different contexts: untouched tree nodes, nodes on the amplification path, and the ones generated as part of node splitting.
+%\end{itemize}
+%Based on these two observations, in fact, bottom-up amplification determines the choice of modal context by giving the boundries of \textit{any claimable knowledge} (e.g.
+$\varphi_{\textsf{amplified}}$, $\varphi_{\textsf{untouched}}$): a subtree bounded by the cursor node.
+
 \subsection{Nominalization}

### 2025-06-16 paper.tex
+

@@
 
-\subsection{Nominalization}
+\subsection{Nominals for Systems Resource Contexts}
 Finally, another design point is the question of whether or not resource element assertions must explicitly track

### 2025-06-16 paper.tex
+

@@
 \paragraph{Virtualization of Disk-Blocks in Filesystems}
-
+\textsf{ModFS} is a copy-on-write (\textsf{CoW}) filesystem, which as described in the introduction performs updates by copying modified data to \emph{new and unused} locations on disk rather than updating in place, ensuring that in the absence of physical storage corruption, there is \emph{always} a fully consistent filesystem on disk, removing the need for recovery from journals in the case of a crash, which can be very expensive for large volumes.
+%\textsf{ModFS} is implemented in \textsc{Go}~\cite{donovan2015go}, using the FUSE~\cite{fuse} interface to run a filesystem in usermode.
+\textsf{ModFS}'s code is translated to \textsf{Roqc} using the \textsf{Goose}~\cite{chajed2019verifying,chajed2021gojournal} tool, which translates a significant subset of \textsf{Go} to \textsf{Roqc} to permit formal verification of \textsc{Go} code using \textsf{Iris}~\cite{jung2018iris}.
+\textsf{Goose} generates code in an extension of \textsf{Iris}'s \textsf{HeapLang} with libraries for disk manipulation and \textsc{Go} primitive types (e.g., maps) as shown in Figure \ref{fig:modfslayout}.
+
+%On the left part of Figure \ref{fig:modfslayout}, we see the shaded region -- \textsf{ModFs} storage layer-- under the yellow layer --- the \textsf{ModFs} system call layer. 
+%We have verified the parts of the shaded storage layer involved in creating a new on-disk snapshot, up to an axiomatized specification of of the \textsf{Bcache} component that implements a simple in-memory cache of disk blocks.
+%This primarily involves the \textbf{write\_amplification} function on btree nodes (from the \textsf{Btree} module), and its associated dependencies, including the \textbf{split}, \textbf{put}, and \textbf{del}(ete) operations on nodes, and the \textbf{write} serialization function.
+%This also includes snapshot operations including \textbf{snap\_initialize} and \textbf{snap\_alloc} functions (among others) in the \textsf{Snapshot} module, and freelist operations (\textbf{free} and \textbf{allocate}) from \textsf{Freelist}, as well as the supporting functions from the \textsf{Chunk} (roughly a structured disk block abstraction), \textsf{PageMap} (raw disk block abstraction), \textsf{and Lock} (thread synchronization via a global lock map) modules.
+%\textsf{Storage} (a data structure of runtime pointers\colinsays{doesn't really explain it, I don't know what this component is/does}), \textsf{BtrItr} (an iterator library), \textsf{inode}, and \textsf{Btree} have not yet been fully verified, though the components involved in write amplification have been verified.
+The filesystem does run and standard filesystem functionality (core file and directory operations) work.
+%This paper focuses on the verification of write amplification, which is verified up to a small number of axioms for \textsf{Go} data structures\colinsays{update w/ very short summary of remaining admits} and the unverified \textsf{Bcache} specification.
+%\colinsaysinl{How many lines of code are covered by the verification?
+(excluding the bcache) This might help assuage reviewer concerns that we're not done, by making the scope of this project more explicit so ``why didn't you finish'' has a self-evident answer of ``because they already did a ton of work''}
+
+%\textsf{ModFs} is a \textsf{CoW} file system where the mapping of the logical locations, e.g., file pages indexed by file offsets, to disk locations by employing a \textsf{CoW} tree \cite{Astrahan1976SystemR,shadowclone} structure.
+In Figure \ref{fig:cow_index_tree}, we see that accessing and updating the $\textsf{page}_n$ of file \textsf{"foo"} referred by the file offset $\textsf{offset}$
+%\[ \textsf{write}(\textit{foo},\textsf{offset})\]
+%requires locating the corresponding physical disk block of the in-memory $\textsf{page}_n$ whose mapping is establised the \textit{copy-on-write} index tree (in the middle layer).
+Like Btrfs ~\cite{}, \textsf{ModFs} indexes all filesystem objects (e.g.
+files and directories) as a nested trees, and to map a logical page to a physical page, the file system first locates the root of the nested index-tree belonging to \textsf{"foo"}.
+Semantically, in \textsf{ModFS} (as in other \textsf{CoW} filesystems~\cite{young1992episode,bonwick2003zettabyte,rodeh2013btrfs,Rosenblum1992LFS,hitz1994file}) any update to file data or metadata is realized on  unused disk locations by treating the tree as a persistent data structure.
+As an example, Figure \ref{fig:cow_index} shows that an update to the file $\textsf{"foo"}$ at $\textsf{"offset"}$ ends up with locating the related disk chunk (a set of blocks) in the gray region, copying it to newly allocated disk chunks (dotted chunks on the right side of the bold long-dash-point line of the middle layer withing dotted section), and performing the update on these newly-allocated disk blocks.
+This requires further updates to those metadata blocks which previously referred to the old block locations to instead reference the newly-allocated blocks.
+This must be repeated until the allocation/update of the new chunks percolates up to the root (including the root), through a process called \textsf{\textit{write-amplification}}.
+
+\begin{figure}
+    \includegraphics[width=0.75\columnwidth]{cow_index.png}
+    \caption{A Filesystem called \textsf{Modfs} Indexing its Physical Pages for File Pages with CoW (copy-on-write) Tree}
+    \label{fig:cowindex}
+\end{figure}
+
+After this process, traversing the filesystem from either the old or new root yields consistent views of the filesystem at different points in time, each commonly referred to as a \emph{snapshot}.
+These snapshots overlap on the unmodified portions of the filesystem (the left-most grid section), with the reclaimed the chunks (the middle gray section) becoming available for reuse once the new snapshot becomes the official current state of the filesystem.\footnote{Many of these filesystems permit keeping archived snapshots of the filesystem state, e.g., a snapshot of the system every hour up to a certain number of days, to support user data recovery.
+In these cases only disk regions that are not still in use by an archived snapshot are reclaimed.}.
+This permits the filesystem to atomically switch between consistent on-disk states, by updating the oldest so-called \emph{super} chunk at a fixed location on disk.\footnote{Filesystems must take care to ensure that possibly-flaky hardware truly does perform an atomic update.} Super chunks, as we see in Figure \ref{fig:cow_index} as ($\textsf{Super}_0$ and $\textsf{Super}_1$), are the first two blocks of the disk dedicated to keep the state of the file system which is composed of some meta-data on the filesystem, and most importantly to the root chunk of the file system index-tree.
+Therefore, in case of a crash during the update to the \textsf{file "foo"}, we can reboot the file system from the previous persisted root, \textsf{root chunk}, via reading the $\textsf{Super}_0$.
+This means the whole disk consistency checks are unnecessary when recovering from a crash~\cite{bonwick2003zettabyte,hitz1994file}; recovery is immediate, as opposed to requiring hours or days~\cite{henson2006reducing,agrawal2009speedy} (or longer~\cite{henson2006chunkfs}) for recovering a large volume’s consistency from logs.
+An important additional subtlety with this kind of filesystem is \emph{when} blocks in the reclaimed region can actually be reused: only \emph{after} the snapshot that made those blocks ``free'' has been completely written to disk.
+Allowing reuse before this could result in in-place updates: part of the new snapshot could attempt to store data in space used by the prior snapshot, allowing inconsistent on-disk structures.
+This means the filesystem must carefully track not only what blocks are reclaimed, but \emph{when} they are safe to reuse, making the \textsf{Freelist} component of \textsf{Modfs} and other \textsf{Roqc} filesystems non-trivial.
+With their intricate semantics and critical place in modern systems, formal verification of functional correctness for storage systems (including filesystems and other kinds of storage) have drawn much attention~\cite{ntzik-gardner-15,ntzik2018concurrent,chen-crash-16}\todo{lots more cites}.
+Unlike our \textsc{CoW} filesystem \textsf{ModFs} which allocates new chunks and rewrites all the nodes from leaf to root on the write-amplification path, some filesystems allows in-place updates to handle the heavy workloads more efficiently in principle.
+Unlike \textsc{CoW} filesystems, filesystems with in-place updates need to implement some form of \emph{journalling} to ensure that a crash partway through an in-place update can be rolled back (using an undo journal that records the original contents before modification) or completed (using a write-ahead journal that first writes new data in unused space before \emph{also} updating in place).\footnote{\textsf{CoW} filesystems can be viewed as a twist on write-ahead journals that uses the journal as the actual data once written to a fresh location, thereby reclaiming the original location.
+This results in a performance and complexity trade-off: journalling writes twice for every updated block and requires recovery logic to roll back or re-execute the log, while \textsf{CoW} techniques write each update only once and require slightly more intricate allocation strategies to compact fragmentation.}
+\paragraph{In-Place Updates in Filesystems}In-place updating filesystems with journalling components have been specified and verified with an extended Hoare Logic to specify not only the pre/post conditions of the operations successful operations in the filesystem, but also the crash condition~\cite{chen-crash-16,chen2017verifying}, with later work extending this for concurrent filesystem implementations~\cite{chajed2021gojournal}.
+\textsf{Perennial}~\cite{chajed2019verifying} extends the weakest precondition of the \textsf{Iris} separation logic~\cite{jung2018iris,krebbers2017essence} to allow asserting facts about disk state at crashes (the \emph{crash condition}) and apply this extension to a journalling filesystem \textsf{GoJournal}~\cite{chajed2021gojournal}.
+Unlike \textsf{GoJournal}, \textsf{ModFs} relies on \textsf{CoW} semantics for consistency, eliminating the need for recovery logic.
+Essentially, successfully writing the new superblock for a new snapshot is analagous to completing a valid journal entry, except it \emph{also} has the effect of updating the live filesystem data due to making the ``journal entries'' valid filesystem data and metadata blocks (an idea that goes back to early work on log-structured filesystems~\cite{Rosenblum1992LFS}).
+In fact, we can think of the atomicity of the recovery from the crash state conceptually equal to the single block disk write persisting the change from $\textsf{Super}_0$ to $\textsf{Super}_1$ in Figure \ref{fig:cow_index_tree_update}, therefore, considering \textsc{CoW} semantics, we do not need to consider the specification and proof of crash condition separately; instead a global never-violated invariant (not even temporarily broken) of the filesystem ensures the on-disk state is always consistent.
 \paragraph{Consistency of Disk Layout} \todo{}
+
 \paragraph{Memory Reclamation in Client Code}\todo{iso:speak about Amal's stuff}

### 2025-06-16 paper.tex
+

@@
 This results in a performance and complexity trade-off: journalling writes twice for every updated block and requires recovery logic to roll back or re-execute the log, while \textsf{CoW} techniques write each update only once and require slightly more intricate allocation strategies to compact fragmentation.}
-\paragraph{In-Place Updates in Filesystems}In-place updating filesystems with journalling components have been specified and verified with an extended Hoare Logic to specify not only the pre/post conditions of the operations successful operations in the filesystem, but also the crash condition~\cite{chen-crash-16,chen2017verifying}, with later work extending this for concurrent filesystem implementations~\cite{chajed2021gojournal}.
+\begin{figure}
+    \includegraphics[width=\columnwidth]{journalling.png}
+    \caption{A Filesystem Utilizing \emph{Journalling} for Consistent Disk-Layout ~\cite{chajed2021gojournal}}
+    \label{fig:journal}
+\end{figure}
+
+\paragraph{Different Models in Ensuring Disk Consistency of Filesystems}In-place updating filesystems with journalling components have been specified and verified with an extended Hoare Logic to specify not only the pre/post conditions of the operations successful operations in the filesystem, but also the crash condition~\cite{chen-crash-16,chen2017verifying}, with later work extending this for concurrent filesystem implementations~\cite{chajed2021gojournal}.
 \textsf{Perennial}~\cite{chajed2019verifying} extends the weakest precondition of the \textsf{Iris} separation logic~\cite{jung2018iris,krebbers2017essence} to allow asserting facts about disk state at crashes (the \emph{crash condition}) and apply this extension to a journalling filesystem \textsf{GoJournal}~\cite{chajed2021gojournal}.
@@
 In fact, we can think of the atomicity of the recovery from the crash state conceptually equal to the single block disk write persisting the change from $\textsf{Super}_0$ to $\textsf{Super}_1$ in Figure \ref{fig:cow_index_tree_update}, therefore, considering \textsc{CoW} semantics, we do not need to consider the specification and proof of crash condition separately; instead a global never-violated invariant (not even temporarily broken) of the filesystem ensures the on-disk state is always consistent.
-\paragraph{Consistency of Disk Layout} \todo{}
-
-\paragraph{Memory Reclamation in Client Code}\todo{iso:speak about Amal's stuff}
+\paragraph{Memory Reclamation in Client Code} Allocating memory locations for a client code leaves the 
 \paragraph{Consistency of Reclaimed Memory}\todo{iso: speak about the operator for which the hybrid logic stuff is usefull }
-
-\subsection{Speculative Observation in Systems Software}
-\label{sec:speculativeobservation}
-\section{Introduction}\label{sec:intro}
-\todo[inline]{Colin says: These comments aren't the only places things need to change, but they're
-specific places I have suggestions.
-One thing to keep in mind is that the CfP lists two kinds of papers:
-short papers are for "original research or case studies, maybe in the incubation phase,"
-while long papers are for "original research, case studies, or surveys."
-This is clearly in the survey category, so we (a) should take advantage of the extra space to
-explain things clearly (15 pages, plus 2 pages for bibliography), and (b) lean into the survey nature explicitly.
-}
-
-\todo[inline]{I think one of the biggest things we can do to reframe this is to explain the broader vision
-here before jumping into contributions.
-This is an audience that knows dynamic logic and modal logic more broadly, but perhaps not everyone is familiar
-with hybrid logic, and many will not be familiar with systems concepts.
-This is the place to provide high-level intuition about the domain we're surveying, tailored for this audience.
-I'd talk about dynamic logic in general, but then highlight the need for reasoning specific to certain
-compute resources to produce manageable specifications and proofs.
-Think big picture: why is this space interesting to logicians?
-}
 
@@
 
-\subsection{Speculation in Consistency}
+\subsection{Speculation in Consistency} \todo{speak about the weak-mem modalities in here.}
 %\subsection{Sharing} Sharing is one of the fundamental reasons for kernels to {virtualize} resources,

### 2025-06-16 paper.tex
+

@@
 In fact, we can think of the atomicity of the recovery from the crash state conceptually equal to the single block disk write persisting the change from $\textsf{Super}_0$ to $\textsf{Super}_1$ in Figure \ref{fig:cow_index_tree_update}, therefore, considering \textsc{CoW} semantics, we do not need to consider the specification and proof of crash condition separately; instead a global never-violated invariant (not even temporarily broken) of the filesystem ensures the on-disk state is always consistent.
-\paragraph{Memory Reclamation in Client Code} Allocating memory locations for a client code leaves the 
-\paragraph{Consistency of Reclaimed Memory}\todo{iso: speak about the operator for which the hybrid logic stuff is usefull }
-
+\paragraph{Memory Reclamation in Client Code} In a typical configuration, allocating, deallocating and referencing memory locations for a client code leave the memory layout in a form of DAG (directed-asyclic-graph) with a unique root which resembles the first stack pointer referencing the allocated memory.
+To ease picturing the memory layout after a set of requests from the client code (for allocation, referencing etc.), we can think of the unique roots of DAGs as \emph{super} block addresses -- this time multiple ones, not only two -- in Figure \ref{fig:cow_index}, from which a path is created for a stack pointer reference to the allocated memory.
+To reclaim the allocated memory location that is no longer being accessed, systems uses a well-known method called \emph{reference-counting} in which a new reference to the allocated memory increases the reference counter on it w.r.t a certain regime.
+\paragraph{Consistency of Reclaimed Memory} A sound memory reclamation is based on the condition where there is no \emph{reachability} to the reference pointing to the allocated memory reference, and it is not \emph{shared}.
+Once the requirements are satisfied then the memory location can be reclaimed \emph{soundly} -- i.e., without violating any access through a reference to it.
 \subsection{Conclusion \& Contributions}

### 2025-06-16 paper.tex
+

@@
 To reclaim the allocated memory location that is no longer being accessed, systems uses a well-known method called \emph{reference-counting} in which a new reference to the allocated memory increases the reference counter on it w.r.t a certain regime.
-\paragraph{Consistency of Reclaimed Memory} A sound memory reclamation is based on the condition where there is no \emph{reachability} to the reference pointing to the allocated memory reference, and it is not \emph{shared}.
+\paragraph{Consistency of Reclaimed Memory} A sound memory reclamation is based on the conditions where the unique is not \emph{shared}, and there is no \emph{reachability} from the root to any the references pointing to the allocated memory.
 Once the requirements are satisfied then the memory location can be reclaimed \emph{soundly} -- i.e., without violating any access through a reference to it.
+\paragraph{Buffering} Buffering is a technique used in variety of systems (filesystem page caching, write buffer in multi-processor architectures, translation-lookaside-buffers for translated addresses etc.) to delay the costly operation (committing file pages, flushing buffered writes from write-cache to memory etc.) making updates consistent.
+\paragraph{Consistency of Reads of Buffered Writes} Consistency of reads of any value potentially buffered depends on the semantics of the operation making them \emph{observable}.
+A typical architecture such as Intel-x86 considers the \emph{flush} of buffered writes \emph{non-deterministically} -- i.e., at any time at any number of buffered writes.
+On the other hand, a typical filesystem can follow a certain policy to commit the in-memory file pages to the disk.
 \subsection{Conclusion \& Contributions}
-
 \paragraph{Conclusion} Low-level systems exhibit certain patterns in their designs, especially when interacting with computing resources.

### 2025-06-17 paper.tex
+

@@
 \label{sec:background}
+One of the most common forms of abstraction provided by systems software is \emph{virtualization},
+which abstracts the relationship between conceptual and physical computing resources. 
+Operating System (OS) kernels virtualize memory locations and quantity (via virtual memory and paging~\cite{Denning1970VM}).
+Distributed language runtimes may virtualize addresses, even when processes may migrate across machines~\cite{jul1988fine}.
+Filesystems virtualize locations on disk~\cite{rodeh2013btrfs,hitz1994file,Rosenblum1992LFS,bonwick2003zettabyte}.
+Programs built on top of the corresponding systems software layer work logically at the level of these virtualized resources,
+and it makes sense to specify the systems software directly in terms of those abstractions.
+\paragraph{Access by Translation}
+Accessing virtualized resources via translation is a common way to virtualize notions of location (e.g.,
+virtual memory addresses, inodes or object IDs instead of disk addresses~\cite{bonwick2003zettabyte,hitz1994file,Rosenblum1992LFS,rodeh2013btrfs}).
+%Accessing the hardware resources in the presence of sharing means having more resource references (references to the virtualized resources) to access the \emph{mapped} resources that are physical and potentially fewer.
+B-trees, page tables, and related structures both behave like maps, when the corresponding physical resources exist as such
+just in a different location.
+Control over the lookup process (e.g, handling the case of a missing translation entry) allows for additional flexibility,
+such as filling holes in sparse files, or demand paging (both from disk or lazily populating anonymous initially-zeroed mappings).
+Although the realization of these maps may differ from a system to system based on the context (and sometimes hardware details),
+they are semantically --- logically --- partial maps, worth treating as such in verification.
 \paragraph{Virtualization of Memory Locations}
@@
 Allowing reuse before this could result in in-place updates: part of the new snapshot could attempt to store data in space used by the prior snapshot, allowing inconsistent on-disk structures.
-This means the filesystem must carefully track not only what blocks are reclaimed, but \emph{when} they are safe to reuse, making the \textsf{Freelist} component of \textsf{Modfs} and other \textsf{Roqc} filesystems non-trivial.
-With their intricate semantics and critical place in modern systems, formal verification of functional correctness for storage systems (including filesystems and other kinds of storage) have drawn much attention~\cite{ntzik-gardner-15,ntzik2018concurrent,chen-crash-16}\todo{lots more cites}.
-Unlike our \textsc{CoW} filesystem \textsf{ModFs} which allocates new chunks and rewrites all the nodes from leaf to root on the write-amplification path, some filesystems allows in-place updates to handle the heavy workloads more efficiently in principle.
-Unlike \textsc{CoW} filesystems, filesystems with in-place updates need to implement some form of \emph{journalling} to ensure that a crash partway through an in-place update can be rolled back (using an undo journal that records the original contents before modification) or completed (using a write-ahead journal that first writes new data in unused space before \emph{also} updating in place).\footnote{\textsf{CoW} filesystems can be viewed as a twist on write-ahead journals that uses the journal as the actual data once written to a fresh location, thereby reclaiming the original location.
-This results in a performance and complexity trade-off: journalling writes twice for every updated block and requires recovery logic to roll back or re-execute the log, while \textsf{CoW} techniques write each update only once and require slightly more intricate allocation strategies to compact fragmentation.}
 \begin{figure}
@@
 
-\paragraph{Different Models in Ensuring Disk Consistency of Filesystems}In-place updating filesystems with journalling components have been specified and verified with an extended Hoare Logic to specify not only the pre/post conditions of the operations successful operations in the filesystem, but also the crash condition~\cite{chen-crash-16,chen2017verifying}, with later work extending this for concurrent filesystem implementations~\cite{chajed2021gojournal}.
-\textsf{Perennial}~\cite{chajed2019verifying} extends the weakest precondition of the \textsf{Iris} separation logic~\cite{jung2018iris,krebbers2017essence} to allow asserting facts about disk state at crashes (the \emph{crash condition}) and apply this extension to a journalling filesystem \textsf{GoJournal}~\cite{chajed2021gojournal}.
-Unlike \textsf{GoJournal}, \textsf{ModFs} relies on \textsf{CoW} semantics for consistency, eliminating the need for recovery logic.
+\paragraph{Different Models in Ensuring Disk Consistency of Filesystems}
+With their intricate semantics and critical place in modern systems, formal verification of functional correctness for storage systems (including filesystems and other kinds of storage) have drawn much attention~\cite{ntzik-gardner-15,ntzik2018concurrent,chen-crash-16}\todo{lots more cites}.
+Unlike our \textsc{CoW} filesystem \textsf{ModFs} which allocates new chunks and rewrites all the nodes from leaf to root on the write-amplification path, some filesystems allows in-place updates to handle the heavy workloads more efficiently in principle.
+Unlike \textsc{CoW} filesystems, filesystems with in-place updates need to implement some form of \emph{journalling} to ensure that a crash partway through an in-place update can be rolled back (using an undo journal that records the original contents before modification) or completed (using a write-ahead journal that first writes new data in unused space before \emph{also} updating in place).\footnote{\textsf{CoW} filesystems can be viewed as a twist on write-ahead journals that uses the journal as the actual data once written to a fresh location, thereby reclaiming the original location.
+This results in a performance and complexity trade-off: journalling writes twice for every updated block and requires recovery logic to roll back or re-execute the log, while \textsf{CoW} techniques write each update only once and require slightly more intricate allocation strategies to compact fragmentation.}
+
 Essentially, successfully writing the new superblock for a new snapshot is analagous to completing a valid journal entry, except it \emph{also} has the effect of updating the live filesystem data due to making the ``journal entries'' valid filesystem data and metadata blocks (an idea that goes back to early work on log-structured filesystems~\cite{Rosenblum1992LFS}).
-In fact, we can think of the atomicity of the recovery from the crash state conceptually equal to the single block disk write persisting the change from $\textsf{Super}_0$ to $\textsf{Super}_1$ in Figure \ref{fig:cow_index_tree_update}, therefore, considering \textsc{CoW} semantics, we do not need to consider the specification and proof of crash condition separately; instead a global never-violated invariant (not even temporarily broken) of the filesystem ensures the on-disk state is always consistent.
+In fact, we can think of the atomicity of the recovery from the crash state conceptually equal to the single block disk write persisting the change from $\textsf{Super}_0$ to $\textsf{Super}_1$ in Figure \ref{fig:cow_index}, therefore, considering \textsc{CoW} semantics, we do not need to consider the specification and proof of crash condition separately; instead a global never-violated invariant (not even temporarily broken) of the filesystem ensures the on-disk state is always consistent.
 \paragraph{Memory Reclamation in Client Code} In a typical configuration, allocating, deallocating and referencing memory locations for a client code leave the memory layout in a form of DAG (directed-asyclic-graph) with a unique root which resembles the first stack pointer referencing the allocated memory.
@@
 \end{itemize}
-\section{Resources in Systems Software}
-\label{sec:systemsoft}
-In this section, we argue the position that \emph{modalities}  
-should be a go-to approach when specifying and verifying low-level systems code.
-We explain how the concept of a \emph{resource context} helps guide the design of new modalities
-for verification of systems code, and
-we justify our perspective by discussing prior systems that have used modalities for systems verification
-successfully, arguing that they fit into the verification design pattern we articulate,
-and explaining how this approach might apply to other systems verification challenges.
-
-To explain our ideas in the general systems understanding, we briefly recap some of the background and themes our ideas build on again,casting them in a certain way to bring out the relevance of our philosophy.
-
-\subsection{Virtualization}
-One of the most common forms of abstraction provided by systems software is \emph{virtualization},
-which abstracts the relationship between conceptual and physical computing resources. 
-Operating System (OS) kernels virtualize memory locations and quantity (via virtual memory and paging~\cite{Denning1970VM}).
-Distributed language runtimes may virtualize addresses, even when processes may migrate across machines~\cite{jul1988fine}.
-Filesystems virtualize locations on disk~\cite{rodeh2013btrfs,hitz1994file,Rosenblum1992LFS,bonwick2003zettabyte}.
-Programs built on top of the corresponding systems software layer work logically at the level of these virtualized resources,
-and it makes sense to specify the systems software directly in terms of those abstractions.
-\paragraph{Access by Translation}
-Accessing virtualized resources via translation is a common way to virtualize notions of location (e.g.,
-virtual memory addresses, inodes or object IDs instead of disk addresses~\cite{bonwick2003zettabyte,hitz1994file,Rosenblum1992LFS,rodeh2013btrfs}).
-%Accessing the hardware resources in the presence of sharing means having more resource references (references to the virtualized resources) to access the \emph{mapped} resources that are physical and potentially fewer.
-B-trees, page tables, and related structures both behave like maps, when the corresponding physical resources exist as such
-just in a different location.
-Control over the lookup process (e.g, handling the case of a missing translation entry) allows for additional flexibility,
-such as filling holes in sparse files, or demand paging (both from disk or lazily populating anonymous initially-zeroed mappings).
-Although the realization of these maps may differ from a system to system based on the context (and sometimes hardware details),
-they are semantically --- logically --- partial maps, worth treating as such in verification.
-
-\subsection{Speculation in Consistency} \todo{speak about the weak-mem modalities in here.}
+%\section{Resources in Systems Software}
+%\label{sec:systemsoft}
+
+%\subsection{Speculation in Consistency} \todo{speak about the weak-mem modalities in here.}
 %\subsection{Sharing} Sharing is one of the fundamental reasons for kernels to {virtualize} resources,
@@
 \section{Contingency Decomposition of a System}
+In this section, we argue the position that \emph{modalities}  
+should be a go-to approach when specifying and verifying low-level systems code.
+We explain how the concept of a \emph{resource context} helps guide the design of new modalities
+for verification of systems code, and
+we justify our perspective by discussing prior systems that have used modalities for systems verification
+successfully, arguing that they fit into the verification design pattern we articulate,
+and explaining how this approach might apply to other systems verification challenges.
+
+To explain our ideas in the general systems understanding, we briefly recap some of the background and themes our ideas build on again,casting them in a certain way to bring out the relevance of our philosophy.
+
 \todo[inline]{Colin says: I don't think this section needs to change much,
@@
 Write amplification traverses the overall tree depth-first, transforming the tree from the bottom up, one full subtree at a time.
-Each time a node is split, it results in disconnecting some nodes from the next snapshot, sharing other nodes, and linking some newly-allocated nodes into the new snapshot, as in Figure \ref{fig:cow_index_tree_update}.
+Each time a node is split, it results in disconnecting some nodes from the next snapshot, sharing other nodes, and linking some newly-allocated nodes into the new snapshot, as in Figure \ref{fig:cow_index}.
 Nodes are split (or deemed not to require splitting) one at a time, and considered only after all child nodes have been amplified.
@@
 $\varphi_{\textsf{amplified}}$, $\varphi_{\textsf{untouched}}$): a subtree bounded by the cursor node.
-
 \subsection{Nominals for Systems Resource Contexts}

### 2025-06-17 paper.tex
+

@@
 Write amplification traverses the overall tree depth-first, transforming the tree from the bottom up, one full subtree at a time.
-Each time a node is split, it results in disconnecting some nodes from the next snapshot, sharing other nodes, and linking some newly-allocated nodes into the new snapshot, as in Figure \ref{fig:cow_index}.
-Nodes are split (or deemed not to require splitting) one at a time, and considered only after all child nodes have been amplified.
 Concepually this means the amplification code's proof must track the fact that at any point in time, there are regions of the tree where each node has already been amplified (e.g., $\phi_{\textsf{amplified}}$ definitely holds for that node), and regions where nodes may or may not require amplification, and this will not be known until those nodes are visited (i.e., some basic validity $\phi_{\textsf{valid}}$ is known to hold, but not the stronger $\phi_{\textsf{amplified}}$).
@@
 As alluded to earlier, our $\phi_{\textsf{amplified}}$ and company are in fact \emph{predicates} parameterized by a choice of node, though by lifing \textsf{Iris}'s BI structure they can use all of the standard separation logic connectives.
-The rest of this subsection develops this modality in full detail.
 %\colinsaysinl{I cut some text here about the three-way split (commented out) that I wasn't sure how to use, but maybe I deleted something important?}

### 2025-06-17 paper.tex
colin can look at now

@@
 %
-\paragraph{Modal Abstractions for Recursive Trees Structure with Nominals}
+\subsection{Nominals for Systems Resource Contexts}
+Finally, another design point is the question of whether or not resource element assertions must explicitly track
+their corresponding context, or if they implicitly pick up their context from where they are used.
+
+\emph{Strong nominalization} is the case where resource elements must explicitly include
+the identity of their intended context, while \emph{weak nominalization} occurs when the resource
+elements implicitly pick up the relevant context from how they are used.
+The first three modalities in Table \ref{table:decomposition} are strongly nominalized,
+with the resource elements generally carrying identifiers of a specific modality usage.\footnote{The Post-Crash modality does not look like this in the presentation here; technically the definition of $\diamond$ quantifies over names $\overline{\gamma}$ internally, dealing with sets of possible contexts.}
+The last three are weakly nominalized.
+
+This choice trades off complexity against flexibility and scoping constraints.
+Strongly nominalized elements track additional specifier/prover-visible book-keeping data.
+But in exchange for carrying those identifiers of their context with them, 
+strongly nominalized elements can be used together under any modality. For example, one can
+use the StackRegion modality to talk about two different stack frames simultaneously for code
+which accesses multiple stack frames: $\fbox{n} \; \ell \mapsto v \ast \fbox{n+1} \; \ell' \mapsto v'$.
+Using a given strongly nominalized assertion element under different modalities for different
+frames does not change its meaning.
+
+By contrast, weakly nominalized elements are more concise, but then make talking about multiple 
+contexts together marginally more complex: changing which modality an assertion is used with drastically changes its meaning. In the case of the Ref-Count modality, $@_\ell(\ell_1\mapsto v)$
+says that $\ell$ points to a reference count wrapping $\ell_1$, while placing the $\ell_1\mapsto v$ under a jump modality for a different location entails talking about a different region of memory.
+
+In general, use cases where code frequently manipulates small parts of multiple contexts together
+should prefer strong nominalization, while use cases where usually larger portions of a single context
+are at issue should prefer weak nominalization.
+\section{Universal Properties over Resource Contexts with Nominals:  An Filesystem Recursive Trees Structure}
 \label{sec:overviewnode}
+Up until now, we discuss the concept of resource and resource context related under a certain \emph{nominalization} scheme.
+These definitions establishes the foundation for local properties of modal patterns.
+In this section, we discuss how we would be discussing the universal properties over the modal context themselves through an filesystem tree example.
+By doing so, we are able to show how the local properties defined over a resource and its context can be interfaced and exposed universally through reasoning over resource contexts.
 
@@
 $\varphi_{\textsf{amplified}}$, $\varphi_{\textsf{untouched}}$): a subtree bounded by the cursor node.
-\subsection{Nominals for Systems Resource Contexts}
-Finally, another design point is the question of whether or not resource element assertions must explicitly track
-their corresponding context, or if they implicitly pick up their context from where they are used.
-
-\emph{Strong nominalization} is the case where resource elements must explicitly include
-the identity of their intended context, while \emph{weak nominalization} occurs when the resource
-elements implicitly pick up the relevant context from how they are used.
-The first three modalities in Table \ref{table:decomposition} are strongly nominalized,
-with the resource elements generally carrying identifiers of a specific modality usage.\footnote{The Post-Crash modality does not look like this in the presentation here; technically the definition of $\diamond$ quantifies over names $\overline{\gamma}$ internally, dealing with sets of possible contexts.}
-The last three are weakly nominalized.
-
-This choice trades off complexity against flexibility and scoping constraints.
-Strongly nominalized elements track additional specifier/prover-visible book-keeping data.
-But in exchange for carrying those identifiers of their context with them, 
-strongly nominalized elements can be used together under any modality. For example, one can
-use the StackRegion modality to talk about two different stack frames simultaneously for code
-which accesses multiple stack frames: $\fbox{n} \; \ell \mapsto v \ast \fbox{n+1} \; \ell' \mapsto v'$.
-Using a given strongly nominalized assertion element under different modalities for different
-frames does not change its meaning.
-
-By contrast, weakly nominalized elements are more concise, but then make talking about multiple 
-contexts together marginally more complex: changing which modality an assertion is used with drastically changes its meaning. In the case of the Ref-Count modality, $@_\ell(\ell_1\mapsto v)$
-says that $\ell$ points to a reference count wrapping $\ell_1$, while placing the $\ell_1\mapsto v$ under a jump modality for a different location entails talking about a different region of memory.
-
-In general, use cases where code frequently manipulates small parts of multiple contexts together
-should prefer strong nominalization, while use cases where usually larger portions of a single context
-are at issue should prefer weak nominalization.
-
 %$\paragraph{Strong Nominalization} Some resource elements require strong \emph{association} with their contexts as \emph{up-to} proper identification of resource context.

### 2025-06-17 paper.tex
+

@@
 Unlike our \textsc{CoW} filesystem \textsf{ModFs} which allocates new chunks and rewrites all the nodes from leaf to root on the write-amplification path, some filesystems allows in-place updates to handle the heavy workloads more efficiently in principle.
-Unlike \textsc{CoW} filesystems, filesystems with in-place updates need to implement some form of \emph{journalling} to ensure that a crash partway through an in-place update can be rolled back (using an undo journal that records the original contents before modification) or completed (using a write-ahead journal that first writes new data in unused space before \emph{also} updating in place).\footnote{\textsf{CoW} filesystems can be viewed as a twist on write-ahead journals that uses the journal as the actual data once written to a fresh location, thereby reclaiming the original location.
+Unlike \textsc{CoW} filesystems, filesystems with in-place updates need to implement some form of \emph{journalling} to ensure that a crash partway through an in-place update can be rolled back (using an undo journal that records the original contents before modification \ref{fig:journal}) or completed (using a write-ahead journal that first writes new data in unused space before \emph{also} updating in place).\footnote{\textsf{CoW} filesystems can be viewed as a twist on write-ahead journals that uses the journal as the actual data once written to a fresh location, thereby reclaiming the original location.
 This results in a performance and complexity trade-off: journalling writes twice for every updated block and requires recovery logic to roll back or re-execute the log, while \textsf{CoW} techniques write each update only once and require slightly more intricate allocation strategies to compact fragmentation.}

### 2025-06-17 paper.tex
+

--- 2025-06-17 paper.tex
@@
 Unlike our \textsc{CoW} filesystem \textsf{ModFs} which allocates new chunks and rewrites all the nodes from leaf to root on the write-amplification path, some filesystems allows in-place updates to handle the heavy workloads more efficiently in principle.
-Unlike \textsc{CoW} filesystems, filesystems with in-place updates need to implement some form of \emph{journalling} to ensure that a crash partway through an in-place update can be rolled back (using an undo journal that records the original contents before modification) or completed (using a write-ahead journal that first writes new data in unused space before \emph{also} updating in place).\footnote{\textsf{CoW} filesystems can be viewed as a twist on write-ahead journals that uses the journal as the actual data once written to a fresh location, thereby reclaiming the original location.
+Unlike \textsc{CoW} filesystems, filesystems with in-place updates need to implement some form of \emph{journalling} to ensure that a crash partway through an in-place update can be rolled back (using an undo journal that records the original contents before modification as abstarcted in Figure \ref{fig:journal}) or completed (using a write-ahead journal that first writes new data in unused space before \emph{also} updating in place).\footnote{\textsf{CoW} filesystems can be viewed as a twist on write-ahead journals that uses the journal as the actual data once written to a fresh location, thereby reclaiming the original location.
 This results in a performance and complexity trade-off: journalling writes twice for every updated block and requires recovery logic to roll back or re-execute the log, while \textsf{CoW} techniques write each update only once and require slightly more intricate allocation strategies to compact fragmentation.}

### 2025-06-23 paper.tex
comments

@@
   \end{comment}
+  \todo[inline,color=cyan]{Ismail, why did you comment out my suggestion and keep the original?}
 \end{abstract}
@@
 Although they differ in the functionality they offer, low-level systems exhibit certain patterns of design and utilization of computing resources.
-  \todo[inline]{this section is already nice for introducing the ideas and what you mean by "systems software."
-  This section mostly needs a little more background on systems concepts like virtual memory, or what it means
-  to virtualize locations of blocks on disk}
-Systems software, in general, interfaces with an underlying computing architecture such that any software system at any higher level in the software stack can (at least indirectly) utilize the resources of the machine.
+Systems software, in general, interfaces with an underlying computing substrate such that any software system at any higher level in the software stack can (at least indirectly) utilize the resources of the machine.
 The last layer of software before the hardware is naturally critical to the correctness of an overall system,
@@
 the construction of higher layers of the stack.
+
+% \subsection{Conclusion \& Contributions}
+% \paragraph{Conclusion} 
+Low-level systems software exhibits certain patterns in their designs, especially when interacting with computing resources. 
+Exploiting certain patterns while designing software has been an important field of study.
+This survey examines these common patterns in systems software, and their relationships
+to modalities. We believe that certain properties of modalities enable us to understand and 
+address the verification challenges of these systems by tailoring specification and proof
+to the same patterns used to design the systems.
+
+\paragraph{Contributions} We argue that modal abstractions can be used to identify and abstract system verification challenges.
+We justify our perspective by discussing prior systems that have successfully used modalities for system verification, arguing that they fit into the verification design pattern we articulate, and explaining how this approach might apply to other systems' verification challenges.
+\begin{itemize}
+\item \textbf{Identifying System Verification Challenges}: We start with identifying common patterns in system verification: \emph{virtualization, sharing, and translation}.
+\item \textbf{Introducing the Concept of \emph{Resource}}: Then we discuss the concept of \emph{resource} which has already been an essential concept in the design of systems.
+Inspired by the concept of resource in the systems, we define what a resource and its context are in the modal abstractions.
+\item \textbf{Introducing the Concept of \emph{Nominals for System Resources}}:
+Nominalization enables identifying a resource in a context. 
+For example, a transaction is a context of resources of in-memory updated disk blocks. i
+The transaction identifier is used to associate a transaction with a disk-block to be persisted so that, 
+in case of a crash while persisting updated disk-blocks, the filesystem can rollback the 
+\emph{already persisted} disk-blocks of the transaction, and reach to the previous consistent disk state. 
+To be able to do so, both the updated in-memory disk blocks and the transaction must refer to the transaction 
+identifier -- \emph{strong nominalization}. In another example, virtual memory references (resources) in 
+an address space (resource context), which can be uniquely identified with a root address (the nominal) 
+of its page-table tree, are \emph{agnostic} to the address space that they are in. 
+However, they can only be accessed (be valid) in the address space to which they are agnostic. 
+However, then an address space switch happens, the virtual memory references of the previous address 
+space must be made inaccessible. To be able to do so, although virtual memory references do not hold any 
+piece of information related to their address space, we still have to associate them. 
+We call this kind of unilateral nominalization -- \emph{weak nominalization}.
+\todo[inline,color=cyan]{I left this paragraph alone, mostly, but it's too detailed for the new intro section position here}
+\item \textbf{Taxonomy of Current Modal Approaches in System Verification}: Based on these concepts defined, we summarize contemporary verification efforts using modal abstractions.
+We choose them from different domains, for example, reasoning about weak memory models and storage persistence ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2024modalabstractionsvirtualizingmemory,amalreal2024} because we would like to justify that our \emph{definitions} are not domain dependent.
+\end{itemize}
+
+\section{Background on Modal Logics for Programs, and Low-Level Systems Software}
+
+This section briefly recalls related concepts from modal logic of particular relevance
+to our taxonomy, which should be familiar but benefit from fixing notation (Section \ref{sec:modalbackground}),
+and enough general background on the systems software concepts organized by our taxonomy
+for the connections to modalities to be clear (Section \ref{sec:background}).
+
 \subsection{Hybrid Logic Dynamic Logic and Nominals in Program Logics}
@@
 \paragraph{Hybrid Logic} Hybrid logics~\cite{blackburn1995hybrid,goranko1996hierarchies,gargov1993modal,areces2001hybrid,brauner2010hybrid} are a class of modal logics that add two primary new concepts to a logic.
-Nominals $\loc \in \textsf{Loc}$ uniquely identify points in a model (e.g., a particular state), so there is exactly one point in the model where a nominal $\loc$ is true.
+Nominals $\l \in \textsf{Loc}$ uniquely identify points in a model (e.g., a particular state), so there is exactly one point in the model where a nominal $\loc$ is true.
 Satisfaction operators are modal operators indexed by nominals, which enable claims about the truth of another proposition at some arbitrary point in the model identified by a nominal.
@@
 However, most prior work on hybrid logics works in classical logics rather than a substructural setting like separation logic.
-We adapt these ideas to assertions of the form $@[l]\varphi$asserts that $\varphi$ is true in a data structure with \emph{logical} state $\sigma$ \emph{at location} $l$.
+We adapt these ideas to assertions of the form $@[l]\varphi$ asserts that $\varphi$ is true in a data structure with \emph{logical} state $\sigma$ \emph{at location} $l$.\todo{seems out of place?}
 In practice we must design variants of this for each kind of structure we wish to apply this to, e.g. tree nodes, free list, or snapshots.
@@
 Protocol assertions are \emph{annotated with the name of the last (abstract) state at which the protocol is ensured}.
-An example more familiar to the systems community is Halpern et al.'s history of adapting modal logics of
+One classic application of this idea to systems is 
+Halpern et al.'s work adapting modal logics of
 knowledge to deal with distributed systems~\cite{halpern1990knowledge,halpern1989modelling,halpern2017reasoning}.
@@
 such as filling holes in sparse files, or demand paging (both from disk or lazily populating anonymous initially-zeroed mappings).
-Although the realization of these maps may differ from a system to system based on the context (and sometimes hardware details),
+Although the realization of these maps may differ from system to system based on the context (and sometimes hardware details),
 they are semantically --- logically --- partial maps, worth treating as such in verification.
@@
 
-On the x86-64 architecture, the input to the \textsc{MMU}'s address translation is a sparse hierarchical set of tables calle d \emph{page tables} (referring to pages of memory).
+\todo[inline,color=cyan]{Ismail, is this text copy-pasted from the VMM paper? 
+It has phrasing like ``In this paper...''
+That's actually a major issue, self-plagiarism of another peer-reviewed piece of work can get
+you into a lot of trouble.
+}
+On the x86-64 architecture, the input to the \textsc{MMU}'s address translation is a sparse hierarchical set of tables called \emph{page tables} (referring to pages of memory).
 As Figure \ref{fig:pagetables} shows, address translation proceeds by repeatedly taking designated slices of the virtual address and indexing into a table.
@@
 
-\begin{figure}
+\begin{figure}[t]
     \includegraphics[width=0.75\columnwidth]{cow_index.png}
@@
 \paragraph{Different Models in Ensuring Disk Consistency of Filesystems}
+\todo[inline,color=cyan]{This looks like it's verbatim copy-pasted from the ModFS draft; if this exact text
+it used in this publication, it means we'd need to completely rewrite the relevant text in that draft.}
 With their intricate semantics and critical place in modern systems, formal verification of functional correctness for storage systems (including filesystems and other kinds of storage) have drawn much attention~\cite{ntzik-gardner-15,ntzik2018concurrent,chen-crash-16}\todo{lots more cites}.
@@
 On the other hand, a typical filesystem can follow a certain policy to commit the in-memory file pages to the disk.
-\subsection{Conclusion \& Contributions}
-\paragraph{Conclusion} Low-level systems exhibit certain patterns in their designs, especially when interacting with computing resources.
-Exploiting certain patterns while designing software has been an important field of study.
-In this regard, we think that certain properties of modalities enable us to understand and do the verification challenges that show certain patterns.
-
-\paragraph{Contributions} We argue how modal abstractions can be used to identify and abstract system verification challenges.
-We justify our perspective by discussing prior systems that have successfully used modalities for system verification, arguing that they fit into the verification design pattern we articulate, and explaining how this approach might apply to other systems' verification challenges.
-\begin{itemize}
-\item \textbf{Identifying System Verification Challenges}: We start with identifying common patterns in system verification: \emph{virtualization, sharing, and translation}.
-\item \textbf{Introducing the Concept of \emph{Resource}}: Then we discuss the concept of \emph{resource} which has already been an essential concept in the design of systems.
-Inspired by the concept of resource in the systems, we define what a resource and its context are in the modal abstractions.
-\item \textbf{Introducing the Concept of \emph{Nominals}}:
-\todo[inline]{Colin says: I'd reframe the way nominals are introduced. This discussion is good,
-but is a "systems-first" explanation. At least some of this audience will already know
-what classic nominals are, so need to be shown naming specific resources as an application
-of what they already know. (And some readers will need a very short intro/refresher.)
-}
-Nominalization enables identifying a resource in a context.
-For example, a transaction is a context of resources of in-memory updated disk blocks.
-The transaction identifier is used to associate a transaction with a disk-block to be persisted so that, in case of a crash while persisting updated disk-blocks, the filesystem can rollback the \emph{already persisted} disk-blocks of the transaction, and reach to the previous consistent disk state.
-To be able to do so, both the updated in-memory disk blocks and the transaction must refer to the transaction identifier -- \emph{strong nominalization}.
-In another example, virtual memory references (resources) in an address space (resource context), which can be uniquely identified with a root address (the nominal) of its page-table tree, are \emph{agnostic} to the address space that they are in.
-However, they can only be accessed (be valid) in the address space to which they are agnostic.
-However, then an address space switch happens, the virtual memory references of the previous address space must be made inaccessible.
-To be able to do so, although virtual memory references do not hold any piece of information related to their address space, we still have to associate them.
-We call this kind of unilateral nominalization -- \emph{weak nominalization}.
-\item \textbf{Taxonomy of the Current Modal Approaches in System Verification}: Based on these concepts defined, we summarize contemporary verification efforts using modal abstractions.
-We choose them from different domains, for example, weak memory verification, storage persistence ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2024modalabstractionsvirtualizingmemory,amalreal2024} because we would like to justify that our \emph{definitions} are not domain dependent.
-\end{itemize}
+
 %\section{Resources in Systems Software}
@@
 
-To explain our ideas in the general systems understanding, we briefly recap some of the background and themes our ideas build on again,casting them in a certain way to bring out the relevance of our philosophy.
+To explain our ideas in the general systems understanding, we briefly recap some of the background and themes our ideas build on again, casting them in a certain way to bring out the relevance of our philosophy.
 
@@
 \label{sec:modal_in_iris}
+\todo[inline,color=cyan]{this also seems taken from the VMM work}
 Classically modal operators are given direct semantics in terms of an underlying model, which for \textsf{Iris} would be a collection of \textsc{CMRA}s; this is the case for \textsf{Iris}'s one built-in modality for step-indexing~\cite{krebbers2017essence}.
@@
 \label{sec:definitions}
-\todo[inline]{Colin says: I dropped the strong/weak nominalization column for space and readability;
-it can be explicit in the text, or maybe we can mark the system names with daggers or similar to indicate.}
-\begin{table*}[h]
+\begin{table*}[t]
 \centering

### 2025-06-23 paper.tex
started working on comments

@@
  Although they differ in the functionality they offer, low-level systems exhibit certain patterns of design and utilization of computing resources.
-
+\begin{comment}
   In this paper, we argue the position that \emph{modalities}, in the sense of modal logic, 
@@
   and explaining how this approach might apply to other systems verification challenges.
-
+\end{comment}
   \todo[inline]{Colin says: how about this framing instead of the second paragraph above:}
-  \begin{comment}
+
   In this paper we examine how modalities have emerged as a common structure in formal verification
@@
   of filesystems code.
-  \end{comment}
   \todo[inline,color=cyan]{Ismail, why did you comment out my suggestion and keep the original?}
+  \todo[inline,color=pink]{Sorry, I thought I did the other way}
 \end{abstract}
@@
 However, most prior work on hybrid logics works in classical logics rather than a substructural setting like separation logic.
-We adapt these ideas to assertions of the form $@[l]\varphi$ asserts that $\varphi$ is true in a data structure with \emph{logical} state $\sigma$ \emph{at location} $l$.\todo{seems out of place?}
+%We adapt these ideas to assertions of the form $@[l]\varphi$ asserts that $\varphi$ is true in a data structure with \emph{logical} state $\sigma$ \emph{at location} $l$.
+\todo{seems out of place?} \todo[inline,color=pink]{addressed}
 In practice we must design variants of this for each kind of structure we wish to apply this to, e.g. tree nodes, free list, or snapshots.
@@
 you into a lot of trouble.
-}
+}\todo{inline,color=pink}{I am well-aware what counts as as (including self-plagiarism). These were annotated with todos including cautions for changing the text.}
 On the x86-64 architecture, the input to the \textsc{MMU}'s address translation is a sparse hierarchical set of tables called \emph{page tables} (referring to pages of memory).

### 2025-06-23 paper.tex
+

@@
 \todo{change the text}
-Memory addresses seen by programs running on modern computers are \emph{virtualized}: the address observed by a running program generally will not correspond directly to the physical location in memory, and may not even correspond to a physical location that \emph{exists} in the machine.
-Instead, these \emph{virtual} addresses are translated to \emph{physical} addresses that correspond directly to locations in RAM.
-On most modern architectures, this translation is performed through cooperation of the hardware and OS kernel: while executing an instruction that dereferences a (virtual) address, the CPU's \emph{memory management unit} (\textsc{MMU}) implements an architecture-specific process of \emph{address translation}, resulting in a physical address used to access the cache\footnote{Technically, for performance reasons most caches are indexed with parts of the virtual address, but tagged with the physical data addresses, so cache lookups and address translations can proceed in parallel.} and/or memory-bus.
-
+\todo[inline,color=pink]{Based on the note I took above, I changed the text}
+A typical general purpose computers make the memory resources in RAM as \emph{virtually} available.
+In other words, a program asking for a memory unit from an operating system kernel is served with a memory address that is \emph{virtualized} such that it may be \emph{mapped to} a physical RAM location or not.
+As our informal definition suggests, the common technique for \emph{virtualization} of memory locations in OS kernels relies \emph{translating}(\emph{mapping}) virtual addresses into physical ones.
+The conceptual address-translation map is implemented with processor's(CPU) page-table trees as showin in Figure \ref{fig:pagetables2.png}.
 \begin{figure}

### 2025-06-23 paper.tex
+

@@
 As our informal definition suggests, the common technique for \emph{virtualization} of memory locations in OS kernels relies \emph{translating}(\emph{mapping}) virtual addresses into physical ones.
-The conceptual address-translation map is implemented with processor's(CPU) page-table trees as showin in Figure \ref{fig:pagetables2.png}.
+The conceptual address-translation map is implemented with processor's(CPU) page-table trees as shown in Figure \ref{fig:pagetables2.png}.
+In this Figure, we see a typical 4 levels of page tables, a virtual address on the top indexing into the different levels of the page-table tree.
+Fabrication of virtual addresses out of limited physical addresses is provided with entries in the page-table tree where each points to a physical address aligned to 4KB (4096 byte) boundaries.
+An address translation requires traversal of a series of tables starting from level 4, and it is conventionally referred as a \emph{page-table walk}.
+The traversal ends in the level 1 page table with the final lookup to the actual page of physical memory holding the requested data, and the low-order 12 bits being used to index into this page.
 \begin{figure}
@@
 }\todo{inline,color=pink}{I am well-aware what counts as as (including self-plagiarism). These were annotated with todos including cautions for changing the text.}
-On the x86-64 architecture, the input to the \textsc{MMU}'s address translation is a sparse hierarchical set of tables called \emph{page tables} (referring to pages of memory).
-As Figure \ref{fig:pagetables} shows, address translation proceeds by repeatedly taking designated slices of the virtual address and indexing into a table.
-On x86-64, standard configurations use 4 levels of page tables, labelled levels 4 through 1, with lookups in the level 1 page table resulting in the actual page of physical memory holding the requested data, and the low-order 12 bits being used to index into this page.\footnote{Technically levels 1--3 have designated names originating when there were fewer levels.
-We omit those names to avoid confusion, and for greater consistency with the recent addition of a 5th level on the most recent processors, where levels 4 and 5 are simply known by their numbers.}  Because address translation traverses a series of tables, the translation process or algorithm is sometimes referred to as a \emph{page-table walk}.
-While Figure \ref{fig:pagetables} and most of our details are specific to the x86-64 architecture, ARMv8 (a.k.a.\ \texttt{aarch64}), RISC-V, and PowerPC use similar hierarchical page tables for address translation.
-
-The entries of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set), and a range of additional bits which can be used to control caching, write-through, and more.
-This paper will only consider the present bit (0), the read-write bit (1), and the accessed bit (5), which is set to 1 by the hardware every time an entry is used, allowing the OS to determine which pieces of memory are used or unused.
-
+\paragraph{Virtual Memory Managers with Multiple Address-Spaces}
+\label{sec:backgroundonvmm}
+\todo{change the text}
 The page tables mapping virtual addresses to physical addresses are managed by the OS.
@@
 
-If an instruction is executed that accesses a virtual address that either has no mapping, or does not have a mapping permitting the kind of access that was performed (e.g., the instruction was a memory write, but the relevant address range was marked read-only in the relevant page table entry), the hardware triggers a \emph{page fault}, transferring control to a \emph{page fault handler} registered with the hardware by the OS, allowing it to take corrective action (described in the next section).
-
-\paragraph{Virtual Memory Managers with Multiple Address-Spaces}
-\label{sec:backgroundonvmm}
-\todo{change the text}
 Most OS kernels have a component called a \emph{Virtual Memory Manager} (VMM)\footnote{Not to be confused with Virtual Machine Monitor.

### 2025-06-23 paper.tex
+

@@
 As our informal definition suggests, the common technique for \emph{virtualization} of memory locations in OS kernels relies \emph{translating}(\emph{mapping}) virtual addresses into physical ones.
-The conceptual address-translation map is implemented with processor's(CPU) page-table trees as shown in Figure \ref{fig:pagetables2.png}.
+The conceptual address-translation map is implemented with processor's(CPU) page-table trees as shown in Figure \ref{fig:pagetables}.
 In this Figure, we see a typical 4 levels of page tables, a virtual address on the top indexing into the different levels of the page-table tree.
@@
 \todo{change the text}
-The page tables mapping virtual addresses to physical addresses are managed by the OS.
-Typically each process has its on virtul-to-physical mappings represented by its own page table, which the OS registers with the CPU by storing it in a specific register (\texttt{cr3}) as part of switching to a new process.
+An intuition for the abstraction that \emph{contains} a page-table tree is called an \emph{address-space} of a process whose virtual addresses to physical addresses.
+A typical OS manages multiple processes at once .
+Each process is identified by an \emph{unique root address} of its page-table tree.
+As part of switching to a new process, i.e., loading the address-space mappings as the current view of the memory, the CPU stores the unique root address of the page-table tree in a specific register (\texttt{cr3}) as shown in Figure \ref{fig:pagetables}.
 Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section) is how the OS ensures memory isolation between processes.
-
-Most OS kernels have a component called a \emph{Virtual Memory Manager} (VMM)\footnote{Not to be confused with Virtual Machine Monitor.
-In this paper we focus on non-virtualized scenarios, but hardware virtualization extensions for both x86-64 and ARM make use of an additional set of page tables translating what a \emph{guest} considers to be its (virutalized) physical memory to actual physical memory.
-Our contributions should offer value in this scenario as well.}
-which is responsible for setting up page table mappings and for taking action when a page fault occurs.
-Often, a page fault indicates a bug in a program --- for example, a null pointer dereference crashes a program not because the hardware designates it an error, but because most OSes refuse to map the first 4K worth of virtual addresses, so dereferencing \texttt{NULL} results in a page fault.
-In such cases, the OS will terminate the program.
-
-In other cases the OS uses page faults to implement specialized functionality and optimizations.
-A key example of this is (confusingly) called \emph{paging}: saving room in physical RAM by waiting until memory addresses for code are accessed before copying them to RAM (saving process startup time); or copying memory that has not recently been accessed to disk (i.e., in a \emph{swap} file or partition) and marking those page table entries invalid so a future page fault will give the OS the chance to copy the relevant data back into memory when the program tries to access it again.
-
-The key pieces of functionality a VMM must implement itself, or have available in another OS component (depending on design), are
-adding a new page mapping (whether the mapped page contains zeros, file data, or swap data), and removing an existing page mapping.
-While this initially sounds like relatively modest functionality whose implementation may be complicated by hardware subtleties, a number of more serious complexities must be handled.
-%
-%For one, virtual addresses may \emph{alias} when more than one is mapped to the same physical memory.
-This can occur when a process requests it (e.g., via \texttt{mmap} or \texttt{shm\_open} on UNIX-style kernels).
-It also occurs automatically in most general-purpose kernels: while for normal programs the OS uses the MMU to isolate and abstract physical memory, the kernel's own functionality is often easier to implement if the kernel can directly access memory given its physical address, for example when interacting with hardware devices.
-For this reason, most kernels contain an \emph{identity map} of all physical memory in a machine.
-In some cases this is a literal identity map, where for every virtual address $p$, interpreting $p$ as a virtual address instead will map back to $p$ as a physical address.
-In other cases there is an offset involved, where to access physical address $p$ the kernel can access virtual address $p+\mathit{offset}$.
-This identity map saves the kernel the performance cost of having to continually add and remove transient mappings simply to briefly access a specific physical memory region.
-But this coexists with additional virtual addresses (those used by most kernel code and data structures), in a form of virtual address aliasing.
-Note that virtual address aliasing is both prevalent in general-purpose kernels, and violates the core assumption of most memory models assumed by verified compilers like CompCert, which assume no virtual address aliasing.
-%
-%Another complicating factor is that portions of page tables can be shared, i.e., trees of page tables can overlap.
-One common way this used to happen was for one subtree of page tables to map the kernel if linked in appropriately to a higher-level page table, and then to reuse that kernel mapping portion in every process's page tables.
-This was mostly phased out in the wake of Spectre
-%\mytodo{Colin finish}
-
-The OS scheduler, since it is responsible for dealing with multiple address spaces, must keep track of which virtual addresses are valid (and in what way) in which address spaces, where some virtual addresses are valid in only a single address space (e.g., a code address for a particular usermode process), while others are valid in all address spaces (e.g., kernel data structure pointers).
-The VMM must maintain some of these assumptions on behalf of the rest of the kernel, for example by guaranteeing that a certain range of virtual addresses (corresponding to the kernel's code and data) are valid in every address space.
-
+Virtual memory managers are the modules in OS kernels handling this kind of tasks.
+The treat address-spaces as containers of virtual addresses so that they can bookkeep which which virtual addresses are valid (and in what way) in which address spaces.
 \paragraph{Virtualization of Disk-Blocks in Filesystems}

### 2025-06-23 paper.tex
+

@@
 }\todo{inline,color=pink}{I am well-aware what counts as as (including self-plagiarism). These were annotated with todos including cautions for changing the text.}
-\paragraph{Virtual Memory Managers with Multiple Address-Spaces}
+\paragraph{Virtual Memory Managers(VMM) with Multiple Address-Spaces}
 \label{sec:backgroundonvmm}
@@
 \paragraph{Virtualization of Disk-Blocks in Filesystems}
+Another important computer resource computer resource that is \emph{virtualized} for access is disk blocks.
+Filesystems are computer programs that serve to other programs for their disk resource need.
+Like VMMs using page-table trees to fabricate more memory addresses than what actually exists, filesystems uses indexing maps for fabricating more \emph{virtual disk-page} addresses which are conventionally called \emph{disk-pages} to be distinguished from physical \emph{disk-block} addresses.
 \textsf{ModFS} is a copy-on-write (\textsf{CoW}) filesystem, which as described in the introduction performs updates by copying modified data to \emph{new and unused} locations on disk rather than updating in place, ensuring that in the absence of physical storage corruption, there is \emph{always} a fully consistent filesystem on disk, removing the need for recovery from journals in the case of a crash, which can be very expensive for large volumes.

### 2025-06-23 paper.tex
+

@@
 Like VMMs using page-table trees to fabricate more memory addresses than what actually exists, filesystems uses indexing maps for fabricating more \emph{virtual disk-page} addresses which are conventionally called \emph{disk-pages} to be distinguished from physical \emph{disk-block} addresses.
+
+Filesystems do not only handle the address-translation from a disk-page to a disk-block but also need to handle the different \emph{modes} of computing in which addressing disk resources occur at different consistency modes: a disk block persisted on the disk and a disk-page which may not be persisted on the disk yet.
+To handle the consistency of the disk resources in the existence of different modes of execution, filesystems employ different policies when indexing the disk resource.
+Which data-structure is going to be used to implement \emph{mappi
 \textsf{ModFS} is a copy-on-write (\textsf{CoW}) filesystem, which as described in the introduction performs updates by copying modified data to \emph{new and unused} locations on disk rather than updating in place, ensuring that in the absence of physical storage corruption, there is \emph{always} a fully consistent filesystem on disk, removing the need for recovery from journals in the case of a crash, which can be very expensive for large volumes.

### 2025-06-23 paper.tex
+

@@
 To handle the consistency of the disk resources in the existence of different modes of execution, filesystems employ different policies when indexing the disk resource.
-Which data-structure is going to be used to implement \emph{mappi
-\textsf{ModFS} is a copy-on-write (\textsf{CoW}) filesystem, which as described in the introduction performs updates by copying modified data to \emph{new and unused} locations on disk rather than updating in place, ensuring that in the absence of physical storage corruption, there is \emph{always} a fully consistent filesystem on disk, removing the need for recovery from journals in the case of a crash, which can be very expensive for large volumes.
+Which data-structure is going to be used to implement address-traslation from disk-pages to disk-blocks?
+What is going to be the \emph{update-policy} to the indexing map implementing address-translation which defines the \emph{consistency-policy} for the disk resources?
+Different filesystem implementations give different answers to these questions.
+One example answer to updating policy is copy-on-write (\textsf{CoW}).
+A (\textsf{CoW}) filesystem performs updates by copying modified data to \emph{new and unused} locations on disk rather than updating \emph{in-place}, ensuring that in the absence of physical storage corruption, there is \emph{always} a fully consistent filesystem on disk, removing the need for recovery from journals in the case of a crash, which can be very expensive for large volumes.
 %\textsf{ModFS} is implemented in \textsc{Go}~\cite{donovan2015go}, using the FUSE~\cite{fuse} interface to run a filesystem in usermode.
@@
 files and directories) as a nested trees, and to map a logical page to a physical page, the file system first locates the root of the nested index-tree belonging to \textsf{"foo"}.
-Semantically, in \textsf{ModFS} (as in other \textsf{CoW} filesystems~\cite{young1992episode,bonwick2003zettabyte,rodeh2013btrfs,Rosenblum1992LFS,hitz1994file}) any update to file data or metadata is realized on  unused disk locations by treating the tree as a persistent data structure.
-As an example, Figure \ref{fig:cow_index} shows that an update to the file $\textsf{"foo"}$ at $\textsf{"offset"}$ ends up with locating the related disk chunk (a set of blocks) in the gray region, copying it to newly allocated disk chunks (dotted chunks on the right side of the bold long-dash-point line of the middle layer withing dotted section), and performing the update on these newly-allocated disk blocks.
+Semantically, in \textsf{ModFS}~\cite{} (as in other \textsf{CoW} filesystems~\cite{young1992episode,bonwick2003zettabyte,rodeh2013btrfs,Rosenblum1992LFS,hitz1994file}) any update to file data or metadata is realized on  unused disk locations by treating the tree as a persistent data structure.
+As an example, Figure \ref{fig:cowindex} shows that an update to the file $\textsf{"foo"}$ at $\textsf{"offset"}$ ends up with locating the related disk chunk(multiple disk-pages) in the gray region, copying it to newly allocated disk chunks (dotted chunks on the right side of the bold long-dash-point line of the middle layer withing dotted section), and performing the update on these newly-allocated disk blocks.
 This requires further updates to those metadata blocks which previously referred to the old block locations to instead reference the newly-allocated blocks.
@@
 
-After this process, traversing the filesystem from either the old or new root yields consistent views of the filesystem at different points in time, each commonly referred to as a \emph{snapshot}.
+Like VMMs navigating between different address-spaces, a \textsf{CoW} filesystem switches from either the old to the new root yields consistent views of the filesystem at different points in time, each commonly referred to as a \emph{snapshot}.
 These snapshots overlap on the unmodified portions of the filesystem (the left-most grid section), with the reclaimed the chunks (the middle gray section) becoming available for reuse once the new snapshot becomes the official current state of the filesystem.\footnote{Many of these filesystems permit keeping archived snapshots of the filesystem state, e.g., a snapshot of the system every hour up to a certain number of days, to support user data recovery.
 In these cases only disk regions that are not still in use by an archived snapshot are reclaimed.}.
-This permits the filesystem to atomically switch between consistent on-disk states, by updating the oldest so-called \emph{super} chunk at a fixed location on disk.\footnote{Filesystems must take care to ensure that possibly-flaky hardware truly does perform an atomic update.} Super chunks, as we see in Figure \ref{fig:cow_index} as ($\textsf{Super}_0$ and $\textsf{Super}_1$), are the first two blocks of the disk dedicated to keep the state of the file system which is composed of some meta-data on the filesystem, and most importantly to the root chunk of the file system index-tree.
+This permits the filesystem to atomically switch between consistent on-disk states, by updating the oldest so-called \emph{super} chunk at a fixed location on disk.\footnote{Filesystems must take care to ensure that possibly-flaky hardware truly does perform an atomic update.} Super chunks, as we see in Figure \ref{fig:cowindex} as ($\textsf{Super}_0$ and $\textsf{Super}_1$), are the first two blocks of the disk dedicated to keep the state of the file system which is composed of some meta-data on the filesystem, and most importantly to the root chunk of the file system index-tree.
 Therefore, in case of a crash during the update to the \textsf{file "foo"}, we can reboot the file system from the previous persisted root, \textsf{root chunk}, via reading the $\textsf{Super}_0$.

### 2025-06-23 paper.tex
+

@@
 Like VMMs navigating between different address-spaces, a \textsf{CoW} filesystem switches from either the old to the new root yields consistent views of the filesystem at different points in time, each commonly referred to as a \emph{snapshot}.
-These snapshots overlap on the unmodified portions of the filesystem (the left-most grid section), with the reclaimed the chunks (the middle gray section) becoming available for reuse once the new snapshot becomes the official current state of the filesystem.\footnote{Many of these filesystems permit keeping archived snapshots of the filesystem state, e.g., a snapshot of the system every hour up to a certain number of days, to support user data recovery.
-In these cases only disk regions that are not still in use by an archived snapshot are reclaimed.}.
-This permits the filesystem to atomically switch between consistent on-disk states, by updating the oldest so-called \emph{super} chunk at a fixed location on disk.\footnote{Filesystems must take care to ensure that possibly-flaky hardware truly does perform an atomic update.} Super chunks, as we see in Figure \ref{fig:cowindex} as ($\textsf{Super}_0$ and $\textsf{Super}_1$), are the first two blocks of the disk dedicated to keep the state of the file system which is composed of some meta-data on the filesystem, and most importantly to the root chunk of the file system index-tree.
+Like VMMs
+These snapshots overlap on the unmodified portions of the filesystem (the left-most grid section), with the reclaimed the chunks (the middle gray section) becoming available for reuse once the new snapshot becomes the official current state of the filesystem.
+This permits the filesystem to atomically switch between consistent on-disk states, by updating the oldest so-called \emph{super} chunk at a fixed location on disk.
+Super chunks, as we see in Figure \ref{fig:cowindex} as ($\textsf{Super}_0$ and $\textsf{Super}_1$), are the first two blocks of the disk dedicated to keep the state of the file system which is composed of some meta-data on the filesystem, and most importantly to the root chunk of the file system index-tree.
 Therefore, in case of a crash during the update to the \textsf{file "foo"}, we can reboot the file system from the previous persisted root, \textsf{root chunk}, via reading the $\textsf{Super}_0$.

### 2025-06-23 paper.tex
+

@@
 Like VMMs navigating between different address-spaces, a \textsf{CoW} filesystem switches from either the old to the new root yields consistent views of the filesystem at different points in time, each commonly referred to as a \emph{snapshot}.
-Like VMMs
-These snapshots overlap on the unmodified portions of the filesystem (the left-most grid section), with the reclaimed the chunks (the middle gray section) becoming available for reuse once the new snapshot becomes the official current state of the filesystem.
-This permits the filesystem to atomically switch between consistent on-disk states, by updating the oldest so-called \emph{super} chunk at a fixed location on disk.
+Like VMMs loading a process's memory mappings to be the current view of the memory by setting the control register to show the unique root address of the page-table tree of the process, the filesystem to atomically switch between consistent on-disk states, by updating the oldest so-called \emph{super} chunk at a fixed location on disk.
 Super chunks, as we see in Figure \ref{fig:cowindex} as ($\textsf{Super}_0$ and $\textsf{Super}_1$), are the first two blocks of the disk dedicated to keep the state of the file system which is composed of some meta-data on the filesystem, and most importantly to the root chunk of the file system index-tree.
-Therefore, in case of a crash during the update to the \textsf{file "foo"}, we can reboot the file system from the previous persisted root, \textsf{root chunk}, via reading the $\textsf{Super}_0$.
-This means the whole disk consistency checks are unnecessary when recovering from a crash~\cite{bonwick2003zettabyte,hitz1994file}; recovery is immediate, as opposed to requiring hours or days~\cite{henson2006reducing,agrawal2009speedy} (or longer~\cite{henson2006chunkfs}) for recovering a large volume’s consistency from logs.
-An important additional subtlety with this kind of filesystem is \emph{when} blocks in the reclaimed region can actually be reused: only \emph{after} the snapshot that made those blocks ``free'' has been completely written to disk.
-Allowing reuse before this could result in in-place updates: part of the new snapshot could attempt to store data in space used by the prior snapshot, allowing inconsistent on-disk structures.
 \begin{figure}
@@
   it used in this publication, it means we'd need to completely rewrite the relevant text in that draft.}
+Therefore, in case of a crash during the update to the \textsf{file "foo"}, we can reboot the file system from the previous persisted root, \textsf{root chunk}, via reading the $\textsf{Super}_0$.
+This means the whole disk consistency checks are unnecessary when recovering from a crash~\cite{bonwick2003zettabyte,hitz1994file}; recovery is immediate, as opposed to requiring hours or days~\cite{henson2006reducing,agrawal2009speedy} (or longer~\cite{henson2006chunkfs}) for recovering a large volume’s consistency from logs.
+An important additional subtlety with this kind of filesystem is \emph{when} blocks in the reclaimed region can actually be reused: only \emph{after} the snapshot that made those blocks ``free'' has been completely written to disk.
+Allowing reuse before this could result in in-place updates: part of the new snapshot could attempt to store data in space used by the prior snapshot, allowing inconsistent on-disk structures.
+
 With their intricate semantics and critical place in modern systems, formal verification of functional correctness for storage systems (including filesystems and other kinds of storage) have drawn much attention~\cite{ntzik-gardner-15,ntzik2018concurrent,chen-crash-16}\todo{lots more cites}.

### 2025-06-23 paper.tex
+

@@
 \todo[inline,color=cyan]{this also seems taken from the VMM work}
+\todo[inline,color=pink]{this text has never been part of any text that is either submitted or made publicly available.}
 Classically modal operators are given direct semantics in terms of an underlying model, which for \textsf{Iris} would be a collection of \textsc{CMRA}s; this is the case for \textsf{Iris}'s one built-in modality for step-indexing~\cite{krebbers2017essence}.

### 2025-06-23 paper.tex
+

@@
 The treat address-spaces as containers of virtual addresses so that they can bookkeep which which virtual addresses are valid (and in what way) in which address spaces.
-\paragraph{Virtualization of Disk-Blocks in Filesystems}
+\paragraph{Virtualized Resources in Filesystems}
 Another important computer resource computer resource that is \emph{virtualized} for access is disk blocks.
@@
 Like VMMs using page-table trees to fabricate more memory addresses than what actually exists, filesystems uses indexing maps for fabricating more \emph{virtual disk-page} addresses which are conventionally called \emph{disk-pages} to be distinguished from physical \emph{disk-block} addresses.
-
+\paragraph{Overview of Policies in Filesystems}
 Filesystems do not only handle the address-translation from a disk-page to a disk-block but also need to handle the different \emph{modes} of computing in which addressing disk resources occur at different consistency modes: a disk block persisted on the disk and a disk-page which may not be persisted on the disk yet.
@@
 files and directories) as a nested trees, and to map a logical page to a physical page, the file system first locates the root of the nested index-tree belonging to \textsf{"foo"}.
+\paragraph{An Example to a Policy for Updating Disk Index Tree}
 Semantically, in \textsf{ModFS}~\cite{} (as in other \textsf{CoW} filesystems~\cite{young1992episode,bonwick2003zettabyte,rodeh2013btrfs,Rosenblum1992LFS,hitz1994file}) any update to file data or metadata is realized on  unused disk locations by treating the tree as a persistent data structure.
@@
 \end{figure}
-
+\paragraph{Examples to Different Policies on Navigating on Consistent Views of Disk}
 Like VMMs navigating between different address-spaces, a \textsf{CoW} filesystem switches from either the old to the new root yields consistent views of the filesystem at different points in time, each commonly referred to as a \emph{snapshot}.
+A \textsf{CoW} filesystems treat the snapshot abstraction as VMMs treat the address-space abstraction, container of the relevant resources.
 Like VMMs loading a process's memory mappings to be the current view of the memory by setting the control register to show the unique root address of the page-table tree of the process, the filesystem to atomically switch between consistent on-disk states, by updating the oldest so-called \emph{super} chunk at a fixed location on disk.
-Super chunks, as we see in Figure \ref{fig:cowindex} as ($\textsf{Super}_0$ and $\textsf{Super}_1$), are the first two blocks of the disk dedicated to keep the state of the file system which is composed of some meta-data on the filesystem, and most importantly to the root chunk of the file system index-tree.
+Super chunks, as we see in Figure \ref{fig:cowindex} as ($\textsf{Super}_0$ and $\textsf{Super}_1$), are the first two blocks of the disk dedicated to keep the state of the file system which is composed of some meta-data on the filesystem, and most importantly to the root chunk of the filesystem index tree.
 \begin{figure}
@@
 \end{figure}
-
-\paragraph{Different Models in Ensuring Disk Consistency of Filesystems}
 \todo[inline,color=cyan]{This looks like it's verbatim copy-pasted from the ModFS draft; if this exact text
@@
 
-With their intricate semantics and critical place in modern systems, formal verification of functional correctness for storage systems (including filesystems and other kinds of storage) have drawn much attention~\cite{ntzik-gardner-15,ntzik2018concurrent,chen-crash-16}\todo{lots more cites}.
-Unlike our \textsc{CoW} filesystem \textsf{ModFs} which allocates new chunks and rewrites all the nodes from leaf to root on the write-amplification path, some filesystems allows in-place updates to handle the heavy workloads more efficiently in principle.
-Unlike \textsc{CoW} filesystems, filesystems with in-place updates need to implement some form of \emph{journalling} to ensure that a crash partway through an in-place update can be rolled back (using an undo journal that records the original contents before modification as abstarcted in Figure \ref{fig:journal}) or completed (using a write-ahead journal that first writes new data in unused space before \emph{also} updating in place).\footnote{\textsf{CoW} filesystems can be viewed as a twist on write-ahead journals that uses the journal as the actual data once written to a fresh location, thereby reclaiming the original location.
-This results in a performance and complexity trade-off: journalling writes twice for every updated block and requires recovery logic to roll back or re-execute the log, while \textsf{CoW} techniques write each update only once and require slightly more intricate allocation strategies to compact fragmentation.}
-
-Essentially, successfully writing the new superblock for a new snapshot is analagous to completing a valid journal entry, except it \emph{also} has the effect of updating the live filesystem data due to making the ``journal entries'' valid filesystem data and metadata blocks (an idea that goes back to early work on log-structured filesystems~\cite{Rosenblum1992LFS}).
-In fact, we can think of the atomicity of the recovery from the crash state conceptually equal to the single block disk write persisting the change from $\textsf{Super}_0$ to $\textsf{Super}_1$ in Figure \ref{fig:cow_index}, therefore, considering \textsc{CoW} semantics, we do not need to consider the specification and proof of crash condition separately; instead a global never-violated invariant (not even temporarily broken) of the filesystem ensures the on-disk state is always consistent.
+On the other hand, filesystems with in-place updates need to implement some form of \emph{journalling} to ensure that a crash partway through an in-place update can be rolled back using an undo journal that records the original contents before modification as abstacted in Figure \ref{fig:journal} or completed using a write-ahead journal that first writes new data in unused space before \emph{also} updating in place.
+In other words, unlike our ability to distinguish all resources (disk-pages and diskblocks) either being consistent under a snapshot or not in filesystems with \textsf{CoW}, with in-place updates, journalling mechanism does not allow such a claim.
+In other words, as we see in Figure \ref{fig:journal}, the middle level disk-layout is composed of the updates to the disk
+
+Unlike Essentially, successfully writing the new superblock for a new snapshot is analagous to completing a valid journal entry, except it \emph{also} has the effect of updating the live filesystem data due to making the ``journal entries'' valid filesystem data and metadata blocks (an idea that goes back to early work on log-structured filesystems~\cite{Rosenblum1992LFS}).
+In fact, we can think of the atomicity of the recovery from the crash state conceptually equal to the single block disk write persisting the change from $\textsf{Super}_0$ to $\textsf{Super}_1$ in Figure \ref{fig:cowindex}, therefore, considering \textsf{CoW} semantics, we do not need to consider the specification and proof of crash condition separately; instead a global never-violated invariant (not even temporarily broken) of the filesystem ensures the on-disk state is always consistent.
 \paragraph{Memory Reclamation in Client Code} In a typical configuration, allocating, deallocating and referencing memory locations for a client code leave the memory layout in a form of DAG (directed-asyclic-graph) with a unique root which resembles the first stack pointer referencing the allocated memory.

### 2025-06-23 paper.tex
+

@@
 you into a lot of trouble.
-}\todo{inline,color=pink}{I am well-aware what counts as as (including self-plagiarism). These were annotated with todos including cautions for changing the text.}
+}\todo[inline,color=pink]{I am well-aware what counts as plagiarism (including self-plagiarism). These were annotated with todos including cautions for changing the text.}
 \paragraph{Virtual Memory Managers(VMM) with Multiple Address-Spaces}
@@
 On the other hand, filesystems with in-place updates need to implement some form of \emph{journalling} to ensure that a crash partway through an in-place update can be rolled back using an undo journal that records the original contents before modification as abstacted in Figure \ref{fig:journal} or completed using a write-ahead journal that first writes new data in unused space before \emph{also} updating in place.
-In other words, unlike our ability to distinguish all resources (disk-pages and diskblocks) either being consistent under a snapshot or not in filesystems with \textsf{CoW}, with in-place updates, journalling mechanism does not allow such a claim.
-In other words, as we see in Figure \ref{fig:journal}, the middle level disk-layout is composed of the updates to the disk
-
-Unlike Essentially, successfully writing the new superblock for a new snapshot is analagous to completing a valid journal entry, except it \emph{also} has the effect of updating the live filesystem data due to making the ``journal entries'' valid filesystem data and metadata blocks (an idea that goes back to early work on log-structured filesystems~\cite{Rosenblum1992LFS}).
-In fact, we can think of the atomicity of the recovery from the crash state conceptually equal to the single block disk write persisting the change from $\textsf{Super}_0$ to $\textsf{Super}_1$ in Figure \ref{fig:cowindex}, therefore, considering \textsf{CoW} semantics, we do not need to consider the specification and proof of crash condition separately; instead a global never-violated invariant (not even temporarily broken) of the filesystem ensures the on-disk state is always consistent.
+In other words, unlike our ability to distinguish all resources (disk-pages and dis-kblocks) either being consistent under a snapshot or not in filesystems with \textsf{CoW}, with in-place updates, journalling mechanism does not allow such a claim.
+In other words, as we see in Figure \ref{fig:journal}, unlike \textsf{CoW}'s uniform persistency claim over disk resources under a snapshot as a resource container, elaborative bookkeeping mechanism of journalling leaves the filesystem treat disk resources in one of the resource groups in non-uniform fashion: disk-pages loaded into memory, updated disk-page not logged yet, updated and being logged or already logged disk-pages.
+Being able to treat the resources in a uniform fashion determines wwhether we need to consider the specification and proof of crash condition separately; instead a global never-violated invariant (not even temporarily broken) of the filesystem ensures the on-disk state is always consistent.
+
 \paragraph{Memory Reclamation in Client Code} In a typical configuration, allocating, deallocating and referencing memory locations for a client code leave the memory layout in a form of DAG (directed-asyclic-graph) with a unique root which resembles the first stack pointer referencing the allocated memory.
-To ease picturing the memory layout after a set of requests from the client code (for allocation, referencing etc.), we can think of the unique roots of DAGs as \emph{super} block addresses -- this time multiple ones, not only two -- in Figure \ref{fig:cow_index}, from which a path is created for a stack pointer reference to the allocated memory.
+To ease picturing the memory layout after a set of requests from the client code (for allocation, referencing etc.), we can think of the unique roots of DAGs as \emph{super} block addresses -- this time multiple ones, not only two -- in Figure \ref{fig:cowindex}, from which a path is created for a stack pointer reference to the allocated memory.
 To reclaim the allocated memory location that is no longer being accessed, systems uses a well-known method called \emph{reference-counting} in which a new reference to the allocated memory increases the reference counter on it w.r.t a certain regime.

### 2025-06-24 paper.tex
revisions and comments up through 3.0

@@
 \end{comment}
-  \todo[inline]{Colin says: how about this framing instead of the second paragraph above:}
+  % \todo[inline]{Colin says: how about this framing instead of the second paragraph above:}
 
@@
   of filesystems code.
-  \todo[inline,color=cyan]{Ismail, why did you comment out my suggestion and keep the original?}
-  \todo[inline,color=pink]{Sorry, I thought I did the other way}
+  % \todo[inline,color=cyan]{Ismail, why did you comment out my suggestion and keep the original?}
+  % \todo[inline,color=pink]{Sorry, I thought I did the other way}
 \end{abstract}
@@
 \item \textbf{Identifying System Verification Challenges}: We start with identifying common patterns in system verification: \emph{virtualization, sharing, and translation}.
-\item \textbf{Introducing the Concept of \emph{Resource}}: Then we discuss the concept of \emph{resource} which has already been an essential concept in the design of systems.
+\item \textbf{Introducing Modal \emph{Resource Contexts}}: Then we discuss the concept of \emph{resource} which has already been an essential concept in the design of systems.
 Inspired by the concept of resource in the systems, we define what a resource and its context are in the modal abstractions.
-\item \textbf{Introducing the Concept of \emph{Nominals for System Resources}}:
-Nominalization enables identifying a resource in a context. 
-For example, a transaction is a context of resources of in-memory updated disk blocks. i
+\item \textbf{Introducing \emph{Nominals for System Resources}}:
+Existing logics for systems software must frequently identify
+individual resources within a given context ---
+nominalization (in the sense of hybrid logic) enables identifying a resource in its context. 
+For example, a filesystem transaction is a context of resources for in-memory updated disk blocks
+which have not yet been saved to disk.
 The transaction identifier is used to associate a transaction with a disk-block to be persisted so that, 
@@
 \emph{already persisted} disk-blocks of the transaction, and reach to the previous consistent disk state. 
-To be able to do so, both the updated in-memory disk blocks and the transaction must refer to the transaction 
-identifier -- \emph{strong nominalization}. In another example, virtual memory references (resources) in 
-an address space (resource context), which can be uniquely identified with a root address (the nominal) 
-of its page-table tree, are \emph{agnostic} to the address space that they are in. 
-However, they can only be accessed (be valid) in the address space to which they are agnostic. 
-However, then an address space switch happens, the virtual memory references of the previous address 
-space must be made inaccessible. To be able to do so, although virtual memory references do not hold any 
-piece of information related to their address space, we still have to associate them. 
-We call this kind of unilateral nominalization -- \emph{weak nominalization}.
-\todo[inline,color=cyan]{I left this paragraph alone, mostly, but it's too detailed for the new intro section position here}
+We discuss other examples to contextualize the existing systems surveyed.
+% to be able to do so, both the updated in-memory disk blocks and the transaction must refer to the transaction 
+% identifier -- \emph{strong nominalization}. 
+% in another example, virtual memory references (resources) in 
+% an address space (resource context), which can be uniquely identified with a root address (the nominal) 
+% of its page-table tree, are \emph{agnostic} to the address space that they are in. 
+% However, they can only be accessed (be valid) in the address space to which they are agnostic. 
+% However, then an address space switch happens, the virtual memory references of the previous address 
+% space must be made inaccessible. To be able to do so, although virtual memory references do not hold any 
+% piece of information related to their address space, we still have to associate them. 
+% We call this kind of unilateral nominalization -- \emph{weak nominalization}.
+% \todo[inline,color=cyan]{I left this paragraph alone, mostly, but it's too detailed for the new intro section position here}
 \item \textbf{Taxonomy of Current Modal Approaches in System Verification}: Based on these concepts defined, we summarize contemporary verification efforts using modal abstractions.
@@
 
-\subsection{Hybrid Logic Dynamic Logic and Nominals in Program Logics}
+\subsection{Hybrid Logic, Dynamic Logic, and Nominals in Program Logics}
 \label{sec:modalbackground}
 
-\paragraph{Dynamic Logic's Satisfaction Operator for Program Logics}
+\paragraph{Dynamic Logics as Program Logics}
 We believe that for the kinds of informal reasoning and the sorts of data structures discussed in the previous section, ideas from modal logic are a promising approach to formalization and verification.
@@
 The latter is an example of \emph{multimodal} logic, where there is an family of modalities (modal operators) parameterized by some dimension of interest (there, participants).
-A closely related variant of multimodal logic is \emph{dynamic logic}~\cite{pratt1976semantical}, a logic of weakest preconditions~\cite{dijkstra-75} which works with modalities of the form $[p](P)$, which states that \emph{in the current program state}, \emph{if} program $p$ is run then afterwards $P$ will hold (modulo non-termination).
-This same idea is used to encode Hoare triples in the Iris program logic~\cite{krebbers2017essence}, using the same encoding as in Pratt's original presentation, where a Hoare triple $\{P\}C\{Q\}$ is encoded as $P\rightarrow[C](Q)$.
-A hallmark of a unary logical operator $M$ being a modality is if $M$ satisfies a property akin to $(\varphi\rightarrow\psi)\rightarrow M(\varphi)\rightarrow M(\psi)$, which roughly states that modus ponens holds under the modality.\footnote{Afficionados of modal logic will note that this property is not quite Axiom K (which requires the initial implication to also hold under $M$), but follows from K and a necessitation rule $\varphi\rightarrow M(\varphi)$.
+
+\emph{Dynamic logic}~\cite{pratt1976semantical}, a (multi)modal logic of weakest preconditions~\cite{dijkstra-75} which works with modalities of the form $[p](P)$, which states that \emph{in the current program state}, \emph{if} program $p$ is run then afterwards $P$ will hold (modulo non-termination).
+This same idea is used to encode Hoare triples in the Iris program logic~\cite{jung2018iris}, using the same encoding as in Pratt's original presentation, where a Hoare triple $\{P\}C\{Q\}$ is encoded as $P\rightarrow[C](Q)$.
+Iris is now one of the most successful foundational program logics, and its rich support for
+extending the base logic with additional modalities~\cite{krebbers2017essence} means that
+modal logics, and in particular dynamic logic, now play an outsized role in program verification ---
+many challenging proofs of program correctness use a substructural dynamic logic containing
+many additional modalities which interact with the weakest precondition modality.
+
+% A hallmark of a unary logical operator $M$ being a modality is if $M$ satisfies a property akin to $(\varphi\rightarrow\psi)\rightarrow M(\varphi)\rightarrow M(\psi)$, which roughly states that modus ponens holds under the modality.\footnote{Afficionados of modal logic will note that this property is not quite Axiom K (which requires the initial implication to also hold under $M$), but follows from K and a necessitation rule $\varphi\rightarrow M(\varphi)$.
 Non-necessitive modalities typically satisfy the weaker property we call out above.} This and related laws allow transporting general reasoning into the specific contexts or hypotheticals modeled by the modalities.
@@
 However, most prior work on hybrid logics works in classical logics rather than a substructural setting like separation logic.
+Newer work surveyed here adapts this idea to deal with different structures specified in
+separation logic.
 %We adapt these ideas to assertions of the form $@[l]\varphi$ asserts that $\varphi$ is true in a data structure with \emph{logical} state $\sigma$ \emph{at location} $l$.
-\todo{seems out of place?} \todo[inline,color=pink]{addressed}
-In practice we must design variants of this for each kind of structure we wish to apply this to, e.g. tree nodes, free list, or snapshots.
+% In practice we must design variants of this for each kind of structure we wish to apply this to, e.g. tree nodes, free list, or snapshots.
 %This section provides details of and models for each modal abstraction, focusing on how each data structure can act as a model for a data-structure-specific logic.
 %We continue to motivate the development with aspects of the write amplification process described in the previous section; Section \ref{sec:evaluation} provides further examples using the modal abstractions in proofs.
-\paragraph{Temporal Operators}
-The best known class of modal logics in the systems community is undoubtedly temporal logics (whether \textsc{LTL}~\cite{pnueli1977temporal}, \textsc{CTL}~\cite{emerson1982using}, \textsc{TLA}~\cite{lamport1994temporal}, or others),
-due in part to Lamport's influence~\cite{lamport2002specifying} leading to its not-infrequent use in specifying
-distributed algorithms \cite{ongaro2014search,ailijiang2019wpaxos}.
+
+% \paragraph{Temporal Operators}
+% The best known class of modal logics in the systems community is undoubtedly temporal logics (whether \textsc{LTL}~\cite{pnueli1977temporal}, \textsc{CTL}~\cite{emerson1982using}, \textsc{TLA}~\cite{lamport1994temporal}, or others),
+% due in part to Lamport's influence~\cite{lamport2002specifying} leading to its not-infrequent use in specifying
+% distributed algorithms \cite{ongaro2014search,ailijiang2019wpaxos}.
+
+\todo[inline,color=cyan]{the hybrid logic paragraph above is good, but this first paragraph about nominalization repeats
+some of the same things. We should merge them.}
 \paragraph{Nominalization}
@@
 
-An example utilization of state naming explicitly on the assertion appears in program logics such as Iris~\cite{jung2018iris}, which enables encoding of usage protocols (e.g.
-state transition system) resembling typestate~\cite{strom1986typestate,garcia2014foundations} as specifications.
+An example use of state naming explicitly in the assertion appears in program logics 
+with usage protocols (e.g. state transition systems resembling typestate~\cite{strom1986typestate,garcia2014foundations}, which ensure
+the operations applied to some data occur in a specified order, and that invariants for various conceptual states of the data
+--- such as a file being open or closed --- are respected). 
 Protocol assertions are \emph{annotated with the name of the last (abstract) state at which the protocol is ensured}.
-One classic application of this idea to systems is 
+A classic application of this idea to systems is 
 Halpern et al.'s work adapting modal logics of
@@
 one would need to separately provide custom reasoning to conclude $\mathsf{minCommitIndex}(3)$.
+As our survey shows, the value of the modal view (which grants both intuitive specifications and natural
+reasoning principles) is becoming apparent in systems software verification.
 
@@
 which abstracts the relationship between conceptual and physical computing resources.
+Upper-levels of the software stack work with these conceptual --- \emph{virtual} ---
+resources, while the lower levels of systems software deal with the translation
+of requests expressed in terms of virtual resources into operations on physical resources.
+
 Operating System (OS) kernels virtualize memory locations and quantity (via virtual memory and paging~\cite{Denning1970VM}).
@@
 and it makes sense to specify the systems software directly in terms of those abstractions.
+
 \paragraph{Access by Translation}
@@
 they are semantically --- logically --- partial maps, worth treating as such in verification.
+
 \paragraph{Virtualization of Memory Locations}
 \label{sec:backgroundonmachinemodel}
-\todo{change the text}
-\todo[inline,color=pink]{Based on the note I took above, I changed the text}
-A typical general purpose computers make the memory resources in RAM as \emph{virtually} available.
-In other words, a program asking for a memory unit from an operating system kernel is served with a memory address that is \emph{virtualized} such that it may be \emph{mapped to} a physical RAM location or not.
-As our informal definition suggests, the common technique for \emph{virtualization} of memory locations in OS kernels relies \emph{translating}(\emph{mapping}) virtual addresses into physical ones.
-The conceptual address-translation map is implemented with processor's(CPU) page-table trees as shown in Figure \ref{fig:pagetables}.
-In this Figure, we see a typical 4 levels of page tables, a virtual address on the top indexing into the different levels of the page-table tree.
-Fabrication of virtual addresses out of limited physical addresses is provided with entries in the page-table tree where each points to a physical address aligned to 4KB (4096 byte) boundaries.
-An address translation requires traversal of a series of tables starting from level 4, and it is conventionally referred as a \emph{page-table walk}.
+A typical general purpose computer virtualizes  memory resources in RAM.
+A program asking for a memory unit from an operating system kernel is served with a memory address that is \emph{virtualized}
+such that it may be \emph{mapped to} a physical RAM location or not. 
+As our informal definition suggests, the common technique for \emph{virtualization} of memory locations in OS 
+kernels relies \emph{translating} (\emph{mapping}) virtual addresses into physical ones. 
+The conceptual address-translation map is implemented with processor's page-table trees as shown in 
+Figure \ref{fig:pagetables}. In this Figure, we see a typical 4 levels of page tables, a virtual address on the 
+top indexing into the different levels of the page-table tree (the larger rectangles along the bottom
+are blocks of memory which serve as nodes of the tree along a particular lookup path). 
+Fabrication of virtual addresses out of limited physical addresses is provided with entries 
+in the page-table tree where each points to a physical address aligned to 4KB (4096 byte) boundaries. 
+An address translation requires traversal of a series of tables starting from level 4, and it is 
+conventionally referred as a \emph{page-table walk}. 
 The traversal ends in the level 1 page table with the final lookup to the actual page of physical memory holding the requested data, and the low-order 12 bits being used to index into this page.
@@
 Filesystems are computer programs that serve to other programs for their disk resource need.
-Like VMMs using page-table trees to fabricate more memory addresses than what actually exists, filesystems uses indexing maps for fabricating more \emph{virtual disk-page} addresses which are conventionally called \emph{disk-pages} to be distinguished from physical \emph{disk-block} addresses.
+Similar to how VMMs using page-table trees to fabricate more memory addresses than what actually exists, filesystems uses indexing maps for fabricating more \emph{virtual disk-page} addresses which are conventionally called \emph{disk-pages} to be distinguished from physical \emph{disk-block} addresses.
 \paragraph{Overview of Policies in Filesystems}
@@
 \paragraph{An Example to a Policy for Updating Disk Index Tree}
+\todo[inline,color=cyan]{This gets too detailed, and is talking about the details of ModFS itself. 
+If we're casting this
+as a survey, we probably don't want to talk about ModFS except in passing (like towards the end of the paper), but want to focus on things that are at least publicly available.
+This level of detail is also way more than later parts of the paper require for understanding --- generally
+we only want a background section to cover enough for the later parts of the paper to make sense. Any
+extra depth just ends up burning reader patience.}
 Semantically, in \textsf{ModFS}~\cite{} (as in other \textsf{CoW} filesystems~\cite{young1992episode,bonwick2003zettabyte,rodeh2013btrfs,Rosenblum1992LFS,hitz1994file}) any update to file data or metadata is realized on  unused disk locations by treating the tree as a persistent data structure.
@@
 
-\todo[inline]{Colin says: the nominals section needs another shift for the new audience:
-You don't need to define modalities here, but do need to give initially the classic notion of nominal
-(name for a Kripke world) before bridging the gap to using the concept of naming
-for something a little different.
-}
-
 \section{Contingency Decomposition of a System}
@@
 
-\todo[inline]{Colin says: I don't think this section needs to change much,
-but it should be read carefully with the new audience in mind,
-possibly giving more details on what systems concepts each system is dealing with, for a non-systems audience.}
-\paragraph{Shallow Modalities in Iris Separation Logic}
+\paragraph{Shallow Modalities in the Iris Separation Logic}
 \label{sec:modal_in_iris}
-\todo[inline,color=cyan]{this also seems taken from the VMM work}
-\todo[inline,color=pink]{this text has never been part of any text that is either submitted or made publicly available.}
-Classically modal operators are given direct semantics in terms of an underlying model, which for \textsf{Iris} would be a collection of \textsc{CMRA}s; this is the case for \textsf{Iris}'s one built-in modality for step-indexing~\cite{krebbers2017essence}.
+Classically modal operators are given direct semantics in terms of an underlying model, 
+which for \textsf{Iris}~\cite{jung2018iris} would be a collection of \textsc{CMRA}s ---
+an algebraic cousin of a BI-algebra~\cite{ohearn1999bunched}.
+This is the case for \textsf{Iris}'s one built-in modality for step-indexing~\cite{nakano-00}.
 However in general, many modalities can be given proof-theoretic semantics directly within a logic. 
-This is the case for \textsf{Iris}'s weakest precondition modality (which is defined within the \textsf{Iris} logic itself), as well as for a wide range of other modalities in the literature~\cite{restall2002introduction}.
-These modalities are typically multi-modal, with a form of roughly $M~\overline{ctxt}~\phi\triangleq\ldots\rightarrow \phi~\overline{ctxt}$.
-In a substructural setting~\cite{restall1993modalities,krebbers2017essence} the implication $\rightarrow$ is typically replaced with substructural implication $\sepimp$.
+This is the case for \textsf{Iris}'s weakest precondition modality (which is defined within the \textsf{Iris} 
+logic itself), as well as for a wide range of other modalities in the 
+literature~\cite{restall2002introduction}. These modalities are typically multi-modal, 
+with a form of roughly $M~\overline{ctxt}~\phi\triangleq\ldots\rightarrow \phi~\overline{ctxt}$. 
+In a substructural setting~\cite{restall1993modalities,krebbers2017essence} the implication $\rightarrow$ 
+is typically replaced with substructural implication $\sepimp$.
 
 This definition also reveals a subtlety about the denotation of assertions $\phi$ --- they are effectively \emph{predicates} of some kind, functions from some possible world into some logical algebra such as a BI-algebra~\cite{ohearn1999bunched}.
-Taking this interpretation literally gives rise to a convenient way to embed new modalities in \textsf{Iris} in a way well-suited to our use cases.
-If $B$ is a type with a BI-algebra structure, then any space of predicate functions (e.g., $\mathsf{val}\rightarrow B$ as a simplification) also carries a BI-algebra structure, lifted from $B$, often called the \emph{pointwise lifting} of the algebra.
-We will exploit this repeatedly, lifting the BI-algebra structure from \textsf{Iris} itself to \textsf{Iris} predicates used as assertions within our own modalities.
-Connecting this to \textsf{Iris}'s typeclasses for interactive proofs~\cite{krebbers-17} then permits working within the embedded logic as in \textsf{Iris}.
+Taking this interpretation literally gives rise to a convenient way to embed new modalities in \textsf{Iris} 
+in a way well-suited to many use cases. 
+If $B$ is a type with a BI-algebra structure, then any space of predicate functions (e.g., $\mathsf{val}\rightarrow B$ as a simplification) also carries a BI-algebra structure, lifted from $B$, often called the
+\emph{pointwise lifting} of the algebra. A number of 
+projects for for systems software verification now take advantage of this,
+explicitly or implicitly lifting the BI-algebra structure from \textsf{Iris} itself to \textsf{Iris} predicates 
+used as assertions within modalities. 
+Connecting this to \textsf{Iris}'s typeclasses for interactive proofs~\cite{krebbers2018mosel,krebbers-17} then permits working within the embedded logic as in \textsf{Iris}.
 
 In addition, the modal definitions themselves may represent ownership of resources in a substructural setting~\cite{restall1993modalities,dovsen1992modal,kamide2002kripke}. 
-This will allow our modalities to simultaneously represent ownership of certain resources used for interpreting the modality itself, as well as whatever resource ownership is implied by the modal argument.
+This allows modalities to simultaneously represent ownership of certain resources used for interpreting the modality itself, as well as whatever resource ownership is implied by the modal argument.
 

### 2025-06-25 paper.tex
pull ModFS and CoW material from most of the paper

@@
 
-\paragraph{Virtualization of Memory Locations}
+\subsubsection{Virtualization of Memory Locations}
 \label{sec:backgroundonmachinemodel}
@@
 
-\todo[inline,color=cyan]{Ismail, is this text copy-pasted from the VMM paper? 
-It has phrasing like ``In this paper...''
-That's actually a major issue, self-plagiarism of another peer-reviewed piece of work can get
-you into a lot of trouble.
-}\todo[inline,color=pink]{I am well-aware what counts as plagiarism (including self-plagiarism). These were annotated with todos including cautions for changing the text.}
-\paragraph{Virtual Memory Managers(VMM) with Multiple Address-Spaces}
-\label{sec:backgroundonvmm}
-\todo{change the text}
-An intuition for the abstraction that \emph{contains} a page-table tree is called an \emph{address-space} of a process whose virtual addresses to physical addresses.
-A typical OS manages multiple processes at once .
-Each process is identified by an \emph{unique root address} of its page-table tree.
-As part of switching to a new process, i.e., loading the address-space mappings as the current view of the memory, the CPU stores the unique root address of the page-table tree in a specific register (\texttt{cr3}) as shown in Figure \ref{fig:pagetables}.
-Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section) is how the OS ensures memory isolation between processes.
-Virtual memory managers are the modules in OS kernels handling this kind of tasks.
-The treat address-spaces as containers of virtual addresses so that they can bookkeep which which virtual addresses are valid (and in what way) in which address spaces.
-\paragraph{Virtualized Resources in Filesystems}
+A page table tree is referred to as a (virtual) \emph{address space}, as it presents an abstraction
+of how different memory addresses could be related in space.
+This functionality is used by operating systems to present fictional collocation or separation of resources:
+a range of logically (virtually) contiguous memory addresses may be translated to a range of physically disparate locations
+in actual memory.
+The address space then acts as a kind of container for virtualized memory resources tied to that container.
+
+% \paragraph{Virtual Memory Managers(VMM) with Multiple Address-Spaces}
+% \label{sec:backgroundonvmm}
+% \todo{change the text}
+% An intuition for the abstraction that \emph{contains} a page-table tree is called an \emph{address-space} of a process whose virtual addresses to physical addresses. 
+A typical OS manages multiple address spaces at once.
+Each program is associated with a unique page table tree, which is stored in a specific hardware register (\texttt{cr3} in Figure \ref{fig:pagetables}). 
+Using different mappings, which map only disjoint portions of physical memory in the map's codomain, 
+is how the OS ensures memory isolation between processes.
+
+\subsubsection{Virtualized Resources in Filesystems}
 Another important computer resource computer resource that is \emph{virtualized} for access is disk blocks. 
-Filesystems are computer programs that serve to other programs for their disk resource need.
-Similar to how VMMs using page-table trees to fabricate more memory addresses than what actually exists, filesystems uses indexing maps for fabricating more \emph{virtual disk-page} addresses which are conventionally called \emph{disk-pages} to be distinguished from physical \emph{disk-block} addresses.
-\paragraph{Overview of Policies in Filesystems}
-Filesystems do not only handle the address-translation from a disk-page to a disk-block but also need to handle the different \emph{modes} of computing in which addressing disk resources occur at different consistency modes: a disk block persisted on the disk and a disk-page which may not be persisted on the disk yet.
+Filesystems are software components that abstract disk access on behalf of normal programs
+Similar to how VMMs using page-table trees to fabricate more memory addresses than what actually exists, 
+filesystems uses indexing maps for fabricating more \emph{virtual disk-page} addresses which are conventionally called \emph{disk-pages} 
+to be distinguished from physical \emph{disk-block} addresses.
+When a regular program accesses a file, it requests a specific logical offset from the start of that file's contents.
+As with virtual memory, the filesystem may map adjacent logical blocks of a single file to physically disparate parts of the disk (or even
+to different disks), though this translation mapping is implemented entirely in software (as opposed to the hardware cooperation that exists
+for memory virtualization).
+
+% \paragraph{Overview of Policies in Filesystems}
+Filesystems not only handle the address-translation from a disk-page to a disk-block but also need to handle different
+\emph{modes} of computing where to find disk resources: a disk block may live on a disk in stable storage,
+or the most up-to-date version of a block may live only in memory (for the filesystem to write to disk at a later time, batched for efficiency).
 To handle the consistency of the disk resources in the existence of different modes of execution, filesystems employ different policies when indexing the disk resource.
-Which data-structure is going to be used to implement address-traslation from disk-pages to disk-blocks?
+Which data-structure is going to be used to implement address-translation from disk-pages to disk-blocks? 
 What is going to be the \emph{update-policy} to the indexing map implementing address-translation which defines the \emph{consistency-policy} for the disk resources? 
 Different filesystem implementations give different answers to these questions.
-One example answer to updating policy is copy-on-write (\textsf{CoW}).
+
+% One example answer to updating policy is copy-on-write (\textsf{CoW}).
 A (\textsf{CoW}) filesystem performs updates by copying modified data to \emph{new and unused} locations on disk rather than updating \emph{in-place}, ensuring that in the absence of physical storage corruption, there is \emph{always} a fully consistent filesystem on disk, removing the need for recovery from journals in the case of a crash, which can be very expensive for large volumes.
+
 %\textsf{ModFS} is implemented in \textsc{Go}~\cite{donovan2015go}, using the FUSE~\cite{fuse} interface to run a filesystem in usermode.
@@
 files and directories) as a nested trees, and to map a logical page to a physical page, the file system first locates the root of the nested index-tree belonging to \textsf{"foo"}.
-\paragraph{An Example to a Policy for Updating Disk Index Tree}
-\todo[inline,color=cyan]{This gets too detailed, and is talking about the details of ModFS itself. 
-If we're casting this
-as a survey, we probably don't want to talk about ModFS except in passing (like towards the end of the paper), but want to focus on things that are at least publicly available.
-This level of detail is also way more than later parts of the paper require for understanding --- generally
-we only want a background section to cover enough for the later parts of the paper to make sense. Any
-extra depth just ends up burning reader patience.}
-Semantically, in \textsf{ModFS}~\cite{} (as in other \textsf{CoW} filesystems~\cite{young1992episode,bonwick2003zettabyte,rodeh2013btrfs,Rosenblum1992LFS,hitz1994file}) any update to file data or metadata is realized on  unused disk locations by treating the tree as a persistent data structure.
-As an example, Figure \ref{fig:cowindex} shows that an update to the file $\textsf{"foo"}$ at $\textsf{"offset"}$ ends up with locating the related disk chunk(multiple disk-pages) in the gray region, copying it to newly allocated disk chunks (dotted chunks on the right side of the bold long-dash-point line of the middle layer withing dotted section), and performing the update on these newly-allocated disk blocks.
+
+% \paragraph{An Example to a Policy for Updating Disk Index Tree}
+% \todo[inline,color=cyan]{This gets too detailed, and is talking about the details of ModFS itself. 
+% If we're casting this
+% as a survey, we probably don't want to talk about ModFS except in passing (like towards the end of the paper), but want to focus on things that are at least publicly available.
+% This level of detail is also way more than later parts of the paper require for understanding --- generally
+% we only want a background section to cover enough for the later parts of the paper to make sense. Any
+% extra depth just ends up burning reader patience.}
+% Semantically, in \textsf{ModFS}~\cite{} (as in other \textsf{CoW} filesystems~\cite{young1992episode,bonwick2003zettabyte,rodeh2013btrfs,Rosenblum1992LFS,hitz1994file}) any update to file data or metadata is realized on  unused disk locations by treating the tree as a persistent data structure.
+% As an example, Figure \ref{fig:cowindex} shows that an update to the file $\textsf{"foo"}$ at $\textsf{"offset"}$ ends up with locating the related disk chunk(multiple disk-pages) in the gray region, copying it to newly allocated disk chunks (dotted chunks on the right side of the bold long-dash-point line of the middle layer withing dotted section), and performing the update on these newly-allocated disk blocks.
 This requires further updates to those metadata blocks which previously referred to the old block locations to instead reference the newly-allocated blocks.
@@
 
-\begin{figure}[t]
-    \includegraphics[width=0.75\columnwidth]{cow_index.png}
-    \caption{A Filesystem called \textsf{Modfs} Indexing its Physical Pages for File Pages with CoW (copy-on-write) Tree}
-    \label{fig:cowindex}
-\end{figure}
-\paragraph{Examples to Different Policies on Navigating on Consistent Views of Disk}
-Like VMMs navigating between different address-spaces, a \textsf{CoW} filesystem switches from either the old to the new root yields consistent views of the filesystem at different points in time, each commonly referred to as a \emph{snapshot}.
+% \begin{figure}[t]
+%     \includegraphics[width=0.75\columnwidth]{cow_index.png}
+%     \caption{A Filesystem called \textsf{Modfs} Indexing its Physical Pages for File Pages with CoW (copy-on-write) Tree}
+%     \label{fig:cowindex}
+% \end{figure}
+% \paragraph{Examples to Different Policies on Navigating on Consistent Views of Disk}
+% Like VMMs navigating between different address-spaces, a \textsf{CoW} filesystem switches from either the old to the new root yields consistent views of the filesystem at different points in time, each commonly referred to as a \emph{snapshot}.
 A \textsf{CoW} filesystems treat the snapshot abstraction as VMMs treat the address-space abstraction, container of the relevant resources.
@@
 Super chunks, as we see in Figure \ref{fig:cowindex} as ($\textsf{Super}_0$ and $\textsf{Super}_1$), are the first two blocks of the disk dedicated to keep the state of the file system which is composed of some meta-data on the filesystem, and most importantly to the root chunk of the filesystem index tree.
-\begin{figure}
-    \includegraphics[width=\columnwidth]{journalling.png}
-    \caption{A Filesystem Utilizing \emph{Journalling} for Consistent Disk-Layout ~\cite{chajed2021gojournal}}
-    \label{fig:journal}
-\end{figure}
-\todo[inline,color=cyan]{This looks like it's verbatim copy-pasted from the ModFS draft; if this exact text
-  it used in this publication, it means we'd need to completely rewrite the relevant text in that draft.}
-Therefore, in case of a crash during the update to the \textsf{file "foo"}, we can reboot the file system from the previous persisted root, \textsf{root chunk}, via reading the $\textsf{Super}_0$.
+% \begin{figure}
+%     \includegraphics[width=\columnwidth]{journalling.png}
+%     \caption{A Filesystem Utilizing \emph{Journalling} for Consistent Disk-Layout ~\cite{chajed2021gojournal}}
+%     \label{fig:journal}
+% \end{figure}
+% \todo[inline,color=cyan]{This looks like it's verbatim copy-pasted from the ModFS draft; if this exact text
+%   it used in this publication, it means we'd need to completely rewrite the relevant text in that draft.}
+% Therefore, in case of a crash during the update to the \textsf{file "foo"}, we can reboot the file system from the previous persisted root, \textsf{root chunk}, via reading the $\textsf{Super}_0$.
 This means the whole disk consistency checks are unnecessary when recovering from a crash~\cite{bonwick2003zettabyte,hitz1994file}; recovery is immediate, as opposed to requiring hours or days~\cite{henson2006reducing,agrawal2009speedy} (or longer~\cite{henson2006chunkfs}) for recovering a large volume’s consistency from logs.
-An important additional subtlety with this kind of filesystem is \emph{when} blocks in the reclaimed region can actually be reused: only \emph{after} the snapshot that made those blocks ``free'' has been completely written to disk.
+% An important additional subtlety with this kind of filesystem is \emph{when} blocks in the reclaimed region can actually be reused: only \emph{after} the snapshot that made those blocks ``free'' has been completely written to disk.
 Allowing reuse before this could result in in-place updates: part of the new snapshot could attempt to store data in space used by the prior snapshot, allowing inconsistent on-disk structures.
 
-On the other hand, filesystems with in-place updates need to implement some form of \emph{journalling} to ensure that a crash partway through an in-place update can be rolled back using an undo journal that records the original contents before modification as abstacted in Figure \ref{fig:journal} or completed using a write-ahead journal that first writes new data in unused space before \emph{also} updating in place.
-In other words, unlike our ability to distinguish all resources (disk-pages and dis-kblocks) either being consistent under a snapshot or not in filesystems with \textsf{CoW}, with in-place updates, journalling mechanism does not allow such a claim.
+While exceptions exist~\cite{bonwick2003zettabyte,rodeh2013btrfs,Rosenblum1992LFS,hitz1994file},
+most filesystems will continue to store a given block of a given file in the same physical disk block for as long as the file exists.
+These filesystems with in-place updates need to implement some form of \emph{journalling}~\cite{hagmann1987reimplementing,chutani1992episode} to ensure that a crash partway through an in-place update
+can be rolled back using an undo journal that records the original contents before in-place modification (so original contents could be restored after a crash) 
+or completed using a write-ahead journal that first writes new data in unused space before \emph{also} updating in place (so the journal entries could be re-applied after a crash). 
+% \hrule
+% In other words, unlike our ability to distinguish all resources (disk-pages and dis-kblocks) either being consistent under a snapshot or not in filesystems with \textsf{CoW}, with in-place updates, journalling mechanism does not allow such a claim.
 In other words, as we see in Figure \ref{fig:journal}, unlike \textsf{CoW}'s uniform persistency claim over disk resources under a snapshot as a resource container, elaborative bookkeeping mechanism of journalling leaves the filesystem treat disk resources in one of the resource groups in non-uniform fashion: disk-pages loaded into memory, updated disk-page not logged yet, updated and being logged or already logged disk-pages.
 Being able to treat the resources in a uniform fashion determines wwhether we need to consider the specification and proof of crash condition separately; instead a global never-violated invariant (not even temporarily broken) of the filesystem ensures the on-disk state is always consistent.
-
-\paragraph{Memory Reclamation in Client Code} In a typical configuration, allocating, deallocating and referencing memory locations for a client code leave the memory layout in a form of DAG (directed-asyclic-graph) with a unique root which resembles the first stack pointer referencing the allocated memory.
-To ease picturing the memory layout after a set of requests from the client code (for allocation, referencing etc.), we can think of the unique roots of DAGs as \emph{super} block addresses -- this time multiple ones, not only two -- in Figure \ref{fig:cowindex}, from which a path is created for a stack pointer reference to the allocated memory.
-To reclaim the allocated memory location that is no longer being accessed, systems uses a well-known method called \emph{reference-counting} in which a new reference to the allocated memory increases the reference counter on it w.r.t a certain regime.
-\paragraph{Consistency of Reclaimed Memory} A sound memory reclamation is based on the conditions where the unique is not \emph{shared}, and there is no \emph{reachability} from the root to any the references pointing to the allocated memory.
-Once the requirements are satisfied then the memory location can be reclaimed \emph{soundly} -- i.e., without violating any access through a reference to it.
-\paragraph{Buffering} Buffering is a technique used in variety of systems (filesystem page caching, write buffer in multi-processor architectures, translation-lookaside-buffers for translated addresses etc.) to delay the costly operation (committing file pages, flushing buffered writes from write-cache to memory etc.) making updates consistent.
-\paragraph{Consistency of Reads of Buffered Writes} Consistency of reads of any value potentially buffered depends on the semantics of the operation making them \emph{observable}.
-A typical architecture such as Intel-x86 considers the \emph{flush} of buffered writes \emph{non-deterministically} -- i.e., at any time at any number of buffered writes.
-On the other hand, a typical filesystem can follow a certain policy to commit the in-memory file pages to the disk.
+Because disk blocks associated with a file could now be in many states (on-disk, journaled but not updated in place, or both)
+the specification of these systems is complex. And because some states are only possible after a system failure,
+each line of filesystem code has two conceptual post-conditions: one for continued execution, and one for the stable
+state established in the event of a hardware failure immediately after that operation.
+
+\subsubsection{Reference-Counting Memory Reclamation}
+In a typical configuration, allocating, deallocating and referencing memory locations for client code 
+leaves the memory layout in a form of DAG (directed acyclic graph) with a unique root, which resembles the first stack pointer referencing the allocated memory.
+% To ease picturing the memory layout after a set of requests from the client code (for allocation, referencing etc.), we can think of the unique roots of DAGs as 
+% \emph{super} block addresses -- this time multiple ones, not only two -- in Figure \ref{fig:cowindex}, 
+% from which a path is created for a stack pointer reference to the allocated memory. 
+\todo[inline,color=cyan]{
+Why is this a DAG? Normally programs can easily create cycles, so the DAG bit requires a little more explanation.
+Do we mean that it is ``often'' the case that it's a DAG?}
+
+It is safe to reclaim memory (for later reuse) when that memory allocation will no longer be used by the client program.
+Since that property itself is undecidable, sound heuristics are used instead.
+A common proxy for the program never accessing an allocation again is to check that there is no
+access path from data a program can trivially access (e.g., local variables stored on the stack)
+to a memory allocation --- that is, when the allocation is no longer \emph{reachable}.
+In that case the program \emph{cannot} access the memory again, so the memory location can be reclaimed \emph{soundly} ---
+i.e., without violating any access through a reference to it.
+
+To reclaim the allocated memory location that is no longer being accessed, many systems use a well-known method called \emph{reference-counting} in which 
+each memory allocation includes not only its data but also a counter of how many pointers to that memory exist in the rest of the program.
+Any time a new reference to the allocated memory is created, the  counter is incremented,
+and every time an existing reference is destroyed (by overwriting or dropping) the counter is decremented.\footnote{This may be done
+using library functions for a reference counting system, or the operations may be inserted by a compiler.}
+These reference counts are a sound (but incomplete) way to approximate reachability: if a decrement of a reference count
+decrements the count to 0, no other references to that memory exist in the program, and the memory can be reclaimed.
+
+Reasoning about the correctness of such code requires reasoning about these reachability relations and reference counts, relative to specific
+stack frames.
+
+% \paragraph{Buffering} Buffering is a technique used in variety of systems (filesystem page caching, write buffer in multi-processor architectures, 
+% translation-lookaside-buffers for translated addresses etc.) to delay the costly operation (committing file pages, flushing buffered writes from write-cache 
+% to memory etc.) making updates consistent.
+% \paragraph{Consistency of Reads of Buffered Writes} Consistency of reads of any value potentially buffered depends on the semantics of 
+% the operation making them \emph{observable}. A typical architecture such as Intel-x86 considers the \emph{flush} of buffered writes 
+% \emph{non-deterministically} --- i.e., at any time at any number of buffered writes. On the other hand, a typical filesystem can follow a certain 
+% policy to commit the in-memory file pages to the disk.
+
+\subsubsection{Weak Memory Models}
+% I reworked this to be less general, and more about just weak memory, as thats all the survey section discusses.
+Multi-core and multi-processor systems do not directly write every memory update back to memory (or even a coherent hardware cache)
+individually, as this would be extremely slow. Instead, they effectively buffer reads and writes to and from memory
+locally within each CPU core\footnote{This is a layer below the better-known L1 cache, which is coherent across all CPUs in the system.}
+so that operations can be partially reordered and batched to improve performance. For most code --- where the different CPUs are doing
+independent work --- this is transparent to software. But fine-grained concurrent systems code which has multiple CPU cores
+simultaneously coordinating access to shared memory must reason about these (effective) bufferings and reorderings, whose details vary
+across CPU architectures.
+
+Most of this coordination comes through reasoning about what \emph{will be true after local operations are published to memory},
+or what \emph{will be true after remote operations become visible locally}. Each of these flavors of contingent reasoning
+depend on use of certain \emph{fence} or \emph{memory barrier} operations, which block execution of a CPU until either all incoming or outgoing
+memory transactions are complete.
+\todo[inline,color=cyan]{I feel like we need citations here}
 
@@
 are at issue should prefer weak nominalization.
-\section{Universal Properties over Resource Contexts with Nominals:  An Filesystem Recursive Trees Structure}
-\label{sec:overviewnode}
-Up until now, we discuss the concept of resource and resource context related under a certain \emph{nominalization} scheme.
+
+% \section{Universal Properties over Resource Contexts with Nominals:  An Filesystem Recursive Trees Structure}
+% \label{sec:overviewnode}
+% Up until now, we discuss the concept of resource and resource context related under a certain \emph{nominalization} scheme.
 These definitions establishes the foundation for local properties of modal patterns.
@@
 
-To reason about states of trees and subtrees in particular, we work with a modality for stating an assertion that is true \emph{of all nodes within a tree}, identifying the tree by a root note.
+% To reason about states of trees and subtrees in particular, we work with a modality for stating an assertion that is true \emph{of all nodes within a tree}, identifying the tree by a root note.
 We refer to the resources that provide the context of these modal assertions as the \emph{modal context}; for a recursively defined tree like those in \textsf{Modfs}, these contexts are recursively defined in terms of the tree structure.
 
-%We define the modal context as the resources (e.g.
+% %We define the modal context as the resources (e.g.
 a btree node tree, free page id list) which are accessed under some relation defined within the modal definition (e.g.
@@
 
-Write amplification traverses the overall tree depth-first, transforming the tree from the bottom up, one full subtree at a time.
+% Write amplification traverses the overall tree depth-first, transforming the tree from the bottom up, one full subtree at a time.
 Concepually this means the amplification code's proof must track the fact that at any point in time, there are regions of the tree where each node has already been amplified (e.g., $\phi_{\textsf{amplified}}$ definitely holds for that node), and regions where nodes may or may not require amplification, and this will not be known until those nodes are visited (i.e., some basic validity $\phi_{\textsf{valid}}$ is known to hold, but not the stronger $\phi_{\textsf{amplified}}$).
 
-%\setlength\intextsep{0pt}
-\begin{wrapfigure}{r}{0.25\textwidth}
-  \centering  \includegraphics[width=0.25\textwidth]{entail.png}
-  \caption{Understanding the level of $\varphi$ satisfaction in the \textsc{CoW} tree of modal nodes}
-   \label{fig:nodeentail} 
-\end{wrapfigure}
-We use a modality similar to $\text{\faSitemap}@[\ell]\phi$ to indicate that $\phi$ is true of the tree node rooted at $\ell$, as well as at \emph{all reachable children} of that node.
-This means that establishing a particular $\phi$ throughout a tree must be done in a bottom-up manner (see Figure \ref{fig:nodeentail}).
-Capturing this modally, rather than with more typical ad-hoc hand-crafted invariants, offers advantages in a number of proof steps, in addition to concisely capturing intuitive reasoning about subtrees with all nodes satisfying a particular property.
+% %\setlength\intextsep{0pt}
+% \begin{wrapfigure}{r}{0.25\textwidth}
+%   \centering  \includegraphics[width=0.25\textwidth]{entail.png}
+%   \caption{Understanding the level of $\varphi$ satisfaction in the \textsc{CoW} tree of modal nodes}
+%    \label{fig:nodeentail} 
+% \end{wrapfigure}
+% We use a modality similar to $\text{\faSitemap}@[\ell]\phi$ to indicate that $\phi$ is true of the tree node rooted at $\ell$, as well as at \emph{all reachable children} of that node.
+% This means that establishing a particular $\phi$ throughout a tree must be done in a bottom-up manner (see Figure \ref{fig:nodeentail}).
+% Capturing this modally, rather than with more typical ad-hoc hand-crafted invariants, offers advantages in a number of proof steps, in addition to concisely capturing intuitive reasoning about subtrees with all nodes satisfying a particular property.
 One basic example is that while the actual disk sync code requires the full filesystem tree to be in an amplified state (hence write amplification), most of the filesystem works with more general trees.
 Since $\phi_{\textsf{amplified}}\vdash\phi_{\textsf{valid}}$, using our \faSitemap modality we can conclude $\text{\faSitemap}@[\ell]\phi_{\textsf{amplified}}\vdash\text{\faSitemap}@[\ell]\phi_{\textsf{valid}}$ by an analogue of Axiom K.
-There are also other places in the code where a subtree satisfying a stronger assertion at all nodes must be linked into a larger tree satisfying only a weaker assertion overall: predicate representing the valid tree as the tree of nodes which are backed by resources allocated pages, i.e.
+% There are also other places in the code where a subtree satisfying a stronger assertion at all nodes must be linked into a larger tree satisfying only a weaker assertion overall: predicate representing the valid tree as the tree of nodes which are backed by resources allocated pages, i.e.
 $\phi_{\textsf{valid}}$, is weaker than the predicate representing the tree of nodes being currently amplified in the current snapshot.

### 2025-06-25 paper.tex
../..

@@
 \usepackage[utf8]{inputenc}
+\usepackage{rotating}
 
@@
 \label{sec:definitions}
-\begin{table*}[t]
+\begin{sidewaystable*}[t]
 \centering
@@
   \toprule
-{\scriptsize Modality}& {\scriptsize Context} & {\scriptsize Elements}  &  {\scriptsize Nominalization} & {\scriptsize Context Steps} \\ \midrule
-\scriptsize Post-Crash~\cite{tejthesis,perennialgit,tejperennial19} & \scriptsize $\Diamond \; P $  & \scriptsize $  \ell \mapsto_{n}^{\overline{\gamma}} v $ & \scriptsize Strong &  \scriptsize Crash Recovery    \\
+{ Modality}& { Context} & { Elements}  &  { Nominalization} & { Context Steps} \\ \midrule
+ Post-Crash~\cite{tejthesis,perennialgit,tejperennial19} &  $\Diamond \; P $  &  $  \ell \mapsto_{n}^{\overline{\gamma}} v $ &  Strong &   Crash Recovery    \\ 
 %GC-Modality &      & \multicolumn{2}{c}{\multirow{2}{*}{Because why not}} & 8                                   & & ~\cite{}\\
-\scriptsize NextGen~\cite{larsnextgen25} &  \scriptsize $\overset{t}{\hookrightarrow} \; P$  &\scriptsize  Own (t(a))  & \scriptsize Strong & \scriptsize Determined Based on the Model$^{*}$   \\
-\scriptsize StackRegion$^{*}$  ~\cite{larsnextgen25} & \scriptsize $\overset{ICut^{n}}{\hookrightarrow} \; P$      & \scriptsize $\fbox{n} \; \ell \mapsto v$                                 & \scriptsize Strong & \scriptsize Alloc and Return to/from stack\\
-\scriptsize Memory-Fence~\cite{fsl,fsl++,derekrustbelt20} & \scriptsize $\triangle_{\pi}$ and $\triangledown_{\pi}$      & \scriptsize $\ell \mapsto v$                                 & \scriptsize Weak & \scriptsize Fence Acquire and Release  \\
-\scriptsize Address-Space~\cite{kuru2024modalabstractionsvirtualizingmemory} & \scriptsize [r]P     &  \scriptsize $\ell \mapsto v$ & \scriptsize Weak                         & \scriptsize Address-Space Switch  \\
-\scriptsize Ref-Count~\cite{amalreal2024}&\scriptsize @$_{\ell}$ P    & \scriptsize $\ell_1 \mapsto v$ & \scriptsize Weak  & \scriptsize Allocating, Dropping and Sharing a Reference \\ \bottomrule
+ NextGen~\cite{larsnextgen25} &   $\overset{t}{\hookrightarrow} \; P$  &  Own (t(a))  &  Strong &  Determined Based on the Model$^{*}$   \\
+ StackRegion$^{*}$  ~\cite{larsnextgen25} &  $\overset{ICut^{n}}{\hookrightarrow} \; P$      &  $\fbox{n} \; \ell \mapsto v$                                 &  Strong &  Alloc and Return to/from stack\\
+ Memory-Fence~\cite{fsl,fsl++,derekrustbelt20} &  $\triangle_{\pi}$ and $\triangledown_{\pi}$      &  $\ell \mapsto v$                                 &  Weak &  Fence Acquire and Release  \\
+ Address-Space~\cite{kuru2024modalabstractionsvirtualizingmemory} &  [r]P     &   $\ell \mapsto v$ &  Weak                         &  Address-Space Switch  \\ 
+ Ref-Count~\cite{amalreal2024}& @$_{\ell}$ P    &  $\ell_1 \mapsto v$ &  Weak  &  Allocating, Dropping and Sharing a Reference \\ \bottomrule
 \end{tabular}
@@
 %and the $\Diamond$ Figure for Post-Crash Modality uses notationis taken from NextGen Modality paper ~\cite{larsnextgen25}.
-\end{table*}
+\end{sidewaystable*}
 
@@
 \label{sec:conclusion}
-\todo[inline]{this definitely needs adjusting for both audience and emphasis on this being a survey}
-We have argued the essential patterns that help to choose the modalities when working out how to specify different kinds of systems code based on recent successes.
-It also captures (a slightly more organized version of) our own thinking in coming up with designs for specifying distributed systems~\cite{gordon2019modal}, virtual memory managers~\cite{kuru2024modalabstractionsvirtualizingmemory}, and in ongoing work using different modalities for different parts of a copy-on-write filesystem (e.g., for assertions true in different snapshots).
-So not only are the modalities useful
-for conducting proofs, the pieces we have identified seem to be an effective way of working out
-a specific modality for a specific use-case.
-
-As our main conclusion and future work, we think that these essential patterns in the designs of modal abstractions discussed constitute the fundamentals of \emph{verification patterns} when working out how to specify different kinds of systems code.
+% \todo[inline]{this definitely needs adjusting for both audience and emphasis on this being a survey}
+We have argued that there is a systematic pattern emerging in the use of modalities in the verification of low-level systems software,
+and explained how a range of modern work fits into this pattern.
+% We have argued the essential patterns that help to choose the modalities when working out how to specify different kinds of systems code based on recent successes.
+While we have focused our attention on public verification work carried out in Iris,
+it also captures (a slightly more organized version of) our own thinking in coming up with designs for specifying distributed systems~\cite{gordon2019modal}, 
+% virtual memory managers~\cite{kuru2024modalabstractionsvirtualizingmemory}, 
+and in ongoing work using different modalities for different parts of a copy-on-write filesystem (e.g., for assertions true in different snapshots). 
+So not only are the modalities observably present, but they are demonstrably useful
+for conducting proofs of systems software.
+%  the pieces we have identified seem to be an effective way of working out
+% a specific modality for a specific use-case.
+We believe that these essential patterns in the designs of modal abstractions for systems verification constitute a fundamental
+concept
+when working out how to specify different kinds of systems code.
+
 \bibliographystyle{splncs04}

### 2025-06-25 paper.tex
citations to weak memory bg

@@
 individually, as this would be extremely slow. Instead, they effectively buffer reads and writes to and from memory
-locally within each CPU core\footnote{This is a layer below the better-known L1 cache, which is coherent across all CPUs in the system.}
+locally within each CPU core~\cite{owens2009better}\footnote{This is a layer below the better-known L1 cache, which is coherent across all CPUs in the system.}
 so that operations can be partially reordered and batched to improve performance. For most code --- where the different CPUs are doing
@@
 Most of this coordination comes through reasoning about what \emph{will be true after local operations are published to memory},
-or what \emph{will be true after remote operations become visible locally}. Each of these flavors of contingent reasoning
+or what \emph{will be true after remote operations become visible locally}~\cite{sieczkowski2015separation,dang2022compass,fsl}. Each of these flavors of contingent reasoning
 depend on use of certain \emph{fence} or \emph{memory barrier} operations, which block execution of a CPU until either all incoming or outgoing
 memory transactions are complete.
-\todo[inline,color=cyan]{I feel like we need citations here}
 

### 2025-06-25 paper.tex
handle a redundancy todo

@@
 
-\todo[inline,color=cyan]{the hybrid logic paragraph above is good, but this first paragraph about nominalization repeats
-some of the same things. We should merge them.}
 \paragraph{Nominalization}
 Some classes of assertions benefit specifically from \emph{naming} the explicit conditions where they are true (as opposed to simply requiring them to be true \emph{somewhere} or \emph{everywhere} as in the most classic modal operators).
-This naming generally resembles
-the \emph{satisfaction} operator of 
-\emph{hybrid logic}~\cite{blackburn1995hybrid,areces2001hybrid}: $@_\iota P$ which evaluates the truth of $P$ at the named (Kripke model) state $\iota$.
-For this reason we refer to the general idea of naming circumstances explicitly as \emph{nominalization},
-even though the examples we discuss are not necessarily actually hybrid logics.
+% This naming generally resembles
+% the \emph{satisfaction} operator of 
+% \emph{hybrid logic}~\cite{blackburn1995hybrid,areces2001hybrid}: $@_\iota P$ which evaluates the truth of $P$ at the named (Kripke model) state $\iota$. 
+Because these named conditions strongly resemble the nominals of hybrid logic,
+we refer to the general idea of naming circumstances explicitly as \emph{nominalization},
+even though not all of the examples we discuss are actually hybrid logics.
 

### 2025-06-25 paper.tex
a version for weak-stuff

@@
 
-\paragraph{Weak-Memory} When considering weak memory models, we also want points-to information (address-value mappings).
+\paragraph{Weak-Memory} When considering weak memory models, we also want points-to information (address-value mappings) of thread-local views of events (an operationalization of the release-acquire + nonatomic
+portion of the repaired C11 memory model~\cite{lahav2017repairing}): there modalities $\Delta_\pi(P)$ and $\nabla_\pi(P)$
+represent that $P$ held before or will hold after certain memory fence operations by thread $\pi$.
+The definitions of those specific modalities existentially quantify over other views, related to the ``current'' view (the one where
+the current thread's assertions are evaluated), and evaluate $P$ with respect to those other views. This approach to parameterizing
+assertion semantics by a point of evaluation, and evaluating modalized assertions at other points, is what it means
+to have a modality at all.
+It is \emph{not}, however, an instance of hybrid logic, which is specifically demarcated by an assertion language where
+\emph{assertions}, not their semantics, choose and name the evaluation points for modal assertions.
+A hybrid extension of the aforementioned logics would include assertions which named specific views at which to evaluate
+$P$, in the syntax of the assertion (e.g., $\Delta_\pi^v(P):=\lambda\_\ldotp (P\;v)$) rather than the 
+$\Delta_\pi(P):= \lambda v\ldotp (\exists v_{rel}\ldotp \ownGhost{\pi}{\mathsf{RelV}(v_{rel})\;v} \ast (P\;v_{rel})))$ actually used.
+Note the hybrid version takes the place to evaluate $P$ as a parameter, and therefore allows the \emph{derived} (modal) logic to explicitly
+reason in terms of evaluation points, rather than hiding all points of evaluation in the internal definitions of modalities.
 

### 2025-06-25 paper.tex
weak stuff

@@
  \usepackage{array}
+ \usepackage{iris}
 \newcolumntype{O}[1]{>{\raggedright\arraybackslash}m{#1}}
@@
 represent that $P$ held before or will hold after certain memory fence operations by thread $\pi$.
-The definitions of those specific modalities existentially quantify over other views, related to the ``current'' view (the one where
-the current thread's assertions are evaluated), and evaluate $P$ with respect to those other views. This approach to parameterizing
-assertion semantics by a point of evaluation, and evaluating modalized assertions at other points, is what it means
-to have a modality at all.
-It is \emph{not}, however, an instance of hybrid logic, which is specifically demarcated by an assertion language where
-\emph{assertions}, not their semantics, choose and name the evaluation points for modal assertions.
-A hybrid extension of the aforementioned logics would include assertions which named specific views at which to evaluate
-$P$, in the syntax of the assertion (e.g., $\Delta_\pi^v(P):=\lambda\_\ldotp (P\;v)$) rather than the 
-$\Delta_\pi(P):= \lambda v\ldotp (\exists v_{rel}\ldotp \ownGhost{\pi}{\mathsf{RelV}(v_{rel})\;v} \ast (P\;v_{rel})))$ actually used.
-Note the hybrid version takes the place to evaluate $P$ as a parameter, and therefore allows the \emph{derived} (modal) logic to explicitly
-reason in terms of evaluation points, rather than hiding all points of evaluation in the internal definitions of modalities.
 
@@
 Using a given strongly nominalized assertion element under different modalities for different
-frames does not change its meaning.
+frames does not change its meaning. Regarding the definitions of weak-memory specific modalities existentially quantify over other views, related to the ``current'' view (the one where
+the current thread's assertions are evaluated), and evaluate $P$ with respect to those other views. This approach to parameterizing
+assertion semantics by a point of evaluation, and evaluating modalized assertions at other points, is what it means
+to have a modality at all.
+It is \emph{not}, however, an instance of hybrid logic, which is specifically demarcated by an assertion language where
+\emph{assertions}, not their semantics, choose and name the evaluation points for modal assertions.
+A hybrid extension of the aforementioned logics would include assertions which named specific views at which to evaluate
+$P$, in the syntax of the assertion (e.g., $\Delta_\pi^v(P):=\lambda\_\ldotp (P\;v)$) rather than the 
+$\Delta_\pi(P):= \lambda v\ldotp (\exists v_{rel}\ldotp \ownGhost{\pi}{\mathsf{RelV}(v_{rel})\;v} \ast (P\;v_{rel})))$ actually used.
+Note the hybrid version takes the place to evaluate $P$ as a parameter, and therefore allows the \emph{derived} (modal) logic to explicitly
+reason in terms of evaluation points, rather than hiding all points of evaluation in the internal definitions of modalities.
 

### 2025-06-25 paper.tex
fix option clash
some tweak

--- 2025-06-25 paper.tex
@@
 \usepackage[fleqn]{amsmath}
-\usepackage{mathtools}
+\usepackage[fleqn]{mathtools}
 \usepackage{amssymb}
@@
 
-\paragraph{Weak-Memory} When considering weak memory models, we also want points-to information (address-value mappings).
+\paragraph{Weak-Memory} When considering weak memory models, we consider address-value mappings of thread-local views in C11 memory model~\cite{lahav2017repairing}).
+Modalities $\Delta_\pi(P)$ and $\nabla_\pi(P)$ with $P$ represent the memory address-value mappings that held before or will hold after certain memory fence operations by thread $\pi$.
 
@@
 Using a given strongly nominalized assertion element under different modalities for different
-frames does not change its meaning.
+frames does not change its meaning. Likewise,  weak-memory specific modalities existentially quantify over other views, related to the ``current'' view (the one where
+the current thread's assertions are evaluated), and evaluate a resource (an thread-local view of an address-value mapping) with respect to those other views.
+Although it looks like the pattern we see in the VMM case,
+It is \emph{not}, however, an instance of hybrid logic, which is specifically demarcated by an assertion language where
+\emph{assertions}, not their semantics, choose and name the evaluation points for modal assertions.
+A hybrid extension of the aforementioned logics would include assertions which named specific views at which to evaluate
+$P$, in the syntax of the assertion (e.g., $\Delta_\pi^v(P):=\lambda\_\ldotp (P\;v)$) rather than the 
+$\Delta_\pi(P):= \lambda v\ldotp (\exists v_{rel}\ldotp \ownGhost{\pi}{\mathsf{RelV}(v_{rel})\;v} \ast (P\;v_{rel})))$ actually used.
+Note the hybrid version takes the place to evaluate $P$ as a parameter, and therefore allows the \emph{derived} (modal) logic to explicitly
+reason in terms of evaluation points, rather than hiding all points of evaluation in the internal definitions of modalities.
 

### 2025-06-25 paper.tex
i think this is ok

@@
 In a typical configuration, allocating, deallocating and referencing memory locations for client code 
-leaves the memory layout in a form of DAG (directed acyclic graph) with a unique root, which resembles the first stack pointer referencing the allocated memory.
+leaves the memory layout \emph{ideally} in a form of DAG (directed acyclic graph) with a unique root, which resembles the first stack pointer referencing the allocated memory.
+DAG abstraction instead of recursive decent tree allows non-duplication on the reference counts for the identical memory objects.
 % To ease picturing the memory layout after a set of requests from the client code (for allocation, referencing etc.), we can think of the unique roots of DAGs as 

### 2025-06-25 paper.tex
+

@@
 %
-\title{Modal Verification Patterns for Systems}
+\title{Survey: Modal Verification Patterns for Systems}
 %
@@
 %
-\author{Ismail Kuru \orcidID{0000-0002-5796-2150} \and
-Colin S. Gordon \orcidID{0000-0002-9012-4490}}
+%\author{Ismail Kuru \orcidID{0000-0002-5796-2150} \and
+%Colin S. Gordon \orcidID{0000-0002-9012-4490}}
 %%%Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
 %
-\authorrunning{Kuru and Gordon}
+%\authorrunning{Kuru and Gordon}
 % First names are abbreviated in the running head.
@@
 %
-\institute{Drexel University \\ %\and
+%\institute{Drexel University \\ %\and
 %%%Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-\email{\{ik335,csgordon\}@drexel.edu}}%\\
+%\email{\{ik335,csgordon\}@drexel.edu}}%\\
 %%%\url{http://www.springer.com/gp/computer-science/lncs} \and
@@
 the current thread's assertions are evaluated), and evaluate a resource (an thread-local view of an address-value mapping) with respect to those other views.
-Although it looks like the pattern we see in the VMM case,
-It is \emph{not}, however, an instance of hybrid logic, which is specifically demarcated by an assertion language where
-\emph{assertions}, not their semantics, choose and name the evaluation points for modal assertions.
-A hybrid extension of the aforementioned logics would include assertions which named specific views at which to evaluate
-$P$, in the syntax of the assertion (e.g., $\Delta_\pi^v(P):=\lambda\_\ldotp (P\;v)$) rather than the 
-$\Delta_\pi(P):= \lambda v\ldotp (\exists v_{rel}\ldotp \ownGhost{\pi}{\mathsf{RelV}(v_{rel})\;v} \ast (P\;v_{rel})))$ actually used.
-Note the hybrid version takes the place to evaluate $P$ as a parameter, and therefore allows the \emph{derived} (modal) logic to explicitly
-reason in terms of evaluation points, rather than hiding all points of evaluation in the internal definitions of modalities.
 

### 2025-06-25 paper.tex
+
views finalized

@@
 %
-\title{Survey: Modal Verification Patterns for Systems}
+\title{Modal Verification Patterns for Systems}
 %
@@
 % from which a path is created for a stack pointer reference to the allocated memory. 
-\todo[inline,color=cyan]{
-Why is this a DAG? Normally programs can easily create cycles, so the DAG bit requires a little more explanation.
-Do we mean that it is ``often'' the case that it's a DAG?}
+%\todo[inline,color=cyan]{
+%Why is this a DAG? Normally programs can easily create cycles, so the DAG bit requires a little more explanation.
+%Do we mean that it is ``often'' the case that it's a DAG?}
 
@@
 frames does not change its meaning. Likewise,  weak-memory specific modalities existentially quantify over other views, related to the ``current'' view (the one where
-the current thread's assertions are evaluated), and evaluate a resource (an thread-local view of an address-value mapping) with respect to those other views.
+the current thread's assertions are evaluated), and evaluate a resource (a thread-local view of an address-value mapping) with respect to those other views.
+Given that views are identified by \emph{nominals}, referring to all the other views makes weak-memory modalities strongly nominalized.
 

### 2025-09-02 paper.tex
+
plos files

@@
-%\documentclass[a4paper,UKenglish]{lipics-v2016}
-% For two-column layout for PPoPP, change 'acmsmall' to 'sigplan'.  Probably also turn on the
-% 'review' option, and the 'anonymous' option if it's DBR
-\documentclass[runningheads]{llncs}
-%This is a template for producing LIPIcs articles.
-%See lipics-manual.pdf for further information.
-%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
-%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
-% for section-numbered lemmas etc., use "numberwithinsect"
-\usepackage[export]{adjustbox}
-
+%%
+%% This is file `sample-sigconf-authordraft.tex',
+%% generated with the docstrip utility.
+%%
+%% The original source files were:
+%%
+%% samples.dtx  (with options: `all,proceedings,bibtex,authordraft')
+%% 
+%% IMPORTANT NOTICE:
+%% 
+%% For the copyright see the source file.
+%% 
+%% Any modified versions of this file must be renamed
+%% with new filenames distinct from sample-sigconf-authordraft.tex.
+%% 
+%% For distribution of the original source see the terms
+%% for copying and modification in the file samples.dtx.
+%% 
+%% This generated file may be distributed as long as the
+%% original source files, as listed above, are part of the
+%% same distribution. (The sources need not necessarily be
+%% in the same archive or directory.)
+%%
+%%
+%% Commands for TeXCount
+%TC:macro \cite [option:text,text]
+%TC:macro \citep [option:text,text]
+%TC:macro \citet [option:text,text]
+%TC:envir table 0 1
+%TC:envir table* 0 1
+%TC:envir tabular [ignore] word
+%TC:envir displaymath 0 word
+%TC:envir math 0 word
+%TC:envir comment 0 0
+%%
+%% The first command in your LaTeX source must be the \documentclass
+%% command.
+%%
+%% For submission and review of your manuscript please change the
+%% command to \documentclass[manuscript, screen, review]{acmart}.
+%%
+%% When submitting camera ready or to TAPS, please change the command
+%% to \documentclass[sigconf]{acmart} or whichever template is required
+%% for your publication.
+%%
+%%
+\documentclass[sigconf,10pt]{acmart}
+%%
+%% \BibTeX command to typeset BibTeX logo in the docs
+\AtBeginDocument{%
+  \providecommand\BibTeX{{%
+    Bib\TeX}}}
+
+%% Rights management information.  This information is sent to you
+%% when you complete the rights form.  These commands have SAMPLE
+%% values in them; it is your responsibility as an author to replace
+%% the commands and values with those provided to you when you
+%% complete the rights form.
+\setcopyright{acmlicensed}
+\copyrightyear{2018}
+\acmYear{2018}
+\acmDOI{XXXXXXX.XXXXXXX}
+%% These commands are for a PROCEEDINGS abstract or paper.
+\acmConference[Conference acronym 'XX]{Make sure to enter the correct
+  conference title from your rights confirmation email}{June 03--05,
+  2018}{Woodstock, NY}
+%%
+%%  Uncomment \acmBooktitle if the title of the proceedings is different
+%%  from ``Proceedings of ...''!
+%%
+%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
+%%  June 03--05, 2018, Woodstock, NY}
+\acmISBN{978-1-4503-XXXX-X/2018/06}
+
+%%
+%% Submission ID.
+%% Use this when submitting an article to a sponsored event. You'll
+%% receive a unique submission ID from the organizers
+%% of the event, and this ID should be used as the parameter to this command.
+%%\acmSubmissionID{123-A56-BU3}
+
+%%
+%% For managing citations, it is recommended to use bibliography
+%% files in BibTeX format.
+%%
+%% You can then either use BibTeX with the ACM-Reference-Format style,
+%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
+%% support for advanced citation of software artefact from the
+%% biblatex-software package, also separately available on CTAN.
+%%
+%% Look at the sample-*-biblatex.tex files for templates showcasing
+%% the biblatex styles.
+%%
+
+%%
+%% The majority of ACM publications use numbered citations and
+%% references.  The command \citestyle{authoryear} switches to the
+%% "author year" style.
+%%
+%% If you are preparing content for an event
+%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
+%% citations and references.
+%% Uncommenting
+%% the next command will enable that style.
+%%\citestyle{acmauthoryear}
+
+%\usepackage{todonotes}
 \usepackage{booktabs}
 \usepackage{multirow}
+\usepackage{rotating}
  \usepackage{proof}
@@
 \newcolumntype{S}{>{\centering\arraybackslash}m{2cm}}
-
-\usepackage{microtype}%if unwanted, comment out or use option "draft"
-%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
-%\usepackage{t1enc}
-%\usepackage[strings]{underscore}
-%%%%%\usepackage{geometry}
-\usepackage{wrapfig}
-\usepackage[rawfloats=true]{floatrow} 
-\restylefloat{figure}    
-\DeclareMathSizes{4}{4}{4}{4}
-\usepackage{varwidth}
-\usepackage{lmodern}
-\usepackage{amsfonts}
-\usepackage[fleqn]{amsmath}
-\usepackage[fleqn]{mathtools}
-\usepackage{amssymb}
-\usepackage{hyperref}
-\usepackage{proof}
-\usepackage{stmaryrd}
-\usepackage{xspace}
-\usepackage{listings}
-\usepackage{comment}
-\usepackage{longtable}
-\usepackage{float}
-\usepackage[section]{placeins}
-\usepackage[toc,page]{appendix}
-\usepackage[export]{adjustbox}
-\usepackage{amssymb}
-\usepackage{graphics,graphicx}
-\usepackage{mathpartir}
-\usepackage{array}
-\usepackage{tikz}
-\usetikzlibrary{shapes.geometric}
-\usetikzlibrary{arrows.meta,arrows}
-\usetikzlibrary{positioning}
-\usetikzlibrary{calc, shapes, backgrounds}
-\usepackage{verbatim}
-\usepackage{tikz-qtree}
-\usepackage{pgfplots}
-\usepackage{caption,capt-of}
-\usepackage{color}
-\usepackage{soul}
-\setul{1ex}{0.8ex}
-\definecolor{orange}{rgb}{1,0.5,0}
-\setulcolor{orange}
-\usepackage{subcaption}
-\usepackage{lipsum}
-\usepackage{afterpage}
-\usepackage{fontawesome}
-\usepackage[utf8]{inputenc}
-\usepackage{rotating}
-
-%%sequeezeeee
-\addtolength{\textfloatsep}{-10pt}
-\addtolength{\dbltextfloatsep}{-10pt}
-
-\newcommand{\ptree}{\text{\faSitemap}@[\ell]}
-\newcommand{\mtree}{\varupphi\text{\faSitemap}@[\ell]^{\Pi}_{\pi}(\sigma)}
-\newcommand{\oprtree}[4]{#3\;\text{\faSitemap}@[\textsf{#1_{#2}}]^{\Pi}_{\pi}(#4)}
-
-\newcommand{\osubtree}[4]{#2\;\text{\faSitemap}@[\textsf{node}_{#1}]^{\Pi}_{\pi}(#3_#4)}
-\newcommand{\otree}[3]{#2\;\text{\faSitemap}@[\textsf{node}_{#1}]^{\Pi}_{\pi}(#3)}
-\newcommand{\oltree}[3]{#2\;\text{\faSitemap}@[#1]^{\Pi}_{\pi}(#3)}
-
-\newcommand{\defeq}{\mathbin{\stackrel{\text{def}}{=}}}
-% some stuff for prettier proof outlines
-\newcommand{\havoc}{\textsf{havoc}}
-\newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
-\newcommand{\assertion}[1]{\textsf{assert} {\color{red}\left\{#1\right\}}}
-\newcommand{\proofstepnr}[5][1pt/3pt]{\small \left.
-\begin{array}{l}
-\hdashline[#1]
-\ensuremath{\specline{#2}}\\
-\begin{array}{l}
-#3
-\end{array}\\
-
-\ensuremath{\specline{#5}}\\
-\hdashline[1pt/3pt]
-\end{array}
-\color{green}\right)\hspace{-6pt}{\color{green}-}
-\begin{array}{@{}l@{}}
-#4
-\end{array}
-}
-\newcommand{\proofoutstepnr}[5][1pt/3pt]{\small \left.
-\begin{array}{@{}l@{}}
-\hdashline[#1]
-\ensuremath{\specline{#2}}\\
-\begin{array}{@{\;\;\;}l}
-#3
-\end{array}\\
-\ensuremath{\specline{#5}}\\
-\hdashline[1pt/3pt]
-\end{array}
-\color{green}\right)\hspace{-6pt}{\color{green}-}
-\begin{array}{@{}l@{}}
-#4
-\end{array}
-}
-% \proofoutstep[dash stuff]{pre}{C}{rule}{post}
-\newcommand{\proofoutstep}[5][1pt/3pt]{
-\proofoutstepnr[#1]{#2}
-{#3}
-{\rotatebox[origin=c]{90}{$\begin{array}{@{}c@{}}{}#4\end{array}$}}
-{#5}
-}
+% Useful macros from Isaac Oscar Gariano: a very simple library for author-specific change notations:
+%  After calling \newAuthor{<name>}{<colour>}
+% You can use \<name>{<text>} to add a comment
+% Or \<name>A{<text>} to  write stuff in your colour (to denote additional text)
+% Or \<name>B{<text>} is similar, but changes the background (to highlight stuff?)
+% I may add more commands later
+% (there's a 'changes' package with similar functionality, but it has a habit of breaking when you use it with  the Define environment)
+\newcommand\newAuthor[2]{%
+        \ExpandArgs{c}\newcommand{#1}[1]{\todo[color=#2!10, bordercolor=#2, inline]{\textbf{#1:}~##1}}
+	\ExpandArgs{c}\newcommand{#1A}[1]{\textcolor{#2}{##1}}
+        \ExpandArgs{c}\newcommand{#1B}[1]{\colorbox{#2!30}{\textcolor{black}{##1}}}}
+\newtheorem*{remark}{Remark}
+\newcommand{\loc}{\textsf{Loc}}
 \newcommand{\sepimp}{\mathrel{-\mkern-6mu*}}
-\newcommand{\NULL}{\texttt{NULL}}
-\newcommand{\NEW}{\texttt{new}}
-\newcommand{\SKIP}{\texttt{skip}}
-\newcommand{\RESULT}{\texttt{result}}
-\newcommand{\class}{\texttt{class}}
-\newcommand{\where}{\texttt{where}}
-\newcommand{\readable}{\texttt{readable}}
-\newcommand{\writable}{\texttt{writable}}
-\newcommand{\isolated}{\texttt{isolated}}
-\newcommand{\immutable}{\texttt{immutable}}
-\newcommand{\consume}[1]{\texttt{consume}(#1)}
-\newcommand{\consumeiso}[1]{\mathsf{RemIso}(#1)}
-\newcommand{\INT}{\texttt{int}}
-\newcommand{\bool}{\texttt{bool}}
-\newcommand{\this}{\texttt{this}}
-\newcommand{\immutdecl}[1]{\mathsf{ImmutableDeclared}(#1)}
-\newcommand{\IsoOrImm}{\textsf{IsoOrImm}}
-\newcommand{\NonWrit}{\textsf{NoWrit}}
-\newcommand{\db}[1]{\llbracket#1\rrbracket}
-\newcommand{\nil}{\textsf{null}}
-\newcommand{\atom}{\textsf{atom} \mapsto \mstatename \mapsto \mstatename}
-\newcommand{\loc}{\textsf{Loc}}
-\newcommand{\val}{\textsf{Val}}
-\newcommand{\mstatename}{\textsf{MState}}
-\newcommand{\superset}{\mathcal{P}}
-\newcommand{\mstatedecl}{\sigma \in \mstatename \overset{def}{=} s \times h \times l \times rt \times R \times B  \times F }
-\newcommand{\stack}{s \in \var \times \tid \rightharpoonup \loc}
-\newcommand{\heap}{h \in \loc \times \fname \rightharpoonup \val}
-\newcommand{\wlock}{l \in \tid \uplus \textsf{unlocked}}
-\newcommand{\rlock}{R \in \superset(\tid)}
-\newcommand{\fset}{F : \loc \mapsto \superset(\tid) }
-\newcommand{\atomeval}{\db{-}:\atom}
-\newcommand{\rcuwrite}{\textsf{RCUWrite} \text{ x.f as y in }\, \{ C \}}
-\newcommand{\rcuread}{\textsf{RCURead} \text{ x.f as y in }\, \{ C \}}
-  \newcommand{\desugarread}{ \rcuread \overset{\Delta}{=} \texttt{\textsf{ReadBegin}}; y = x.f;\{C'\}  ;\textsf{ReadEnd} }
-  \newcommand{\desugarwrite}{\rcuwrite \overset{\Delta}{=} \texttt{\textsf{WriteBegin}} ; y = x.f ; \{C'\}  ;\textsf{WriteEnd}}
-    \newcommand{\desugarsync}{\textsf{Sync} \overset{\Delta}{=} \texttt{\textsf{SyncStart};\textsf{SyncStop}}}
-  %\newcommand{\boundpath}{\bar f_{i}^{k} = \forall_{k \in \mathbb{N},i \in \fnameenum} \ldotp  \underbrace{f_{i}^{0} \ldots f_{i}^{k}}_\text{$\mid$ $f_{i}^{0}$ \ldots $f_{i}^{k}$ $\mid$ = k} }
-\newcommand{\subt}{\prec:}
-%typing rules
- \newcommand{\fname}{\textsf{FName}}
-\newcommand{\fmname}{\textsf{FMap}}
-\newcommand{\uiglobal}{\textsf{UIGlobal}}
-\newcommand{\udef}{\textsf{undef}}
-\newcommand{\unlinked}{\textsf{unlinked}}
-\newcommand{\fresh}{\textsf{fresh}}
-\newcommand{\var}{\textsf{Var}}
-\newcommand{\tid}{\textsf{TID}}
-
-\newcommand{\fmapfieldemp}{\N_{f,\emptyset} = \N \setminus \{f \rightharpoonup \_ \}}
-\newcommand{\fmapname}{\N = \{ f_0|\ldots|f_n\rightharpoonup  \{y\} \mid f_{i} \in \fname \land \; 0\leq i\leq n  \land \; (y \in \var \lor y \in \{null\}) \}}
-\newcommand{\fmapextend}{  \N(f\rightharpoonup x \setminus y) = \N \setminus \{f \rightharpoonup x\} \cup \{f \rightharpoonup y\} }
-\newcommand{\fmapextendn}{  \N(\cup_{f\rightharpoonup y}) = \N \cup \{f \rightharpoonup y\} }
-\newcommand{\fmapdiff}{\N(\setminus_{f\rightharpoonup y}) = \N - \{f\rightharpoonup y\}}
-\newcommand{\fmapwithf}{\N([f\rightharpoonup  y]) = \N \text{   where  } f \rightharpoonup y \in \N }
-\newcommand{\fmapempty}{\N_{\emptyset} =  \{ \}}
-\newcommand{\codom}{\cod(\N)=\dom(\Gamma)}
-\newcommand{\cod}{\texttt{cod}}
-\newcommand{\dom}{\texttt{dom}}
-%\newcommand{\types}{\tau ::= \textsf{rcu} | \textsf{normal}}
-
-\newcommand{\rpth}{f \in \textsf{FName} 
-     \left\{\begin{array}{cl}
-     \epsilon &    \\
-     f  &  f \in \fname\\
-   f_1 | \ldots | f_n &\forall_{1 \leq i \leq n} \ldotp f_i \in \fname 
-     \end{array}\right\}
-     \quad \text{  where } \textsf{FType(f) = \textsf{RCU}}
-}
-\newcommand{\rextendswith}{\rho^{k}[i] \mapsto f_i \text{  where  } \forall_{0\leq i \leq k} \ldotp f_i \in \fname }
-\newcommand{\extendswith}{\rho[i] \mapsto \rho_{i}^{k} \text{  where  } \forall_{0\leq i \leq n} \ldotp \rho_i^{k} \in \textsf{RelativePath}}
-
-\newcommand{\pth}{\rho \in \textsf{Path}  \qquad \rho ::= \rho_1 \ldots\rho_n  \text{   where  } \forall_{0\leq i \leq n}\ldotp \rho_i \in \textsf{RelativePath} \text{ or } \rho_i \in \textsf{ConcretePath}}
-\newcommand{\cpath}{\forall_{0 \leq i \leq n} \ldotp f_i \in \mathsf{FName} \land l_i \in \textsf{Label} \land \exists_{k\geq0} \ldotp f_i^{l_i} = \overbrace{f_i\ldots f_i}^{k} \land (\forall_{0\leq j \leq n} \ldotp j\neq i \implies l_i \neq l_j) }
-
-\newcommand{\rpath}{\rho \in \textsf{RelativePath} \quad \rho = (f)^{*} \text{   where  } f \in \textsf{FName} \text{  and  } \exists_{k>=0} \ldotp (f)^{*} = \overbrace{f\ldots f}^\text{k} }
-\newcommand{\falt}{f \in \textsf{AlternatingFName} = f_1 | f_2 | \ldots | f_{n-1} | f_n \text{  where  } \forall_{0\leq i \leq n}\ldotp f_i \in \textsf{FName} }
-
-\newcommand{\lenmap}{\mathcal{L} \in \textsf{LPath} : \rho \mapsto \mathbb{N}}
-\newcommand{\etadecl}{\eta = \{f \mapsto y \mid f \in \fname \land y\in \var \}  }
-\newcommand{\rcunfd}{\textsf{rcuFresh} \,  \N}
-\newcommand{\rcunfdecl}{x : \textsf{rcuFresh} \;\; \N }
-\newcommand{\rcunf}[3]{#1 : \textsf{rcuFresh} \;\; \N(\cup_{ #2 \rightharpoonup #3}) } %x, f , y
-\newcommand{\rcuf}[1]{#1 : \textsf{rcuFresh} \;\;\N } %x, f , y
-\newcommand{\rcunextfresh}[1]{\textsf{rcuFresh} \;\; \N #1 }
-\newcommand{\ftypedecl}{\textsf{FType} : \fname \mapsto \tau}
-\newcommand{\ftype}[2]{\textsf{FType}(f_{#1}) \mapsto \textsf{#2}}
-%%%
-\newcommand{\typedepthdecl}{\textsf{TypeDepth}(x:\text{T}) =  \mid\rho\mid \;\; \texttt{where T} = \textsf{rcuItr} \underset{y}{\mid} \rho \,\, \N }
-%%%
-\newcommand{\rcuonly}{\textsf{RCUOnly}(x) \overset{\texttt{def}}{=} x:T \text{ where } T = \textsf{rcuItr}\_}
-\newcommand{\rcumodonlydecl}{\textsf{RCUModifiedOnly}(C) \overset{\texttt{def}}{=} \forall_{x \in \textsf{Modifies}(C)}\ldotp \textsf{RcuOnly}(x) }
-\newcommand{\rcufilterb}{\textsf{RemRcu} : \Gamma \mapsto \Gamma}
-\newcommand{\rcufilterr}[1]{\textsf{RemRcu} ( #1 ) = \texttt{filter} \, ( \lambda x \ldotp x \neq \textsf{rcuItr \_ } ) \,  #1 }
-\newcommand{\rcufiltera}[1]{\textsf{RemRcu} ( #1 ) }
-%%%Alias
-\newcommand{\aliaswith}[2]{#1 \in \textsf{AliasWith}(#2)}
-\newcommand{\aliaswithdecl}{\textsf{AliasWith}(x) = \{y \mid y:\textsf{rcuItr}\underset{r',k'}{\mid} \rho' \;\; \N' \land \rho = \rho' \land \N = \N' \}  \texttt{  where  }  \rcuitrT{x}{r}{k}{kk}{\_} }
-%%%Framing Type Contexts
-    \newcommand{\wellframed}{
-      \mathsf{WellFramed}(\Gamma,\Delta)\overset{def}{=}
-      \left\{\begin{array}{l}
-
-        \forall_{z\in dom(\Gamma),f,\rho,y }\ldotp                           z\notin dom(\Delta) \land            z:\textsf{rcuItr}\,\rho\,\N([f\rightharpoonup y]) \implies \\
-                      \left\{ 
-	              \begin{array}{l}
-                        y\notin dom(\Delta) \land y\in dom(\Gamma) \land \forall_{n\in dom(\Delta),\rho',f',m}\ldotp n\notin dom(\Gamma) \land 	\\
-            \left\{
-            \begin{array}{l}
-              n:\textsf{rcuItr}\,\rho'\,\N([f'\rightharpoonup m]) \implies \\
-              \left\{\begin{array}{l}
-              m \notin dom(\Gamma) \land \\
-              m \in dom(\Delta) \land
-              \rho \neq \rho'                    
-              \end{array} \right\}
-            \end{array}\right\}
-     \end{array}\right\}
-         \end{array}\right.
-
-}
-
-%%%Framing Type Contexts ends
-\newcommand{\deltaepsilon}{\Delta >= 0 \qquad 0\leq \epsilon \leq \Delta }
-\newcommand{\reduce}[2]{\textsf{InReduction}(#1,#2) = \{ \Gamma' \mid  \Gamma' \preceq_{\varepsilon} \Gamma \} }
-\newcommand{\rename}[2]{\textsf{ContextRename}(#1,#2) = #2 \mapsto #1 }
-%%Sep RCU
-\newcommand{\seprcu}[4]{\textsf{SepLoop}(#1,#2,#3,#4) \overset{\texttt{def}}{=} \text{dom}(#1)\cap \text{dom}(#2) = \emptyset \, \land \, #1\uplus#2= #4 \, \land \, \textsf{RcuModifiedOnly}(#2,#3)}
-\newcommand{\seprcud}[4]{\textsf{SepLoop}(#1,#2,#3,#4)}
-%%
-\newcommand{\seprcub}[5]{\textsf{SepBranch}(#1,#2,#3,#4,#5) \overset{\texttt{def}}{=}
-   \left\{\begin{array}{cl}
-  \text{dom}(#2)\cap \text{dom}(#3) = \emptyset \, \land \, #2\uplus#3=#1 \land \\
-  \textsf{RcuModifiedOnly}(#2,#4) \,\land\, \textsf{RcuModifiedOnly}(#2,#5)\,\land \\
-
- \end{array}\right\}
-}
-\newcommand{\seprcubd}[5]{\textsf{SepBranch}(#1,#2,#3,#4,#5)}
-%%
-\newcommand{\precgamma}[2]{#1  \preceq #2 \overset{\texttt{def}}{=} \forall_{(x)\in #1}\ldotp x \in #2 \implies  #1 \preceq_{x} #2}
-%%%if branches
-\newcommand{\iffree}{\textsf{if}(\text{e})\{\text{C}\}}
-\newcommand{\ifelse}{\textsf{if}(\text{e})\{\text{C}_1\}\textsf{else}\{\text{C}_2\}}
-%%var  root,k  , k+1 mappedvar
-\newcommand{\rcuitertype}[4]{\textsf{rcuItr}(#1,#2)\;#3\;#4}
-\newcommand{\rcuit}[4]{#1:\textsf{rcuItr}\;\; #3 \;\; #4 }
-\newcommand{\rcuitrT}[5]{#1:\textsf{rcuItr}\;\; \rho \;\; \N }
-\newcommand{\rcuitrn}[5]{#1:\textsf{rcuItr}\;\; \rho \, f \;\; \N }
-\newcommand{\rcuitr}[5]{#1:\textsf{rcuItr}\;\; \rho \;\; \N[f  \rightharpoonup #5] }
-\newcommand{\rcustack}[2]{#1 : \textsf{rcuItr}\underset{#2}{\mid}}
-\newcommand{\afree}[1]{\textsf{asyncDelayedFree}(#1)}
-\newcommand{\allocate}{\textsf{asyncDelayedFree}}
-\newcommand{\typelocal}{\tau_{local}}
-\newcommand{\typeglobal}{\tau_{global}::=\textsf{rcuRoot}  }
-\newcommand{\rcutypes}{ \typelocal ::=   \textsf{undef} \mid \textsf{rcuItr} \,\, \rho \,\, \N \mid \rcunfd \mid \textsf{unlinked}  \mid \textsf{freeable}}
-\newcommand{\typelocalmap}{\textsf{TypeLocal} : \var \mapsto \typelocal}
-\newcommand{\command}[1]{\db{#1} \mstate \commandeq}
-\newcommand{\commandm}[2]{\db{#1} #2 \commandeq}
-
-\newcommand{\mstate}{(s,h,l,R,F)}
-\newcommand{\commandeq}{\;\;\Downarrow_{\mathrm{tid}}\;\;\overset{\Delta}{=}}
-\newcommand{\op}[2]{\begin{array}{rll}\command{#1} & #2\end{array}}
-\newcommand{\opm}[3]{\begin{array}{rll}\commandm{#1}{#2} & #3\end{array}}
-%%%
-\newcommand{\return}[1]{\mathsf{return~\mathit{#1}}}
-\newcommand{\bind}[2]{\mathsf{Bind}(#1;#2)}
-\newcommand{\result}[1]{\mathsf{y = \mathsf{result}}}
-\newcommand{\mbody}[2]{\mathsf{mbody}(#1,#2)}
-\newcommand{\pre}{\textsf{pre}}
-\newcommand{\assume}[1]{\textsf{assume}(#1)}
-\newcommand{\lift}[1]{\lceil #1\rceil}
-\newcommand{\erase}[1]{\lfloor #1\rfloor}
-%%%
-\newcommand{\M}{\ensuremath{\mathcal{M}}}
-\newcommand{\s}{\ensuremath{\mathcal{S}}}
-\newcommand{\R}{\ensuremath{\mathcal{R}}}
-\newcommand{\N}{\ensuremath{\mathcal{N}}}
-\newcommand{\condexpr}{e \in \textsf{CondExpr}}
-%function definitions for paths
-
-\newcommand{\lengthofpath}{$\mid \rho\mid \;=\;
-  \begin{cases}
-      0+ \mid \rho'\mid &  \rho = \rho_{i}^{k}\rho' \text{   where  } 0\leq i  \text{   and   } \rho_{i}^{k} = \epsilon \\
-         k + \mid \rho'\mid & \rho = \rho_{i}^{k}\rho' \text{   where  } 0\leq i
-  \end{cases}$
-
-}
-%%\textsf{Stabilize}(\Gamma,e) = \forall x \in \N[\fname] \ldotp \exists_{f_1,f_2 \in \fname}\ldotp \N[f_1] =  \N[f_2] \implies \N[f_1,f^{*}] \land \N[f_2,f^{*}]
-
-\newcommand{\stabilizenext}{\textsf{StabilizeNext}(a,b,f)}
-\newcommand{\stabilizepath}{\textsf{StabilzePath}(b,f,k)}
-\newcommand{\stabilizecontext}{\textsf{StabilizeContext}(\Gamma,e) =
-    \left\{\begin{array}{cl}
-    e \in \textsf{PointsToExpr} \land
-    \exists_{a,b,f}\ldotp  (a,b,f) = \textsf{PointsToExprMap}(e) \land \\
-  \pi_{\N}(a) = \textsf{StabilizeNextMap}(a,b,f) \land \\
-  \pi_{\rho}(b) = \textsf{StabilizePath}(b,f, \mid \pi_{\rho}(b)\mid ) \land\\
-  \forall_{c \in \Gamma \land \textsf{TypeDepth}(c) >= \textsf{TypeDepth}(b)} \ldotp \pi_{\rho}(c) = \textsf{StabilizePath}(c,f,\mid\pi_{\rho}(c)\mid)
-  \end{array}\right\}
-}
-%%%frame branch
-\newcommand{\alternatingvars}[1]{\textsf{AbstractVars}(#1)}
-\newcommand{\alternatingvarsdecl}{\textsf{AbstractVars}(\Gamma) \;=\; \{ x,y \mid  (x,y)\in \textsf{AbstractPairs}(\Gamma)\}}
-\newcommand{\alternatetype}{ \text{x:}\textsf{rcuItr}\;\; \rho \;\; \N[f^{*}\mapsto y] \qquad   \text{y:}\textsf{rcuItr}\;\; \rho.\rho^{k} \;\; \N \text{    where    } \rho^{k}[k]=f^{*}}
-\newcommand{\alternatingpath}[1]{f^{*} \; = \; f_1 \mid f_2  \; \text{where} \;  (f=f_1) \;\oplus \; ( f=f_2 )  }
-\newcommand{\alternatingpathsdecl}[1]{\textsf{AbstractPairs}(#1) = \; \{ (\text{x},\text{y}) \mid \text{x} \in #1 \land \text{y} \in #1\}}
-\newcommand{\alternatingpaths}[1]{\textsf{AbstractPairs}(#1)}
-\newcommand{\pointstotuple}{\textsf{PointsToTuple}}
-\newcommand{\pointstoexpr}{\textsf{PointsToExpr} = \{ x.f==y  \in \textsf{Expr} \mid \text{x} \in \var \; \land \; \text{y} \in \var \; \land \; f \in \fname \}}
-\newcommand{\pointstopair}[1]{\textsf{PointsTo}(\text{#1})}
-\newcommand{\projectfieldecl}{\pi_{f} : (\text{x},\text{y},f) \mapsto f}
-\newcommand{\projectfield}[1]{\pi_{f}(#1)}
-\newcommand{\projectpairdecl}{\pi_{\var} :(\text{x},\text{y},f) \mapsto (\text{x},\text{y})}
-\newcommand{\projectpair}[1]{\pi_{\var}(#1)} % tuple,f.
-\newcommand{\pointstoexprdecl}{\textsf{PointsToExpr} = \{e \mid  e = (a.f == b) \land \condexpr \land a,b \in \var \land f \in \fname  \} }
-\newcommand{\pointstostmtdecl}{\textsf{PointsToExprMap}: \textsf{PointsToExpr}\mapsto \var \times \var \times \fname}
-\newcommand{\pointstostmt}[1]{\textsf{PointsToExprMap}(#1) }
-\newcommand{\freductionindexdecl}{\textsf{FieldReductionIndex}(\text{C},\text{x},\text{y},\text{irdc})\; = \;
-  \left\{\begin{array}{cl}
-    \forall_{0 \leq  i \leq n} \ldotp &  (x',y',f')_{i}  \in \pointstostmt{C} \land \\
-    &(\text{x},\text{y})= \projectpair{(\text{x}',\text{y}',f')_{i}} \implies\\
-    & \left\{\begin{array}{cl}
-    \exists_{ firdc} \ldotp  & firdc \leq i \land  (firdc = 0 \implies firdc=irdc) \land \\
-     & \left\{\begin{array}{cl}
-    \forall_{0 \leq none \le firdc} \ldotp & (\text{x},\text{y}) \neq \projectpair{(\text{x}',\text{y}',f')_{none})}  \land \\
-    &  firdc = \text{irdc}
-      \end{array}\right\}
-    \end{array}\right\}
-  \end{array}\right\}
-}
-
-\newcommand{\projectsharpdecl}{\Pi_{\varphi_{e}}(f^{*}) \mapsto f_1 \qquad \Pi_{\lnot\varphi_{e}}(f^{*}) \mapsto f_2 }
-\newcommand{\projectsharpdeclback}{\Pi(\varphi_{e})\mapsto e}
-
-\newcommand{\frenamefielddecl}{\textsf{FieldRename}(\text{x},f,f')  \; = \; \text{x:T}[f/f'] }
-
-\newcommand{\freductiondecl}{\textsf{FieldReduction}(\Gamma,\varphi_{e}) \; = \;
-\left\{\begin{array}{cl}
-  \textsf{FieldRename}(\text{x},f^{*},\Pi_{\varphi_{e}}(f^{*})),
-  \textsf{FieldRename}(\text{y},f^{*},\Pi_{\varphi_{e}}(f^{*})) \mid (\text{x},\text{y}) =\projectpair{\pointstostmt{\Pi(\varphi_{e})}}
-\end{array}\right\}
-}
-
-\newcommand{\freductiondeclneq}{\textsf{FieldReduction}(\Gamma,\lnot\varphi_{e}) \; = \;
-\left\{\begin{array}{cl}
-  \textsf{FieldRename}(\text{x},f^{*},\Pi_{\lnot\varphi_{e}}(f^{*})),
-  \textsf{FieldRename}(\text{y},f^{*},\Pi_{\lnot\varphi_{e}}(f^{*})) \mid (\text{x},\text{y}) = \projectpair{\pointstostmt{\Pi(\varphi_{e})}}
-\end{array}\right\}
-  }
-
-\newcommand{\gammafieldreduction}[4]{\textsf{ReduceAbstractPaths}(#3,#4,#2,#1) = \forall_{x,y \in \textsf{AbstractVars}(#4)} \ldotp \textsf{FieldReduction}(\text{x},\text{y},\varphi_{#3}) = #2 \land \textsf{FieldReduction}(\text{x},\text{y},\varphi_{#3}) = #1  } % x,y,gamma,c
-\newcommand{\alterinxys}{\text{a:}\textsf{rcuItr}\;\; \rho \;\; \N[f_1\mapsto b] \qquad
-  \text{b:}\textsf{rcuItr}\;\; \rho.f_1 \;\; \N  \text{  or }\text{b:}\textsf{rcuItr}\underset{r}{\mid} \rho.\{...f_1\} \;\; \N \\
-  \text{a:}\textsf{rcuItr}\;\; \rho \;\; \N[f_2\mapsto b] \qquad
-  \text{b:}\textsf{rcuItr}\;\; \rho.f_2 \;\; \N \text{  or  } \text{b:}\textsf{rcuItr}\;\; \rho.\{...f_2\} \;\; \N }
-\newcommand{\reducedpathsdecl}[2]{\textsf{ReducedPaths}(#1,#2)\; = \; \left\{\begin{array}{cll}
-  (a,b) \mid  &\text{a:}\textsf{rcuItr}\;\; \rho \;\; \N[f_1\mapsto b] \in #1 & \\
-  &  \land  \text{b:}\textsf{rcuItr}\;\; \rho.\rho^{k} \;\; \N \in #1  & \text{     where    } \rho^{k}[k]=f_1 \\
-  &  \land  \text{a:}\textsf{rcuItr}\;\; \rho \;\; \N[f_2\mapsto b] \in #2 & \\
-  &  \land \text{b:}\textsf{rcuItr}\;\; \rho.\rho^{k} \;\; \N  \in #2 & \text{     where    } \rho^{k}[k]=f_2
-  \end{array}\right\} }
-\newcommand{\exprTopredDecl}{\textsf{ExprToPredicate}(e) \mapsto \varphi_{e}}
-\newcommand{\reducedvars}{\textsf{ReducedVars}(\Gamma_1,\Gamma_2)}
-\newcommand{\reducedvarsdecl}{\textsf{ReducedVars}(\Gamma_1,\Gamma_2)\;=\; \{\forall_{a,b \in \Gamma_1 \cup \Gamma_2} \ldotp (a,b) \in \textsf{ReducedPaths}(\Gamma_1,\Gamma_2) \}}
-\newcommand{\reducedpaths}[2]{\textsf{ReducedPaths}(#1,#2)}
-\newcommand{\abstractpaths}[4]{\textsf{AbstractPaths}(#1,#2,#3,#4) \; = \; \left\{\begin{array}{cl} a ,b \mid
-
-  \begin{array}{cl}
- \forall_{a,b \in #1}\ldotp & \textsf{FieldRename}(a,#2,\rewritealternate{#2}{#3}{#4}) \land \\
-  &\textsf{FieldRename}(a,#3,\rewritealternate{#2}{#3}{#4}) \land\\
-  &\textsf{FieldRename}(b,#2,\rewritealternate{#2}{#3}{#4}) \land \\
-  &\textsf{FieldRename}(b,#3,\rewritealternate{#2}{#3}{#4})
-  \end{array}
-   \end{array}\right\}
- }
-\newcommand{\rewritealternatedecl}[3]{\textsf{AbstractField}(#1,#2,#3) = f^{*} \; \texttt{    where   } f^{*} \;=\;  \frac{\varphi_{#3}}{#1 \mid #2} }
-  \newcommand{\rewritealternate}[3]{\textsf{AbstractField}(#1,#2,#3)}
-\newcommand{\composerename}{\circ \text{  provides composing renaming of types ... removes duplicates? [TODO : May change this, keep this simple, now it is nothing, same as ;]. }}
-
-%%%%%%%
-\newcommand{\p}{\rho}
-
-\newcommand{\ptriple}[3]{\{#1\}#2\{#3\}}
-\newcommand{\atriple}[3]{\begin{array}{l}\{#1\}\\#2\\\{#3\}\end{array}}
-\newcommand{\assert}[1]{\{#1\}}
-\newcommand{\modifies}[1]{\textsf{Modifies}(#1)}
-
-%\usepackage[usenames,dvipsnames]{color}
-\definecolor{framed}{gray}{0.50}
-\newcommand{\framed}[1]{\textcolor{framed}{\assert{#1}}}
-
-% Make sure these are all done
-\usepackage{todonotes}
-
-\newcommand{\iso}[1]{\textrm{\color{magenta}\ensuremath{\llbracket}\textrm{Iso says:} #1\ensuremath{\rrbracket}}}
-\newcommand{\colin}[1]{\textrm{\color{cyan}\ensuremath{\llbracket}\textrm{Colin says:} #1\ensuremath{\rrbracket}}}
-%\newcommand{\tocite}[1]{\textrm{\color{green}$\llbracket$CITE: #1$\rrbracket$}}
-
-\newcommand{\mypar}[1]{\textbf{#1}~\xspace}
-
-\newcommand{\hallocprovided}{
-  \texttt{ provided } s[(y,tid)\mapsto o], \;  h[o \mapsto \texttt{new}] \text{ and } \texttt{ default }(\textsf{FType}(f))\\
-  \lambda o', f  \ldotp  \text{ if }  o = o' \;  \text{ skip } \text{ else }  h(o',f)
-}
-\newcommand{\gcprovided}{
-  \texttt{ provided }  s[(x,tid)\mapsto o], \; (\forall f, o' \ldotp o\neq o' \Rightarrow h(o',f) = h'(o',f))) \\
-  \text{ and } \forall f \ldotp h'(o,f)\; \textsf{ undefined }
-}
-\newcommand{\rcuwbeginprovided}{
- \textrm{provided }tid \notin R
-}
-\newcommand{\rcurbeginprovided}{
- \textrm{provided }tid \neq l
-}
-\newcommand{\rcurendprovided}{
- \textrm{provided } \; F \setminus tid \; \textrm{ as } \; \{ (T \setminus tid \},o) \mid (T,o) \in F \}
-}
-\newcommand{\gc}{
-  \inferrule*[left={\scriptsize(Garbage-Collect)}]{}
-           {
-             \op{\texttt{gcStep}}{(s,h',l,R,F)}
-           }
-}
-%%%
-\newcommand{\rcuwbegin}{
-  \inferrule*[left={\scriptsize(RCU-WriteBegin)}]{}
-           {
-            \opm{\texttt{WriteBegin}}{(s,h,\mathrm{unlocked},R,F)}{(s,h,\mathrm{tid},R,F)}
-           }
-}
-\newcommand{\rcuwend}{
-  \inferrule*[left={\scriptsize(RCU-WriteEnd)}]{}
-           {
-            \opm{\texttt{WriteEnd}}{(s,h,\mathrm{tid},R,F)}{(s,h,\textsf{unlocked},R,F)}
-           }
-}
-\newcommand{\rcurbegin}{
-  \inferrule*[left={\scriptsize(RCU-ReadBegin)}]{}{
-    \opm{\texttt{ReadBegin}}{(s,h,l,R,F)}{(s,h,l,R \uplus \{ \mathrm{tid}\},F)}
-  }
-}
-\newcommand{\rcurend}{
-  \inferrule*[left={\scriptsize(RCU-ReadEnd)}]{}{
-    \opm{\texttt{ReadEnd}}{( s,h,l,R \uplus \{ \mathrm{tid} \},F)}{(s,h,l,R,F \setminus \mathrm{tid}) }
-  }
-}
-
-%%%
-\newcommand{\asyncfree}{
-\inferrule*[left={\scriptsize(Async-Free)}]{}
-           {
-            \op{\texttt{asyncDelayedFree}(x)}{(s,h,l,R,F \uplus \{ (R,S(x))\} )}
-           }
-
-}
-\newcommand{\hallocate}{
-  \inferrule*[left={\scriptsize(Heap-Allocate)}]{}
-             {
-                \op{\texttt{y=new}}{(s,h,l,R,F)}
-             }
-}
-\newcommand{\supdate}{
-  \inferrule*[left={\scriptsize(Stack-Update)}]{}
-           {
-             \op{\texttt{y=x}}{(s[(y,tid) \mapsto (x,tid)],h,l,R,F)}
-           }
-}
-\newcommand{\hread}{
-  \inferrule*[left={\scriptsize(Heap-Read)}]{}
-             {
-               \op{\texttt{y=x.f}}{((s[(y,tid) \mapsto h(s(x,tid),f)],h,l,R,F)}
-             }
-}
-\newcommand{\hupdate}{
-  \inferrule*[left={\scriptsize(Heap-Update)}]{}
-             {
-               \op{\texttt{x.f=y}}{(s,h[s(x,tid),f \mapsto s(y,tid)],l,R,F)}
-             }
-}
-\newcommand{\tuigread}[1][T-UniqueGlobalRead]{
-\infer[\textsc{#1}]{
-  \Gamma \, , y:\_ \vdash_{M,R} y = G \dashv y: \textsf{rcuItr}\;\; \rho \;\; \N_{\emptyset} , \Gamma
-}{
-\begin{array}{c}
-   G : \uiglobal \\
-\end{array}
-}
-}
-
-\newcommand{\tuisread}[1][T-UniqueStackRead]{
-\infer[\textsc{#1}]{
-  \begin{array}{c}
-  \Gamma,\rcuitrn{y}{G}{0}{k+1}{\_} , x:\textsf{\_}  \vdash_{M,R} x = y \dashv  \rcuitrn{x}{y}{1}{2}{\_} , \rcuitrn{y}{G}{0}{1}{\_}, \Gamma \\
-  \end{array}
-}{
-\begin{array}{c}
-   G : \uiglobal \\
-\end{array}
-}
-}
-
-\newcommand{\toverwc}[1][T-OverWrite]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma, x:\textsf{rcuItr}\;\; \rho \;\; \N([f \rightharpoonup \_])  \vdash_{M,R} x = x.f \dashv   x:\textsf{rcuItr}\;\; \rho' \;\; \N(\setminus_{f\rightharpoonup \_})  ,\,  \Gamma \\
-\end{array}
-}{
-\begin{array}{c}
-  \ftype{ }{\textsf{rcu}}  \qquad \mid \rho \mid +1 = \mid \rho' \mid  \\
-\end{array}
-}
-}
-%\qquad \aliaswith{z}{x} \rcuitr{z}{y}{k}{k+1}{x}\,
-
-\newcommand{\treadstack}[1][T-ReadStack]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma, z:{\_}, \rcuitrT{x}{y}{k}{k+1}{\_}   \vdash_{M,R} z=x \dashv  \rcuitrT{x}{y}{k}{k+1}{\_}, \rcuitrT{z}{y}{k}{k+1}{\_} , \Gamma \\
-\end{array}
-}{
-\begin{array}{c}
- \\
-\end{array}
-}
-}
-
-\newcommand{\treadnext}[1][T-ReadNext]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma, z:{\_},  x:\textsf{rcuItr}\;\; \rho \;\; \N([f\rightharpoonup \_])   \vdash_{M,R} z=x.f \dashv  x:\textsf{rcuItr}\;\; \rho \;\; \N(f\rightharpoonup \_ \setminus z),\;\;z:\textsf{rcuItr}\;\; \rho' \;\; \N \;, \Gamma \\
-\end{array}
-}{
-\begin{array}{c}
-  \ftype{ }{\textsf{rcu}} \qquad \mid \rho \mid + 1 = \mid \rho' \mid \\
-\end{array}
-}
-}
-%%%This is read and assignment for loop
-
-\newcommand{\tloverwc}[1][T-OverLoopWrite]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma,  x:\textsf{rcuItr}\;\; \rho\{...f_{loop}\}_{\Delta} \,\, \N   \vdash_{M,R} x = x.f \dashv   \textsf{rcuItr}\;\; \rho\{...f\}_{\Delta} \,\, \N ,\,  \Gamma \\
-\end{array}
-}{
-\begin{array}{c}
-  \ftype{ }{\textsf{rcu}}  \\
-\end{array}
-}
-}
-\newcommand{\tlreadstackext}[1][T-ReadLoopStack1]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma, z:{\_}, x:\textsf{rcuItr}\;\; \rho\{...f_{loop}\}_{\Delta} \,\, \N \\  \vdash_{M,R} z=x \dashv \\ x:\textsf{rcuItr}\;\; \rho\{...f_{loop}\}_{\Delta} \,\, \N, z:\textsf{rcuItr}\;\; \rho\{...f_{loop}\}_{\Delta} \,\, \N \, , \Gamma \\
-\end{array}
-}{
-\begin{array}{c}
- \\
-\end{array}
-}
-}
-\newcommand{\tlreadstackinit}[1][T-ReadLoopStackInit]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma, z:\textsf{rcuItr}\;\; \rho\{...f_{loop}\}_{\Delta} \,\, \N[f \mapsto x], x:\textsf{rcuItr}\;\; \rho f \{...f_{loop}\}_{\Delta} \,\, \N \\  \vdash_{M,R} z=x \dashv \\ x:\textsf{rcuItr}\;\; \rho f \{...f_{loop}\}_{\Delta} \,\, \N, z:\textsf{rcuItr}\;\; \rho\{...f\}_{\Delta} \,\, \N \, , \Gamma \\
-\end{array}
-}{
-\begin{array}{c}
-  \ftype{ }{\textsf{rcu}} \qquad f_{loop} = \epsilon \\
-\end{array}
-}
-}
-\newcommand{\tlreadnextone}[1][T-ReadLoopNextAbstract]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma, z:\textsf{rcuItr}\;\; \rho f_1\{...f_{loop}\}_{\Delta} \,\, \N, x:\textsf{rcuItr}\;\; \rho\{...f_1\}_{\Delta} \,\, \N[f_1\mapsto z] \\  \vdash_{M,R} z=x.f_2 \dashv \\  x:\textsf{rcuItr}\;\; \rho\{...f_1\}_{\Delta} \,\, \N[\frac{}{f_1 \mid f_2}\mapsto z],\, z:\textsf{rcuItr}\;\; \rho\{...f_2\}_{\Delta} \,\, \N \,, \Gamma \\
-\end{array}
-}{
-\begin{array}{c}
-  \ftype{ }{\textsf{rcu}} \qquad f_{loop}=\epsilon  \\
-\end{array}
-}
-}
-\newcommand{\tlreadnexttwo}[1][T-ReadLoopNextExtend]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma, z:\textsf{rcuItr}\;\; \rho f_1\{...f_{loop}\}_{\Delta} \,\, \N, x:\textsf{rcuItr}\;\; \rho\{...f_1\}_{\Delta} \,\, \N[f_1\mapsto z] \\  \vdash_{M,R} z=x.f_1 \dashv \\  x:\textsf{rcuItr}\;\; \rho\{...f_1\}_{\Delta} \,\, \N[f_1\mapsto z],\, z:\textsf{rcuItr}\;\; \rho\{...f_1\}_{\Delta} \,\, \N \,, \Gamma \\
-\end{array}
-}{
-\begin{array}{c}
-  \ftype{ }{\textsf{rcu}} \qquad f_{loop}=\epsilon  \\
-\end{array}
-}
-}
-\newcommand{\tlreadnextthree}[1][T-ReadLoopNext3]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma, z:{\_}, x:\textsf{rcuItr}\;\; \rho\{...f_{loop}\}_{\Delta} \,\, \N \\  \vdash_{M,R} z=x.f \dashv \\  x:\textsf{rcuItr}\;\; \rho\{...f_{loop}\}_{\Delta} \,\, \N[f\mapsto z],\, z:\textsf{rcuItr}\;\; \rho\{...f\}_{\Delta} \,\, \N \,, \Gamma \\
-\end{array}
-}{
-\begin{array}{c}
-  \ftype{ }{\textsf{rcu}} \\
-\end{array}
-}
-}
-\newcommand{\approximatedecl}{\textsf{Approximate}(x:\textsf{rcuItr}\;\; \rho \;\; \N, \Delta)\mapsto x:\textsf{rcuItr} \;\;\rho\{...\} \;\;\N}
-\newcommand{\approxtype}[2]{\textsf{Approximate}(#1,#2)}
-\newcommand{\gammaapproxdecl}{\Gamma_1 \approx_{\Delta} \Gamma_2 \text{   where  } \Gamma_2 = \{\approxtype{x:\text{T}}{\Delta} \mid \forall_{x:\text{T}\in\Gamma_1}  \}}
-\newcommand{\gammapprox}[3]{#1 \approx_{#3} #2}
-% \qquad \aliaswith{t}{x} \rcuitr{t}{y}{k}{k+1}{z}
-\newcommand{\tunlinknode}[1][T-UnlinkANode]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-\Gamma, \,
-x:\textsf{rcuItr}\;\; \rho \;\; \N([f_1\rightharpoonup z]) , \;\;\;
-z:\textsf{rcuItr} \;\; \rho' \;\; \N'([f_2\rightharpoonup r]) , \;\;\; r:\textsf{rcuItr}\;\; \rho'' \;\; \N'' \\
-\vdash_{M} x.f=r \dashv \\
-z:\unlinked, \,
-x:\textsf{rcuItr}\;\; \rho \;\; \N(f_1\rightharpoonup z \setminus r), \;\;\;
-r:\textsf{rcuItr}\;\; \rho' \;\; \N'' \;\;
-\Gamma \\
-\end{array}
-}{
-\begin{array}{c}
-  \ftype{ }{\textsf{rcu}} \qquad \mid\rho\mid + 1= \mid \rho' \mid  \qquad \mid \rho' \mid + 1 = \mid \rho'' \mid  \\ \forall_{f\in dom(\N')} \ldotp f\neq f_2 \implies \N'([f\rightharpoonup \textsf{Null}])  \\
-\end{array}
-}
-}
-%\qquad  \aliaswith{t}{z} t: \unlinked , \,
-\newcommand{\tafree}[1][T-AsyncFree]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma, x:\unlinked \vdash_{M} \afree{x} \dashv x:\udef , \Gamma \\
-\end{array}
-}{
-\begin{array}{c}
-   \\
-\end{array}
-}
-}
-
-\newcommand{\tallocate}[1][T-Allocate]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma, x:\_ \vdash_{M} x = \NEW \dashv x:\textsf{rcuFresh} \;\; \N_{\emptyset} , \, \Gamma \\
-\end{array}
-}{
-\begin{array}{c}
-   \\
-\end{array}
-}
-}
-
-\newcommand{\tsetfieldfresh}[1][T-SetFieldFresh]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma, \rcuf{p} \;\;, x:\textsf{rcuItr}\;\; \rho \;\; \N([f\rightharpoonup z]) \;\; \\
-  \vdash_{M} p.f = x.f \dashv \\
-  \rcunf{p}{f}{z} ,  \;\;, x:\textsf{rcuItr}\;\; \rho \;\; \N([f\rightharpoonup z])\;\;, \Gamma \\
-\end{array}
-}{
-\begin{array}{c}
-   \\ f \notin \N
-\end{array}
-}
-}
-
-\newcommand{\tlinkfreshf}[1][T-LinkFreshNext]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma, \:\;
-z:\textsf{rcuItr}\;\; \rho \;\; \N([f\rightharpoonup x]) \;\;,
-x:\textsf{rcuItr}\;\; \rho' \;\; \N \;\;,
-r:\textsf{rcuFresh} \;\; \N([f\rightharpoonup x]) \\
-  \vdash_{M} z.f = r \dashv \\
-z:\textsf{rcuItr}\;\; \rho \;\; \N(f\rightharpoonup x\setminus r) \;\;,
-r:\textsf{rcuItr}\;\; \rho' \;\; \N([f\rightharpoonup x]) \;\; ,
-x:\textsf{rcuItr}\;\; \rho'' \;\; \N
- \;\; , \Gamma \\
-\end{array}
-}{
-\begin{array}{c}
-   \ftype{ }{\textsf{rcu}} \qquad \mid\rho\mid + 1= \mid \rho' \mid  \qquad \mid \rho' \mid + 1 = \mid \rho'' \mid \\
-\end{array}
-}
-}
-
-\newcommand{\tlinkfresh}[1][T-LinkFresh]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma, \;\;
-  p:\textsf{rcuItr}\;\; \rho \;\; \N([f\rightharpoonup o]) \;\;,
-  o:\textsf{rcuItr}\;\; \rho' \;\; \N' \;\;, n:\textsf{rcuFresh} \;\; \N'' \\
-  \vdash_{M} p.f = n \dashv \\
-  p:\textsf{rcuItr}\;\; \rho \;\; \N(f \rightharpoonup o \setminus n) \;\;,
-  n:\textsf{rcuItr}\;\; \rho' \;\; \N'' \;\;,
-  o:\unlinked, \;\;  \Gamma
-\end{array}
-}{
-\begin{array}{c}
-   \ftype{ }{\textsf{rcu}}  \qquad \mid \rho \mid + 1= \mid \rho' \mid \qquad \N' = \N'' \\
-\end{array}
-}
-}
-\newcommand{\tloop}[1][T-Loop]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-   \Gamma \vdash \text{C}^{*} \dashv  \Gamma \\
-\end{array}
-}{
-  \begin{array}{c}
-     \Gamma \vdash C \dashv \Gamma
-\end{array}
-}
-}
-\newcommand{\tloopold}[1][T-LoopOld]{
-\infer[\textsc{#1}]{
-  \begin{array}{c}
- \Gamma_1',\Gamma_1'' \vdash \text{C}^{\Delta} \dashv   \Gamma_1'''',\Gamma_1''    \\
-%  \tloopweaken\\
-\end{array}
-}{
-  \begin{array}{c}
-    \tframeloop \\
-       \Delta = 0 \implies \Gamma_1' =\Gamma_1''''  \land \\
-       \Delta >0 \implies
-       \left\{\begin{array}{cl}
-       \Gamma_1'\approx_{\Delta}\Gamma_1''' \land \left\{\begin{array}{cl}
-       \exists_{0<\varepsilon \leq\Delta} \ldotp 0<\varepsilon<\Delta  \implies  \Gamma_1'''\vdash C \dashv \textsf{ContexRename}(\Delta,\Gamma_1''',\Gamma_1'''') \land \\
-       \varepsilon = \Delta  \implies \Gamma_1'''\vdash C \dashv \Gamma_1''''
-       \end{array}\right\}
-       \end{array}\right\} \\
-\end{array}
-}
-}
-
-% \qquad \rcuitrT{y}{G}{0}{k+1}{\_} \\  \rcuonly{\rcufiltera{\Gamma_1}}{k}{y} \\ \rcuonly{\rcufiltera{\Gamma_2}}{k'}{y} \\   \Gamma_1 \vdash C \dashv   \Gamma_2
-
-\newcommand{\tpar}[1][T-Par]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma_1 \; \Gamma_2  \vdash_{M,R} \bar{s_1} || \bar{s_2} \dashv  \Gamma'_1 \; , \Gamma'_2 \\
-\end{array}
-}{
-\begin{array}{c}
-   \Gamma_1 \vdash_{R} \bar{s_1} \dashv   \Gamma'_1  \qquad   \Gamma_2 \vdash_{M,R} \bar{s_2} \dashv   \Gamma'_2\\
-\end{array}
-}
-}
-\newcommand{\tllinkfresh}[1][T-LoopExtLinkFresh]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \{\text{TODO}\} \\
-\end{array}
-}{
-\begin{array}{c}
- \{\text{TODO}\}\\
-\end{array}
-}
-}
-\newcommand{\tllinkfreshnext}[1][T-LoopExtLinkFreshNext]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-\{\text{TODO}\}\\
-\end{array}
-}{
-\begin{array}{c}
-\{\text{TODO}\} \\
-\end{array}
-}
-}
-\newcommand{\tlunlink}[1][T-LoopExtUnlink]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-   \{\text{TODO}\} \\
-\end{array}
-}{
-\begin{array}{c}
- \{\text{TODO}\}\\
-\end{array}
-}
-}
-\newcommand{\tskip}[1][T-Skip]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-   \Gamma \vdash \textsf{skip} \dashv   \Gamma \\
-\end{array}
-}{
-\begin{array}{c}
- \\
-\end{array}
-}
-}
-\begin{comment}
-\newcommand{\tframebranch}[1][T-FrameRCUBranch]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
- \Gamma_{1\text{C}_1}, \, \Gamma'_{1\text{C}_1}  \vdash C_1 \dashv \Gamma'_{23R},\, \Gamma_{abs} \qquad  \Gamma_{1\text{C}_1}, \, \Gamma'_{1\text{C}_2}  \vdash C_2 \dashv  \Gamma'_{23R},\, \Gamma_{abs}
-\end{array}
-}{
-  \begin{array}{c}
-    \Gamma_1  \vdash \text{if e C_1 else C_2}  \dashv  \Gamma_4 \\
-    \exists_{\Gamma'_1} \ldotp \Gamma'_1= \textsf{AbstractVars}(\Gamma_1) \qquad \exists_{\Gamma''_1}\ldotp \Gamma_1 = \Gamma'_1 \uplus \Gamma''_1\\
-    \Gamma''_1,\Gamma'_1\vdash C_1 \dashv   \Gamma_2 \qquad \Gamma''_1,\Gamma'_1 \vdash C_2 \dashv \Gamma_3
-
-\end{array}
-}
-}
-\end{comment}
-
-\newcommand{\tbranch}[1][T-Branch]{
-\infer[\textsc{#1}]{
-  \begin{array}{c}
- \Gamma,x:\textsf{rcuItr}\;\; \rho \;\; \N([f_1\mid f_2 \rightharpoonup z]) \vdash \text{if e C$_1$ else C$_2$} \dashv \Gamma_4
-\end{array}
-}{
-  \begin{array}{c}
-    e \in \textsf{PointsToExpr} \qquad e \text{ is } x.f_1==z \\
- \Gamma,x:\textsf{rcuItr}\;\; \rho \;\; \N([f_1 \rightharpoonup  z]) \vdash C_1 \dashv\Gamma_4 \qquad \Gamma,x:\textsf{rcuItr}\;\; \rho \;\; \N([f_2 \rightharpoonup z]) \vdash C_4 \dashv \Gamma_4
-\end{array}
-}
-}
-
-\newcommand{\tframe}[1][T-Frame]{
-\infer[\textsc{#1}]{
-  \begin{array}{c}
- \Gamma, \; \Gamma_1 \vdash C \dashv \Gamma, \; \Gamma_2
-\end{array}
-}{
-  \begin{array}{c}
- \Gamma_1 \vdash C\dashv \Gamma_2  \qquad \textsf{RcuModifiedOnly}(C)
-\end{array}
-}
-}
-
-%%\seprcubd{\Gamma_1}{\Gamma_2}{\Gamma_3}{\Gamma_4}{C_1}{C_2}{\Gamma}
-
-%\Gamma''_1,\Gamma'_3\vdash \text{C$_1$} \dashv   \Gamma_2 \qquad \Gamma''_1,\Gamma'_1 \vdash \text{C$_2$} \dashv \Gamma_3 \qquad \Gamma_{23} = \Gamma_2 \cup \Gamma_3  \\
-   %  \Gamma_{23R} = \textsf{ReducedVars}(\Gamma_2,\Gamma_3) \qquad \Gamma'_{23R}=\Gamma_{23} - \Gamma_{23R} \qquad \Gamma_{abs} = \textsf{AbstractPaths}(\Gamma_{23R},\text{f$_1$},\text{f$_2$},e) \qquad \Gamma_4 =  \Gamma'_{23R},\, \Gamma_{abs}
-
-\newcommand{\tchoosef}[1][T-Choose-First]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma_1  \vdash C_1 \dashv  \Gamma_2 \\
-\end{array}
-}{
-\begin{array}{c}
-   \Gamma_1 \vdash C_1 \dashv   \Gamma_2  \qquad   \Gamma_1 \vdash C_2 \dashv   \Gamma_3\\
-\end{array}
-}
-}
-\newcommand{\tchooses}[1][T-Choose-Second]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma_1  \vdash C_2 \dashv  \Gamma_3 \\
-\end{array}
-}{
-\begin{array}{c}
-   \Gamma_1 \vdash C_1 \dashv   \Gamma_2  \qquad   \Gamma_1 \vdash C_2 \dashv   \Gamma_3\\
-\end{array}
-}
-}
-\newcommand{\tseq}[1][T-Seq]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \Gamma_1  \vdash \bar{s_1} \; ; \; \bar{s_2} \dashv  \Gamma_3 \\
-\end{array}
-}{
-\begin{array}{c}
-   \Gamma_1 \vdash \bar{s_1} \dashv   \Gamma_2  \qquad   \Gamma_2 \vdash \bar{s_2} \dashv   \Gamma_3\\
-\end{array}
-}
-}
-
-\newcommand{\tshuffle}[1][T-Exchange]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-   \Gamma, \, x:T , \, y:T' , \, \Gamma' \vdash \bar{s} \dashv   \Gamma'' \\
-\end{array}
-}{
-\begin{array}{c}
-   \Gamma, \, y:T' , \, x:T, \, \Gamma' \vdash \bar{s} \dashv   \Gamma'' \\
-\end{array}
-}
-}
-\newcommand{\tconsequence}[1][T-Conseq]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-   \Gamma \vdash C \dashv \Gamma''' \\
-\end{array}
-}{
-\begin{array}{c}
-   \Gamma \subt \Gamma'  \qquad \Gamma' \vdash C \dashv \Gamma'' \qquad \Gamma'' \subt \Gamma''' \\
-\end{array}
-}
-}
-
-%%Sub typing
-\newcommand{\nsubone}[1][N-Sub-F1]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-   \vdash \, \N \subt \N'  \\
-\end{array}
-}{
-\begin{array}{c}
-   \text{T } = \textsf{rcuItr}\;\; \rho \;\; \N([f_1 \rightharpoonup y]) \\
-   \text{T}'=\textsf{rcuItr}\;\; \rho \;\; \N'([f^{*} \rightharpoonup y])
-\end{array}
-}
-}
-
-\newcommand{\nsubtwo}[1][N-Sub-F2]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-   \vdash \, \N \subt \N'  \\
-\end{array}
-}{
-\begin{array}{c}
-   \text{T} = \textsf{rcuItr}\;\; \rho \;\; \N([f_2 \rightharpoonup y]) \\
-   \text{T}'=\textsf{rcuItr}\;\; \rho \;\; \N'([f^{*} \rightharpoonup y])
-\end{array}
-}
-}
-\newcommand{\nsubthree}[1][N-Sub]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-   \vdash \, \N_1 \subt \N \, \land \, \N_2 \subt \N  \\
-\end{array}
-}{
-\begin{array}{c}
-  \N_1[f_2\mapsto x] \qquad  \N_2[f_1\mapsto x] \qquad  \N[f^{*} \mapsto x] \qquad f^{*} = \frac{\varphi_{e}}{f_1 \mid f_2}\\
-\end{array}
-}
-}
-
-\newcommand{\psubone}[1][P-Sub]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-   \vdash  \rho' \subt \rho'' \\
-\end{array}
-}{
-  \begin{array}{c}
-  \rho'' = \rho'\rho^{k}
-\end{array}
-}
-}
-
-\newcommand{\psubtwo}[1][P-Sub-Ref]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-   \vdash  \rho \subt \rho  \\
-\end{array}
-}{
-  \begin{array}{c}
-  \\
-\end{array}
-}
-}
-
-\newcommand{\psubthree}[1][P-Sub-Trans]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-   \vdash  \rho' \subt \rho'''  \\
-\end{array}
-}{
-  \begin{array}{c}
-    \vdash  \rho' \subt \rho'' \qquad
-    \vdash  \rho'' \subt \rho'''
-\end{array}
-}
-}
-
-\newcommand{\csubfour}[1][C-Sub-Emp]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-   \vdash  \epsilon \subt \epsilon  \\
-\end{array}
-}{
-  \begin{array}{c}
-\end{array}
-}
-}
-
-\newcommand{\tsubase}[1][S-Base]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
- \vdash \, \text{T} \prec_{\varepsilon = 0} \text{T'} \qquad \text{T} =_{\varepsilon = 0} \text{T'} \\
-\end{array}
-}{
-\begin{array}{c}
- \varepsilon = 0 \\
-\end{array}
-}
-}
-\newcommand{\tsubreflexive}[1][S-Reflexive]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
- \vdash \, \text{T} \preceq_{\varepsilon} \text{T} \\
-\end{array}
-}{
-\begin{array}{c}
-  \\
-\end{array}
-}
-}
-\newcommand{\tsubtransitive}[1][S-Transitive]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-  \vdash \, \text{T}_1 \preceq_{2\varepsilon} \text{T}_3 \\
-\end{array}
-}{
-\begin{array}{c}
-    \vdash \, \text{T}_1 \preceq_{\varepsilon} \text{T}_2 \qquad \vdash \, \text{T}_2 \preceq_{\varepsilon} \text{T}_3 \\
-\end{array}
-}
-}
-\newcommand{\tsubempty}[1][S-Empty]{
-\infer[\textsc{#1}]{
-\begin{array}{c}
-   \vdash \, \epsilon \preceq_{\varepsilon} \epsilon \\
-\end{array}
-}{
-\begin{array}{c}
-    \\
-\end{array}
-}
-}
-\newcommand{\subcontextone}[1][C-Sub-Red-notsure]{
-\infer[\textsc{#1}]{
-  \begin{array}{c}
-        \vdash \Gamma_1  \subt \Gamma \, \land \, \Gamma_2 \subt \Gamma \\
-\end{array}
-}{
-  \begin{array}{c}
-    \textsf{ReduceAbstractPaths}(e,\Gamma,\Gamma_1,\Gamma_2) \\
-\end{array}
-}
-}
-\newcommand{\subeqcontexttwo}[1][C-Sub-Abs-notsure]{
-\infer[\textsc{#1}]{
-  \begin{array}{c}
-        \vdash \Gamma'  \subt \Gamma \\
-\end{array}
-}{
-  \begin{array}{c}
-    \textsf{ReducedVars}(\Gamma_1,\Gamma_2) = \Gamma' \qquad \textsf{AbstractPaths}(\Gamma',f_1,f_2,e) =\Gamma
-\end{array}
-}
-}
-\newcommand{\subeqcontextthree}[1][C-Sub]{
-\infer[\textsc{#1}]{
-  \begin{array}{c}
-        \vdash \Gamma,\, x:\text{T}  \subt \Gamma',\, x:\text{T'} \\
-\end{array}
-}{
-  \begin{array}{c}
-    \vdash\Gamma \subt \Gamma' \qquad
-    \vdash \text{T}  \subt \text{T'}
-\end{array}
-}
-}
-
-\begin{comment}
-\newcommand{\tsubcons}[1][S-Cons]{
-\infer[\textsc{#1}]{
-  \begin{array}{c}
-        \vdash \Gamma, \, \text{x:T}  \prec \Gamma',\, \text{x':T'}  \\
-\end{array}
-}{
-\begin{array}{c}
-      \vdash \Gamma \prec \Gamma'  \qquad \text{T} \prec \text{T'}  \\
-\end{array}
-}
-}
-
-\newcommand{\tsubdrop}[1][S-Drop]{
-\infer[\textsc{#1}]{
-  \begin{array}{c}
-      \vdash \Gamma \prec \Gamma' \\
-\end{array}
-}{
-\begin{array}{c}
-    \vdash \Gamma, \, \text{x:T}  \prec \Gamma'  \\
-\end{array}
-}
-}
-\end{comment}
-%%%%
-\newcommand{\tsmovewrite}{
-\infer[\textsc{Move-To-RCUWrite}]{
-\begin{array}{c}
-  \Gamma \vdash \textsf{RCUWrite}\, x.f \textsf{ as }y\textsf{ in }\{C\} \\
-\end{array}
-}{
- \begin{array}{c}
- \Gamma, y : \textsf{rcuItr} \vdash_M C \dashv \Gamma' \qquad
- \ftype{}{\textsf{RCU}} \\
- \textsf{NoFresh}(\Gamma') \qquad
-\textsf{NoUnlinked}(\Gamma')
-\end{array}
-}
-}
-\newcommand{\tsmoveread}{
-  \infer[\textsc{Move-To-RCURead}]{
-\begin{array}{c}
-\Gamma \vdash \textsf{RCURead}\, x.f \textsf{ as }y\textsf{ in }\{C\} \\
-\end{array}
-}{
-\begin{array}{c}
-\ftype{}{\textsf{rcu}} \\
-\Gamma, y : \textsf{rcuItr} \vdash_R C \dashv \Gamma' \\
-\end{array}
-}
-}
-
-\newcommand{\meta}{
-\[
-\begin{array}{rclcl}
-a & \textsf{atoms} \\
-s & \textsf{statement}  & & s::= & a \mid skip \mid s;s \mid s+s \mid s \parallel s \mid s^{*} \\
-n,x,y,z,o,p & \textsf{variables}  & &  \Gamma::= & . \mid \Gamma,x:t \\
-t & \textsf{types}  & & t::= & \mathsf{int} \mid \mathsf{bool} \mid \mathsf{struct} \; sn \{ \overline{fld}\} \mid T\\
-T & \textsf{Qualified types} & & T::= & t<rt> \\
-f & \textsf{field name}  & & fld::= & t \; f \mid  t<rt> \; f\\
-rb & \textsf{rcu blocks}  & & rb::= & \mathsf{ReadBegin} \, ; \bar{s}  \, ; \mathsf{ReadEnd} \mid \mathsf{WriteBegin} \, ; \bar{s}  \, ;  \mathsf{WriteEnd} \\
-fld & \textsf{field decl}  & & rt::= & \mathsf{rcu} \mid \mathsf{rcuItr} \mid \mathsf{rcuRoot} \\
-sn & \textsf{struct name}   & &  i,j & \mathsf{nat (indices)}  \\
-
-\end{array}
-\]\\
-\[
-\begin{array}{rclcl}
-\alpha &::= & \mathsf{skip} \mid \mathsf{x.f = y} \mid \mathsf{y=x} \mid \mathsf{y =x.f}  \mid \mathsf{y =new}\mathsf{Free}\mathsf{(\textsf{x})} \mid \mathsf{Sync}
- \end{array}
-\]
-}
-
-\newcommand{\grammar}{
-\[
-\begin{array}{rclcl}
-\alpha &::= & \mathsf{skip} \mid \mathsf{x.f = y} \mid \mathsf{y=x} \mid \mathsf{y =x.f}  \mid \mathsf{y =new} \mid  \mathsf{Free}\mathsf{(\textsf{x})} \mid \mathsf{Sync}                   
-\end{array}\quad
-\textsf{Sync} \overset{\Delta}{=} \texttt{\textsf{SyncStart};\textsf{SyncStop}}
-\]
-}
-
-\lstset{numbersep=1.7pt,xrightmargin=-0.7cm,xleftmargin=0.7em,numbers=left,columns=fullflexible,escapechar=`,mathescape,language=Java,basicstyle=\small\ttfamily,keywordstyle=\color{blue},morekeywords={ReadBegin,ReadEnd,WriteBegin,WriteEnd,SyncStart,SyncStop,Free,foreach,in,var,consume,readable,r,isolated,writable,immutable}}
-
-%\lstset{
-%  xleftmargin=1em,      % <-- change this to a suitable length
-%  xrightmargin=-0.5cm,  % <-- change this to a suitable negative length
-%  numbersep=2pt,        % default 10pt
-%  numbers=left,
-%  numberstyle=\footnotesize,
-%  basicstyle=\ttfamily,
-%  breaklines=true}
-\usepackage{lipsum} % for dummy text
-\usepackage{comment}
-
-\renewcommand{\and}{\\}
-
-\usepackage{microtype}%if unwanted, comment out or use option "draft"
-%\bibliographystyle{plainurl}% the recommended bibstyle
-
-%\bibliography{references/paper}
+\newcommand{\blackdiamond}{\rotatebox[origin=c]{45}{$\vcenter{\hbox{$\scriptscriptstyle\blacksquare$}}$}}
+
+\newAuthor{Ismail}{blue}
+\newAuthor{Colin}{cyan}
+
+%% The "title" command has an optional parameter,
+\usepackage{amsthm}
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
+%%
+%% end of the preamble, start of the body of the document source.
 \begin{document}
-%
-\title{Modal Verification Patterns for Systems}
-%
-%\titlerunning{RCU Types}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-%\author{Ismail Kuru \orcidID{0000-0002-5796-2150} \and
-%Colin S. Gordon \orcidID{0000-0002-9012-4490}}
-%%%Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
-%
-%\authorrunning{Kuru and Gordon}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-%\institute{Drexel University \\ %\and
-%%%Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-%\email{\{ik335,csgordon\}@drexel.edu}}%\\
-%%%\url{http://www.springer.com/gp/computer-science/lncs} \and
-%%%ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-%%%\email{\{abc,lncs\}@uni-heidelberg.de}}
-%
-\maketitle              % typeset the header of the contribution
-
-%\input{chapters/squeezing.tex}
+
+%%
+%% The "title" command has an optional parameter,
+%% allowing the author to define a "short title" to be used in page headers.
+\title{On Verification Patterns for Low-Level Systems via Modal Abstractions}
+
+%%
+%% The "author" command and its associated commands are used to define
+%% the authors and their affiliations.
+%% Of note is the shared affiliation of the first two authors, and the
+%% "authornote" and "authornotemark" commands
+%% used to denote shared contribution to the research.
+% Commentout
+%\author{Ben Trovato}
+%\authornote{Both authors contributed equally to this research.}
+%\email{trovato@corporation.com}
+%\orcid{1234-5678-9012}
+%\author{G.K.M. Tobin}
+%\authornotemark[1]
+%\email{webmaster@marysville-ohio.com}
+%\affiliation{%
+ % \institution{Institute for Clarity in Documentation}
+ % \city{Dublin}
+ % \state{Ohio}
+ % \country{USA}
+%}
+
+%%
+%% By default, the full list of authors will be used in the page
+%% headers. Often, this list is too long, and will overlap
+%% other information printed in the page headers. This command allows
+%% the author to define a more concise list
+%% of authors' names for this purpose.
+\renewcommand{\shortauthors}{}
+
+%%
+%% The abstract is a short summary of the work to be presented in the
+%% article.
 \begin{abstract}
   Although they differ in the functionality they offer, low-level systems exhibit certain patterns of design and utilization of computing resources.
-\begin{comment}
-  In this paper, we argue the position that \emph{modalities}, in the sense of modal logic, 
-  should be a go-to approach when specifying and verifying low-level systems code.
-  We explain how the concept of a \emph{resource context} helps guide the design of new modalities
-  for verification of systems code, and
-  we justify our perspective by discussing prior systems that have used modalities for systems verification
-  successfully, arguing that they fit into the verification design pattern we articulate,
-  and explaining how this approach might apply to other systems verification challenges.
-\end{comment}
-  % \todo[inline]{Colin says: how about this framing instead of the second paragraph above:}
 
@@
 %% information and builds the first part of the formatted document.
-%\maketitle
-
+\maketitle
 \section{Introduction}
@@
 
-\paragraph{Contributions} We argue that modal abstractions can be used to identify and abstract system verification challenges.
+\textit{Contributions} We argue that modal abstractions can be used to identify and abstract system verification challenges.
 We justify our perspective by discussing prior systems that have successfully used modalities for system verification, arguing that they fit into the verification design pattern we articulate, and explaining how this approach might apply to other systems' verification challenges.
@@
 \item \textbf{Identifying System Verification Challenges}: We start with identifying common patterns in system verification: \emph{virtualization, sharing, and translation}.
-\item \textbf{Introducing Modal \emph{Resource Contexts}}: Then we discuss the concept of \emph{resource} which has already been an essential concept in the design of systems.
+\item \textbf{Introducing Modal \emph{Resource Contexts}}: Then we discuss the concept of \emph{resource}, which has already been an essential concept in the design of systems.
 Inspired by the concept of resource in the systems, we define what a resource and its context are in the modal abstractions.
@@
 \label{sec:modalbackground}
-
-\paragraph{Dynamic Logics as Program Logics}
+\textit{Dynamic Logics as Program Logics}
 We believe that for the kinds of informal reasoning and the sorts of data structures discussed in the previous section, ideas from modal logic are a promising approach to formalization and verification.
-Broadly speaking modal logics incorporate \emph{modal operators}, which take as arguments a proposition expected to be true in another time~\cite{pnueli1977temporal}, place~\cite{gordon2019modal,goranko1996hierarchies,areces2001hybrid,gargov1993modal,murphy2004symmetric}, circumstance or point-of-view~\cite{hintikka1962knowledge,halpern1985guide}, and result in a proposition true in the \emph{current} time, place, or circumstance at which the truth of the use of the modal operator is being evaluated.
+Broadly speaking, modal logics incorporate \emph{modal operators}, which take as arguments a proposition expected to be true in another time~\cite{pnueli1977temporal}, place~\cite{gordon2019modal,goranko1996hierarchies,areces2001hybrid,gargov1993modal,murphy2004symmetric}, circumstance or point-of-view~\cite{hintikka1962knowledge,halpern1985guide}, and result in a proposition true in the \emph{current} time, place, or circumstance at which the truth of the use of the modal operator is being evaluated.
 Classic examples include modal necessity $\square P$ describing that $P$ is \emph{necessarily} true, $\mathsf{G}~P$ meaning $P$ is true \emph{globally} (i.e., forever from this time onwards in temporal logics), or $K_i(P)$ describing that a particular participant $i$ \emph{knows} that $P$ is true.
@@
 
-\emph{Dynamic logic}~\cite{pratt1976semantical}, a (multi)modal logic of weakest preconditions~\cite{dijkstra-75} which works with modalities of the form $[p](P)$, which states that \emph{in the current program state}, \emph{if} program $p$ is run then afterwards $P$ will hold (modulo non-termination).
+\emph{Dynamic logic}~\cite{pratt1976semantical} is a (multi)modal logic of weakest preconditions~\cite{dijkstra-75} which works with modalities of the form $[p](P)$, which states that \emph{in the current program state}, \emph{if} program $p$ is run then afterwards $P$ will hold (modulo non-termination).
 This same idea is used to encode Hoare triples in the Iris program logic~\cite{jung2018iris}, using the same encoding as in Pratt's original presentation, where a Hoare triple $\{P\}C\{Q\}$ is encoded as $P\rightarrow[C](Q)$.
@@
 
-\paragraph{Hybrid Logic} Hybrid logics~\cite{blackburn1995hybrid,goranko1996hierarchies,gargov1993modal,areces2001hybrid,brauner2010hybrid} are a class of modal logics that add two primary new concepts to a logic.
-Nominals $\l \in \textsf{Loc}$ uniquely identify points in a model (e.g., a particular state), so there is exactly one point in the model where a nominal $\loc$ is true.
+\textit{Hybrid Logic} Hybrid logics~\cite{blackburn1995hybrid,goranko1996hierarchies,gargov1993modal,areces2001hybrid,brauner2010hybrid} are a class of modal logics that add two primary new concepts to a logic.
+Nominals $\ell \in \textsf{Loc}$ uniquely identify points in a model (e.g., a particular state), so there is exactly one point in the model where a nominal $\loc$ is true.
 Satisfaction operators are modal operators indexed by nominals, which enable claims about the truth of another proposition at some arbitrary point in the model identified by a nominal.
@@
 However, most prior work on hybrid logics works in classical logics rather than a substructural setting like separation logic.
-Newer work surveyed here adapts this idea to deal with different structures specified in
+Newer work ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2024modalabstractionsvirtualizingmemory,amalreal2024}  surveyed here adapts this idea to deal with different structures specified in
 separation logic.
@@
 %We continue to motivate the development with aspects of the write amplification process described in the previous section; Section \ref{sec:evaluation} provides further examples using the modal abstractions in proofs.
-
 % \paragraph{Temporal Operators}
@@
 
-\paragraph{Nominalization}
+\textit{Nominalization}
 Some classes of assertions benefit specifically from \emph{naming} the explicit conditions where they are true (as opposed to simply requiring them to be true \emph{somewhere} or \emph{everywhere} as in the most classic modal operators).
@@
 involving facts about specific named entities in the system.
-\\
 
@@
 
-\paragraph{Access by Translation}
+\textit{Access by Translation}
 Accessing virtualized resources via translation is a common way to virtualize notions of location (e.g.,
@@
 kernels relies \emph{translating} (\emph{mapping}) virtual addresses into physical ones. 
-The conceptual address-translation map is implemented with processor's page-table trees as shown in 
-Figure \ref{fig:pagetables}. In this Figure, we see a typical 4 levels of page tables, a virtual address on the 
-top indexing into the different levels of the page-table tree (the larger rectangles along the bottom
-are blocks of memory which serve as nodes of the tree along a particular lookup path). 
+The conceptual address-translation map is implemented with processor's page-table trees. A virtual address indexing into the different levels of the page-table tree 
+are blocks of memory which serve as nodes of the tree along a particular lookup path. 
 Fabrication of virtual addresses out of limited physical addresses is provided with entries 
 in the page-table tree where each points to a physical address aligned to 4KB (4096 byte) boundaries. 
-An address translation requires traversal of a series of tables starting from level 4, and it is 
-conventionally referred as a \emph{page-table walk}. 
+An address translation requires traversal of a series of tables starting from top level, and it is  conventionally referred as a \emph{page-table walk}. 
 The traversal ends in the level 1 page table with the final lookup to the actual page of physical memory holding the requested data, and the low-order 12 bits being used to index into this page.
-\begin{figure}
-    \includegraphics[width=\columnwidth]{pagetables2.png}
-    \caption{x86-64 page table lookups.}
-    \label{fig:pagetables}
-\end{figure}
+%\begin{figure}
+%    \includegraphics[width=\columnwidth]{pagetables2.png}
+%    \caption{x86-64 page table lookups.}
+%    \label{fig:pagetables}
+%\end{figure}
 
@@
 A typical OS manages multiple address spaces at once.
-Each program is associated with a unique page table tree, which is stored in a specific hardware register (\texttt{cr3} in Figure \ref{fig:pagetables}). 
+Each program is associated with a unique page table tree, which is stored in a specific hardware register (e.g., \texttt{cr3}). 
 Using different mappings, which map only disjoint portions of physical memory in the map's codomain, 
@@
 Another important computer resource computer resource that is \emph{virtualized} for access is disk blocks. 
-Filesystems are software components that abstract disk access on behalf of normal programs
-Similar to how VMMs using page-table trees to fabricate more memory addresses than what actually exists, 
-filesystems uses indexing maps for fabricating more \emph{virtual disk-page} addresses which are conventionally called \emph{disk-pages} 
-to be distinguished from physical \emph{disk-block} addresses.
+Filesystems are software components that abstract disk access on behalf of normal programs.
+Similarly to how VMMs use page-table trees to fabricate more memory addresses than actually exist, filesystems use indexing maps to fabricate more virtual disk-page addresses (which are conventionally called disk-pages) to be distinguished from physical disk-block addresses.
 When a regular program accesses a file, it requests a specific logical offset from the start of that file's contents.
@@
 % \paragraph{Overview of Policies in Filesystems}
-Filesystems not only handle the address-translation from a disk-page to a disk-block but also need to handle different
-\emph{modes} of computing where to find disk resources: a disk block may live on a disk in stable storage,
+Filesystems not only handle the address-translation from a disk-page to a disk-block but also need to handle different modes of computing, representing where to find disk resources: a disk block may live on a disk in stable storage,
 or the most up-to-date version of a block may live only in memory (for the filesystem to write to disk at a later time, batched for efficiency).
@@
 leaves the memory layout \emph{ideally} in a form of DAG (directed acyclic graph) with a unique root, which resembles the first stack pointer referencing the allocated memory.
-DAG abstraction instead of recursive decent tree allows non-duplication on the reference counts for the identical memory objects.
+DAG abstraction instead of recursive descent tree allows non-duplication on the reference counts for the identical memory objects.
 % To ease picturing the memory layout after a set of requests from the client code (for allocation, referencing etc.), we can think of the unique roots of DAGs as 
@@
 
-\paragraph{Shallow Modalities in the Iris Separation Logic}
+\textit{Shallow Modalities in the Iris Separation Logic}
 \label{sec:modal_in_iris}
 Classically modal operators are given direct semantics in terms of an underlying model, 
-which for \textsf{Iris}~\cite{jung2018iris} would be a collection of \textsc{CMRA}s ---
+which for \textsf{Iris}~\cite{jung2018iris} would be a collection of \textsc{CMRA}s (complete metric resource algebras) ---
 an algebraic cousin of a BI-algebra~\cite{ohearn1999bunched}.
 This is the case for \textsf{Iris}'s one built-in modality for step-indexing~\cite{nakano-00}.
-However in general, many modalities can be given proof-theoretic semantics directly within a logic. 
+However, in general, many modalities can be given proof-theoretic semantics directly within a logic. 
 This is the case for \textsf{Iris}'s weakest precondition modality (which is defined within the \textsf{Iris} 
@@
 \emph{pointwise lifting} of the algebra. A number of 
-projects for for systems software verification now take advantage of this,
+projects for systems software verification ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2024modalabstractionsvirtualizingmemory,amalreal2024} now take advantage of this,
 explicitly or implicitly lifting the BI-algebra structure from \textsf{Iris} itself to \textsf{Iris} predicates 
@@
 \label{sec:definitions}
-\begin{sidewaystable*}[t]
+\begin{table*}[t]\scriptsize
 \centering
@@
 %and the $\Diamond$ Figure for Post-Crash Modality uses notationis taken from NextGen Modality paper ~\cite{larsnextgen25}.
-\end{sidewaystable*}
+\end{table*}
 
@@
 We discuss a range of examples for each of these in turn.
-
 Table \ref{table:decomposition} gives additional examples of systems and their corresponding resource elements and contexts where these elements reside, though none of the work in that table analyzes itself according to the structure we are giving.
@@
 
-\paragraph{Stack Regions} When reasoning about stack frame contents, the resource element would be a stack-memory points-to assertion ( $\fbox{n} \; \ell \mapsto v$)
+\textit{Stack Regions} When reasoning about stack frame contents, the resource element would be a stack-memory points-to assertion ( $\fbox{n} \; \ell \mapsto v$)
 indicating that a certain offset into stack region $n$ holds value $v$.
+
 %which is an instance of ownership assertion ($\textsf{Own(t(a))}$) for which a well-behaved transformation ($\textsf{t}$), e.g. $\textsf{ICut}^{n}$ algebra quantifying over valid stack region ids. 
-\paragraph{Virtual Memory} For virtual memory management, a \textit{virtual-points-to} ownership assertion pairing a virtual address ($\ell$) with data ($v$) in an address space is natural.
+\textit{Virtual Memory} For virtual memory management, a \textit{virtual-points-to} ownership assertion pairing a virtual address ($\ell$) with data ($v$) in an address space is natural.
 A process's address space with the root address \textsf{r} is an abstraction that is treated as a container for virtual address mappings, $\ell \mapsto v$;
 
-\paragraph{Weak-Memory} When considering weak memory models, we consider address-value mappings of thread-local views in C11 memory model~\cite{lahav2017repairing}).
+\textit{Weak-Memory} When considering weak memory models, we consider address-value mappings of thread-local views in C11 memory model~\cite{lahav2017repairing}).
 Modalities $\Delta_\pi(P)$ and $\nabla_\pi(P)$ with $P$ represent the memory address-value mappings that held before or will hold after certain memory fence operations by thread $\pi$.
 
-\paragraph{Reference Counting} When dealing with reference-counting APIs, we may care to specify reachability of memory nodes ($\ell \mapsto v$) in a certain context defined by a shared root address.
+\textit{Reference Counting} When dealing with reference-counting APIs, we may care to specify reachability of memory nodes ($\ell \mapsto v$) in a certain context defined by a shared root address.
 A shared memory address $\ell$ can be the root of the graph that can be a container of memory nodes ($\ell_1 \mapsto v$) that are reachable from the root $\ell$.
 
-\paragraph{Post-Crash}
+\textit{Post-Crash}
 %\Colin{Ismail, not sure how to describe this. We don't want to go into the weeds on the NextGen}
@@
 \subsection{Nominals for Systems Resource Contexts}
-Finally, another design point is the question of whether or not resource element assertions must explicitly track
-their corresponding context, or if they implicitly pick up their context from where they are used.
+Finally, another design point is the question of whether or not resource element assertions must explicitly track their corresponding contexts, or if they implicitly pick up their context from where they are used.
 
@@
 %\Colin{So far I like how this is shaping up as a kind of ``design recipe'' for approaching this style of verification!}
-
-\section{Conclusion}
+\textbf{Conclusion}
 \label{sec:conclusion}
 % \todo[inline]{this definitely needs adjusting for both audience and emphasis on this being a survey}
-We have argued that there is a systematic pattern emerging in the use of modalities in the verification of low-level systems software,
+A systematic pattern emerging in the use of modalities in the verification of low-level systems software,
 and explained how a range of modern work fits into this pattern.
-% We have argued the essential patterns that help to choose the modalities when working out how to specify different kinds of systems code based on recent successes.
-While we have focused our attention on public verification work carried out in Iris,
-it also captures (a slightly more organized version of) our own thinking in coming up with designs for specifying distributed systems~\cite{gordon2019modal}, 
-% virtual memory managers~\cite{kuru2024modalabstractionsvirtualizingmemory}, 
-and in ongoing work using different modalities for different parts of a copy-on-write filesystem (e.g., for assertions true in different snapshots). 
 So not only are the modalities observably present, but they are demonstrably useful
@@
 when working out how to specify different kinds of systems code.
-
-\bibliographystyle{splncs04}
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{software}
-\newpage
+
+%%
+%% If your work has an appendix, this is the place to put it.
 \appendix
@@
 \end{document}
-\grid
+\endinput
+%%
+%% End of file `sample-sigconf-authordraft.tex'.

### 2025-09-04 paper.tex
+
switch from VMM tech report to OOPSLA citation

@@
 % Commentout
-%\author{Ben Trovato}
+\author{Ismail Kuru}
 %\authornote{Both authors contributed equally to this research.}
-%\email{trovato@corporation.com}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
+\email{ik335@drexel.edu}
+\orcid{0000-0002-5796-2150}
+\affiliation{%
+  \institution{Drexel University}
+  \city{Philadelphia}
+  \state{PA}
+  \country{USA}
+}
+\author{Colin Gordon}
 %\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
-%\affiliation{%
- % \institution{Institute for Clarity in Documentation}
- % \city{Dublin}
- % \state{Ohio}
- % \country{USA}
-%}
+\email{csg63@drexel.edu}
+\orcid{0000-0002-9012-4490}
+\affiliation{%
+  \institution{Drexel University}
+  \city{Philadelphia}
+  \state{PA}
+  \country{USA}
+}
 
@@
 \item \textbf{Taxonomy of Current Modal Approaches in System Verification}: Based on these concepts defined, we summarize contemporary verification efforts using modal abstractions.
-We choose them from different domains, for example, reasoning about weak memory models and storage persistence ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2024modalabstractionsvirtualizingmemory,amalreal2024} because we would like to justify that our \emph{definitions} are not domain dependent.
+We choose them from different domains, for example, reasoning about weak memory models and storage persistence ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2025modalabstractionsvirtualizingmemory,amalreal2024} because we would like to justify that our \emph{definitions} are not domain dependent.
 \end{itemize}
@@
 However, most prior work on hybrid logics works in classical logics rather than a substructural setting like separation logic.
-Newer work ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2024modalabstractionsvirtualizingmemory,amalreal2024}  surveyed here adapts this idea to deal with different structures specified in
+Newer work ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2025modalabstractionsvirtualizingmemory,amalreal2024}  surveyed here adapts this idea to deal with different structures specified in
 separation logic.
@@
 \emph{pointwise lifting} of the algebra. A number of 
-projects for systems software verification ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2024modalabstractionsvirtualizingmemory,amalreal2024} now take advantage of this,
+projects for systems software verification ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2025modalabstractionsvirtualizingmemory,amalreal2024} now take advantage of this,
 explicitly or implicitly lifting the BI-algebra structure from \textsf{Iris} itself to \textsf{Iris} predicates 
@@
  Memory-Fence~\cite{fsl,fsl++,derekrustbelt20} &  $\triangle_{\pi}$ and $\triangledown_{\pi}$      &  $\ell \mapsto v$                                 &  Weak &  Fence Acquire and Release  \\
- Address-Space~\cite{kuru2024modalabstractionsvirtualizingmemory} &  [r]P     &   $\ell \mapsto v$ &  Weak                         &  Address-Space Switch  \\ 
+ Address-Space~\cite{kuru2025modalabstractionsvirtualizingmemory} &  [r]P     &   $\ell \mapsto v$ &  Weak                         &  Address-Space Switch  \\ 
  Ref-Count~\cite{amalreal2024}& @$_{\ell}$ P    &  $\ell_1 \mapsto v$ &  Weak  &  Allocating, Dropping and Sharing a Reference \\ \bottomrule
@@
 the unique memory address of the root page-table tree therefore, a virtual-address-points-to ownership assertion does not need to be aware (or nominalized with) of the nominal of the address-space which is the unique memory address of the address table.
-Logically, specifying an address-space switch ~\cite{kuru2024modalabstractionsvirtualizingmemory} from the address space with the root address $r2$ to the one with the root address $r1$
+Logically, specifying an address-space switch ~\cite{kuru2025modalabstractionsvirtualizingmemory} from the address space with the root address $r2$ to the one with the root address $r1$
 %\[

### 2025-09-11 paper.tex
sigpln
+

@@
 %%
-\documentclass[sigconf,10pt]{acmart}
+\documentclass[sigplan,10pt]{acmart}
 %%
@@
   Although they differ in the functionality they offer, low-level systems exhibit certain patterns of design and utilization of computing resources.
-
   In this paper we examine how modalities have emerged as a common structure in formal verification
@@
 % \paragraph{Conclusion} 
-Low-level systems software exhibits certain patterns in their designs, especially when interacting with computing resources. 
-Exploiting certain patterns while designing software has been an important field of study.
+Interacting with computing resources is the essential point which shapes the design of low-level systems software. 
+Exploiting these design choices has been an important field of study.
 This survey examines these common patterns in systems software, and their relationships
@@
 \item \textbf{Identifying System Verification Challenges}: We start with identifying common patterns in system verification: \emph{virtualization, sharing, and translation}.
-\item \textbf{Introducing Modal \emph{Resource Contexts}}: Then we discuss the concept of \emph{resource}, which has already been an essential concept in the design of systems.
+\item \textbf{Introducing Modal \emph{Resource Contexts}}: Then we discuss the concept of \emph{resource} which has already been an essential concept in the design of systems.
 Inspired by the concept of resource in the systems, we define what a resource and its context are in the modal abstractions.

### 2025-09-11 paper.tex
+

@@
 \item \textbf{Introducing Modal \emph{Resource Contexts}}: Then we discuss the concept of \emph{resource} which has already been an essential concept in the design of systems.
-Inspired by the concept of resource in the systems, we define what a resource and its context are in the modal abstractions.
+Inspired by the concept of resource in systems, we define what a resource and its context are in the modal abstractions.
+For example, the \emph{transaction} abstraction, which behaves like a container for in-memory dirty disk pages, can be considered as a resrouce context for any dirty in-memory disk page resource associated with it.
 \item \textbf{Introducing \emph{Nominals for System Resources}}:

### 2025-09-11 paper.tex
+

@@
 Systems software, in general, interfaces with an underlying computing substrate such that any software system at any higher level in the software stack can (at least indirectly) utilize the resources of the machine.
+
 The last layer of software before the hardware is naturally critical to the correctness of an overall system,
@@
 
-\section{Background on Modal Logics for Programs, and Low-Level Systems Software}
-
-This section briefly recalls related concepts from modal logic of particular relevance
-to our taxonomy, which should be familiar but benefit from fixing notation (Section \ref{sec:modalbackground}),
-and enough general background on the systems software concepts organized by our taxonomy
-for the connections to modalities to be clear (Section \ref{sec:background}).
-
 \subsection{Hybrid Logic, Dynamic Logic, and Nominals in Program Logics}
 \label{sec:modalbackground}
-\textit{Dynamic Logics as Program Logics}
+
+\paragraph{Dynamic Logics as Program Logics}
 We believe that for the kinds of informal reasoning and the sorts of data structures discussed in the previous section, ideas from modal logic are a promising approach to formalization and verification.
-Broadly speaking, modal logics incorporate \emph{modal operators}, which take as arguments a proposition expected to be true in another time~\cite{pnueli1977temporal}, place~\cite{gordon2019modal,goranko1996hierarchies,areces2001hybrid,gargov1993modal,murphy2004symmetric}, circumstance or point-of-view~\cite{hintikka1962knowledge,halpern1985guide}, and result in a proposition true in the \emph{current} time, place, or circumstance at which the truth of the use of the modal operator is being evaluated.
+Broadly speaking modal logics incorporate \emph{modal operators}, which take as arguments a proposition expected to be true in another time~\cite{pnueli1977temporal}, place~\cite{gordon2019modal,goranko1996hierarchies,areces2001hybrid,gargov1993modal,murphy2004symmetric}, circumstance or point-of-view~\cite{hintikka1962knowledge,halpern1985guide}, and result in a proposition true in the \emph{current} time, place, or circumstance at which the truth of the use of the modal operator is being evaluated.
 Classic examples include modal necessity $\square P$ describing that $P$ is \emph{necessarily} true, $\mathsf{G}~P$ meaning $P$ is true \emph{globally} (i.e., forever from this time onwards in temporal logics), or $K_i(P)$ describing that a particular participant $i$ \emph{knows} that $P$ is true.
@@
 
-\emph{Dynamic logic}~\cite{pratt1976semantical} is a (multi)modal logic of weakest preconditions~\cite{dijkstra-75} which works with modalities of the form $[p](P)$, which states that \emph{in the current program state}, \emph{if} program $p$ is run then afterwards $P$ will hold (modulo non-termination).
+\emph{Dynamic logic}~\cite{pratt1976semantical}, a (multi)modal logic of weakest preconditions~\cite{dijkstra-75} which works with modalities of the form $[p](P)$, which states that \emph{in the current program state}, \emph{if} program $p$ is run then afterwards $P$ will hold (modulo non-termination).
 This same idea is used to encode Hoare triples in the Iris program logic~\cite{jung2018iris}, using the same encoding as in Pratt's original presentation, where a Hoare triple $\{P\}C\{Q\}$ is encoded as $P\rightarrow[C](Q)$.
@@
 
-\textit{Hybrid Logic} Hybrid logics~\cite{blackburn1995hybrid,goranko1996hierarchies,gargov1993modal,areces2001hybrid,brauner2010hybrid} are a class of modal logics that add two primary new concepts to a logic.
+\paragraph{Hybrid Logic} Hybrid logics~\cite{blackburn1995hybrid,goranko1996hierarchies,gargov1993modal,areces2001hybrid,brauner2010hybrid} are a class of modal logics that add two primary new concepts to a logic.
 Nominals $\ell \in \textsf{Loc}$ uniquely identify points in a model (e.g., a particular state), so there is exactly one point in the model where a nominal $\loc$ is true.
 Satisfaction operators are modal operators indexed by nominals, which enable claims about the truth of another proposition at some arbitrary point in the model identified by a nominal.
-Traditionally $@_l\varphi$ asserts that $\varphi$ is true in the state identified by $l$.
-Conceptually, given our context, we might like $l$ to indicate truth in a particular data structure $l$.
+Traditionally $@_{\ell}\varphi$ asserts that $\varphi$ is true in the state identified by $\ell$.
+Conceptually, given our context, we might like $l$ to indicate truth in a particular data structure $\ell$.
 However, most prior work on hybrid logics works in classical logics rather than a substructural setting like separation logic.
-Newer work ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2025modalabstractionsvirtualizingmemory,amalreal2024}  surveyed here adapts this idea to deal with different structures specified in
+Newer work surveyed here adapts this idea to deal with different structures specified in
 separation logic.
@@
 %We continue to motivate the development with aspects of the write amplification process described in the previous section; Section \ref{sec:evaluation} provides further examples using the modal abstractions in proofs.
+
 % \paragraph{Temporal Operators}
@@
 
-\textit{Nominalization}
+\paragraph{Nominalization}
 Some classes of assertions benefit specifically from \emph{naming} the explicit conditions where they are true (as opposed to simply requiring them to be true \emph{somewhere} or \emph{everywhere} as in the most classic modal operators).
@@
 involving facts about specific named entities in the system.
+\\
 
@@
 because this follows from standard properties of modal operators as described above.
-In contrast, if verification instead used a custom assertion $\mathsf{minCommitIndex}(5)$ to represent the former knowledge,
-one would need to separately provide custom reasoning to conclude $\mathsf{minCommitIndex}(3)$.
+In contrast, if verification instead used a custom assertion \[ \mathsf{minCommitIndex}(5)\] to represent the former knowledge,
+one would need to separately provide custom reasoning to conclude \[ \mathsf{minCommitIndex}(3)\].
 As our survey shows, the value of the modal view (which grants both intuitive specifications and natural
@@
 
-\textit{Access by Translation}
+\paragraph{Access by Translation}
 Accessing virtualized resources via translation is a common way to virtualize notions of location (e.g.,
@@
 kernels relies \emph{translating} (\emph{mapping}) virtual addresses into physical ones. 
-The conceptual address-translation map is implemented with processor's page-table trees. A virtual address indexing into the different levels of the page-table tree 
-are blocks of memory which serve as nodes of the tree along a particular lookup path. 
+The conceptual address-translation map is implemented with processor's page-table trees as shown in 
+Figure \ref{fig:pagetables}. In this Figure, we see a typical 4 levels of page tables, a virtual address on the 
+top indexing into the different levels of the page-table tree (the larger rectangles along the bottom
+are blocks of memory which serve as nodes of the tree along a particular lookup path). 
 Fabrication of virtual addresses out of limited physical addresses is provided with entries 
 in the page-table tree where each points to a physical address aligned to 4KB (4096 byte) boundaries. 
-An address translation requires traversal of a series of tables starting from top level, and it is  conventionally referred as a \emph{page-table walk}. 
+An address translation requires traversal of a series of tables starting from level 4, and it is 
+conventionally referred as a \emph{page-table walk}. 
 The traversal ends in the level 1 page table with the final lookup to the actual page of physical memory holding the requested data, and the low-order 12 bits being used to index into this page.
-%\begin{figure}
-%    \includegraphics[width=\columnwidth]{pagetables2.png}
-%    \caption{x86-64 page table lookups.}
-%    \label{fig:pagetables}
-%\end{figure}
+\begin{figure}
+    \includegraphics[width=\columnwidth]{pagetables2.png}
+    \caption{x86-64 page table lookups.}
+    \label{fig:pagetables}
+\end{figure}
 
@@
 A typical OS manages multiple address spaces at once.
-Each program is associated with a unique page table tree, which is stored in a specific hardware register (e.g., \texttt{cr3}). 
+Each program is associated with a unique page table tree, which is stored in a specific hardware register (\texttt{cr3} in Figure \ref{fig:pagetables}). 
 Using different mappings, which map only disjoint portions of physical memory in the map's codomain, 
@@
 Another important computer resource computer resource that is \emph{virtualized} for access is disk blocks. 
-Filesystems are software components that abstract disk access on behalf of normal programs.
-Similarly to how VMMs use page-table trees to fabricate more memory addresses than actually exist, filesystems use indexing maps to fabricate more virtual disk-page addresses (which are conventionally called disk-pages) to be distinguished from physical disk-block addresses.
+Filesystems are software components that abstract disk access on behalf of normal programs
+Similar to how VMMs using page-table trees to fabricate more memory addresses than what actually exists, 
+filesystems uses indexing maps for fabricating more \emph{virtual disk-page} addresses which are conventionally called \emph{disk-pages} 
+to be distinguished from physical \emph{disk-block} addresses.
 When a regular program accesses a file, it requests a specific logical offset from the start of that file's contents.
@@
 % \paragraph{Overview of Policies in Filesystems}
-Filesystems not only handle the address-translation from a disk-page to a disk-block but also need to handle different modes of computing, representing where to find disk resources: a disk block may live on a disk in stable storage,
+Filesystems not only handle the address-translation from a disk-page to a disk-block but also need to handle different
+\emph{modes} of computing where to find disk resources: a disk block may live on a disk in stable storage,
 or the most up-to-date version of a block may live only in memory (for the filesystem to write to disk at a later time, batched for efficiency).
@@
 leaves the memory layout \emph{ideally} in a form of DAG (directed acyclic graph) with a unique root, which resembles the first stack pointer referencing the allocated memory.
-DAG abstraction instead of recursive descent tree allows non-duplication on the reference counts for the identical memory objects.
+DAG abstraction instead of recursive decent tree allows non-duplication on the reference counts for the identical memory objects.
 % To ease picturing the memory layout after a set of requests from the client code (for allocation, referencing etc.), we can think of the unique roots of DAGs as 
@@
 
-\textit{Shallow Modalities in the Iris Separation Logic}
+\paragraph{Shallow Modalities in the Iris Separation Logic}
 \label{sec:modal_in_iris}
 Classically modal operators are given direct semantics in terms of an underlying model, 
-which for \textsf{Iris}~\cite{jung2018iris} would be a collection of \textsc{CMRA}s (complete metric resource algebras) ---
+which for \textsf{Iris}~\cite{jung2018iris} would be a collection of \textsc{CMRA}s ---
 an algebraic cousin of a BI-algebra~\cite{ohearn1999bunched}.
 This is the case for \textsf{Iris}'s one built-in modality for step-indexing~\cite{nakano-00}.
-However, in general, many modalities can be given proof-theoretic semantics directly within a logic. 
+However in general, many modalities can be given proof-theoretic semantics directly within a logic. 
 This is the case for \textsf{Iris}'s weakest precondition modality (which is defined within the \textsf{Iris} 
@@
 \emph{pointwise lifting} of the algebra. A number of 
-projects for systems software verification ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2025modalabstractionsvirtualizingmemory,amalreal2024} now take advantage of this,
+projects for for systems software verification now take advantage of this,
 explicitly or implicitly lifting the BI-algebra structure from \textsf{Iris} itself to \textsf{Iris} predicates 
@@
 \label{sec:definitions}
-\begin{table*}[t]\scriptsize
+\begin{table*}[t]\small
 \centering
@@
 
-{\scriptsize *The StackRegion Modality is an instance of NextGen (called the Independence Modality in \cite{larsnextgen25})}.
+{*The StackRegion Modality is an instance of NextGen (called the Independence Modality in \cite{larsnextgen25})}.
 %and the $\Diamond$ Figure for Post-Crash Modality uses notationis taken from NextGen Modality paper ~\cite{larsnextgen25}.
@@
 We discuss a range of examples for each of these in turn.
+
 Table \ref{table:decomposition} gives additional examples of systems and their corresponding resource elements and contexts where these elements reside, though none of the work in that table analyzes itself according to the structure we are giving.
@@
 
-\textit{Stack Regions} When reasoning about stack frame contents, the resource element would be a stack-memory points-to assertion ( $\fbox{n} \; \ell \mapsto v$)
+\paragraph{Stack Regions} When reasoning about stack frame contents, the resource element would be a stack-memory points-to assertion ( $\fbox{n} \; \ell \mapsto v$)
 indicating that a certain offset into stack region $n$ holds value $v$. 
-
 %which is an instance of ownership assertion ($\textsf{Own(t(a))}$) for which a well-behaved transformation ($\textsf{t}$), e.g. $\textsf{ICut}^{n}$ algebra quantifying over valid stack region ids. 
-\textit{Virtual Memory} For virtual memory management, a \textit{virtual-points-to} ownership assertion pairing a virtual address ($\ell$) with data ($v$) in an address space is natural.
+\paragraph{Virtual Memory} For virtual memory management, a \textit{virtual-points-to} ownership assertion pairing a virtual address ($\ell$) with data ($v$) in an address space is natural.
 A process's address space with the root address \textsf{r} is an abstraction that is treated as a container for virtual address mappings, $\ell \mapsto v$;
 
-\textit{Weak-Memory} When considering weak memory models, we consider address-value mappings of thread-local views in C11 memory model~\cite{lahav2017repairing}).
+\paragraph{Weak-Memory} When considering weak memory models, we consider address-value mappings of thread-local views in C11 memory model~\cite{lahav2017repairing}).
 Modalities $\Delta_\pi(P)$ and $\nabla_\pi(P)$ with $P$ represent the memory address-value mappings that held before or will hold after certain memory fence operations by thread $\pi$.
 
-\textit{Reference Counting} When dealing with reference-counting APIs, we may care to specify reachability of memory nodes ($\ell \mapsto v$) in a certain context defined by a shared root address.
+\paragraph{Reference Counting} When dealing with reference-counting APIs, we may care to specify reachability of memory nodes ($\ell \mapsto v$) in a certain context defined by a shared root address.
 A shared memory address $\ell$ can be the root of the graph that can be a container of memory nodes ($\ell_1 \mapsto v$) that are reachable from the root $\ell$.
 
-\textit{Post-Crash}
+\paragraph{Post-Crash}
 %\Colin{Ismail, not sure how to describe this. We don't want to go into the weeds on the NextGen}
@@
 \subsection{Nominals for Systems Resource Contexts}
-Finally, another design point is the question of whether or not resource element assertions must explicitly track their corresponding contexts, or if they implicitly pick up their context from where they are used.
+Finally, another design point is the question of whether or not resource element assertions must explicitly track
+their corresponding context, or if they implicitly pick up their context from where they are used.
 
@@
 the unique memory address of the root page-table tree therefore, a virtual-address-points-to ownership assertion does not need to be aware (or nominalized with) of the nominal of the address-space which is the unique memory address of the address table.
-Logically, specifying an address-space switch ~\cite{kuru2025modalabstractionsvirtualizingmemory} from the address space with the root address $r2$ to the one with the root address $r1$
+Logically, specifying an address-space switch ~\cite{kuru2024modalabstractionsvirtualizingmemory} from the address space with the root address $r2$ to the one with the root address $r1$
 %\[
@@
 %\Colin{So far I like how this is shaping up as a kind of ``design recipe'' for approaching this style of verification!}
-\textbf{Conclusion}
+
+\section{Conclusion}
 \label{sec:conclusion}
 % \todo[inline]{this definitely needs adjusting for both audience and emphasis on this being a survey}
-A systematic pattern emerging in the use of modalities in the verification of low-level systems software,
+We have argued that there is a systematic pattern emerging in the use of modalities in the verification of low-level systems software,
 and explained how a range of modern work fits into this pattern.
+% We have argued the essential patterns that help to choose the modalities when working out how to specify different kinds of systems code based on recent successes.
+While we have focused our attention on public verification work carried out in Iris,
+it also captures (a slightly more organized version of) our own thinking in coming up with designs for specifying distributed systems~\cite{gordon2019modal}, 
+% virtual memory managers~\cite{kuru2024modalabstractionsvirtualizingmemory}, 
+and in ongoing work using different modalities for different parts of a copy-on-write filesystem (e.g., for assertions true in different snapshots). 
 So not only are the modalities observably present, but they are demonstrably useful
@@
 \bibliography{software}
-
 %%
@@
 \appendix
-
 \end{document}
-\endinput
+%\endinput
 %%

### 2025-09-11 paper.tex
+

@@
 
+\section{Background on Modal Logics for Programs, and Low-Level Systems Software}
+
+This section briefly recalls related concepts from modal logic of particular relevance
+to our taxonomy, which should be familiar but benefit from fixing notation (Section \ref{sec:modalbackground}),
+and enough general background on the systems software concepts organized by our taxonomy
+for the connections to modalities to be clear (Section \ref{sec:background}).
+
 \subsection{Hybrid Logic, Dynamic Logic, and Nominals in Program Logics}
@@
 Traditionally $@_{\ell}\varphi$ asserts that $\varphi$ is true in the state identified by $\ell$.
-Conceptually, given our context, we might like $l$ to indicate truth in a particular data structure $\ell$.
+Conceptually, given our context, we might like $l$ to indicate truth in a particular data structure $l$.
 However, most prior work on hybrid logics works in classical logics rather than a substructural setting like separation logic.
@@
 because this follows from standard properties of modal operators as described above.
-In contrast, if verification instead used a custom assertion \[ \mathsf{minCommitIndex}(5)\] to represent the former knowledge,
-one would need to separately provide custom reasoning to conclude \[ \mathsf{minCommitIndex}(3)\].
+In contrast, if verification instead used a custom assertion \[\mathsf{minCommitIndex}(5) \] to represent the former knowledge,
+one would need to separately provide custom reasoning to conclude \[\mathsf{minCommitIndex}(3)\].
 As our survey shows, the value of the modal view (which grants both intuitive specifications and natural

### 2025-09-11 paper.tex
+

@@
 Traditionally $@_{\ell}\varphi$ asserts that $\varphi$ is true in the state identified by $\ell$.
-Conceptually, given our context, we might like $l$ to indicate truth in a particular data structure $l$.
+Conceptually, given our context, we might like $l$ to indicate truth in a particular data structure $\ell$.
 However, most prior work on hybrid logics works in classical logics rather than a substructural setting like separation logic.
@@
 (for example, a Raft node may ``know'' a lower bound on the commit index).
-Alternatively, a modality $@_i(P)$ may represent that $P$ is true of/at the specific node $i$~\cite{gordon2019modal} (e.g., that node $i$ has stored a certain
+Alternatively, a modality $@_i(P)$ may represent that $P$ is true of/at the specific node $i$~\cite{gordon2019modal} (e.g., that node $i$ has physically stored a certain
 piece of data to reliable storage).
@@
 
-In each of these cases, the fact that these facts are described using modalities with the core modal property $P \Rightarrow Q$, then $M(P) \Rightarrow M(Q)$ means, for example, that if a
+In each of these cases, the fact that these facts are described using modalities with the core modal property that if $P \Rightarrow Q$, then $M(P) \Rightarrow M(Q)$.
+This means, for example, that if a
 process $p$ knows that the commit index is greater than 5 (e.g., $\mathcal{K}_p(\mathsf{commitIndex} > 5)$)
@@
 kernels relies \emph{translating} (\emph{mapping}) virtual addresses into physical ones. 
-The conceptual address-translation map is implemented with processor's page-table trees as shown in 
+The conceptual address-translation map is implemented with processor's page table trees as shown in 
 Figure \ref{fig:pagetables}. In this Figure, we see a typical 4 levels of page tables, a virtual address on the 
-top indexing into the different levels of the page-table tree (the larger rectangles along the bottom
+top indexing into the different levels of the page table tree (the larger rectangles along the bottom
 are blocks of memory which serve as nodes of the tree along a particular lookup path). 
 Fabrication of virtual addresses out of limited physical addresses is provided with entries 
-in the page-table tree where each points to a physical address aligned to 4KB (4096 byte) boundaries. 
+in the page table tree where each points to a physical address aligned to 4KB (4096 byte) boundaries. 
 An address translation requires traversal of a series of tables starting from level 4, and it is 
-conventionally referred as a \emph{page-table walk}. 
+conventionally referred as a \emph{page table walk}. 
 The traversal ends in the level 1 page table with the final lookup to the actual page of physical memory holding the requested data, and the low-order 12 bits being used to index into this page.
@@
 Filesystems are software components that abstract disk access on behalf of normal programs
-Similar to how VMMs using page-table trees to fabricate more memory addresses than what actually exists, 
+Similar to how VMMs using page table trees to fabricate more memory addresses than what actually exists, 
 filesystems uses indexing maps for fabricating more \emph{virtual disk-page} addresses which are conventionally called \emph{disk-pages} 

### 2025-09-12 paper.tex
+

@@
 kernels relies \emph{translating} (\emph{mapping}) virtual addresses into physical ones. 
-The conceptual address-translation map is implemented with processor's page table trees as shown in 
+The conceptual address translation map is implemented with processor's page table trees as shown in 
 Figure \ref{fig:pagetables}. In this Figure, we see a typical 4 levels of page tables, a virtual address on the 
@@
 
-% \paragraph{Virtual Memory Managers(VMM) with Multiple Address-Spaces}
+% \paragraph{Virtual Memory Managers(VMM) with Multiple Address Spaces}
 % \label{sec:backgroundonvmm}
 % \todo{change the text}
-% An intuition for the abstraction that \emph{contains} a page-table tree is called an \emph{address-space} of a process whose virtual addresses to physical addresses. 
+% An intuition for the abstraction that \emph{contains} a page-table tree is called an \emph{address space} of a process whose virtual addresses to physical addresses. 
 A typical OS manages multiple address spaces at once.
@@
 % \paragraph{Overview of Policies in Filesystems}
-Filesystems not only handle the address-translation from a disk-page to a disk-block but also need to handle different
+Filesystems not only handle the address translation from a disk-page to a disk-block but also need to handle different
 \emph{modes} of computing where to find disk resources: a disk block may live on a disk in stable storage,
@@
 To handle the consistency of the disk resources in the existence of different modes of execution, filesystems employ different policies when indexing the disk resource.
-Which data-structure is going to be used to implement address-translation from disk-pages to disk-blocks? 
-What is going to be the \emph{update-policy} to the indexing map implementing address-translation which defines the \emph{consistency-policy} for the disk resources? 
+Which data-structure is going to be used to implement address translation from disk-pages to disk-blocks? 
+What is going to be the \emph{update policy} to the indexing map implementing address translation which defines the \emph{consistency policy} for the disk resources? 
 Different filesystem implementations give different answers to these questions.
@@
 % \paragraph{Examples to Different Policies on Navigating on Consistent Views of Disk}
-% Like VMMs navigating between different address-spaces, a \textsf{CoW} filesystem switches from either the old to the new root yields consistent views of the filesystem at different points in time, each commonly referred to as a \emph{snapshot}.
-A \textsf{CoW} filesystems treat the snapshot abstraction as VMMs treat the address-space abstraction, container of the relevant resources.
+% Like VMMs navigating between different address spaces, a \textsf{CoW} filesystem switches from either the old to the new root yields consistent views of the filesystem at different points in time, each commonly referred to as a \emph{snapshot}.
+A \textsf{CoW} filesystems treat the snapshot abstraction as VMMs treat the address space abstraction, container of the relevant resources.
 Like VMMs loading a process's memory mappings to be the current view of the memory by setting the control register to show the unique root address of the page-table tree of the process, the filesystem to atomically switch between consistent on-disk states, by updating the oldest so-called \emph{super} chunk at a fixed location on disk.
@@
 the specification of these systems is complex. And because some states are only possible after a system failure,
-each line of filesystem code has two conceptual post-conditions: one for continued execution, and one for the stable
-state established in the event of a hardware failure immediately after that operation.
+each line of filesystem code has two conceptual post-conditions: one referring to the successfully stored disk pages from a disk page container (e.g, a disk tree) , which can be uniquely identified by the last successful commit identifer (e.g., a transaction identifier or snapshot identifier), as the current view of the disk in the specification for continued execution, and one referring to the previous view of the disk in the specification for the stable state established in the event of a hardware failure immediately after that operation.
 
@@
  Memory-Fence~\cite{fsl,fsl++,derekrustbelt20} &  $\triangle_{\pi}$ and $\triangledown_{\pi}$      &  $\ell \mapsto v$                                 &  Weak &  Fence Acquire and Release  \\
- Address-Space~\cite{kuru2025modalabstractionsvirtualizingmemory} &  [r]P     &   $\ell \mapsto v$ &  Weak                         &  Address-Space Switch  \\ 
+ Address Space~\cite{kuru2025modalabstractionsvirtualizingmemory} &  [r]P     &   $\ell \mapsto v$ &  Weak                         &  Address Space Switch  \\ 
  Ref-Count~\cite{amalreal2024}& @$_{\ell}$ P    &  $\ell_1 \mapsto v$ &  Weak  &  Allocating, Dropping and Sharing a Reference \\ \bottomrule
@@
 \subsection{Resource}
-Consider first the address-space abstraction in an OS kernel.
-An address-space of a process is a \textsf{container} of \textsf{virtual} addresses referencing data in memory.
+Consider first the address space abstraction in an OS kernel.
+An address space of a process is a \textsf{container} of \textsf{virtual} addresses referencing data in memory.
 One would expect to have \textit{points-to} assertion from separation logic to specify \textit{ownership} of a memory reference pointing to some data.
@@
 %\paragraph{Weak Context Nominalization} Unlike the resources we discussed so far, some resource elements are \emph{agnostic} to the resource context they belong.
-That means the resource context needs a nominalization as it needs to be distinguished from other resource contexts when certain actions in the operational semantics are taken (e.g., address-space switch in Table \ref{table:decomposition}), but the resource elements of the context are unaware of this context nominalization, and they may even not be labeled with any other nominal as well.
+That means the resource context needs a nominalization as it needs to be distinguished from other resource contexts when certain actions in the operational semantics are taken (e.g., address space switch in Table \ref{table:decomposition}), but the resource elements of the context are unaware of this context nominalization, and they may even not be labeled with any other nominal as well.
 For example, virtual memory references in an address space do not know which address space they belong, but the address spaces are distinguished w.r.t.
-the unique memory address of the root page-table tree therefore, a virtual-address-points-to ownership assertion does not need to be aware (or nominalized with) of the nominal of the address-space which is the unique memory address of the address table.
-Logically, specifying an address-space switch ~\cite{kuru2024modalabstractionsvirtualizingmemory} from the address space with the root address $r2$ to the one with the root address $r1$
+the unique memory address of the root page-table tree therefore, a virtual-address-points-to ownership assertion does not need to be aware (or nominalized with) of the nominal of the address space which is the unique memory address of the address table.
+Logically, specifying an address space switch ~\cite{kuru2024modalabstractionsvirtualizingmemory} from the address space with the root address $r2$ to the one with the root address $r1$
 %\[
@@
 %\[P \triangleq \ell_1 \mapsto v ...\]
-%to be in the resource context of the address-space modality with the nominal $r1$, and, once the new address space is loaded --- i.e., cr3 is loaded with the other address-space root address ---, the resource elements ($Q$) that were loaded in the previous view of the memory are now introduced to the resource context with the nominal $r2$.
+%to be in the resource context of the address space modality with the nominal $r1$, and, once the new address space is loaded --- i.e., cr3 is loaded with the other address space root address ---, the resource elements ($Q$) that were loaded in the previous view of the memory are now introduced to the resource context with the nominal $r2$.
 Unlike what we discuss in the strong normalization (e.g.
 StackModality in Return specification) in which nominalized resource elements (stack points-to assertions) are guarded by Independence Modality ~\cite{larsnextgen25} (StackModality row in Table \ref{table:decomposition}) to preserve the soundness of the proof rule.
-However, with weak-nominalization, resource elements don't have nominals (Address-Space Modality Resource Element Column in Table \ref{table:decomposition}), and most importantly, address-space resource context switch, unlike shifted stack region in Return Specification, does not require any explicit independence check because all the resource elements at any time are either in one of two resource contexts or hosted in the ambient logic.
+However, with weak-nominalization, resource elements don't have nominals (Address Space Modality Resource Element Column in Table \ref{table:decomposition}), and most importantly, address space resource context switch, unlike shifted stack region in Return Specification, does not require any explicit independence check because all the resource elements at any time are either in one of two resource contexts or hosted in the ambient logic.
 
@@
 %\end{definition}
-%The address-space switch updating the cr3 register that holds the nominal value for address-space modality's resource contexts and the resource ambiance's (Hoare Triple indexed with resource context nominal) is an example of updating a resource context in a resource ambiance.
+%The address-space switch updating the cr3 register that holds the nominal value for address space modality's resource contexts and the resource ambiance's (Hoare Triple indexed with resource context nominal) is an example of updating a resource context in a resource ambiance.
 %\paragraph{Composition of Contingent Truth} Another example of weak-nominalization is Ref-Counter Modality (jump modality ($@_{l} \; P$ ~\cite{amalreal2024}) shown in Table \ref{table:decomposition}.
@@
 %Although it is almost always the case that when a modality is defined, classical proof rules such as introduction and elimination are also defined and proven.
-However, an interesting perspective emerges when we realize that, in fact, the proof rule for switching address spaces resembles an introduction (also elimination) rule for adding resource elements into an address-space modality's resource context.
+However, an interesting perspective emerges when we realize that, in fact, the proof rule for switching address spaces resembles an introduction (also elimination) rule for adding resource elements into an address space modality's resource context.
 %\Colin{So far I like how this is shaping up as a kind of ``design recipe'' for approaching this style of verification!}

### 2025-09-12 paper.tex
+

@@
 
-\subsubsection{Reference-Counting Memory Reclamation}
+\subsubsection{Reference Counting Memory Reclamation}
 In a typical configuration, allocating, deallocating and referencing memory locations for client code 
@@
 
-To reclaim the allocated memory location that is no longer being accessed, many systems use a well-known method called \emph{reference-counting} in which 
+To reclaim the allocated memory location that is no longer being accessed, many systems use a well-known method called \emph{reference counting} in which 
 each memory allocation includes not only its data but also a counter of how many pointers to that memory exist in the rest of the program.
@@
 
-\paragraph{Reference Counting} When dealing with reference-counting APIs, we may care to specify reachability of memory nodes ($\ell \mapsto v$) in a certain context defined by a shared root address.
+\paragraph{Reference Counting} When dealing with reference counting APIs, we may care to specify reachability of memory nodes ($\ell \mapsto v$) in a certain context defined by a shared root address.
 A shared memory address $\ell$ can be the root of the graph that can be a container of memory nodes ($\ell_1 \mapsto v$) that are reachable from the root $\ell$.

### 2025-09-12 paper.tex
+

@@
 \label{sec:modalbackground}
-
 \paragraph{Dynamic Logics as Program Logics}
@@
 \emph{Dynamic logic}~\cite{pratt1976semantical}, a (multi)modal logic of weakest preconditions~\cite{dijkstra-75} which works with modalities of the form $[p](P)$, which states that \emph{in the current program state}, \emph{if} program $p$ is run then afterwards $P$ will hold (modulo non-termination).
-This same idea is used to encode Hoare triples in the Iris program logic~\cite{jung2018iris}, using the same encoding as in Pratt's original presentation, where a Hoare triple $\{P\}C\{Q\}$ is encoded as $P\rightarrow[C](Q)$.
-Iris is now one of the most successful foundational program logics, and its rich support for
-extending the base logic with additional modalities~\cite{krebbers2017essence} means that
-modal logics, and in particular dynamic logic, now play an outsized role in program verification ---
-many challenging proofs of program correctness use a substructural dynamic logic containing
-many additional modalities which interact with the weakest precondition modality.
-
+%This same idea is used to encode Hoare triples in the Iris program logic~\cite{jung2018iris}, using the same encoding as in Pratt's original presentation, where a Hoare triple $\{P\}C\{Q\}$ is encoded as $P\rightarrow[C](Q)$.
+%Iris is now one of the most successful foundational program logics, and its rich support for
+%extending the base logic with additional modalities~\cite{krebbers2017essence} means that
+%modal logics, and in particular dynamic logic, now play an outsized role in program verification ---
+%many challenging proofs of program correctness use a substructural dynamic logic containing
+%many additional modalities which interact with the weakest precondition modality.
 % A hallmark of a unary logical operator $M$ being a modality is if $M$ satisfies a property akin to $(\varphi\rightarrow\psi)\rightarrow M(\varphi)\rightarrow M(\psi)$, which roughly states that modus ponens holds under the modality.\footnote{Afficionados of modal logic will note that this property is not quite Axiom K (which requires the initial implication to also hold under $M$), but follows from K and a necessitation rule $\varphi\rightarrow M(\varphi)$.
 Non-necessitive modalities typically satisfy the weaker property we call out above.} This and related laws allow transporting general reasoning into the specific contexts or hypotheticals modeled by the modalities.
-
 \paragraph{Hybrid Logic} Hybrid logics~\cite{blackburn1995hybrid,goranko1996hierarchies,gargov1993modal,areces2001hybrid,brauner2010hybrid} are a class of modal logics that add two primary new concepts to a logic.
@@
 % distributed algorithms \cite{ongaro2014search,ailijiang2019wpaxos}.
-
 \paragraph{Nominalization}
@@
 is how the OS ensures memory isolation between processes. 
-
 \subsubsection{Virtualized Resources in Filesystems}
@@
 Allowing reuse before this could result in in-place updates: part of the new snapshot could attempt to store data in space used by the prior snapshot, allowing inconsistent on-disk structures.
-
 While exceptions exist~\cite{bonwick2003zettabyte,rodeh2013btrfs,Rosenblum1992LFS,hitz1994file},
@@
 each line of filesystem code has two conceptual post-conditions: one referring to the successfully stored disk pages from a disk page container (e.g, a disk tree) , which can be uniquely identified by the last successful commit identifer (e.g., a transaction identifier or snapshot identifier), as the current view of the disk in the specification for continued execution, and one referring to the previous view of the disk in the specification for the stable state established in the event of a hardware failure immediately after that operation.
-
 \subsubsection{Reference Counting Memory Reclamation}
@@
 stack frames.
-
 % \paragraph{Buffering} Buffering is a technique used in variety of systems (filesystem page caching, write buffer in multi-processor architectures, 
@@
 % policy to commit the in-memory file pages to the disk.   
-
 \subsubsection{Weak Memory Models}
@@
 memory transactions are complete.
-
 %\section{Resources in Systems Software}
 %\label{sec:systemsoft}
-
 %\subsection{Speculation in Consistency} \todo{speak about the weak-mem modalities in here.}
@@
 %%\Colin{Not sure what you want here?}
-
 \section{Contingency Decomposition of a System}
@@
 To explain our ideas in the general systems understanding, we briefly recap some of the background and themes our ideas build on again, casting them in a certain way to bring out the relevance of our philosophy.
+\subsection{Encoding Modalities in Program Logics}
 
-\paragraph{Shallow Modalities in the Iris Separation Logic}
-\label{sec:modal_in_iris}
-Classically modal operators are given direct semantics in terms of an underlying model, 
-which for \textsf{Iris}~\cite{jung2018iris} would be a collection of \textsc{CMRA}s ---
+\paragraph{Encoding Hoare Triples} With Pratt's original representation, we can encode where a Hoare triple $\{P\}C\{Q\}$ is encoded as $P\rightarrow[C](Q)$.
+However, this modal representation requires to be given a semantics which is in terms of an underlying model.
+
+On the other hand the underlying model for \textsf{Iris}~\cite{jung2018iris} would be a collection of \textsc{CMRA}s ---
 an algebraic cousin of a BI-algebra~\cite{ohearn1999bunched}.
@@
 This is the case for \textsf{Iris}'s weakest precondition modality (which is defined within the \textsf{Iris} 
-logic itself), as well as for a wide range of other modalities in the 
+logic itself) , as well as for a wide range of other modalities in the 
 literature~\cite{restall2002introduction}. These modalities are typically multi-modal, 
 with a form of roughly $M~\overline{ctxt}~\phi\triangleq\ldots\rightarrow \phi~\overline{ctxt}$. 
+\paragraph{Shallow Modalities in Iris Separation Logic}
+\label{sec:modal_in_iris}
 In a substructural setting~\cite{restall1993modalities,krebbers2017essence} the implication $\rightarrow$ 
@@
 in a way well-suited to many use cases. 
-If $B$ is a type with a BI-algebra structure, then any space of predicate functions (e.g., $\mathsf{val}\rightarrow B$ as a simplification) also carries a BI-algebra structure, lifted from $B$, often called the
-\emph{pointwise lifting} of the algebra. A number of 
+If $B$ is a type with a BI-algebra structure, then any space of predicate functions (e.g., $\mathsf{val}\rightarrow B$ as a simplification) also carries a BI-algebra structure, lifted from $B$, often called the \emph{pointwise lifting} of the algebra.
+A number of
 projects for for systems software verification now take advantage of this,
@@
 This allows modalities to simultaneously represent ownership of certain resources used for interpreting the modality itself, as well as whatever resource ownership is implied by the modal argument.
-
 \subsection{Decomposing a System into its Constituents \emph{Contingently}}
@@
 \paragraph{Reference Counting} When dealing with reference counting APIs, we may care to specify reachability of memory nodes ($\ell \mapsto v$) in a certain context defined by a shared root address.
-A shared memory address $\ell$ can be the root of the graph that can be a container of memory nodes ($\ell_1 \mapsto v$) that are reachable from the root $\ell$.
+A shared memory address $\ell$ can be the root of the graph ($@_{\ell}\; ( \ell_1 \mapsto v_1 \;\ast\; ...
+\; \ast \; \ell_j \mapsto v_j )$) that can be a container of memory nodes ($\ell_1 \mapsto v$) that are reachable from the root $\ell$.
 
@@
 are at issue should prefer weak nominalization.
-
 % \section{Universal Properties over Resource Contexts with Nominals:  An Filesystem Recursive Trees Structure}
@@
 $\varphi_{\textsf{amplified}}$, $\varphi_{\textsf{untouched}}$): a subtree bounded by the cursor node.
+
 %$\paragraph{Strong Nominalization} Some resource elements require strong \emph{association} with their contexts as \emph{up-to} proper identification of resource context.

### 2025-09-12 paper.tex
+

@@
 \subsection{Encoding Modalities in Program Logics}
-
+Program logics which would like to enjoy modal propositions need to have an underlying model that allows encoding of the propositions.
+Since .
+Kripke model is the model all program logics directly or indirectly, sits on.
+Kripke model can be roughly seen as a triple of set of worlds (e.g., program states), a binary relation between worlds (e.g., transition relations between states) and a state interpretation function.
 \paragraph{Encoding Hoare Triples} With Pratt's original representation, we can encode where a Hoare triple $\{P\}C\{Q\}$ is encoded as $P\rightarrow[C](Q)$.
@@
 
-On the other hand the underlying model for \textsf{Iris}~\cite{jung2018iris} would be a collection of \textsc{CMRA}s ---
+Conventionally, a Kripke model with its constituent must be picked.
+Gordon give a Kripke model for an actor-based hybrid modal logic ~\cite{} in which the transition relation is indexed by a particular command and a particular actor.
+Wagner et.al.
+takes the similar approach for the design of their program logic for reference counted memory.
+
+The underlying model for \textsf{Iris}~\cite{jung2018iris} would be a collection of \textsc{CMRA}s ---
 an algebraic cousin of a BI-algebra~\cite{ohearn1999bunched}.

### 2025-09-12 paper.tex
+

@@
 Conventionally, a Kripke model with its constituent must be picked.
-Gordon give a Kripke model for an actor-based hybrid modal logic ~\cite{} in which the transition relation is indexed by a particular command and a particular actor.
+Gordon give a Kripke model for an actor-based hybrid modal logic ~\cite{gordon2019modal} in which the transition relation is indexed by a particular command and a particular actor.
 Wagner et.al.
@@
 
-The underlying model for \textsf{Iris}~\cite{jung2018iris} would be a collection of \textsc{CMRA}s ---
-an algebraic cousin of a BI-algebra~\cite{ohearn1999bunched}.
+On the other hand, program logics such as \textsf{Iris}~\cite{jung2018iris} utilizes Kripke model in ordering the separation algebras (a collection of \textsc{CMRA}s as an algebraic cousin of a BI-algebra~\cite{ohearn1999bunched}) both by heap extension and step index.
 This is the case for \textsf{Iris}'s one built-in modality for step-indexing~\cite{nakano-00}.
@@
 with a form of roughly $M~\overline{ctxt}~\phi\triangleq\ldots\rightarrow \phi~\overline{ctxt}$.
-\paragraph{Shallow Modalities in Iris Separation Logic}
+
+\paragraph{Shallow Modalities in Separation Logic}
 \label{sec:modal_in_iris}

### 2025-09-12 paper.tex
+

@@
  Address Space~\cite{kuru2025modalabstractionsvirtualizingmemory} &  [r]P     &   $\ell \mapsto v$ &  Weak                         &  Address Space Switch  \\ 
- Ref-Count~\cite{amalreal2024}& @$_{\ell}$ P    &  $\ell_1 \mapsto v$ &  Weak  &  Allocating, Dropping and Sharing a Reference \\ \bottomrule
+ Ref-Count~\cite{amalreal2024}& @$_{\ell}$ P    &  $\ell_1 \mapsto v$ &  Weak  &  Allocating, Dropping and Sharing a Reference \\
+  Actor~\cite{gordon2019modal}& @$_{\ell}$ P    &  $\ell_1 \mapsto v$ &  Weak  &  Send Message \\\bottomrule
 \end{tabular}

### 2025-09-12 paper.tex
+

--- 2025-09-12 paper.tex
@@
  Address Space~\cite{kuru2025modalabstractionsvirtualizingmemory} &  [r]P     &   $\ell \mapsto v$ &  Weak                         &  Address Space Switch  \\ 
- Ref-Count~\cite{amalreal2024}& @$_{\ell}$ P    &  $\ell_1 \mapsto v$ &  Weak  &  Allocating, Dropping and Sharing a Reference \\ \bottomrule
+ Ref-Count~\cite{amalreal2024}& @$_{\ell}$ P    &  $\ell_1 \mapsto v$ &  Weak  &  Allocating, Dropping and Sharing a Reference \\
+  Actor~\cite{gordon2019modal}& [A(t)]P    &  $(t,x,y)$ &  Weak  &  Send Message \\\bottomrule
 \end{tabular}

### 2025-09-12 paper.tex
+

--- 2025-09-12 paper.tex
@@
  Address Space~\cite{kuru2025modalabstractionsvirtualizingmemory} &  [r]P     &   $\ell \mapsto v$ &  Weak                         &  Address Space Switch  \\ 
- Ref-Count~\cite{amalreal2024}& @$_{\ell}$ P    &  $\ell_1 \mapsto v$ &  Weak  &  Allocating, Dropping and Sharing a Reference \\ \bottomrule
+ Ref-Count~\cite{amalreal2024}& @$_{\ell}$ P    &  $\ell_1 \mapsto v$ &  Weak  &  Allocating, Dropping and Sharing a Reference \\
+  Actor~\cite{gordon2019modal}& [A(t)]P    &  $(t,x,y)$ &  Strong  &  Send Message \\\bottomrule
 \end{tabular}

### 2025-09-13 paper.tex
+

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{On Verification Patterns for Low-Level Systems via Modal Abstractions}
+\title{Modal Verification Patterns for Systems Software}
 
@@
 \item \textbf{Taxonomy of Current Modal Approaches in System Verification}: Based on these concepts defined, we summarize contemporary verification efforts using modal abstractions.
-We choose them from different domains, for example, reasoning about weak memory models and storage persistence ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2025modalabstractionsvirtualizingmemory,amalreal2024} because we would like to justify that our \emph{definitions} are not domain dependent.
+We choose them from different domains, for example, reasoning about weak memory models and storage persistence ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2025modalabstractionsvirtualizingmemory,amalreal2024,gordon2019modal} and using different programming logics such as Iris ~\cite{jung2018iris} and Dafny ~\cite{leino2010dafny} because we would like to justify that our \emph{definitions} are neither domain nor programming logic dependent.
 \end{itemize}
@@
 Conventionally, a Kripke model with its constituent must be picked.
-Gordon give a Kripke model for an actor-based hybrid modal logic ~\cite{gordon2019modal} in which the transition relation is indexed by a particular command and a particular actor.
+Gordon give a Kripke model for an actor-based hybrid modal logic ~\cite{gordon2019modal} in which the transition relation is indexed by a particular command and a particular actor which relates pre- and post-states of a particular actor referenced by a nominal executing the command.
+Gordon encodes this hybrid logic in Dafny program logic ~\cite{leino2010dafny} with the help of two state invariants and rely-guarantee assertions checking the stability of invariants.
 Wagner et.al.
-takes the similar approach for the design of their program logic for reference counted memory.
+takes the similar approach for the design of their separation logic and pick up a Kripke model for reference counted memory reasoning principles.
 
@@
 \; \ast \; \ell_j \mapsto v_j )$) that can be a container of memory nodes ($\ell_1 \mapsto v$) that are reachable from the root $\ell$.
-
+\paragraph{Actors} When dealing with reference counting APIs, we may care to specify reachability of memory nodes ($\ell \mapsto v$) in a certain context defined by a shared root address.
+A shared memory address $\ell$ can be the root of the graph ($@_{\ell}\; ( \ell_1 \mapsto v_1 \;\ast\; ...
+\; \ast \; \ell_j \mapsto v_j )$) that can be a container of memory nodes ($\ell_1 \mapsto v$) that are reachable from the root $\ell$.
 \paragraph{Post-Crash}

### 2025-09-13 paper.tex
+

@@
 % I reworked this to be less general, and more about just weak memory, as thats all the survey section discusses.
-Multi-core and multi-processor systems do not directly write every memory update back to memory (or even a coherent hardware cache)
+Multicore and multiprocessor systems do not directly write every memory update back to memory (or even a coherent hardware cache)
 individually, as this would be extremely slow. Instead, they effectively buffer reads and writes to and from memory
@@
 %%\Colin{Not sure what you want here?}
+\subsubsection{Actor Model} An alternative to deal with the shared mmemory concurrency complication, i.e., data races, is using actors ~\cite{actormodel} which appears as a way of sequentializing the concurrent interleaved execution with some dependency on the \emph{knowledge} about the other actors.
+
 \section{Contingency Decomposition of a System}

### 2025-09-13 paper.tex
+

@@
 %%\Colin{Not sure what you want here?}
-\subsubsection{Actor Model} An alternative to deal with the shared mmemory concurrency complication, i.e., data races, is using actors ~\cite{actormodel} which appears as a way of sequentializing the concurrent interleaved execution with some dependency on the \emph{knowledge} about the other actors.
+\subsubsection{Actor Model} An alternative approach to addressing the shared memory concurrency complication, i.e., data races, is to use actors ~\cite{actormodel} by eliminating shared mutable state and enforcing updates on exclusive local mutable state via exchanging immutable messages among processes.
+This exhibits itself as a way of sequentializing the concurrent interleaved execution with some dependency on the \emph{knowledge} about other actors.
+In a typical consensus algorithm, an operation is committed if nodes reach an agreement on it.
 
+Therefore, as expected, the specification of an actor in the current view would refer to the other actors' states and has to validate that the shared knowledge amongst actors is preserved ~\cite{gordon2019modal}.
 \section{Contingency Decomposition of a System}
@@
 \paragraph{Encoding Hoare Triples}  With Pratt's original representation, we can encode where a Hoare triple $\{P\}C\{Q\}$ is encoded as $P\rightarrow[C](Q)$.
-However, this modal representation requires to be given a semantics which is in terms of an underlying model.
+However, this modal representation requires a semantics, which is defined in terms of an underlying model.
 
 Conventionally, a Kripke model with its constituent must be picked.
-Gordon give a Kripke model for an actor-based hybrid modal logic ~\cite{gordon2019modal} in which the transition relation is indexed by a particular command and a particular actor which relates pre- and post-states of a particular actor referenced by a nominal executing the command.
+Gordon provides a Kripke model for an actor-based hybrid modal logic ~\cite{gordon2019modal}, in which the transition relation is indexed by a particular command and a particular actor, relating pre- and post-states of a specific actor referenced by a nominal executing the command.
 Gordon encodes this hybrid logic in Dafny program logic ~\cite{leino2010dafny} with the help of two state invariants and rely-guarantee assertions checking the stability of invariants.
-Wagner et.al.
-takes the similar approach for the design of their separation logic and pick up a Kripke model for reference counted memory reasoning principles.
+Wagner et al.
+take a similar approach to the design of their separation logic and utilize a Kripke model for reference-counted memory reasoning principles.
 

### 2025-09-13 paper.tex
+

@@
  Ref-Count~\cite{amalreal2024}& @$_{\ell}$ P    &  $\ell_1 \mapsto v$ &  Weak  &  Allocating, Dropping and Sharing a Reference \\
-  Actor~\cite{gordon2019modal}& [A(t)]P    &  $(t,x,y)$ &  Strong  &  Send Message \\\bottomrule
+  Actor~\cite{gordon2019modal}& @$_{t}$ P    &  $(t,x,y)$ &  Strong  &  Send Message \\\bottomrule
 \end{tabular}
@@
 Except for Post-Crash-Modality, one can think of the resource contexts in the first column in Table \ref{table:decomposition} as containers for the corresponding resource elements in the second column.
-
 \paragraph{Stack Regions} When reasoning about stack frame contents, the resource element would be a stack-memory points-to assertion ( $\fbox{n} \; \ell \mapsto v$)
@@
 %which is an instance of ownership assertion ($\textsf{Own(t(a))}$) for which a well-behaved transformation ($\textsf{t}$), e.g. $\textsf{ICut}^{n}$ algebra quantifying over valid stack region ids. 
-\paragraph{Virtual Memory} For virtual memory management, a \textit{virtual-points-to} ownership assertion pairing a virtual address ($\ell$) with data ($v$) in an address space is natural.
-A process's address space with the root address \textsf{r} is an abstraction that is treated as a container for virtual address mappings, $\ell \mapsto v$;
+\paragraph{Virtual Memory} For virtual memory management, a \textit{virtual-points-to}, $\ell \mapsto v$, is an ownership assertion pairing a virtual address ($\ell$) with data ($v$) in an address space, which is a resource context for virtual address mappings, $\ell \mapsto v$, and uniquely identified with a root address.
+Logically, specifying an address space switch ~\cite{kuru2025modalabstractionsvirtualizingmemory} from the address space with the root address $r2$ to the one with the root address $r1$
+\[
+\{[r1] P \ast Q \} cr3 := r1 \{P \ast [r2]Q\}_{\textsf{cr3}}
+\]
+requires having the resource elements loaded into the memory 
+\[P \triangleq \ell_1 \mapsto v ...\]
+to be in the resource context of the address space modality with the root $r1$, and, once the new address space is loaded—i.e., cr3 is loaded with the other address space root address—the resource elements ($Q$) that were loaded in the previous view of the memory are now introduced to the resource context with the root $r2$.
 

### 2025-09-13 paper.tex
reviews

@@
  StackRegion$^{*}$  ~\cite{larsnextgen25} &  $\overset{ICut^{n}}{\hookrightarrow} \; P$      &  $\fbox{n} \; \ell \mapsto v$                                 &  Strong &  Alloc and Return to/from stack\\
+ Actor~\cite{gordon2019modal}& @$_{t}$ P    &  $(t,x,y)$ &  Strong  &  Send Message \\
  Memory-Fence~\cite{fsl,fsl++,derekrustbelt20} &  $\triangle_{\pi}$ and $\triangledown_{\pi}$      &  $\ell \mapsto v$                                 &  Weak &  Fence Acquire and Release  \\
  Address Space~\cite{kuru2025modalabstractionsvirtualizingmemory} &  [r]P     &   $\ell \mapsto v$ &  Weak                         &  Address Space Switch  \\ 
- Ref-Count~\cite{amalreal2024}& @$_{\ell}$ P    &  $\ell_1 \mapsto v$ &  Weak  &  Allocating, Dropping and Sharing a Reference \\
-  Actor~\cite{gordon2019modal}& @$_{t}$ P    &  $(t,x,y)$ &  Strong  &  Send Message \\\bottomrule
+ Ref-Count~\cite{amalreal2024}& @$_{\ell}$ P    &  $\ell_1 \mapsto v$ &  Weak  &  Allocating, Dropping and Sharing a Reference \\ \bottomrule
 \end{tabular}
@@
 to be in the resource context of the address space modality with the root $r1$, and, once the new address space is loaded—i.e., cr3 is loaded with the other address space root address—the resource elements ($Q$) that were loaded in the previous view of the memory are now introduced to the resource context with the root $r2$.
-
 \paragraph{Weak-Memory} When considering weak memory models, we consider address-value mappings of thread-local views in C11 memory model~\cite{lahav2017repairing}).
 Modalities $\Delta_\pi(P)$ and $\nabla_\pi(P)$ with $P$ represent the memory address-value mappings that held before or will hold after certain memory fence operations by thread $\pi$.
-
 \paragraph{Reference Counting} When dealing with reference counting APIs, we may care to specify reachability of memory nodes ($\ell \mapsto v$) in a certain context defined by a shared root address.
@@
 \; \ast \; \ell_j \mapsto v_j )$) that can be a container of memory nodes ($\ell_1 \mapsto v$) that are reachable from the root $\ell$.
-\paragraph{Actors} When dealing with reference counting APIs, we may care to specify reachability of memory nodes ($\ell \mapsto v$) in a certain context defined by a shared root address.
-A shared memory address $\ell$ can be the root of the graph ($@_{\ell}\; ( \ell_1 \mapsto v_1 \;\ast\; ...
-\; \ast \; \ell_j \mapsto v_j )$) that can be a container of memory nodes ($\ell_1 \mapsto v$) that are reachable from the root $\ell$.
+\paragraph{Actors} When specifying the consistency of an actor in the current view,  we need to refer to other actors' state: $@_{t}(P)$ asserts that $P$ is true in the actor reference by a token identifying the message class.
+To be able to validate that the knowledge shared via immutable messages (resource elements), $(t,x,y)$, are preserved, the logic has to acknowledge the message class, $t$ both when referring to the messages and when referring to the other actors' state.
 \paragraph{Post-Crash}

### 2025-09-13 paper.tex
updates for administrativia for camera-ready

@@
 
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmlicensed}
-\copyrightyear{2018}
-\acmYear{2018}
-\acmDOI{XXXXXXX.XXXXXXX}
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[Conference acronym 'XX]{Make sure to enter the correct
-  conference title from your rights confirmation email}{June 03--05,
-  2018}{Woodstock, NY}
-%%
-%%  Uncomment \acmBooktitle if the title of the proceedings is different
-%%  from ``Proceedings of ...''!
-%%
-%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%%  June 03--05, 2018, Woodstock, NY}
-\acmISBN{978-1-4503-XXXX-X/2018/06}
-
+% %% Rights management information.  This information is sent to you
+% %% when you complete the rights form.  These commands have SAMPLE
+% %% values in them; it is your responsibility as an author to replace
+% %% the commands and values with those provided to you when you
+% %% complete the rights form.
+% \setcopyright{acmlicensed}
+% \copyrightyear{2018}
+% \acmYear{2018}
+% \acmDOI{XXXXXXX.XXXXXXX}
+% %% These commands are for a PROCEEDINGS abstract or paper.
+% \acmConference[Conference acronym 'XX]{Make sure to enter the correct
+%   conference title from your rights confirmation email}{June 03--05,
+%   2018}{Woodstock, NY}
+% %%
+% %%  Uncomment \acmBooktitle if the title of the proceedings is different
+% %%  from ``Proceedings of ...''!
+% %%
+% %%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
+% %%  June 03--05, 2018, Woodstock, NY}
+% \acmISBN{978-1-4503-XXXX-X/2018/06}
+\acmYear{2025}
+\copyrightyear{2025}
+\setcopyright{cc}
+\setcctype[4.0]{by}
+\acmConference[PLOS '25]{13th Workshop on Programming Languages and Operating Systems}{October 13--16, 2025}{Seoul, Republic of Korea}
+\acmBooktitle{13th Workshop on Programming Languages and Operating Systems (PLOS '25), October 13--16, 2025, Seoul, Republic of Korea}
+\acmDOI{10.1145/3764860.3768337}
 %%
@@
 % (there's a 'changes' package with similar functionality, but it has a habit of breaking when you use it with  the Define environment)
-\newcommand\newAuthor[2]{%
-        \ExpandArgs{c}\newcommand{#1}[1]{\todo[color=#2!10, bordercolor=#2, inline]{\textbf{#1:}~##1}}
-	\ExpandArgs{c}\newcommand{#1A}[1]{\textcolor{#2}{##1}}
-        \ExpandArgs{c}\newcommand{#1B}[1]{\colorbox{#2!30}{\textcolor{black}{##1}}}}
-\newtheorem*{remark}{Remark}
+% \newcommand\newAuthor[2]{%
+%         \ExpandArgs{c}\newcommand{#1}[1]{\todo[color=#2!10, bordercolor=#2, inline]{\textbf{#1:}~##1}}
+% 	\ExpandArgs{c}\newcommand{#1A}[1]{\textcolor{#2}{##1}}
+%         \ExpandArgs{c}\newcommand{#1B}[1]{\colorbox{#2!30}{\textcolor{black}{##1}}}}
+% \newtheorem*{remark}{Remark}
 \newcommand{\loc}{\textsf{Loc}}
@@
 
-\newAuthor{Ismail}{blue}
-\newAuthor{Colin}{cyan}
+% \newAuthor{Ismail}{blue}
+% \newAuthor{Colin}{cyan}
 
@@
 %% of authors' names for this purpose.
-\renewcommand{\shortauthors}{}
+\renewcommand{\shortauthors}{Kuru and Gordon}
 
@@
 \end{abstract}
+
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10003752.10003790.10003793</concept_id>
+<concept_desc>Theory of computation~Modal and temporal logics</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10003752.10003790.10002990</concept_id>
+<concept_desc>Theory of computation~Logic and verification</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10011007.10010940.10010941.10010949</concept_id>
+<concept_desc>Software and its engineering~Operating systems</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Theory of computation~Modal and temporal logics}
+\ccsdesc[500]{Theory of computation~Logic and verification}
+\ccsdesc[500]{Software and its engineering~Operating systems}
 
@@
 %%% the work being presented. Separate the keywords with commas.
-%\keywords{Do, Not, Us, This, Code, Put, the, Correct, Terms, for, Your, Paper}
+\keywords{separation logic, modal logic, verification, operating systems}
 %% A "teaser" image appears between the author and affiliation

### 2025-09-15 paper.tex
a couple live fixes during review

@@
 Interacting with computing resources is the essential point which shapes the design of low-level systems software. 
-Exploiting these design choices has been an important field of study.
+Exploiting similar design choices has been an important field of study.
 This survey examines these common patterns in systems software, and their relationships
@@
 However, most prior work on hybrid logics works in classical logics rather than a substructural setting like separation logic.
-Newer work surveyed here adapts this idea to deal with different structures specified in
+Newer work~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2025modalabstractionsvirtualizingmemory,amalreal2024}:w
+ surveyed here adapts this idea to deal with different structures specified in
 separation logic.
@@
 Program logics which would like to enjoy modal propositions need to have an underlying model that allows encoding of the propositions. 
-Since .
-Kripke model is the model all program logics directly or indirectly, sits on.
+Since . Kripke model is the model all program logics directly or indirectly, sits on. 
 Kripke model can be roughly seen as a triple of set of worlds (e.g., program states), a binary relation between worlds (e.g., transition relations between states) and a state interpretation function.
@@
 This is the case for \textsf{Iris}'s weakest precondition modality (which is defined within the \textsf{Iris} 
-logic itself) , as well as for a wide range of other modalities in the 
+logic itself), as well as for a wide range of other modalities in the 
 literature~\cite{restall2002introduction}. These modalities are typically multi-modal, 
@@
 A number of
-projects for for systems software verification now take advantage of this,
+projects for systems software verification~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2025modalabstractionsvirtualizingmemory,amalreal2024} now take advantage of this,
 explicitly or implicitly lifting the BI-algebra structure from \textsf{Iris} itself to \textsf{Iris} predicates 
